### 2020-03-09 paper/notes.tex
Adds a lab notes document

% This is an example of using latex for a paper/report of specified
% size/layout. It's useful if you want to provide a PDF that looks
% like it was made in a normal word processor.

% While writing, don't stop for errors
\nonstopmode

% Use the article doc class, with an 11 pt basic font size
\documentclass[11pt, a4paper]{article}

% Makes the main font Nimbus Roman, a Times New Roman lookalike:
%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
% OR use this for proper Times New Roman (from msttcorefonts package
% on Ubuntu). Use xelatex instead of pdflatex to compile:
\usepackage{fontspec}
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
\setmainfont{Times New Roman}

% Set margins
\usepackage[margin=2.5cm]{geometry}

% Multilingual support
\usepackage[english]{babel}

% Nice mathematics
\usepackage{amsmath}

% Left right harpoons for kinetic equations
\usepackage{mathtools}

% Control over maketitle
\usepackage{titling}

% Section styling
\usepackage{titlesec}

% Ability to use colour in text
\usepackage[usenames]{color}

% For the \degree symbol
\usepackage{gensymb}

% Allow includegraphics and nice wrapped figures
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage[outercaption]{sidecap}

% Set formats using titlesec
\titleformat*{\section}{\bfseries\rmfamily}
\titleformat*{\subsection}{\bfseries\itshape\rmfamily}

% thetitle is the number of the section. This sets the distance from
% the number to the section text.
\titlelabel{\thetitle.\hskip0.3em\relax}

% Set title spacing with titlesec, too.  The first {1.0ex plus .2ex
% minus .7ex} sets the spacing above the section title. The second
% {-1.0ex plus 0.2ex} sets the spacing the section title to the
% paragraph.
\titlespacing{\section}{0pc}{1.0ex plus .2ex minus .7ex}{-1.1ex plus 0.2ex}

%% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
%% End language alias trick

%% Any aliases here
\newcommand{\mb}[1]{\mathbf{#1}} % this won't work?
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\mycite}[1]{\cite{#1}}
\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
%% END aliases

% Custom font defs
% fontsize is \fontsize{fontsize}{linespacesize}
\def\authorListFont{\fontsize{11}{11} }
\def\corrAuthorFont{\fontsize{10}{10} }
\def\affiliationListFont{\fontsize{11}{11}\itshape }
\def\titleFont{\fontsize{14}{11} \bfseries }
\def\textFont{\fontsize{11}{11} }
\def\sectionHdrFont{\fontsize{11}{11}\bfseries}
\def\bibFont{\fontsize{10}{10} }
\def\captionFont{\fontsize{10}{10} }

% Caption font size to be small.
\usepackage[font=small,labelfont=bf]{caption}

% Make a dot for the dot product, call it vcdot for 'vector calculus
% dot'. Bigger than \cdot, smaller than \bullet.
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\def\firstAuthorLast{James}

% Affiliations
\def\Address{\\
\affiliationListFont Adaptive Behaviour Research Group, Department of Psychology,
  The University of Sheffield, Sheffield, UK \\
}

% The Corresponding Author should be marked with an asterisk. Provide
% the exact contact address (this time including street name and city
% zip code) and email of the corresponding author
\def\corrAuthor{Seb James}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\corrEmail{seb.james@sheffield.ac.uk}

% Figure out the font for the author list..
\def\Authors{\authorListFont Sebastian James\\[1 ex]  \Address \\
  \corrAuthorFont $^{*}$ Correspondence: \corrEmail}

% No page numbering please
\pagenumbering{gobble}

% A trick to get the bibliography to show up with 1. 2. etc in place
% of [1], [2] etc.:
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

% reduce separation between bibliography items if not using natbib:
\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

% Set correct font for bibliography (doesn't work yet)
%\renewcommand*{\bibfont}{\bibFont}

% No paragraph indenting to match the VPH format
\setlength{\parindent}{0pt}

% Skip a line after paragraphs
\setlength{\parskip}{0.5\baselineskip}
\onecolumn

% titling definitions
\pretitle{\begin{center}\titleFont}
\posttitle{\par\end{center}\vskip 0em}
\preauthor{ % Fonts are set within \Authors
        \vspace{-1.1cm} % Bring authors up towards title
        \begin{center}
        \begin{tabular}[t]{c}
}
\postauthor{\end{tabular}\par\end{center}}

% Define title, empty date and authors
\title {
  Modelling the Retinotectal projection - can competition provide a stopping mechanism?
}
\date{} % No date please
\author{\Authors}

%% END OF PREAMBLE

\begin{document}

\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
\maketitle

\vspace{-1.8cm} % HACK bring the introduction up towards the title. It
                % would be better to do this with titling in \maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

I want to see if the competition mechanism which we demonstrated can form the
murine barrel pattern~\cite{james_modelling_2020} can also help explain the
arrangement of the Retinotectal projection.

The Retinotectal (RT) projection is thought to use pairs of orthogonal
gradients in the retina to specify the eventual location of synapses made on
the surface of the tectum. Similar orthogonal gradients in the tectum provide
a coordinate system which allows the axons to match their prespecified
destination with an actual location on the tectum. Although the Ephrins have
been shown to provide the guidance which sends axons \e{towards} their
destination~\cite{feldheim_genetic_2000}, no mechanism has yet been discovered
which actually \e{halts} the axonal growth cones in the correct location.

The main difference which I envisage between the RT and whisker barrel model
systems is in the way competition is applied to axons growing towards the
destination tissue. In the barrel field, I (notionally) assigned to a `group'
of axons the same interaction parameter, if they originated from the same
thalamic barreloid. The Retiontectal system is, in contrast, continuous;
although I am hypothesizing that the RT axons obtain an interaction parameter
from their location in the retina, just as I did for axons originating in
different barreloids, I will allow each axon to have a unique interaction
parameter, and expect that this leads to the continuous arrangement of axons
expected in the tectum.

To test whether the competetive model (assuming it does reproduce the
retinotopic neural arrangement) is useful, I can consider the following
manipulations:

I'd like to make a comparison with a system in which the stopping criterion is
an absolute measurement of the expression level of the Ephrins in the
tectum. I plan to introduce noise to the measurement that the growth cone
makes of the Ephrin expression level. This will introduce stochasticity in the
movement of the growth cones. The competition mechanism will be supported if
it is more robust to noise than an absolute stopping mechanism.

I will also then need to examine whether the competition mechanism is robust
to the experimental manipulations of the sort which Sperry carried
out~\cite{sperry_chemoaffinity_1963,goodhill_retinotectal_1999,goodhill_development_2005}.

\section{Model}

The initial model is very similar to that in the Barrels paper. However,
instead of providing interacton parameters ($\gamma_{i,j}$) in a configuration
file, I allow the config to specify $N$ neurons originating in the retina; I
automatically arrange these in a radially symmetric pattern, and then
automatically obtain the $gamma$ parameters from the neuron locations within
the pattern.

%
% BIBLIOGRAPHY
%
\selectlanguage{English}
\bibliographystyle{abbrvnotitle}
\bibliography{RetinoTectal}

\end{document}

### 2020-03-10 paper/notes.tex
A few more notes

@@
 The initial model is very similar to that in the Barrels paper. However,
-instead of providing interacton parameters ($\gamma_{i,j}$) in a configuration
+instead of providing interaction parameters ($\gamma_{i,j}$) in a configuration
 file, I allow the config to specify $N$ neurons originating in the retina; I
@@
 the pattern.
+
+\subsection{Configurations}
+
+In the config directory, there are numerous basic configs, such as N12.json,
+N24.json and so on. I find I have to modify some of the parameters as N
+increases. Larger N requires smaller hexes to allow a suitable pattern to
+develop. Larger N also generally requires a smaller diffusion constant
+$D$. I'm experimenting with $\epsilon$, the competition parameter, which, if
+too high causes the retinotopy to degrade during the simulation.
 

### 2020-03-11 paper/notes.tex
Notes on noise

@@
 
-The initial model is very similar to that in the Barrels paper. However,
-instead of providing interaction parameters ($\gamma_{i,j}$) in a configuration
-file, I allow the config to specify $N$ neurons originating in the retina; I
-automatically arrange these in a radially symmetric pattern, and then
-automatically obtain the $gamma$ parameters from the neuron locations within
-the pattern.
+The initial model is derived from the one in the Barrels
+paper~\cite{james_modelling_2020} which is found
+in \code{sim/rd\_barrel.h}. The specialization of \code{RD\_Barrel} can be
+found in \code{sim/rd\_rettec.h}, with main code file \code{rettec.cpp}. Here,
+instead of providing interaction parameters ($\gamma_{i,j}$) in a
+configuration file, I use only the parameter $N$, which in this context is the
+number of neurons originating in the retina. I automatically arrange these $N$
+retinal neurons in a radially symmetric pattern. I then obtain the $gamma$
+parameters from the neuron locations within the pattern.
 
@@
 
+\subsection{Noise}
+
+I wish to examine the effect of noise on the arrangement of axons on the
+tectum that results from axon guidance. I assume that there is no noise in the
+locations of the retinal neurons which serve as the origin of the interaction
+parameters; this remains a symmetric, ordered pattern.
+
+The hypothesis which the model is based on is that the interaction parameters
+are derived from the positions of the neurons on the source tissue (the
+retina) and that there are signalling molecule gradients which provide the
+required coordinate axes. Thus, the first source of noise would be in the
+sampling of these gradients and hence on the interaction parameters which are
+carried with the efferent axons to the tectum. This noise would be expected to
+be fixed after the axons had grown away from the retina.
+
+On arriving at the tectum, axon growth cones must determine their direction of
+movement. The model encapsulates the quantitative hypotheses that growth cone
+movement is a combination of diffusion and guidance. The guidance is subject
+to sampling error and thus a second source of noise can be introduced to the
+model. This noise varies with time during the simulation and can be considered
+to have two contributions; one from variations in the true expression level of
+the signalling gradient and one from sampling error measuring it. We will
+combine these into a single error.
+
+The system equations are modified as follows. Starting from the branching equation
+
+\begin{equation} \label{eq:da}
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i,
+\end{equation}
+
+the term for the divergence of $\mb{J}$ can be written:
+%
+\begin{equation}
+  \label{eq:divJ}
+  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
+\end{equation}
+%
+In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
+definition of a time-independent $\mb{g}_i(\mb{x})$:
+%
+\begin{equation}
+\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
+\end{equation}
+%
+Here, we represent the time-dependence of the signalling molecule expression
+by noting that $\rho$ is now a function of time.  The error which the growth
+cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
+$\xi_\rho(t)$:
+%
+\begin{equation}
+\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
++ \xi_\rho(t) \big).
+\end{equation}
+%
+One problem with this representation is that the equations represent a
+population model. Within one spatial element of the descretized solution of
+the differential equations, the error, $\xi_\rho(t)$ can have only one value at
+time $t$; this value is applied to all the individual axon growth cones that
+are represented by that element. Nonetheless, it's a useful, and fairly simple
+way to represent the guidance error in the system. In practice, we ignore the
+variation in $\rho_j$ and write
+%
+\begin{equation}
+\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
++ \xi_\rho(t) \big).
+\end{equation}
+
+$\xi_\rho(t)$ will be included in the model as a function returning numbers sampled
+from a normal distribution with mean 0 and standard deviation $\sigma_\rho$.
+
+The systematic error in $\gamma_{i,j}$ will be applied at setup time. The
+symmetric gamma values will first be derived from the symmetric layout of the
+retinal neurons, then a normally distributed perturbation will be applied to
+each $\gamma_{i,j}$.
+
+For retinal neurons with two-dimensional coordinates $\mb{r}_i$, $\gamma$ is
+given (in the absence of noise) by
+$\gamma_{i,0} = G\,|\mb{r}_i| \mb{r}_i \cdot \hat{\mb{x}}$ and
+$\gamma_{i,1} = G\,|\mb{r}_i| \mb{r}_i \cdot \hat{\mb{y}}$, where
+$\hat{\mb{x}}$ and $\hat{\mb{y}}$ are unit vectors, and $G$ is simply a gain
+on $\gamma$. With noise, these become
+$\gamma_{i,0} = G\,\big(\xi_\gamma |\mb{r}_i| \mb{r}_i \cdot \hat{\mb{x}} \big)$ and
+$\gamma_{i,1} = G\,\big(\xi_\gamma |\mb{r}_i| \mb{r}_i \cdot \hat{\mb{y}} \big)$
+
+$\xi_\gamma$ is drawn from a normal distribution with mean 0 and standard
+deviation $\sigma_\rho$.
 %

### 2020-03-24 paper/notes.tex
Some recent notes on different manipulations for the models

@@
 \usepackage[outercaption]{sidecap}
+
+% Nice quotes
+\usepackage{csquotes}
 
@@
 deviation $\sigma_\rho$.
+
+\section{Results}
+
+\subsection{Wild type}
+
+The operation of the model without modifiction and with zero or low noise.
+
+\subsection{Manipulations}
+
+Reference \cite{goodhill_development_2005} (from which the headings and quotes
+below are taken) gives a list of features which a model of retinotectal
+development or regeneration should address.
+
+\subsubsection{Ectopic targeting}
+
+\begin{displayquote}
+\emph{Retinal axons entering the tectum via abnormal trajectories can still find
+their appropriate termination sites (Finlay et al. 1979b; Harris 1982,
+1984). This suggests the presence of vector signals throughout the tectum
+pointing axons towards their correct termination sites.}
+\end{displayquote}
+
+The orthogonal gradients in the current model which signal growth direction of
+axonal growth cones naturally provide the vector signals which ensure that
+regardless of entry point, growth cones tend towards their correct location.
+
+\subsubsection{Shifting connections}
+
+\begin{displayquote}
+\emph{In some fish and amphibians, the retina grows by addition of new neurons
+around the ciliary margin, while the tectum grows by addition of cells to its
+caudomedial edge (Gaze et al. 1974; Fraser 1983).The retinotectal map remains
+ordered throughout this time, indicating that the retinotectal projection is
+continually shifting caudally. This suggests that, at least in these species,
+either chemoaffinity gradients do not irreversibly specify the map or the
+gradients must evolve with time.}
+\end{displayquote}
+
+The current model should work under the condition that the underlying domain
+is increased in size, with the competition acting to ensure that the entire
+domain is filled and the ordering maintained.
+
+\subsubsection{Rotation experiments}
+
+\begin{displayquote}
+\emph{If a Xenopus presumptive tectum is rotated early enough during devel- opment,
+a map is formed that is normal relative to the whole animal, whereas later
+rotations lead to a rotated map (Chung \& Cooke 1978).}
+\end{displayquote}
+
+\subsubsection{Map compression and expansion}
+
+\begin{displayquote}
+\emph{The map formed after removal of half the retina initially covers half the
+tectum (Attardi \& Sperry 1963), but then gradually expands to fill the whole
+tectum (Schmidt et al. 1978). The axon terminal density remains the same
+(Schmidt et al.  1978). If the optic nerve is then made to regenerate again,
+an expanded map is immediately formed (Schmidt 1978). If half the tectum is
+ablated, the regenerated map is compressed into the remaining tectal space
+(Yoon 1971; Sharma 1972; Cook 1979; Finlay et al. 1979a).}
+\end{displayquote}
+
+\subsubsection{Translocation}
+
+\begin{displayquote}
+\emph{A}
+\end{displayquote}
+
+\subsubsection{Compound-eye experiments}
+
+\begin{displayquote}
+\emph{B}
+\end{displayquote}
+
 %

### 2020-03-31 paper/notes.tex
Tuesday's later afternoon notes

@@
 paper~\cite{james_modelling_2020} which is found
-in \code{sim/rd\_barrel.h}. The specialization of \code{RD\_Barrel} can be
-found in \code{sim/rd\_rettec.h}, with main code file \code{rettec.cpp}. Here,
-instead of providing interaction parameters ($\gamma_{i,j}$) in a
-configuration file, I use only the parameter $N$, which in this context is the
-number of neurons originating in the retina. I automatically arrange these $N$
-retinal neurons in a radially symmetric pattern. I then obtain the $gamma$
+in \code{sim/rd\_barrel.h}. The specialization of \code{RD\_Barrel} (and a
+sister class, \code{RetArrange}) can be found in \code{sim/rd\_rettec.h}, with
+main code file \code{rettec.cpp}. Here, instead of providing interaction
+parameters ($\gamma_{i,j}$) in a configuration file, I use only the parameter
+$N$, which in this context is the number of neurons originating in the
+retina. I automatically arrange these $N$ retinal neurons in a radially
+symmetric pattern (See \code{RetArrange}). I then obtain the $\gamma$
 parameters from the neuron locations within the pattern.
@@
 
+\subsection{Non-competition based stopping mechanism}
+
+Starting from the branching equation again (but dropping the competition term,
+$\chi_i$):
+
+\begin{equation} \label{eq:da_again}
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t},
+\end{equation}
+
+The spatial part of the system can be halted if $D\rightarrow
+0, \gamma_{i,j}\rightarrow 0$ as
+$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$. That is, if, for
+a particular axon, $i$, a sampling of $\rho_j$ is close to $f(\gamma_{i,j})$
+for each $j$, then $D$ is reduced to zero, and likewise, the second term also
+reduces to zero. $f()$ is probably $1-$ a (possibly noisy) normalized Gaussian
+function, or a capped Gaussian (with height capped by a sigmoid?), with mean
+$(\gamma_{i,1},\gamma_{i,2})$ and a width based on the differences between
+$\gamma_{i\pm 1,j}$. Something like:
+
+\begin{equation} \label{eq:da_stopping}
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f(\gamma_{i,1},\gamma_{i,2},\mb{x}) - \frac{\partial c_i}{\partial t},
+\end{equation}
+
+$f$ could be thought of as a functional:
+$f_i[\gamma_{i,1},\gamma_{i,2}](\mb{x})$ (if that's the right way to write
+it...).
+
+Let's try to write f out. See gaussians.m initially.
+
 \section{Results}
@@
 
-The operation of the model without modifiction and with zero or low noise.
+The operation of the model without modification and with zero or low noise.
 

### 2020-03-31 paper/gaussians.m

@@ -1,0 +1,13 @@
+x = -3:0.01:3;
+
+% Params
+m = 5        % multiplier on Gaussian; make sure it's >1
+sigma = 0.2   % Width of Gaussian
+sh = 3      % sharpness of sigmoid
+
+figure (1); clf; hold on;
+
+g = m .* exp (-0.5 .* power (((x-mu)./sigma), 2));
+plot (x,g);
+gg = 2 .* (-0.5+(1./(1 + exp(-g.*sh))));
+plot (x,1-gg)

### 2020-04-06 paper/notes.tex
Notes on the modelling of axon stopping.

@@
 One problem with this representation is that the equations represent a
-population model. Within one spatial element of the descretized solution of
+population model. Within one spatial element of the discretized solution of
 the differential equations, the error, $\xi_\rho(t)$ can have only one value at
@@
 
-Starting from the branching equation again (but dropping the competition term,
-$\chi_i$):
+Here, we seek to model the `pure' chemospecificity theory in which axons are
+guided by orthogonal morphogen gradients until they reach their destination,
+which they must be able to sense. The axons, guided by the \emph{gradient} of
+the morphogen concentration are assumed also to be able to sample
+the \emph{absolute} concentration of the signalling molecules.  A reduction in
+branching, and possibly an enhancement of synaptogenesis are triggered when
+the concentrations of the two orthogonal morphogens are simultaneously
+measured to be within the axon's `stopping' ranges.
+
+To find a way to implement this in a population model similar to the
+competitive model, we start from the branching equation (dropping the
+competition term, $\chi_i$):
 
@@
 
-The spatial part of the system can be halted if $D\rightarrow
-0, \gamma_{i,j}\rightarrow 0$ as
-$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$. That is, if, for
-a particular axon, $i$, a sampling of $\rho_j$ is close to $f(\gamma_{i,j})$
-for each $j$, then $D$ is reduced to zero, and likewise, the second term also
-reduces to zero. $f()$ is probably $1-$ a (possibly noisy) normalized Gaussian
-function, or a capped Gaussian (with height capped by a sigmoid?), with mean
-$(\gamma_{i,1},\gamma_{i,2})$ and a width based on the differences between
-$\gamma_{i\pm 1,j}$. Something like:
+Spatial transfer of information in the system can be halted if
+$D\rightarrow 0$ and $\gamma_{i,j}\rightarrow 0$.
+%$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$.
+
+This can be achieved with a function of the gammas and $\mb{x}$:
 
@@
 
-$f$ could be thought of as a functional:
-$f_i[\gamma_{i,1},\gamma_{i,2}](\mb{x})$ (if that's the right way to write
-it...).
+Where $f$ is a function which permits branching far away from a given axon's
+target region and tends to zero inside the target region. It represents an
+axon's internal `stopping process'. Start with a Gaussian function:
 
-Let's try to write f out. See gaussians.m initially.
+\begin{equation}\label{eq:g}
+g(\mb{x}) = \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 }
+\end{equation}
+
+where $\mb{\gamma}$ gives the location of the peak and $\sigma$ is the width
+of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:
+
+\begin{equation}\label{eq:f}
+f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s g(\mb{x})}} - 1
+\end{equation}
+
+In one equation:
+
+\begin{equation}\label{eq:gf}
+f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 } }} - 1
+\end{equation}
 
@@
 below are taken) gives a list of features which a model of retinotectal
-development or regeneration should address.
+dfevelopment or regeneration should address.
 

### 2020-04-06 paper/squashgauss.m

@@ -1,0 +1,16 @@
+x = -3:0.01:3;
+
+% Params
+%m = 15       % the sharpness/gauss height combined multiplier
+mu = 1       % mean of Gaussian - supplied by gamma_i,1 or gamma_i,2
+sigma = 0.2  % Width of Gaussian; a parameter
+
+figure (1); clf; hold on;
+
+g = exp (-0.5 .* power (((x-mu)./sigma), 2));
+f = 2 .* (-0.5+(1./(1 + exp(-g.*m))));
+
+plot (x,g,'-b');
+plot (x,1-f,'-r');
+lg = legend('Gaussian','Inverted squashed Gaussian')
+set (lg, 'Location', 'NorthWest')

### 2020-04-06 paper/notes.tex
f is a function of rho, not directly of x.
Further tweaking of the meaning of the vvarious symbols and of what f() actually is and represents

@@
 the signalling gradient and one from sampling error measuring it. We will
-combine these into a single error.
+combine these into a single error. \textbf{Or maybe not:} Because the
+no-competition model (see Section \ref{sec:nocomp}) has a function
+$f(\gamma, \rho)$, which represents the axons' ability to sample the absolute
+value of $\rho$, it is important to represent the variability of the true
+value of $\rho(\mb{x})$.
 
@@
 cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
-$\xi_\rho(t)$:
+$\xi_{\rho'}(t)$:
 %
@@
 \mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
-+ \xi_\rho(t) \big).
++ \xi_{\rho'}(t) \big).
 \end{equation}
@@
 \mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
-+ \xi_\rho(t) \big).
++ \xi_{\rho'}(t) \big).
 \end{equation}
@@
 \subsection{Non-competition based stopping mechanism}
+\label{sec:nocomp}
 
@@
 
-This can be achieved with a function of the gammas and $\mb{x}$:
+This can be achieved with a function of the gammas and rhos:
 
 \begin{equation} \label{eq:da_stopping}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f(\gamma_{i,1},\gamma_{i,2},\mb{x}) - \frac{\partial c_i}{\partial t},
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f\big(\gamma_{i,1},\gamma_{i,2},\tilde{\rho}_{1}(\mb{x}),\tilde{\rho}_{2}(\mb{x})\big) - \frac{\partial c_i}{\partial t},
 \end{equation}
@@
 target region and tends to zero inside the target region. It represents an
-axon's internal `stopping process'. Start with a Gaussian function:
+axon's internal `stopping process'. Note that $\tilde{\rho}_j(\mb{x})$ is the
+axon's measurement of $\rho_j(\mb{x})$ and may include measurement noise,
+$\xi_\rho$: $\tilde{\rho}_j = \rho_j + \xi_\rho$. Given that the morphogen
+gradients are orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector
+field $\vec{\rho}(\mb{x})$. Similarly, we can combine
+$(\gamma_{i,1},\gamma_{i,2})$ into the vector $\vec{\gamma}$.
+
+This makes it easy to start to build $f$ with a two dimensional Gaussian
+function:
 
 \begin{equation}\label{eq:g}
-g(\mb{x}) = \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 }
+g\big(\vec{\rho}(\mb{x})\big) = \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 }
 \end{equation}
 
-where $\mb{\gamma}$ gives the location of the peak and $\sigma$ is the width
+where $\vec{\gamma}$ gives the location of the peak and $w$ is the width
 of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:
@@
 \begin{equation}\label{eq:f}
-f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s g(\mb{x})}} - 1
+f\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1 + \mathrm{e}^{-s g(\vec{\rho}\,)}} - 1
 \end{equation}
@@
 \begin{equation}\label{eq:gf}
-f(\mb{x}) = \frac{2}{1 + \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\mb{x}-\mb{\gamma}}{2\sigma}\big)^2 } }} - 1
+f(\vec{\rho}(\mb{x})) = \frac{2}{1
++ \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }} - 1
 \end{equation}

### 2020-04-06 paper/notes.tex
More notation tweaks

@@
 population model. Within one spatial element of the discretized solution of
-the differential equations, the error, $\xi_\rho(t)$ can have only one value at
+the differential equations, the error, $\xi_{\rho'}(t)$ can have only one value at
 time $t$; this value is applied to all the individual axon growth cones that
@@
 
-$\xi_\rho(t)$ will be included in the model as a function returning numbers sampled
-from a normal distribution with mean 0 and standard deviation $\sigma_\rho$.
+$\xi_{\rho'}(t)$ will be included in the model as a function returning numbers sampled
+from a normal distribution with mean 0 and standard deviation $\sigma_{\rho'}$.
 
@@
 $\xi_\gamma$ is drawn from a normal distribution with mean 0 and standard
-deviation $\sigma_\rho$.
+deviation $\sigma_{\gamma}$.
 
@@
 \begin{equation} \label{eq:da_stopping}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f\big(\gamma_{i,1},\gamma_{i,2},\tilde{\rho}_{1}(\mb{x}),\tilde{\rho}_{2}(\mb{x})\big) - \frac{\partial c_i}{\partial t},
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f_i\big(\gamma_{i,1},\gamma_{i,2},\tilde{\rho}_{1}(\mb{x}),\tilde{\rho}_{2}(\mb{x})\big) - \frac{\partial c_i}{\partial t},
 \end{equation}
 
-Where $f$ is a function which permits branching far away from a given axon's
+Where $f_i$ is a function which permits branching far away from a given axon's
 target region and tends to zero inside the target region. It represents an
@@
 
-This makes it easy to start to build $f$ with a two dimensional Gaussian
+This makes it easy to start to build $f_i$ with a two dimensional Gaussian
 function:
@@
 \begin{equation}\label{eq:g}
-g\big(\vec{\rho}(\mb{x})\big) = \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 }
+g_i\big(\vec{\rho}(\mb{x})\big) = \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 }
 \end{equation}
 
-where $\vec{\gamma}$ gives the location of the peak and $w$ is the width
-of the peak. This is squashed through an inverting sigmoid, with sharpness $s$:
+where $\vec{\gamma}$ gives the location of the peak and the parameter $w$ is
+the width of the peak. This is squashed through an inverting sigmoid, with
+sharpness parameter $s$:
 
 \begin{equation}\label{eq:f}
-f\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1 + \mathrm{e}^{-s g(\vec{\rho}\,)}} - 1
+f_i\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1 + \mathrm{e}^{-s g_i(\vec{\rho}\,)}} - 1
 \end{equation}
@@
 \begin{equation}\label{eq:gf}
-f(\vec{\rho}(\mb{x})) = \frac{2}{1
+f_i\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1
 + \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }} - 1

### 2020-04-07 paper/squashgauss.m
Correctly invert

@@ -1,16 +1,17 @@
 x = -3:0.01:3;
 
 % Params
-%m = 15       % the sharpness/gauss height combined multiplier
+s = 15       % the sharpness/gauss height combined multiplier
 mu = 1       % mean of Gaussian - supplied by gamma_i,1 or gamma_i,2
-sigma = 0.2  % Width of Gaussian; a parameter
+w = 0.2      % Width of Gaussian; a parameter
 
 figure (1); clf; hold on;
 
-g = exp (-0.5 .* power (((x-mu)./sigma), 2));
-f = 2 .* (-0.5+(1./(1 + exp(-g.*m))));
-
+g = exp (- power (((x-mu)./(2.*w)), 2));
+f = 2 ./(1 + exp(-g.*s)) - 1;
+finv = 1 - f;
 plot (x,g,'-b');
-plot (x,1-f,'-r');
-lg = legend('Gaussian','Inverted squashed Gaussian')
+plot (x,f,'-g');
+plot (x,finv,'-r');
+lg = legend('Gaussian','Squashed Gaussian','Inverted squashed Gaussian')
 set (lg, 'Location', 'NorthWest')

### 2020-04-07 paper/notes.tex
Correctly invert
Have tweaked params to get fairly good arrangements of the max graph.

@@
 
-Spatial transfer of information in the system can be halted if
-$D\rightarrow 0$ and $\gamma_{i,j}\rightarrow 0$.
-%$f(\gamma_{i,1},\gamma_{i,2},...,\gamma{i,M})\rightarrow 1$.
-
-This can be achieved with a function of the gammas and rhos:
+Spatial transfer of information in the system can be halted if the flux of
+axonal branching (the term in brackets) tends to 0.  This can be achieved with
+a function of the gammas and rhos:
 
@@
 target region and tends to zero inside the target region. It represents an
-axon's internal `stopping process'. Note that $\tilde{\rho}_j(\mb{x})$ is the
-axon's measurement of $\rho_j(\mb{x})$ and may include measurement noise,
-$\xi_\rho$: $\tilde{\rho}_j = \rho_j + \xi_\rho$. Given that the morphogen
-gradients are orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector
-field $\vec{\rho}(\mb{x})$. Similarly, we can combine
-$(\gamma_{i,1},\gamma_{i,2})$ into the vector $\vec{\gamma}$.
+axon's internal `stopping process'. Note carefully that this internal axonal
+function does \emph{not} depend directly on space; it is a function of the
+interaction parameters $\gamma$ and of the signalling molecule density,
+$\rho$. Note that $\tilde{\rho}_j(\mb{x})$ is the axon's \emph{measurement} of
+$\rho_j(\mb{x})$ and may include measurement noise, $\xi_\rho$:
+$\tilde{\rho}_j = \rho_j + \xi_\rho$. Given that the morphogen gradients are
+orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector field
+$\vec{\rho}\,(\mb{x})$. Similarly, we can combine $(\gamma_{i,1},\gamma_{i,2})$
+into the vector $\vec{\gamma}$.
 
@@
 where $\vec{\gamma}$ gives the location of the peak and the parameter $w$ is
-the width of the peak. This is squashed through an inverting sigmoid, with
-sharpness parameter $s$:
+the width of the peak. This is squashed through a sigmoid, with
+sharpness parameter $s$ and inverted:
 
 \begin{equation}\label{eq:f}
-f_i\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1 + \mathrm{e}^{-s g_i(\vec{\rho}\,)}} - 1
+f_i\big(\vec{\rho}(\mb{x})\big) = 1 - \bigg( \frac{2}{1 + \mathrm{e}^{-s
+g_i(\vec{\rho}\,)}} - 1 \bigg)
 \end{equation}
@@
 \begin{equation}\label{eq:gf}
-f_i\big(\vec{\rho}(\mb{x})\big) = \frac{2}{1
-+ \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }} - 1
+f_i\big(\vec{\rho}(\mb{x})\big) = 2 - \frac{2}{1
++ \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }}
 \end{equation}
@@
 below are taken) gives a list of features which a model of retinotectal
-dfevelopment or regeneration should address.
+development or regeneration should address.
 
@@
 \begin{displayquote}
-\emph{If a Xenopus presumptive tectum is rotated early enough during devel- opment,
+\emph{If a Xenopus presumptive tectum is rotated early enough during development,
 a map is formed that is normal relative to the whole animal, whereas later

### 2020-04-07 sim/rd_retnocomp.h

@@ -1029,12 +1029,13 @@ public:
 
         // (vecrho - vecgamma)^2 = (vecrho - vecgamma) . (vecrho - vecgamma) = |vecrho-vecgamma|^2
         // for each hex:
+        Flt reduce_factor = 0.4;
         for (unsigned int i=0; i<this->N; ++i) {
 #pragma omp parallel for
             for (unsigned int h=0; h<this->nhex; ++h) {
                 // NB: If noise is to be included, replace this->rho with (this->rho + noise)
-                Flt vr_minus_vg_x = this->rho[0][h] - this->gamma[0][i];
-                Flt vr_minus_vg_y = this->rho[1][h] - this->gamma[1][i];
+                Flt vr_minus_vg_x = this->rho[0][h] - (reduce_factor*this->gamma[0][i]);
+                Flt vr_minus_vg_y = this->rho[1][h] - (reduce_factor*this->gamma[1][i]);
                 Flt vr_len_sq = vr_minus_vg_x*vr_minus_vg_x + vr_minus_vg_y*vr_minus_vg_y; // This is |vecrho-vecgamma|^2
                 Flt gausshump = exp (-vr_len_sq/this->two_w_sq);
                 this->f[i][h] = static_cast<Flt>(2.0) - (static_cast<Flt>(2.0)/(static_cast<Flt>(1.0) + exp (-s*gausshump)));

### 2020-04-07 sim/rettec.cpp

@@ -380,8 +380,8 @@ int main (int argc, char **argv)
         for (unsigned int i = 0; i<RD.N; ++i) {
             spatOff[0] = xzero + RD.hg->width() * (i/side);
             spatOff[1] = RD.hg->width() * (i%side);
-            //unsigned int idx = v1.addHexGridVisualMono (RD.hg, spatOff, RD.f[i], scaling2, (float)i/(float)RD.N);
-            unsigned int idx = v1.addHexGridVisual (RD.hg, spatOff, RD.f[i], scaling2);
+            unsigned int idx = v1.addHexGridVisualMono (RD.hg, spatOff, RD.f[i], scaling2, (float)i/(float)RD.N);
+            //unsigned int idx = v1.addHexGridVisual (RD.hg, spatOff, RD.f[i], scaling2);
             fgrids.push_back (idx);
         }
         xzero = spatOff[0] + RD.hg->width();

### 2020-04-08 paper/notes.tex
NOtes about the previous commmit

@@
 
+\subsection{Allowing the stopping mechanism to enhance connection making}
+
+As well as affecting branching, there is no reason why I couldn't have the
+axonal function also affect connection making. Consider the equation governing
+the rate of change of connections.
+
+\begin{equation} \label{eq:dc}
+\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
+\end{equation}
+
+If $\beta$ were to be modulated by $1 - f_i$, then connection making would
+occur much more strongly in the correct location. Would it be too good I
+wonder? Answer: Yes, it's highly effective. To reproduce, set fModulatesBeta
+to true, prop\_modulated to 1 and prop\_unmodulated to 0.
+
+\subsection{Allowing the stopping mechanism to reduce connection decay}
+
+An alternative to the above, is to slow down connection decay within the
+stopping zone.
+
 \section{Results}

### 2020-07-27 paper/notes.tex
A few notes, prior to actually creating configs for all these expts.

@@
 axonal growth cones naturally provide the vector signals which ensure that
-regardless of entry point, growth cones tend towards their correct location.
+regardless of entry point, growth cones tend towards their correct
+location. Can prove with different distributions of initial branching density
+($a(\mb{x},t=0)$).
 
@@
 
-\subsubsection{Rotation experiments}
+To do: implement a HexGrid whose boundary increases with time. That will mean
+some tampering with \code{morph::HexGrid}.
+
+\subsubsection{Rotation experiments)}
 
@@
 \end{displayquote}
+
+These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2B--D and can
+include retinal rotation as well as early/late tectal rotations.
+
+The model as it stands contains an assumption about the orientation of the
+guidance gradients. So rotating the tectum early enough means making no change
+(given that I have no model for \emph{creating} the guidance gradients).
 
@@
 
+I need to examine more carefully the timeline of these experiments.
+
+These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2E and 2F.
+
 \subsubsection{Translocation}
@@
 
+These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2I and 2J.
+
 \subsubsection{Compound-eye experiments}
@@
 \end{displayquote}
+
+This corresponds to \cite{goodhill_retinotectal_1999}, Fig 2H.
 

### 2020-08-06 paper/notes.tex
Adds code to limit a, to prevent simulation crashes.

@@
 On arriving at the tectum, axon growth cones must determine their direction of
-movement. The model encapsulates the quantitative hypotheses that growth cone
+movement. The model encapsulates the quantitative hypothesis that growth cone
 movement is a combination of diffusion and guidance. The guidance is subject
@@
 to have two contributions; one from variations in the true expression level of
-the signalling gradient and one from sampling error measuring it. We will
+the signalling gradient and one from sampling error measuring it. I will
 combine these into a single error. \textbf{Or maybe not:} Because the
@@
 value of $\rho$, it is important to represent the variability of the true
-value of $\rho(\mb{x})$.
+value of $\rho(\mb{x})$. \textbf{Update (July 2020):} Since I was last
+thinking about this, I added length- and amplitude-variable noise to $\rho$,
+so I will use that code here, and additionally consider noise in the sampling
+of $\rho$.
 
@@
 \begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t} + \chi_i,
+\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla
+a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) + \chi_i \right) - \frac{\partial c_i}{\partial t},
 \end{equation}
@@
   \label{eq:divJ}
-  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right).
+  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
+  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})+ \chi_i \right).
 \end{equation}
@@
 %
-Here, we represent the time-dependence of the signalling molecule expression
+Here, I represent the time-dependence of the signalling molecule expression
 by noting that $\rho$ is now a function of time.  The error which the growth
@@
 time $t$; this value is applied to all the individual axon growth cones that
-are represented by that element. Nonetheless, it's a useful, and fairly simple
-way to represent the guidance error in the system. In practice, we ignore the
+are represented by that element. Nonetheless, it's a useful (and fairly simple)
+way to represent the guidance error in the system. In practice, I ignore the
 variation in $\rho_j$ and write
@@
 
-Here, we seek to model the `pure' chemospecificity theory in which axons are
+Here, I seek to model the `pure' chemospecificity theory in which axons are
 guided by orthogonal morphogen gradients until they reach their destination,
@@
 To find a way to implement this in a population model similar to the
-competitive model, we start from the branching equation (dropping the
+competitive model, I start from the branching equation (dropping the
 competition term, $\chi_i$):
@@
 orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector field
-$\vec{\rho}\,(\mb{x})$. Similarly, we can combine $(\gamma_{i,1},\gamma_{i,2})$
-into the vector $\vec{\gamma}$.
+$\vec{\rho}\,(\mb{x})$. Similarly, $(\gamma_{i,1},\gamma_{i,2})$ can be
+combined to give the vector $\vec{\gamma}$.
 
@@
 
+\section{Implementation notes}
+
+\subsection{Stability}
+
+The model can become unstable due to the unbounded nature of $a$. Essentially,
+$a$ can blow up in the computation of $\dvrg J$.
+
+Stability with respect to $D$ is a particular issue, as is stability in the
+presence of noise.
+
+\subsection{Stability for small $D$}
+
+In the basic, $N=56$ wild type simulation, $D$ seems to need to be close to
+0.02. Unfortunately, the simulation is unstable for $\delta t=0.0001$ and
+$d=0.015$. Having fixed a bug which prevented me from reducing $d$ very much,
+and having implemented fixed seed random numbers for repeatable sims,
+I can now try to determine if a smaller $d$ will help the simujlation compute,
+though note that as $d$ is reduced, so $\delta t$ must be reduced, leading to
+longer simulation computations.
+
+Note: $\epsilon=1.2$ here in the following:
+
+D=0.02, dt=0.00005, d=0.007: Crash at step 13803.
+
+D=0.01, dt=0.0001, d=0.015: Crash at step 2629.
+
+D=0.015, dt=0.0001, d=0.015: Crash at step 4864.
+
+D=0.02, dt=0.0001, d=0.015: Crash at step 7280.
+
+D=0.025, dt=0.0001, d=0.015: Crash at step 9649.
+
+D=0.03, dt=0.0001, d=0.015: Crash at step 12394.
+
+This pattern is similar to that shown in Fig.~2 of the Barrels paper. To
+enable $D$ to be low, it's necessary to reduce $\epsilon$, BUT I want to
+explore increased competition, whilst maintaining low $D$.
+
+Thus, some sort of cap on $a$ may be called for.
+
+But, lets first see how much reduction of $\epsilon$ is required. Starting
+from D=0.01, dt=0.0001, d=0.015: Crash at step 2629, with $\epsilon=1.2$,
+reduce $\epsilon$.
+
+$\epsilon=0.8$: Crash step 4779.
+
+$\epsilon=0.4$: Crash step 9802.
+
+$\epsilon=0.1$: Computation makes it to step 24000 with no crash, but that's
+not enough competition for an even result.
+
+\subsection{Bounding $a$}
+
+I've already implemented systems in which $a$ is bounded. Simplest is probably
+best - a piecewise linear transfer function on $a$. Thus, I've added the code
+to rd_barrel.h to compute a transformed version of $a$ called $\bar{a}$ and
+use this in the computation. It mitigates the instabilities without
+making much change to the operation of the model. It adds an additional
+parameter to the model, $a_{max}$, which I've set to 0.5 by default.
+
+\subsection{Stability for noise}
+
+Write me.
+
 %

### 2020-08-06 sim/rd_barrel.h

@@ -94,6 +94,14 @@ public:
     //! divergence of \hat{a}_i(x,t).
     alignas(alignof(std::vector<Flt>)) std::vector<Flt> div_ahat;
 
+    //! \bar{a}_i - result of processing a_i through a sigmoid.
+    alignas(alignof(std::vector<Flt>))
+    std::vector<Flt> abar;
+
+    //! Gradient of \bar{a}
+    alignas(alignof(std::array<std::vector<Flt>, 2>))
+    std::array<std::vector<Flt>, 2> grad_abar;
+
     /*!
      * To hold div(g) / 3d, a static scalar field. There are M vectors of N of these
      * vectors of Flts
@@ -118,6 +126,9 @@ public:
 
     //! The power to which a_i(x,t) is raised in Eqs 1 and 2 in the paper.
     alignas(Flt) Flt k = 3.0;
+
+    //! Maximum allowed value for a_i at the end of \a integrate_a
+    alignas(Flt) Flt a_max = 0.5;
 
 protected:
     //! The diffusion parameter.
@@ -454,6 +465,9 @@ public:
         this->resize_gradient_field (this->grad_ahat);
         this->resize_vector_variable (this->div_ahat);
 
+        this->resize_vector_variable (this->abar);
+        this->resize_gradient_field (this->grad_abar);
+
         // rhomethod is a vector of size M
         this->rhoMethod.resize (this->M);
         for (unsigned int j=0; j<this->M; ++j) {
@@ -501,6 +515,9 @@ public:
         this->zero_vector_variable (this->ahat);
         this->zero_gradient_field (this->grad_ahat);
         this->zero_vector_variable (this->div_ahat);
+
+        this->zero_vector_variable (this->abar);
+        this->zero_gradient_field (this->grad_abar);
 
         // Configure random number generators
         if (this->fixedSeeds == true) {
@@ -855,15 +872,14 @@ public:
     //! The normalization/transfer function.
     virtual inline Flt transfer_a (const Flt& _a, const unsigned int _i)
     {
-#ifdef NORMALIZE_TO_ONE
-        // Divisive normalization to one
-        Flt a_rtn = this->nhex * _a / this->sum_a[_i];
-#else
-        // Divisive norm with initial sum multiplier
-        Flt a_rtn = this->sum_a_init[_i] * _a / this->sum_a[_i];
-#endif
[…]

### 2020-08-06 sim/rettec.cpp

@@ -43,6 +43,7 @@ namespace sighandling {
 
 //! Our Retinotectal reaction diffusion class
 #ifdef AXONCOMP
+#define LINEAR_MAX 1
 # include "rd_rettec.h"
 #else
 # include "rd_rettec_nocomp.h"
@@ -273,6 +274,7 @@ int main (int argc, char **argv)
     RD.beta_ = conf.getDouble ("beta", 20.0);
 #ifdef AXONCOMP
     RD.epsilon = conf.getDouble ("epsilon", 0.2);
+    RD.a_max = conf.getDouble ("a_max", 0.5);
 #else
     RD.s = conf.getDouble ("s", 1.0);
     RD.set_w (conf.getDouble ("w", 1.0));

### 2021-01-04 paper/notes.tex
WIP

@@
 
-\subsubsection{Rotation experiments)}
+\subsubsection{Rotation experiments}
 
@@
 and having implemented fixed seed random numbers for repeatable sims,
-I can now try to determine if a smaller $d$ will help the simujlation compute,
+I can now try to determine if a smaller $d$ will help the simulation compute,
 though note that as $d$ is reduced, so $\delta t$ must be reduced, leading to
@@
 best - a piecewise linear transfer function on $a$. Thus, I've added the code
-to rd_barrel.h to compute a transformed version of $a$ called $\bar{a}$ and
+to rd\_barrel.h to compute a transformed version of $a$ called $\bar{a}$ and
 use this in the computation. It mitigates the instabilities without

### 2021-01-04 sim/rettec.cpp

@@ -66,6 +66,7 @@ namespace sighandling {
 # include <morph/Visual.h>
 // All the Visual models here derive from VisualDataModel
 # include <morph/VisualDataModel.h>
+# include <morph/GraphVisual.h>
 // Alias VisualDataModel<FLT>* as VdmPtr, to neaten code
 typedef morph::VisualDataModel<FLT>* VdmPtr;
 // We're visualizing HexGrids...
@@ -229,6 +230,8 @@ int main (int argc, char **argv)
     v1.setSceneTransXY (conf.getFloat ("x_default", 0.0f), conf.getFloat ("y_default", 0.0f));
     // Make this larger to "scroll in and out of the image" faster
     v1.scenetrans_stepsize = 0.5;
+    // Config can tell the program to finish as soon as the sim is done
+    v1.readyToFinish = conf.getBool ("finish_asap", false);
 #endif // COMPILE_PLOTTING
 
     // Instantiate and set up the model object
@@ -348,7 +351,8 @@ int main (int argc, char **argv)
     // Z position scaling - how hilly/bumpy the visual will be.
     morph::Scale<FLT> zscale; zscale.setParams (_m/10.0f, _c/10.0f);
     // The second is the colour scaling.
-    morph::Scale<FLT> cscale; cscale.setParams (_m, _c);
+    //morph::Scale<FLT> cscale; cscale.setParams (_m, _c);
+    morph::Scale<FLT> cscale; cscale.compute_autoscale (0, 1);
 # ifndef AXONCOMP
     // scaling2
     morph::Scale<FLT> zscale2; zscale2.setParams (1.0f/5.0f, 0.0f);
@@ -376,44 +380,49 @@ int main (int argc, char **argv)
 # endif
 
     // The a variable
-    std::vector<unsigned int> agrids;
+    std::vector<unsigned int> agrids (RD.N, 0);
     unsigned int side = static_cast<unsigned int>(floor (sqrt (RD.N)));
     if (plot_a) {
         spatOff = {xzero, 0.0, 0.0 };
         for (unsigned int i = 0; i<RD.N; ++i) {
             spatOff[0] = xzero + RD.hg->width() * (i/side);
             spatOff[1] = RD.hg->width() * (i%side);
-            unsigned int idx = v1.addVisualModel (new morph::HexGridVisual<FLT> (v1.shaderprog,
-                                                                                 v1.tshaderprog,
-                                                                                 RD.hg,
-                                                                                 spatOff,
-                                                                                 &RD.a[i],
[…]

### 2021-02-03 paper/notes.tex
WIP on the blowup problem

@@
 
+\emph{January 2021.} I did a lot of work on the visualisation code, making it a lot
+faster. This allowed me to get to the bottom of the blow up. The culprit is
+$\hat{a}$ and its divergence. There's a positive feedback loop here. I don't
+think this invalidates the model of using the gradient of $\hat{a}$ as
+competition, because in a real, biological system, $a$, and hence $\hat{a}$
+are bounded by finite resources. I'm thus going to allow bounding of $a_i$.
+
 \subsection{Stability for noise}

### 2021-02-03 sim/rd_rettec.h

@@ -72,7 +72,7 @@ protected:
 
 public:
     //! Override step() to recompute g on each sim step
-    virtual void step()
+    void step() override
     {
         this->stepCount++;
         // 0. Rebuild g, with a new set of random samples, if necessary
@@ -88,7 +88,7 @@ public:
 
 protected:
     //! Build g from the gradient of rho and the gammas, with an option for a noisy calculation.
-    void build_g()
+    void build_g() override
     {
         this->zero_vector_vector_array_vector (this->g, this->N, this->M);
         if (this->sigma_rho == scf(0.0)) {

### 2021-02-18 paper/notes.tex
Some work in progress

@@
 
+(Note, Feb 2021: I think the above mechanism is essentially Honda's
+`servomechanism' \cite{honda_competition_2003,honda_topographic_1998})
+
 This makes it easy to start to build $f_i$ with a two dimensional Gaussian
@@
 competition, because in a real, biological system, $a$, and hence $\hat{a}$
-are bounded by finite resources. I'm thus going to allow bounding of $a_i$.
+are bounded by finite resources. I'm thus going to allow bounding of $a_i$ by
+a transfer function at the end of the \code{integrate\_a} function.
+
+\begin{figure}\label{fig:stipples}
+\includegraphics[width=\textwidth]{images/stipples.png}
+\caption{Numerical artefact visible in $\hat{a}$. This stippling shows up in
+the final pattern (the graph I call `dr').}
+\end{figure}
+
+\begin{figure}\label{fig:stipple_src}
+\includegraphics[width=\textwidth]{images/stipple_source1.png}
+\caption{The source of the numerical artefact; the high frequency
+oscillations in $\dvrg{\hat{a}}$.}
+\end{figure}
+
+This resolves the instabilities and gives a simulation with $\epsilon=4$ that
+shows organisation of 56 or 96 axons which originate on a circular retina into
+an ordered pattern on an elliptical tectum. However, there are some numerical
+artefacts visible in the form of a `stippled' pattern in $\hat{a}$. These
+originate in high frequency power in the computation of $\dvrg{\hat{a}_i}$. At
+considerable computational cost, a smoothing of $\dvrg{\hat{a}_i}$ by Gaussian
+convolution resolves the stippling issue.
 

### 2021-02-18 paper/paper.tex
Some work in progress

--- 2021-01-04 paper/notes.tex
@@
 % Figure out the font for the author list..
-\def\Authors{\authorListFont Sebastian James\\[1 ex]  \Address \\
+\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address \\
   \corrAuthorFont $^{*}$ Correspondence: \corrEmail}
@@
 \title {
-  Modelling the Retinotectal projection - can competition provide a stopping mechanism?
+  Competition can provide a stopping mechanism for the retinotectal
+  projection \\
+  or \\
+  Competition provides a stopping mechanism for the chemoaffinity
+  theory which is robust to noise \\
+  or \\
+  Self-organisation of topographically ordered axon connections can take place
+  in a noisy environment if axons compete
 }
@@
 
+\emph{Abstract here.}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-I want to see if the competition mechanism which we demonstrated can form the
-murine barrel pattern~\cite{james_modelling_2020} can also help explain the
-arrangement of the Retinotectal projection.
+The retinotectal projection has proved to be a deep mine of information for
+the study of how the cells of the central nervous system are accurately
+connected together into functional networks. This projection connects the
+light-gathering cells in the retina to movement related cells in the optic
+tectum (known as the superior colliculus in mammals). Light sources
+originating close to each other in the environment tend to activate retinal
+cells situated close together in the eye so that an image of the environment
+is formed on the retinal surface. It has been discovered that the topography
+of the retina is preserved within brain regions that process this information
+such that cells which are adjacent within the retina primarily excite cells
+adjacent in the tectum. This indicates that during development there must
+exist a mechanism which ensures the correct wiring of the axons which leave
+the retina and connect to cells in the tectum.
 
-The Retinotectal (RT) projection is thought to use pairs of orthogonal
-gradients in the retina to specify the eventual location of synapses made on
-the surface of the tectum. Similar orthogonal gradients in the tectum provide
-a coordinate system which allows the axons to match their prespecified
-destination with an actual location on the tectum. Although the Ephrins have
-been shown to provide the guidance which sends axons \e{towards} their
-destination~\cite{feldheim_genetic_2000}, no mechanism has yet been discovered
-which actually \e{halts} the axonal growth cones in the correct location.
+One reason for the success of the study of the retinotectal projection is its
+capacity to be experimentally manipulated. In some non-mammalian species, both
+the retina and the tectum can be partially ablated, or even physically
+reorganised in-vivo, after which axons regrow to restore the order and
+function of the system for the individual animal. This manipulability was
+exploited in influential work by R. W. Sperry and co-workers during the
+mid-twentieth century, leading to Sperry's 1963 summary of
+the \emph{chemoaffinity theory}~\cite{sperry_chemoaffinity_1963} which
+proposes the existence of morphogenetic gradients that guide axons to their
+destination. The chemoaffinity theory was given robust support by the
+discovery of the ephrin ligands and their receptors~\cite{first_ephrin_paper}
+which have been shown to form into graded expression fields in the
+retina~\cite{ephrin_retina}, tectum~\cite{ephrin_tectum} as well as in other
+sensory systems, such as the somatosensory
+system~\cite{vanderhaegen_as_in_elife_paper}. The Ephrin ligands have a clear
+effect on axonal outgrowth, as shown in in-vitro~\cite{one} and
+in-vivo~\cite{another} studies.
 
-The main difference which I envisage between the RT and whisker barrel model
-systems is in the way competition is applied to axons growing towards the
-destination tissue. In the barrel field, I (notionally) assigned to a `group'
-of axons the same interaction parameter, if they originated from the same
-thalamic barreloid. The Retiontectal system is, in contrast, continuous;
-although I am hypothesizing that the RT axons obtain an interaction parameter
-from their location in the retina, just as I did for axons originating in
-different barreloids, I will allow each axon to have a unique interaction
-parameter, and expect that this leads to the continuous arrangement of axons
-expected in the tectum.
+It is tempting to consider the retinotectal projection well understood. With a
+comprehensive theory supported by a biochemical mechanism, is there anything
+left to understand? That question can be answered by reviewing retinotectal
+modelling papers.
 
-To test whether the competetive model (assuming it does reproduce the
-retinotopic neural arrangement) is useful, I can consider the following
-manipulations:
+Mini-review here, which contrasts some of the modelling papers. The upshot is
+that a central problem with models is not how the axons know how to get closer
+to their destination, but how they know they have \emph{arrived} at their true
+destination.
 
-I'd like to make a comparison with a system in which the stopping criterion is
-an absolute measurement of the expression level of the Ephrins in the
-tectum. I plan to introduce noise to the measurement that the growth cone
-makes of the Ephrin expression level. This will introduce stochasticity in the
-movement of the growth cones. The competition mechanism will be supported if
-it is more robust to noise than an absolute stopping mechanism.
-
-I will also then need to examine whether the competition mechanism is robust
-to the experimental manipulations of the sort which Sperry carried
-out~\cite{sperry_chemoaffinity_1963,goodhill_retinotectal_1999,goodhill_development_2005}.
+In a recent paper, we proposed a self-organising mechanism, based on
+morphogenetic signalling gradients, which can arrange
+axons growing from the thalamus into the somatosensory cortex into the well
+known murine barrel cortex pattern~\cite{james_modelling_2020}. In
+characterising that system, we explored the effect of various types of
+noise. We found that the mechanism was robust to noise in the expression of
+the signalling molecules over a wide range of amplitudes and length-scales,
+and that noise in the interaction parameters, which are obtained by sampling
+the signalling molecules in the source tissue (the thalamic barreloid field)
+could cause topological defects. The question of noise in axon guidance has
+been explored by Goodhill~\cite{goodhill_can_2016}.
 
@@
 
-The initial model is derived from the one in the Barrels
-paper~\cite{james_modelling_2020} which is found
-in \code{sim/rd\_barrel.h}. The specialization of \code{RD\_Barrel} (and a
-sister class, \code{RetArrange}) can be found in \code{sim/rd\_rettec.h}, with
-main code file \code{rettec.cpp}. Here, instead of providing interaction
-parameters ($\gamma_{i,j}$) in a configuration file, I use only the parameter
-$N$, which in this context is the number of neurons originating in the
-retina. I automatically arrange these $N$ retinal neurons in a radially
-symmetric pattern (See \code{RetArrange}). I then obtain the $\gamma$
-parameters from the neuron locations within the pattern.
-
-\subsection{Configurations}
-
-In the config directory, there are numerous basic configs, such as N12.json,
-N24.json and so on. I find I have to modify some of the parameters as N
-increases. Larger N requires smaller hexes to allow a suitable pattern to
-develop. Larger N also generally requires a smaller diffusion constant
-$D$. I'm experimenting with $\epsilon$, the competition parameter, which, if
-too high causes the retinotopy to degrade during the simulation.
-
-\subsection{Noise}
-
-I wish to examine the effect of noise on the arrangement of axons on the
-tectum that results from axon guidance. I assume that there is no noise in the
-locations of the retinal neurons which serve as the origin of the interaction
-parameters; this remains a symmetric, ordered pattern.
-
-The hypothesis which the model is based on is that the interaction parameters
-are derived from the positions of the neurons on the source tissue (the
-retina) and that there are signalling molecule gradients which provide the
-required coordinate axes. Thus, the first source of noise would be in the
-sampling of these gradients and hence on the interaction parameters which are
-carried with the efferent axons to the tectum. This noise would be expected to
-be fixed after the axons had grown away from the retina.
-
-On arriving at the tectum, axon growth cones must determine their direction of
-movement. The model encapsulates the quantitative hypothesis that growth cone
-movement is a combination of diffusion and guidance. The guidance is subject
-to sampling error and thus a second source of noise can be introduced to the
-model. This noise varies with time during the simulation and can be considered
-to have two contributions; one from variations in the true expression level of
-the signalling gradient and one from sampling error measuring it. I will
-combine these into a single error. \textbf{Or maybe not:} Because the
-no-competition model (see Section \ref{sec:nocomp}) has a function
-$f(\gamma, \rho)$, which represents the axons' ability to sample the absolute
-value of $\rho$, it is important to represent the variability of the true
-value of $\rho(\mb{x})$. \textbf{Update (July 2020):} Since I was last
-thinking about this, I added length- and amplitude-variable noise to $\rho$,
-so I will use that code here, and additionally consider noise in the sampling
-of $\rho$.
-
-The system equations are modified as follows. Starting from the branching equation
-
-\begin{equation} \label{eq:da}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla
-a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x}) + \chi_i \right) - \frac{\partial c_i}{\partial t},
-\end{equation}
-
-the term for the divergence of $\mb{J}$ can be written:
-%
-\begin{equation}
-  \label{eq:divJ}
-  \nabla\vcdot\mb{J}_i(\mb{x},t) = \nabla\vcdot\left(D \nabla
-  a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})+ \chi_i \right).
-\end{equation}
-%
-In the barrels model, each $\rho_j(\mb{x})$ was assumed static, allowing the
-definition of a time-independent $\mb{g}_i(\mb{x})$:
-%
-\begin{equation}
-\mb{g}_i(\mb{x}) \equiv \sum_{j=1}^{M} \gamma_{i,j} \nabla\rho_j(\mb{x}).
-\end{equation}
-%
-Here, I represent the time-dependence of the signalling molecule expression
-by noting that $\rho$ is now a function of time.  The error which the growth
-cones make in sampling $\rho$ and estimating $\nabla\rho_j(\mb{x})$ is
-$\xi_{\rho'}(t)$:
-%
-\begin{equation}
-\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x},t)
-+ \xi_{\rho'}(t) \big).
-\end{equation}
-%
-One problem with this representation is that the equations represent a
-population model. Within one spatial element of the discretized solution of
-the differential equations, the error, $\xi_{\rho'}(t)$ can have only one value at
-time $t$; this value is applied to all the individual axon growth cones that
-are represented by that element. Nonetheless, it's a useful (and fairly simple)
-way to represent the guidance error in the system. In practice, I ignore the
-variation in $\rho_j$ and write
-%
-\begin{equation}
-\mb{g}_i(\mb{x},t) \equiv \sum_{j=1}^{M} \gamma_{i,j} \big( \nabla\rho_j(\mb{x})
-+ \xi_{\rho'}(t) \big).
-\end{equation}
-
-$\xi_{\rho'}(t)$ will be included in the model as a function returning numbers sampled
-from a normal distribution with mean 0 and standard deviation $\sigma_{\rho'}$.
-
-The systematic error in $\gamma_{i,j}$ will be applied at setup time. The
-symmetric gamma values will first be derived from the symmetric layout of the
-retinal neurons, then a normally distributed perturbation will be applied to
-each $\gamma_{i,j}$.
-
-For retinal neurons with two-dimensional coordinates $\mb{r}_i$, $\gamma$ is
-given (in the absence of noise) by
-$\gamma_{i,0} = G\,|\mb{r}_i| \mb{r}_i \cdot \hat{\mb{x}}$ and
-$\gamma_{i,1} = G\,|\mb{r}_i| \mb{r}_i \cdot \hat{\mb{y}}$, where
-$\hat{\mb{x}}$ and $\hat{\mb{y}}$ are unit vectors, and $G$ is simply a gain
-on $\gamma$. With noise, these become
-$\gamma_{i,0} = G\,\big(\xi_\gamma |\mb{r}_i| \mb{r}_i \cdot \hat{\mb{x}} \big)$ and
-$\gamma_{i,1} = G\,\big(\xi_\gamma |\mb{r}_i| \mb{r}_i \cdot \hat{\mb{y}} \big)$
-
-$\xi_\gamma$ is drawn from a normal distribution with mean 0 and standard
-deviation $\sigma_{\gamma}$.
-
-\subsection{Non-competition based stopping mechanism}
-\label{sec:nocomp}
-
-Here, I seek to model the `pure' chemospecificity theory in which axons are
-guided by orthogonal morphogen gradients until they reach their destination,
-which they must be able to sense. The axons, guided by the \emph{gradient} of
-the morphogen concentration are assumed also to be able to sample
-the \emph{absolute} concentration of the signalling molecules.  A reduction in
-branching, and possibly an enhancement of synaptogenesis are triggered when
-the concentrations of the two orthogonal morphogens are simultaneously
-measured to be within the axon's `stopping' ranges.
-
-To find a way to implement this in a population model similar to the
-competitive model, I start from the branching equation (dropping the
-competition term, $\chi_i$):
-
-\begin{equation} \label{eq:da_again}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M} \gamma_{i,j}\nabla \rho_j(\mb{x})\right) - \frac{\partial c_i}{\partial t},
-\end{equation}
-
-Spatial transfer of information in the system can be halted if the flux of
-axonal branching (the term in brackets) tends to 0.  This can be achieved with
-a function of the gammas and rhos:
-
-\begin{equation} \label{eq:da_stopping}
-\frac{\partial a_i}{\partial t} = \nabla\vcdot\left(D \nabla a_i-a_i\sum_{j=1}^{M=2} \gamma_{i,j}\nabla \rho_j(\mb{x})\right)\;f_i\big(\gamma_{i,1},\gamma_{i,2},\tilde{\rho}_{1}(\mb{x}),\tilde{\rho}_{2}(\mb{x})\big) - \frac{\partial c_i}{\partial t},
-\end{equation}
-
-Where $f_i$ is a function which permits branching far away from a given axon's
-target region and tends to zero inside the target region. It represents an
-axon's internal `stopping process'. Note carefully that this internal axonal
-function does \emph{not} depend directly on space; it is a function of the
-interaction parameters $\gamma$ and of the signalling molecule density,
-$\rho$. Note that $\tilde{\rho}_j(\mb{x})$ is the axon's \emph{measurement} of
-$\rho_j(\mb{x})$ and may include measurement noise, $\xi_\rho$:
-$\tilde{\rho}_j = \rho_j + \xi_\rho$. Given that the morphogen gradients are
-orthogonal, $(\tilde{\rho}_1,\tilde{\rho}_2)$ defines a vector field
-$\vec{\rho}\,(\mb{x})$. Similarly, $(\gamma_{i,1},\gamma_{i,2})$ can be
-combined to give the vector $\vec{\gamma}$.
-
-This makes it easy to start to build $f_i$ with a two dimensional Gaussian
-function:
-
-\begin{equation}\label{eq:g}
-g_i\big(\vec{\rho}(\mb{x})\big) = \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 }
-\end{equation}
-
-where $\vec{\gamma}$ gives the location of the peak and the parameter $w$ is
-the width of the peak. This is squashed through a sigmoid, with
-sharpness parameter $s$ and inverted:
-
-\begin{equation}\label{eq:f}
-f_i\big(\vec{\rho}(\mb{x})\big) = 1 - \bigg( \frac{2}{1 + \mathrm{e}^{-s
-g_i(\vec{\rho}\,)}} - 1 \bigg)
-\end{equation}
-
-In one equation:
-
-\begin{equation}\label{eq:gf}
-f_i\big(\vec{\rho}(\mb{x})\big) = 2 - \frac{2}{1
-+ \mathrm{e}^{-s \mathrm{e}^{-\big(\frac{\vec{\rho}(\mb{x})-\vec{\gamma}}{2 w}\big)^2 } }}
-\end{equation}
-
-\subsection{Allowing the stopping mechanism to enhance connection making}
-
-As well as affecting branching, there is no reason why I couldn't have the
-axonal function also affect connection making. Consider the equation governing
-the rate of change of connections.
-
-\begin{equation} \label{eq:dc}
-\frac{\partial c_i}{\partial t} =-\alpha c_i +\beta  \left(1 - \sum_{j=1}^{N} c_{j}\right)[a_i]^k.
-\end{equation}
-
-If $\beta$ were to be modulated by $1 - f_i$, then connection making would
-occur much more strongly in the correct location. Would it be too good I
-wonder? Answer: Yes, it's highly effective. To reproduce, set fModulatesBeta
-to true, prop\_modulated to 1 and prop\_unmodulated to 0.
-
-\subsection{Allowing the stopping mechanism to reduce connection decay}
-
-An alternative to the above, is to slow down connection decay within the
-stopping zone.
-
 \section{Results}
 
-\subsection{Wild type}
-
-The operation of the model without modification and with zero or low noise.
-
-\subsection{Manipulations}
-
-Reference \cite{goodhill_development_2005} (from which the headings and quotes
-below are taken) gives a list of features which a model of retinotectal
-development or regeneration should address.
-
-\subsubsection{Ectopic targeting}
-
-\begin{displayquote}
-\emph{Retinal axons entering the tectum via abnormal trajectories can still find
-their appropriate termination sites (Finlay et al. 1979b; Harris 1982,
-1984). This suggests the presence of vector signals throughout the tectum
-pointing axons towards their correct termination sites.}
-\end{displayquote}
-
-The orthogonal gradients in the current model which signal growth direction of
-axonal growth cones naturally provide the vector signals which ensure that
-regardless of entry point, growth cones tend towards their correct
-location. Can prove with different distributions of initial branching density
-($a(\mb{x},t=0)$).
-
-\subsubsection{Shifting connections}
-
-\begin{displayquote}
-\emph{In some fish and amphibians, the retina grows by addition of new neurons
-around the ciliary margin, while the tectum grows by addition of cells to its
-caudomedial edge (Gaze et al. 1974; Fraser 1983).The retinotectal map remains
-ordered throughout this time, indicating that the retinotectal projection is
-continually shifting caudally. This suggests that, at least in these species,
-either chemoaffinity gradients do not irreversibly specify the map or the
-gradients must evolve with time.}
-\end{displayquote}
-
-The current model should work under the condition that the underlying domain
-is increased in size, with the competition acting to ensure that the entire
-domain is filled and the ordering maintained.
-
-To do: implement a HexGrid whose boundary increases with time. That will mean
-some tampering with \code{morph::HexGrid}.
-
-\subsubsection{Rotation experiments}
-
-\begin{displayquote}
-\emph{If a Xenopus presumptive tectum is rotated early enough during development,
-a map is formed that is normal relative to the whole animal, whereas later
-rotations lead to a rotated map (Chung \& Cooke 1978).}
-\end{displayquote}
-
-These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2B--D and can
-include retinal rotation as well as early/late tectal rotations.
-
-The model as it stands contains an assumption about the orientation of the
-guidance gradients. So rotating the tectum early enough means making no change
-(given that I have no model for \emph{creating} the guidance gradients).
-
-\subsubsection{Map compression and expansion}
-
-\begin{displayquote}
-\emph{The map formed after removal of half the retina initially covers half the
-tectum (Attardi \& Sperry 1963), but then gradually expands to fill the whole
-tectum (Schmidt et al. 1978). The axon terminal density remains the same
-(Schmidt et al.  1978). If the optic nerve is then made to regenerate again,
-an expanded map is immediately formed (Schmidt 1978). If half the tectum is
-ablated, the regenerated map is compressed into the remaining tectal space
-(Yoon 1971; Sharma 1972; Cook 1979; Finlay et al. 1979a).}
-\end{displayquote}
-
-I need to examine more carefully the timeline of these experiments.
-
-These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2E and 2F.
-
-\subsubsection{Translocation}
-
-\begin{displayquote}
-\emph{A}
-\end{displayquote}
-
-These correspond to \cite{goodhill_retinotectal_1999}, Figs.~2I and 2J.
-
-\subsubsection{Compound-eye experiments}
-
-\begin{displayquote}
-\emph{B}
-\end{displayquote}
-
-This corresponds to \cite{goodhill_retinotectal_1999}, Fig 2H.
-
-\section{Implementation notes}
-
-\subsection{Stability}
-
-The model can become unstable due to the unbounded nature of $a$. Essentially,
-$a$ can blow up in the computation of $\dvrg J$.
-
-Stability with respect to $D$ is a particular issue, as is stability in the
-presence of noise.
-
-\subsection{Stability for small $D$}
-
-In the basic, $N=56$ wild type simulation, $D$ seems to need to be close to
-0.02. Unfortunately, the simulation is unstable for $\delta t=0.0001$ and
-$d=0.015$. Having fixed a bug which prevented me from reducing $d$ very much,
-and having implemented fixed seed random numbers for repeatable sims,
-I can now try to determine if a smaller $d$ will help the simulation compute,
-though note that as $d$ is reduced, so $\delta t$ must be reduced, leading to
-longer simulation computations.
-
-Note: $\epsilon=1.2$ here in the following:
-
-D=0.02, dt=0.00005, d=0.007: Crash at step 13803.
-
-D=0.01, dt=0.0001, d=0.015: Crash at step 2629.
-
-D=0.015, dt=0.0001, d=0.015: Crash at step 4864.
-
-D=0.02, dt=0.0001, d=0.015: Crash at step 7280.
-
-D=0.025, dt=0.0001, d=0.015: Crash at step 9649.
-
-D=0.03, dt=0.0001, d=0.015: Crash at step 12394.
-
-This pattern is similar to that shown in Fig.~2 of the Barrels paper. To
-enable $D$ to be low, it's necessary to reduce $\epsilon$, BUT I want to
-explore increased competition, whilst maintaining low $D$.
-
-Thus, some sort of cap on $a$ may be called for.
-
-But, lets first see how much reduction of $\epsilon$ is required. Starting
-from D=0.01, dt=0.0001, d=0.015: Crash at step 2629, with $\epsilon=1.2$,
-reduce $\epsilon$.
-
-$\epsilon=0.8$: Crash step 4779.
-
-$\epsilon=0.4$: Crash step 9802.
-
-$\epsilon=0.1$: Computation makes it to step 24000 with no crash, but that's
-not enough competition for an even result.
-
-\subsection{Bounding $a$}
-
-I've already implemented systems in which $a$ is bounded. Simplest is probably
-best - a piecewise linear transfer function on $a$. Thus, I've added the code
-to rd\_barrel.h to compute a transformed version of $a$ called $\bar{a}$ and
-use this in the computation. It mitigates the instabilities without
-making much change to the operation of the model. It adds an additional
-parameter to the model, $a_{max}$, which I've set to 0.5 by default.
-
-\subsection{Stability for noise}
-
-Write me.
+\section{Discussion}
 

### 2021-05-12 paper/paper.tex
Model tweaked to have retina and tectum with receptors and ligands on each

@@
 connected together into functional networks. This projection connects the
-light-gathering cells in the retina to movement related cells in the optic
+light-gathering cells in the retina to movement-related cells in the optic
 tectum (known as the superior colliculus in mammals). Light sources
@@
 in-vivo~\cite{another} studies.
+%
+% A-P:
+%
+% EphA and ephrinAs: Anteroposterior axis of retiontectal projection
+% (Nasal-Temporal on retina; Anteroposterior on tectum).
+%
+% RGM (tectum) Neogenin (retina)
+%
+% D-V:
+%
+% EphB and ephrin B: Dorsoventral on retina; dorsoventral on tectum.
+%
+% En-2: Repels temporal growth cones of xenopus ret-tec axons and attracts
+% nasal ones. Brunet et all, Nature, 2005
+%
+% Ryk/Fz (retina) and Wnt3 (tectum) (Flanahan, 2006)
+%
 
@@
 
+New introduction: Simpson and Goodhill presented a model in which the
+chemoaffinity mechanism is non-biological. We offer a model in which a
+competition mechanism counteracts the drive of axons along the ephrin
+gradient. Will we need the axon-axon interaction mechanism based on
+non-same-Eph repulsion?
+
 \section{Model}

### 2021-05-12 sim/agent/branch.h

@@ -1,27 +1,38 @@
 #pragma once
 
 #include <vector>
-//#include <deque>
 #include <morph/vVector.h>
 #include <morph/Vector.h>
+#include "tissue.h"
 
 // A retinotectal axon branch class. Holds current and historical positions, a preferred
 // termination zone, and the algorithm for computing the next position. Could derive
-// from an 'agent' base class.
-template<typename T>
+// from an 'agent' base class. The branches express N Ephrin receptor types (EphA, EphB etc)
+template<typename T, size_t N>
 struct branch
 {
     // Compute the next position for this branch, using information from all other
-    // branches and the parameters vector, m
-    void compute_next (const std::vector<branch<T>>& branches, const morph::Vector<T, 4>& m)
+    // branches and the parameters vector, m. Will also need location on target tissue,
+    // to get its ephrin values.
+    void compute_next (const std::vector<branch<T, N>>& branches,
+                       const guidingtissue<T, N>* tissue,
+                       const morph::Vector<T, 4>& m)
     {
         // Current location is named b
         morph::Vector<T, 2> b = path.back();
 
         // Chemoaffinity, graded by origin position (i.e termination zone) of each axon
         // in the retina.
-        morph::Vector<T, 2> G = this->interaction + morph::Vector<T,2>({-0.5,-0.5});
-        //std::cout << "G=" << G << std::endl; // range 0,0 to 1,1
+        morph::Vector<T, 2> G;
+        if constexpr (N==4) {
+            // With 4 receptor/ligand pairs, the x component is zeroth minus the 2nd
+            G[0] = this->rcpt[0] * tissue->lgnd[aid][0] - this->rcpt[2] * tissue->lgnd[aid][2];
+            // and y component is the first minus the 3rd
+            G[1] = this->rcpt[1] * tissue->lgnd[aid][1] - this->rcpt[3] * tissue->lgnd[aid][3];
+        } else if constexpr (N==2) {
+            G[0] = this->rcpt[0] * tissue->lgnd[aid][0];
+            G[1] = this->rcpt[1] * tissue->lgnd[aid][1];
+        }
 
         // Competition, C, and Axon-axon interactions, I, computed during the same loop
         // over the other branches
@@ -38,9 +49,10 @@ struct branch
             morph::Vector<T, 2> kb = b - k.path.back();
             T d = kb.length();
             T W = d <= this->two_r ? (T{1} - d/this->two_r) : T{0};
[…]

### 2021-05-12 sim/agent/branchvisual.h

@@ -24,21 +24,23 @@
 
 #include "branch.h"
 
-template <typename Flt>
+template <typename Flt, size_t N>
 class BranchVisual : public morph::VisualModel
 {
 public:
-    BranchVisual(GLuint sp, const morph::Vector<float, 3> _offset,
-                 std::vector<branch<Flt>>* _branches, bool _axonview = false)
+    BranchVisual(GLuint sp, GLuint tsp, const morph::Vector<float, 3> _offset,
+                 std::vector<branch<Flt, N>>* _branches, bool _axonview = false)
     {
         this->branches = _branches;
         this->axonview = _axonview;
         this->shaderprog = sp;
+        this->tshaderprog = tsp;
         this->mv_offset = _offset;
         this->viewmatrix.translate (this->mv_offset);
     }
 
-    morph::Scale<Flt, Flt> EphA_scale;
+    // Receptor and ligand need to be scaled to 0/1 range
+    morph::Scale<Flt, Flt> rcpt_scale;
 
     void initializeVertices()
     {
@@ -46,23 +48,14 @@ public:
 
         VBOint idx = 0;
         // For each branch, draw lines for the path history and a sphere for the current
-        // location, with a second colour for the EphA expression.
+        // location, with a second colour for the rcpt expression.
         for (auto b : *this->branches) {
             // Colour comes from target location.
-            std::array<float, 3> clr = { b.interaction[0], b.interaction[1], 0 };
-            std::array<float, 3> clr2 = { 0, 0, this->EphA_scale.transform_one(b.EphA) };
+            // Prolly need receptor scaling here:
+            std::array<float, 3> clr = { this->rcpt_scale.transform_one(b.rcpt[0]),
+                                         this->rcpt_scale.transform_one(b.rcpt[1]), 0 };
+            std::array<float, 3> clr2 = { 0, 0, this->rcpt_scale.transform_one(b.rcpt[0]) };
             morph::Vector<float, 3> cur = { 0, 0, 0 };
-#if 0
-            morph::Vector<float, 3> last = { 0, 0, 0 };
-            // First draw the path. Problem here is where to *start* in the history
-            for (unsigned int i = 1; i < b.path.size() && i < this->histlen; ++i) {
-                last[0] = b.path[i-1][0];
-                last[1] = b.path[i-1][1];
-                cur[0] = b.path[i][0];
-                cur[1] = b.path[i][1];
[…]

### 2021-05-12 sim/agent/netvisual.h

@@ -27,10 +27,11 @@ template <typename Flt>
 class NetVisual : public morph::VisualModel
 {
 public:
-    NetVisual(GLuint sp, const morph::Vector<float, 3> _offset, net<Flt>* _locations)
+    NetVisual(GLuint sp, GLuint tsp, const morph::Vector<float, 3> _offset, net<Flt>* _locations)
     {
         this->locations = _locations;
         this->shaderprog = sp;
+        this->tshaderprog = tsp;
         this->mv_offset = _offset;
         this->viewmatrix.translate (this->mv_offset);
     }

### 2021-05-12 sim/agent/retvisual.h

@@ -17,12 +17,12 @@
 
 #include "tissue.h"
 
-template <class T>
+template <class T, size_t N>
 class retvisual : public morph::VisualModel
 {
 public:
     //! Single constructor for simplicity
-    retvisual(GLuint sp, GLuint tsp, const guidingtissue<T>* _r, const morph::Vector<float> _offset)
+    retvisual(GLuint sp, GLuint tsp, const guidingtissue<T, N>* _r, const morph::Vector<float> _offset)
     {
         this->shaderprog = sp;
         this->tshaderprog = tsp;
@@ -48,11 +48,11 @@ public:
         unsigned int idx = 0;
 
         // Use the interaction parameter of the retina object to set the colours of the elements
-        this->dcolour.resize (this->ret->interaction.size());
-        this->dcolour2.resize (this->ret->interaction.size());
-        for (unsigned int i = 0; i < this->ret->interaction.size(); ++i) {
-            this->dcolour[i] = this->ret->interaction[i][0];
-            this->dcolour2[i] = this->ret->interaction[i][1];
+        this->dcolour.resize (this->ret->rcpt.size());
+        this->dcolour2.resize (this->ret->rcpt.size());
+        for (unsigned int i = 0; i < this->ret->rcpt.size(); ++i) {
+            this->dcolour[i] = this->ret->rcpt[i][0];
+            this->dcolour2[i] = this->ret->rcpt[i][1];
         }
         this->colourScale.transform (this->dcolour, this->dcolour);
         this->colourScale.autoscaled = false;
@@ -142,7 +142,7 @@ protected:
     }
 
     //! The retina data to visualize
-    const guidingtissue<T>* ret;
+    const guidingtissue<T, N>* ret;
 
     //! Colour mapping
     std::vector<float> dcolour;

### 2021-05-12 sim/agent/tissue.h

@@ -30,86 +30,89 @@ struct tissue
     }
 };
 
-//! Tissue which has guidance parameters in the form of an interaction parameter vector
-template<typename T>
+//! Tissue which has guidance parameters in the form of a 2D vector which stands for the
+//! expression of 2 or possibly 4 receptor molecules. If positive/negative values are
+//! allowed, then it's 4.
+template<typename T, size_t N>
 struct guidingtissue : public tissue<T>
 {
     //! A left-right interaction parameter and an up-down interaction parameter for each
-    //! retinal neuron (hence Vectors with 2 elements, each of which can be +/-)
-    morph::vVector<morph::Vector<T,2>> interaction;
-    morph::vVector<T> EphA;
+    //! piece of guiding tissue requires 4 receptors and 4 ligands. Holds receptor
+    //! expressions for each cell.
+    morph::vVector<morph::Vector<T,N>> rcpt;
 
-    //! Init the wildtype retina's interaction parameters, which stand in for the Ephrins.
+    //! The tissue also has an expression of ligands to interact with receptors of other
+    //! cells/agents
+    morph::vVector<morph::Vector<T,N>> lgnd;
+
+    //! Init the wildtype tissue's receptors and ligands.
     guidingtissue(size_t _w, size_t _h, morph::Vector<T,2> _dx, morph::Vector<T,2> _x0)
         : tissue<T> (_w, _h, _dx, _x0)
     {
-        this->interaction = this->posn - this->x0;
+        this->rcpt.resize (this->posn.size());
+        this->lgnd.resize (this->posn.size());
 
-        this->EphA.resize (this->posn.size());
+        T max_x = this->w * this->dx[0] + this->x0[0];
+        T max_y = this->h * this->dx[1] + this->x0[1];
+
+        // Ligands and receptors are set up as a function of their cell's position in the tissue.
         for (size_t ri = 0; ri < this->posn.size(); ++ri) {
-            this->EphA[ri] = T{1.05} + (T{0.26} * std::exp (T{2.3} * this->posn[ri][0])); // R(x) = 0.26e^(2.3x) + 1.05,
+            if constexpr (N == 4) {
+                // First orthogonal pair of receptors
+                this->rcpt[ri][0] = T{1.05} + (T{0.26} * std::exp (T{2.3} * this->posn[ri][0])); // R(x) = 0.26e^(2.3x) + 1.05,
[…]

### 2021-06-23 paper/paper.tex
Nothing much but a start to the model description

@@
 adjacent in the tectum. This indicates that during development there must
-exist a mechanism which ensures the correct wiring of the axons which leave
-the retina and connect to cells in the tectum.
+exist a mechanism which ensures the correct arrangement of the axons which
+leave the retina and connect to cells in the tectum.
 
@@
 to their destination, but how they know they have \emph{arrived} at their true
-destination.
+destination. Also make the point that many of the models are phenomenological
+in nature and do not elucidate the mechanisms behind the arrangement.
 
@@
 morphogenetic signalling gradients, which can arrange
-axons growing from the thalamus into the somatosensory cortex into the well
+axons growing from the thalamus to the somatosensory cortex into the well
 known murine barrel cortex pattern~\cite{james_modelling_2020}. In
@@
 
+Another approach: We present a model of gradient-following based on graded
+interactions between the growth cones of retinal ganglion cell axons and the
+tectal surface. We show that a combination of exponentially graded expression
+in the retina and logarithmically graded expression on the tectum permits a
+purely local model to reproduce the majority of reported phenomena associated
+with retinotectal organisation.
+
 \section{Model}
+
+Our model consists of agents, representing axonal growth cones, originating
+from a square retina and moving upon a square tectum. Each agent carries a set
+of 4 receptors and 4 ligands which are expressed at differing levels depending
+upon the soma's location on the retina. The expression levels of the receptors
+and ligands form gradients across the retinal surface. Receptors are arranged
+in orthogonal pairs with $r_0$ being orthogonal to $r_1$ and $r_2$, whose
+gradient is opposite to $r_0$ is orthogonal to $r_3$. Ligands are also
+arranged in orthogonal pairs, with...
+%
+Similarly, the tectum expresses the same sets of
+receptors and ligands, also in orthogonal pairs of gadients, though not
+necessarily with the same expression function.
 

### 2021-06-25 paper/paper.tex
A few small changes

@@
 from a square retina and moving upon a square tectum. Each agent carries a set
-of 4 receptors and 4 ligands which are expressed at differing levels depending
-upon the soma's location on the retina. The expression levels of the receptors
-and ligands form gradients across the retinal surface. Receptors are arranged
-in orthogonal pairs with $r_0$ being orthogonal to $r_1$ and $r_2$, whose
-gradient is opposite to $r_0$ is orthogonal to $r_3$. Ligands are also
-arranged in orthogonal pairs, with...
+of 4 receptors and 4 ligands which it expresses at differing levels depending
+upon the soma's origin on the retina. The expression levels of
+the receptors and ligands form gradients across the retinal surface. Receptors
+are arranged in orthogonal pairs with $r_0$ being orthogonal to $r_1$ and
+$r_2$, whose gradient is opposite to $r_0$ is orthogonal to $r_3$. Ligands are
+also arranged in orthogonal pairs, with...
 %
-Similarly, the tectum expresses the same sets of
-receptors and ligands, also in orthogonal pairs of gadients, though not
-necessarily with the same expression function.
+The tectum expresses the same sets of receptors and ligands, also in
+orthogonal pairs of gradients, though not necessarily with the same expression
+function.
 

### 2021-06-28 paper/paper.tex
Paper WIP

@@
 
-New introduction: Simpson and Goodhill presented a model in which the
+\textbf{New introduction:} Simpson and Goodhill presented a model in which the
 chemoaffinity mechanism is non-biological. We offer a model in which a
@@
 
-Another approach: We present a model of gradient-following based on graded
+\textbf{Another approach:} We present a model of gradient-following based on graded
 interactions between the growth cones of retinal ganglion cell axons and the
@@
 from a square retina and moving upon a square tectum. Each agent carries a set
-of 4 receptors and 4 ligands which it expresses at differing levels depending
-upon the soma's origin on the retina. The expression levels of
-the receptors and ligands form gradients across the retinal surface. Receptors
-are arranged in orthogonal pairs with $r_0$ being orthogonal to $r_1$ and
-$r_2$, whose gradient is opposite to $r_0$ is orthogonal to $r_3$. Ligands are
-also arranged in orthogonal pairs, with...
+of 4 receptors, $r$, and 4 ligands, $l$, which it expresses at differing
+levels depending upon the soma's origin on the retina. The expression levels
+of the receptors and ligands form gradients across the retinal
+surface. Receptors are arranged in orthogonal pairs with the gradient of $r_0$
+being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$
+is orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs, with
+$l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$
+orthogonal to $l_0$ and opposing $r_1$. While we do not explicitly name $r_0$,
+$r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes
+EphA, EphB and Ryk~\cite{schmitt_wntryk_2006} receptors and that $l$ includes
+the ephrin-A, ephrin-B and Wnt3~\cite{schmitt_wntryk_2006} ligands.
 %
@@
 orthogonal pairs of gradients, though not necessarily with the same expression
-function.
+function. Figure~\ref{f:1} shows colour maps of receptor and ligand
+expression, in this case experssion maps are linear.
 

### 2021-06-28 paper/paper.tex
WIP - getting my expression profiles all in the right sense

@@
 tectal surface. We show that a combination of exponentially graded expression
-in the retina and logarithmically graded expression on the tectum permits a
-purely local model to reproduce the majority of reported phenomena associated
-with retinotectal organisation.
+in the retina~\cite{reber_relative_2004} and exponentially (for repulsively
+interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
+interacting receptor/ligands) graded expression on the tectum permits a purely
+local model to reproduce the majority of reported phenomena associated with
+retinotectal organisation.
 
@@
 
+Interactions in the model are by receptor-ligand interactions only. Forward
+interactions are those in which the ligand activates the receptor, delivering
+a signal into the receptor-expressing cell. Reverse signals occur when a
+membrane-bound ligand is activated~\cite{wu_role_2019}.
+
 \section{Results}

### 2021-06-28 sim/agent/branch.h

@@ -16,7 +16,15 @@ enum class border_effect
     penned_pushed // Like 'penned in' but also pushed away from the border with 1/r component as in original
 };
 
-// The branch which does chemoaffinity+comp+(forward)interaction
+// The branch which does chemoaffinity + comp + axon-axon interaction
+//
+// For receptor/ligand interactions we define:
+//
+// let receptor 0 interact primarily with ligand 3 [gradients (6,7)]
+// let receptor 1 interact primarily with ligand 2 [gradients (4,5)]
+// let receptor 2 interact primarily with ligand 1 [gradients (2,3)] // opposite to receptor 0
+// let receptor 3 interact primarily with ligand 0 [gradients (0,1)] // opposite to receptor 1
+//
 template<typename T, size_t N>
 struct branch : public branch_base<T,N>
 {
@@ -26,7 +34,6 @@ struct branch : public branch_base<T,N>
     // Signalling ratio parameter
     static constexpr T s = T{1.1};
 
-    static constexpr bool biological_tectum = true;
     // Compute the next position for this branch, using information from all other
     // branches and the parameters vector, m. Will also need location on target tissue,
     // to get its ephrin values.
@@ -43,22 +50,13 @@ struct branch : public branch_base<T,N>
             // First, find the ligand gradients close to the current location b.
             morph::Vector<T, 8> lg = tissue->lgnd_grad_at (b);
             // 4 receptors and 4 ligand gradients. The 4 receptors are in order: r, u, l, d
-            if constexpr (biological_tectum == false) {
-                // If no rotation of tectum wrt retina then:
-                // let receptor 0 interact primarily with ligand gradients (0,1)
-                // let receptor 1 interact primarily with ligand gradients (2,3)
-                // let receptor 2 interact primarily with ligand gradients (4,5) // opposite to receptor 0
-                // let receptor 3 interact primarily with ligand gradients (6,7) // opposite to receptor 1
-                G[0] = this->rcpt[0] * lg[0] + this->rcpt[1] * lg[2] + this->rcpt[2] * lg[4] + this->rcpt[3] * lg[6];
[…]

### 2021-06-28 sim/agent/tissue.h

@@ -71,9 +71,15 @@ enum class tissue_region
     right_half   // The right half
 };
 
-//! Tissue which has guidance parameters in the form of a 2D vector which stands for the
-//! expression of 2 or possibly 4 receptor molecules. If positive/negative values are
-//! allowed, then it's 4.
+/*!
+ * Tissue which has guidance parameters in the form of a 2D vector which stands for the
+ * expression of 2 or possibly 4 receptor molecules. If positive/negative values are
+ * allowed, then it's 4.
+ *
+ * tparam T: Data type for processing. Usually float unless double seems necessary.
+ *
+ * tparam N: There are N receptors and N ligands in the guidingtissue.
+ */
 template<typename T, size_t N>
 struct guidingtissue : public tissue<T>
 {
@@ -124,6 +130,7 @@ struct guidingtissue : public tissue<T>
                 // Second orthogonal pair in opposite sense
                 this->rcpt[ri][2] = this->rcpt_expression_function (max_x-this->posn[ri][0]);
                 this->rcpt[ri][3] = this->rcpt_expression_function (max_y-this->posn[ri][1]);
+
                 this->lgnd[ri][2] = this->lgnd_expression_function (this->posn[ri][0]);
                 this->lgnd[ri][3] = this->lgnd_expression_function (this->posn[ri][1]);
 

### 2021-07-05 paper/paper.tex
WIP on paper

@@
 destination. The chemoaffinity theory was given robust support by the
-discovery of the ephrin ligands and their receptors~\cite{first_ephrin_paper}
+discovery of the ephrin ligands and their receptors~\cite{cheng_complementary_1995,drescher_vitro_1995}
 which have been shown to form into graded expression fields in the
-retina~\cite{ephrin_retina}, tectum~\cite{ephrin_tectum} as well as in other
+retina~\cite{braisted_graded_1997}, tectum~\cite{braisted_graded_1997,feldheim_genetic_2000} as well as in other
 sensory systems, such as the somatosensory
-system~\cite{vanderhaegen_as_in_elife_paper}. The Ephrin ligands have a clear
-effect on axonal outgrowth, as shown in in-vitro~\cite{one} and
-in-vivo~\cite{another} studies.
+system~\cite{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
+effect on axonal outgrowth, as shown in in-vitro~\cite{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
+in-vivo~\cite{rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
 %
@@
 
-Our model consists of agents, representing axonal growth cones, originating
-from a square retina and moving upon a square tectum. Each agent carries a set
-of 4 receptors, $r$, and 4 ligands, $l$, which it expresses at differing
-levels depending upon the soma's origin on the retina. The expression levels
-of the receptors and ligands form gradients across the retinal
-surface. Receptors are arranged in orthogonal pairs with the gradient of $r_0$
-being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$
-is orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs, with
-$l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$
-orthogonal to $l_0$ and opposing $r_1$. While we do not explicitly name $r_0$,
-$r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes
-EphA, EphB and Ryk~\cite{schmitt_wntryk_2006} receptors and that $l$ includes
-the ephrin-A, ephrin-B and Wnt3~\cite{schmitt_wntryk_2006} ligands.
+\subsection{Receptor and ligand expression}
+
+Our model was inspired by the agent-based model of \cite{simpson_simple_2011}
+and, in common with that work, consists of agents representing axonal growth
+cones, originating from a square retina and moving upon a square tectum. In
+our model, each retinal ganglion cell projects $n=8$ growth cone agents which
+carry a set of 4 receptors, $r$, and 4 ligands, $l$, expressed at levels
+determined by the cell's location on the retina. The expression levels of the
+receptors and ligands vary with respect to the cell's position on the retinal
+surface. Each receptor or ligand varies with respect only to one dimension.
+Receptor expression gradients are arranged in orthogonal pairs with the
+gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is
+opposite to $r_0$, is orthogonal to $r_3$. Ligands are also arranged in
+orthogonal pairs, with $l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$
+decreases) and $l_1$ orthogonal to $l_0$ and opposing $r_1$.
+
+Because there is evidence that EphA and EphB receptors are expressed in
+exponentially increasing
+patterns \cite{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
+we use an exponential form for retinal receptor expressions, in common with
+other modelling
+studies~\cite{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
 %
 The tectum expresses the same sets of receptors and ligands, also in
-orthogonal pairs of gradients, though not necessarily with the same expression
-function. Figure~\ref{f:1} shows colour maps of receptor and ligand
-expression, in this case experssion maps are linear.
+orthogonal pairs of gradients. In this work, we did not consider the effect of
+reverse signalling from tectal receptors to RGC ligands so we only modelled
+tectal ligand expression. Although
+several studies model tectal ligand expression with exponential functions~\cite{koulakov_stochastic_2004}, the
+experimental evidence for ligand expression is more ambiguous (Supp. Fig
+X). Correspondingly, we leave open the possibility that ligand expression may
+be best modelled by exponential, linear or logarithmic functions.
+
+While we do not
+explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion
+is that $r$ includes EphA, EphB and Ryk~\cite{schmitt_wntryk_2006} receptors
+and that $l$ includes the ephrin-A, ephrin-B and
+Wnt3~\cite{schmitt_wntryk_2006} ligands.
+
+Figure~\ref{f:1} shows colour maps of receptor and ligand
+expression, in this case tectal ligand expression maps are linear.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expressions_fig.png}
+\caption{Retinal and tectal gene expression.
+%
+\textbf{A} Retinal receptor expression. Four receptors are expressed and
+displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and
+green;$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an
+exponential function of $x$; $r_1$ increases with an exponential function of
+$y$. $r_2$ and $r_3$ have the opposite sense. The expression is modelled with
+exponential functions because there is ample evidence that shows that retinal
+receptor expressions follow this form.
+%
+\textbf{B} Retinal ligand expression. May not be included.
+%
+\textbf{C} Tectal receptor expression. May not be included.
+%
+\textbf{D} Tectal ligand expression. $l_0$ and $l_1$ are red and green; $l_2$
+and $l_3$ are cyan and magenta. These expressions are shown in a linear form,
+as there is less evidence to determine whether tectal ligand expression is
+exponential, linear or logarithmic in form.
+}
+\label{f:1}
+\end{figure}
+
+\subsection{Signalling}
+
+This is where I really want to say ``the only interactions in the model are
+via forward (ligand-activates-receptor) and reverse
+(receptor-activates-ligand) signalling''.
+
+% Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
 
@@
 \selectlanguage{English}
-\bibliographystyle{abbrvnotitle}
+\bibliographystyle{apalike}
 \bibliography{RetinoTectal}

### 2021-07-05 paper/paper.tex
More paper scrawlings

@@
 effect on axonal outgrowth, as shown in in-vitro~\cite{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
-in-vivo~\cite{rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
+in-vivo~\cite{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
 %
@@
 
+There have been many approaches to modelling the organisation of the
+retino-tectal projection. Many of these models are wholly or partially
+phenomenological.
+
 \section{Model}
+
+Our model was inspired by the agent-based model of \cite{simpson_simple_2011}
+and, in common with that work, consists of agents representing axonal growth
+cones, originating from a square retina and moving upon a square tectum. Also
+in common with the Simpson \& Goodhill model, we model three contributions to
+the movement of the growth cones; chemoaffinity, competition and axon-axon
+interaction. In our model, chemoaffinity and axon-axon interactions occur via
+signal transmission between receptors and ligands which are expressed in
+graded patterns on both retina and tectum. This sets our model apart from that
+of Simpson \& Goodhill, in which each growth cone moving on the tectum was
+always able to determine a vector to its designated target and advance along
+that vector. Competition is modelled as a distance based repulsion as
+in \cite{simpson_simple_2011}.
 
@@
 
-Our model was inspired by the agent-based model of \cite{simpson_simple_2011}
-and, in common with that work, consists of agents representing axonal growth
-cones, originating from a square retina and moving upon a square tectum. In
-our model, each retinal ganglion cell projects $n=8$ growth cone agents which
-carry a set of 4 receptors, $r$, and 4 ligands, $l$, expressed at levels
-determined by the cell's location on the retina. The expression levels of the
-receptors and ligands vary with respect to the cell's position on the retinal
-surface. Each receptor or ligand varies with respect only to one dimension.
-Receptor expression gradients are arranged in orthogonal pairs with the
-gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is
-opposite to $r_0$, is orthogonal to $r_3$. Ligands are also arranged in
-orthogonal pairs, with $l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$
-decreases) and $l_1$ orthogonal to $l_0$ and opposing $r_1$.
+Each retinal ganglion cell projects $n=8$ growth cone agents which carry a set
+of 4 receptors, $r$, and 4 ligands, $l$, expressed at levels determined by the
+cell's location on the retina. The expression levels of the receptors and
+ligands vary with respect to the cell's position on the retinal surface. Each
+receptor or ligand varies with respect only to one dimension.  Receptor
+expression gradients are arranged in orthogonal pairs with the gradient of
+$r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to
+$r_0$, is orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs,
+with $l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$
+orthogonal to $l_0$ and opposing $r_1$.
 
-Because there is evidence that EphA and EphB receptors are expressed in
-exponentially increasing
+Because there is convincing evidence that EphA and EphB receptors are
+expressed in exponentially increasing
 patterns \cite{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
@@
 reverse signalling from tectal receptors to RGC ligands so we only modelled
-tectal ligand expression. Although
-several studies model tectal ligand expression with exponential functions~\cite{koulakov_stochastic_2004}, the
+tectal ligand expression. Although several studies model tectal ligand
+expression with exponential functions~\cite{koulakov_stochastic_2004}, the
 experimental evidence for ligand expression is more ambiguous (Supp. Fig
 X). Correspondingly, we leave open the possibility that ligand expression may
-be best modelled by exponential, linear or logarithmic functions.
+be modelled by exponential, linear or logarithmic functions.
 
-While we do not
-explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion
-is that $r$ includes EphA, EphB and Ryk~\cite{schmitt_wntryk_2006} receptors
-and that $l$ includes the ephrin-A, ephrin-B and
-Wnt3~\cite{schmitt_wntryk_2006} ligands.
+While we do not explicitly name $r_0$, $r_1$,
+etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
+EphB, Ryk~\cite{schmitt_wntryk_2006} and
+Neogenin~\cite{rajagopalan_neogenin_2004} receptors and that $l$ includes the
+ephrin-A, ephrin-B, Wnt3~\cite{schmitt_wntryk_2006} and
+RGM~\cite{monnier_rgm_2002} ligands, each of which has been shown to play a
+role in retino-tectal map formation.
 
@@
 
+In this model, the signal transmitted when a ligand binds to a receptor on a
+growth cone can lead to one of two effects; the signal may induce the growth
+cone to climb the gradient of ligand expression, in which case the interaction
+is \emph{attractive} or the cone may descend the gradient of ligand expression
+if the interaction is \emph{repulsive}.
+
 This is where I really want to say ``the only interactions in the model are
 via forward (ligand-activates-receptor) and reverse
-(receptor-activates-ligand) signalling''.
-
-% Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
+(receptor-activates-ligand) signalling''. Somehthing like:
 
@@
 
+% Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
+
 %

### 2021-07-06 paper/paper.tex
Corrects positionining in m_ee.json with e_wt.json

@@
 \title {
-  Competition can provide a stopping mechanism for the retinotectal
-  projection \\
-  or \\
-  Competition provides a stopping mechanism for the chemoaffinity
+  Competition can account for stopping in a gradient-following model
+  of the retinotectal projection \\
+  \emph{or} \\
+  Competition provides stopping for the chemoaffinity
   theory which is robust to noise \\
-  or \\
+  \emph{or} \\
   Self-organisation of topographically ordered axon connections can take place
-  in a noisy environment if axons compete
+  in a noisy environment if axons compete \\
+  Time to stop; modelling chemoaffinity with competition \\
 }
@@
 \author{\Authors}
+
+% Simplified chemoaffinity explains retinotectal mapping via local axon-axon
+% interactions
+%
+% Focus on the chemoaffinity idea for this model. Show that local interactions
+% can provide an account for growth cones finding their targets.
 
@@
 
+Model which carefully deals with receptor
+binding: \cite{naoki_revisiting_2017}.
+
+Gradient based models: \cite{nakomoto_topographically_1996}
+
 \section{Model}
@@
 via forward (ligand-activates-receptor) and reverse
-(receptor-activates-ligand) signalling''. Somehthing like:
+(receptor-activates-ligand) signalling''. Something like:
 
@@
 % Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
+% This relates to the work \cite{naoki_revisiting_2017}
 

### 2021-07-06 sim/agent/agent1.h

@@ -240,7 +240,17 @@ struct Agent1
                                                 const std::string& tag,
                                                 expression_view exview, size_t pair_to_view, bool alt_cmap=false)
     {
-        tissuevisual<float, N>* tv = new tissuevisual<float, N>(this->tvv->shaderprog, this->tvv->tshaderprog, gtissue, offset);
+        return this->createTissueVisual (this->tvv->shaderprog, this->tvv->tshaderprog,
+                                         offset, gtissue, tag, exview, pair_to_view, alt_cmap);
+    }
+
+    //! Create a tissue visual, to reduce boilerplate code in init(). Use either shader/tshader progs
+    tissuevisual<float, N>* createTissueVisual (GLuint shad_prog, GLuint tshad_prog,
+                                                morph::Vector<T,3>& offset, guidingtissue<T, N>* gtissue,
+                                                const std::string& tag,
+                                                expression_view exview, size_t pair_to_view, bool alt_cmap=false)
+    {
+        tissuevisual<float, N>* tv = new tissuevisual<float, N>(shad_prog, tshad_prog, gtissue, offset);
         tv->view = exview;
         tv->pair_to_view = pair_to_view;
         tv->cm.setType (morph::ColourMapType::Duochrome);
@@ -262,6 +272,8 @@ struct Agent1
             ss << tag << " ligand gradient "     << (pair_to_view*2) << "/" << (1+pair_to_view*2) << " x";
         } else if (exview == expression_view::ligand_grad_y) {
             ss << tag << " ligand gradient "     << (pair_to_view*2) << "/" << (1+pair_to_view*2) << " y";
+        } else if (exview == expression_view::cell_positions) {
+            ss << tag << " cell positions";
         }
         tv->addLabel (ss.str(), {0.0f, 1.15f, 0.0f});
 
@@ -580,16 +592,15 @@ struct Agent1
 
             this->ax_centroids.p[ri] += initpos / static_cast<T>(bpa);
 
-            // The target for axon centroids is defined by their origin location on the
-            // retina. However, their target on the tectum is the retinal position
-            // *transformed*. That means rotating the retinal positions by 90 degrees
-            // clockwise. ALSO, if experimental manipulations have been made, then the
-            // target positions will need to be modified accordingly.
+            // "experiment suggests": The target for axon centroids is defined by their
[…]

### 2021-07-06 sim/agent/branch.h

@@ -47,25 +47,25 @@ struct branch : public branch_base<T,N>
             // First, find the ligand gradients close to the current location b.
             morph::Vector<T, 8> lg = tissue->lgnd_grad_at (b);
             // 4 receptors and 4 ligand gradients.
-            // let receptor 0 interact primarily with ligand 1 [gradients (2,3)]
-            // let receptor 1 interact primarily with ligand 0 [gradients (0,1)]
-            // let receptor 2 interact primarily with ligand 3 [gradients (6,7)] // opposite to receptor 0
-            // let receptor 3 interact primarily with ligand 2 [gradients (4,5)] // opposite to receptor 1
-            G[0] = this->rcpt[0] * (source_tissue->forward_interactions[0] == interaction::repulsion ? -lg[2] : lg[2])
-            + this->rcpt[1] * (source_tissue->forward_interactions[1] == interaction::repulsion ? -lg[0] : lg[0])
-            + this->rcpt[2] * (source_tissue->forward_interactions[2] == interaction::repulsion ? -lg[6] : lg[6])
-            + this->rcpt[3] * (source_tissue->forward_interactions[3] == interaction::repulsion ? -lg[4] : lg[4]);
+            // let receptor 0 interact primarily with ligand 0 [gradients (0,1)]
+            // let receptor 1 interact primarily with ligand 1 [gradients (2,3)]
+            // let receptor 2 interact primarily with ligand 2 [gradients (4,5)]
+            // let receptor 3 interact primarily with ligand 3 [gradients (6,7)]
+            G[0] = this->rcpt[0] * (source_tissue->forward_interactions[0] == interaction::repulsion ? -lg[0] : lg[0])
+            + this->rcpt[1] * (source_tissue->forward_interactions[1] == interaction::repulsion ? -lg[2] : lg[2])
+            + this->rcpt[2] * (source_tissue->forward_interactions[2] == interaction::repulsion ? -lg[4] : lg[4])
+            + this->rcpt[3] * (source_tissue->forward_interactions[3] == interaction::repulsion ? -lg[6] : lg[6]);
 
-            G[1] = this->rcpt[0] * (source_tissue->forward_interactions[0] == interaction::repulsion ? -lg[3] : lg[3])
-            + this->rcpt[1] * (source_tissue->forward_interactions[1] == interaction::repulsion ? -lg[1] : lg[1])
[…]

### 2021-07-08 paper/paper.tex
WIP thinking about receptor binding model

@@
 \setmainfont{Times New Roman}
+
+% Nice citations
+\usepackage{natbib}
 
@@
 \newcommand{\e}{\emph}
-\newcommand{\mycite}[1]{\cite{#1}}
 \newcommand{\code}[1]{\textsf{#1}}
@@
 mid-twentieth century, leading to Sperry's 1963 summary of
-the \emph{chemoaffinity theory}~\cite{sperry_chemoaffinity_1963} which
+the \emph{chemoaffinity theory} \citep{sperry_chemoaffinity_1963} which
 proposes the existence of morphogenetic gradients that guide axons to their
 destination. The chemoaffinity theory was given robust support by the
-discovery of the ephrin ligands and their receptors~\cite{cheng_complementary_1995,drescher_vitro_1995}
+discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995}
 which have been shown to form into graded expression fields in the
-retina~\cite{braisted_graded_1997}, tectum~\cite{braisted_graded_1997,feldheim_genetic_2000} as well as in other
+retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other
 sensory systems, such as the somatosensory
-system~\cite{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
-effect on axonal outgrowth, as shown in in-vitro~\cite{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
-in-vivo~\cite{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
+system \citep{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
+effect on axonal outgrowth, as shown in in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
+in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
 %
@@
 axons growing from the thalamus to the somatosensory cortex into the well
-known murine barrel cortex pattern~\cite{james_modelling_2020}. In
+known murine barrel cortex pattern~\citep{james_modelling_2020}. In
 characterising that system, we explored the effect of various types of
@@
 could cause topological defects. The question of noise in axon guidance has
-been explored by Goodhill~\cite{goodhill_can_2016}.
+been explored by \citet{goodhill_can_2016}.
 
@@
 tectal surface. We show that a combination of exponentially graded expression
-in the retina~\cite{reber_relative_2004} and exponentially (for repulsively
+in the retina \citep{reber_relative_2004} and exponentially (for repulsively
 interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
@@
 
-Model which carefully deals with receptor
-binding: \cite{naoki_revisiting_2017}.
+A model which carefully deals with receptor
+binding is given by \citet{naoki_revisiting_2017} \citep[see also][]{mortimer_bayesian_2009}.
 
-Gradient based models: \cite{nakomoto_topographically_1996}
+Gradient based models: \citet{nakomoto_topographically_1996}
 
@@
 
+Based on a Michaelis-Menten model applied to
+receptor-binding, \citet{mortimer_bayesian_2009} give an expression for the
+probability of an individual receptor being bound as a function of the ligand
+expression, $\gamma$, and the ligand gradient, $\mu$;
+\begin{equation}
+P(b=1|\gamma,\mu) = \frac{\gamma(1 + \mu r)}{1 + \gamma(1 + \mu r)},
+\end{equation}
+where $r$ is the distance of the receptor from the centre of a 1D growth
+cone. To determine the sampled direction signal (for one dimension), we assume
+evenly spaced receptors and sample from a uniform distribution for each
+location across the growth cone.
+
 This is where I really want to say ``the only interactions in the model are
@@
 a signal into the receptor-expressing cell. Reverse signals occur when a
-membrane-bound ligand is activated~\cite{wu_role_2019}.
+membrane-bound ligand is activated~\citep{wu_role_2019}.
 

### 2021-07-09 paper/paper.tex
Just work in progress

@@
 is \emph{attractive} or the cone may descend the gradient of ligand expression
-if the interaction is \emph{repulsive}.
+if the interaction is \emph{repulsive}. To determine the strength of the
+effect we could assume a purely linear receptor binding model, and set the
+movement of the growth cone to be $\sum_i \pm r_i \times \grad l_i$. While
+this approach works, it does not allow us to investigate the effect of
+receptor binding noise (\textbf{FIXME}: Be sure to have included a discussion
+on the three types of noise as described by Goodhill, 2016 in the
+introduction) on the ability of the system to generate topographic order.
 
-Based on a Michaelis-Menten model applied to
-receptor-binding, \citet{mortimer_bayesian_2009} give an expression for the
+Instead, we treat the receptor binding at the growth cone with greater
+care. Based on a Michaelis-Menten model applied to
+receptor binding, \citet{mortimer_bayesian_2009} give an expression for the
 probability of an individual receptor being bound as a function of the ligand
@@
 where $r$ is the distance of the receptor from the centre of a 1D growth
-cone. To determine the sampled direction signal (for one dimension), we assume
+cone. To determine the growth cone's estimate of the gradient, we assume
 evenly spaced receptors and sample from a uniform distribution for each
-location across the growth cone.
+location across the growth cone. Activated receptors contribute
 
-This is where I really want to say ``the only interactions in the model are
-via forward (ligand-activates-receptor) and reverse
-(receptor-activates-ligand) signalling''. Something like:
+%This is where I really want to say ``the only interactions in the model are
+%via forward (ligand-activates-receptor) and reverse
+%(receptor-activates-ligand) signalling''. Something like:
 
-Interactions in the model are by receptor-ligand interactions only. Forward
-interactions are those in which the ligand activates the receptor, delivering
-a signal into the receptor-expressing cell. Reverse signals occur when a
-membrane-bound ligand is activated~\citep{wu_role_2019}.
+%Interactions in the model are by receptor-ligand interactions only. Forward
+%interactions are those in which the ligand activates the receptor, delivering
+%a signal into the receptor-expressing cell. Reverse signals occur when a
+%membrane-bound ligand is activated~\citep{wu_role_2019}.
 

### 2021-07-09 sim/agent/branch.h

@@ -5,7 +5,6 @@
 #include <vector>
 #include <morph/vVector.h>
 #include <morph/Vector.h>
-#include <morph/Random.h>
 #include "tissue.h"
 
 // A choice of schemes to deal with axon branches that try to move outside the tissue domain
@@ -31,6 +30,15 @@ struct branch : public branch_base<T,N>
     // Signalling ratio parameter
     static constexpr T s = T{1.1};
 
+    // Estimate the ligand gradient, given the true ligand gradient
+    virtual morph::Vector<T, 2*N> estimate_ligand_gradient(morph::Vector<T,2*N>& true_lgnd_grad,
+                                                           morph::Vector<T,N>& true_lgnd_exp)
+    {
+        // This is the non-stochastic implementation
+        morph::Vector<T, 2*N> lg = true_lgnd_grad;
+        return lg;
+    }
+
     // Compute the next position for this branch, using information from all other
     // branches and the parameters vector, m. Will also need location on target tissue,
     // to get its ephrin values.
@@ -47,11 +55,12 @@ struct branch : public branch_base<T,N>
         // Chemoaffinity, graded by origin position (i.e termination zone) of each retinal axon
         morph::Vector<T, 2> G;
         if constexpr (N==4) {
-            // First, find the ligand gradients close to the current location b.
-            morph::Vector<T, 8> lg = tissue->lgnd_grad_at (b) * rns;
-            // Adding noise could be as simple as multiplying by a Vector of numbers
-            // drawn from a Gaussian distribution with mean 1. But need to think
-            // carefully and be prepared to justify it with reference to the Mortimer paper.
+            // First, find the ligand gradients close to the current location b.  Adding
+            // noise could be as simple as multiplying by a Vector of numbers drawn from
+            // a Gaussian distribution with mean 1. But need to think carefully and be
+            // prepared to justify it with reference to the Mortimer paper. Updated: I
+            // have a possibly better way to do this. See branch_stochastic.h.
+            morph::Vector<T, 8> lg = this->estimate_ligand_gradient (tissue->lgnd_grad_at (b), tissue->lgnd_at (b));
 
             // 4 receptors and 4 ligand gradients.
             // let receptor 0 interact primarily with ligand 0 [gradients (0,1)]
@@ -315,5 +324,39 @@ struct branch : public branch_base<T,N>
[…]

### 2021-07-14 paper/paper.tex
WIP on paper now I know all interactions are receptor-ligand based

@@
 
-\emph{Abstract here.}
+% Note this is the SfN draft abstract, and will need rewriting.
+\emph{In the classic Chemoaffinity theory, the retinotectal axon projection is
+thought to use pairs of orthogonal signalling gradients in the retina to
+specify the eventual location of synapses made on the surface of the
+tectum/superior colliculus. Similar orthogonal gradients in the tectum provide
+a coordinate system which allows the axons to match their prespecified
+destination with the correct location. Although the Ephrins have been shown to
+guide axons toward their destination, there has yet to emerge a complete
+account of the local interactions which halt the axonal growth cones in the
+correct locations to recreate the topography of the retinal cells. The model
+of Simpson & Goodhill (2011) provides an account of the basic topographic
+arrangement of cells on the tectum, as well as reproducing well known surgical
+and genetic manipulation experiments. However, it suffers from the absence of
+a local chemotactic guidance mechanism. Instead, each agent in their model is
+given instantaneous knowledge of the vector that would move it toward its pre
+specified destination. In addition to the globally supervised chemoaffinity
+term, Simpson & Goodhill introduced a competitive interaction for space
+between growth cone agents and a receptor-ligand axon-axon interaction in
+order to account for the full set of experimental manipulations. Here, we
+propose the replacement of the chemoaffinity term with a gradient following
+model consisting of axonal growth cone agents which carry receptor molecule
+expression determined by their soma's location of origin on the retina. Growth
+cones move on the simulated tectum guided by two pairs of opposing, orthogonal
+signalling molecules representing the Ephrin ligands. We show that with only
+the chemoaffinity term and a receptor-ligand based axon-axon interaction term
+(meaning that all growth cone interactions are by receptor-ligand signalling),
+a full range of experimental manipulations to the retinotectal system can be
+reproduced. Furthermore, we show that the observation that competition is not
+and essential requirement for axons to find their way (Gosse et al., 2008) is
+also accounted for by the model, due to the opposing influences of signalling
+gradient pairs. Finally, we demonstrate that, assuming exponentially varying
+receptor expression in the retina, ligand expression should either be
+exponential if the receptor-ligand signal induces repulsion (i.e. gradient
+descent) or logarithmic if the signal induces attraction (gradient
+ascent). Thus, we find that a model analogous to the one we presented in James
+et al (2020) that accounts for murine barrel patterning is also a candidate
+mechanism for the arrangement of the more continuous retinotectal system.}
 
@@
 and, in common with that work, consists of agents representing axonal growth
-cones, originating from a square retina and moving upon a square tectum. Also
-in common with the Simpson \& Goodhill model, we model three contributions to
-the movement of the growth cones; chemoaffinity, competition and axon-axon
-interaction. In our model, chemoaffinity and axon-axon interactions occur via
-signal transmission between receptors and ligands which are expressed in
-graded patterns on both retina and tectum. This sets our model apart from that
-of Simpson \& Goodhill, in which each growth cone moving on the tectum was
-always able to determine a vector to its designated target and advance along
-that vector. Competition is modelled as a distance based repulsion as
-in \cite{simpson_simple_2011}.
+cones, originating from a square retina and moving upon a square tectum.
+%
+\citet{simpson_simple_2011} models three contributions to the movement of the growth
+cones; chemoaffinity, competition for space and axon-axon interactions. The
+first two of these are modelled without reference to a particular mechanism;
+each agent is provided with globally-supervised information about its
+instantaneous location and its target location, allowing it to move along a
+vector towards the target. Competition is implemented as a movement of
+constant distance along the vector between two agents if two agents are within
+a threshold distance of one another. Interaction is signal-induced repulsion,
+which is triggered if two agents are both within the threshold distance of
+each other and have compatible EphA receptor expression.
+
+In our model, space-based competition is omitted and chemoaffinity and
+axon-axon interactions occur only via signal transmission between receptors
+and ligands, which are expressed in graded patterns on both retina and
+tectum. Signals induce attractive or repulsive interactions; for chemoaffinity
+the interaction is to ascend or descend the gradient of ligand expression on
+the tectum; for axon-axon interactions, the strength of short-range repulsion
+(or attraction) is governed by the receptor-ligand expression on pairs of axons.
 
@@
 expressed in exponentially increasing
-patterns \cite{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
+patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
 we use an exponential form for retinal receptor expressions, in common with
 other modelling
-studies~\cite{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
+studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
 %
@@
 tectal ligand expression. Although several studies model tectal ligand
-expression with exponential functions~\cite{koulakov_stochastic_2004}, the
+expression with exponential functions \citep{koulakov_stochastic_2004}, the
 experimental evidence for ligand expression is more ambiguous (Supp. Fig
@@
 etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
-EphB, Ryk~\cite{schmitt_wntryk_2006} and
-Neogenin~\cite{rajagopalan_neogenin_2004} receptors and that $l$ includes the
-ephrin-A, ephrin-B, Wnt3~\cite{schmitt_wntryk_2006} and
-RGM~\cite{monnier_rgm_2002} ligands, each of which has been shown to play a
+EphB, Ryk \citep{schmitt_wntryk_2006} and
+Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $l$ includes the
+ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
+RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
 role in retino-tectal map formation.
@@
 In this model, the signal transmitted when a ligand binds to a receptor on a
-growth cone can lead to one of two effects; the signal may induce the growth
-cone to climb the gradient of ligand expression, in which case the interaction
-is \emph{attractive} or the cone may descend the gradient of ligand expression
-if the interaction is \emph{repulsive}. To determine the strength of the
-effect we could assume a purely linear receptor binding model, and set the
-movement of the growth cone to be $\sum_i \pm r_i \times \grad l_i$. While
-this approach works, it does not allow us to investigate the effect of
-receptor binding noise (\textbf{FIXME}: Be sure to have included a discussion
-on the three types of noise as described by Goodhill, 2016 in the
-introduction) on the ability of the system to generate topographic order.
+growth cone can lead to one of two effects; the signal may induce an
+attraction towards the ligand-expression region or a repulsion away from it.
+%
+For axon-tectum interactions, we implement attraction as a climbing of the
+tectal ligand expression gradient and repulsion as gradient descent.
+%
+To determine the strength of the effect we assume a purely linear receptor
+binding model, and set the movement of the growth cone to be $\sum_i \pm
+r_i \times \grad l_i$.
 
-Instead, we treat the receptor binding at the growth cone with greater
-care. Based on a Michaelis-Menten model applied to
-receptor binding, \citet{mortimer_bayesian_2009} give an expression for the
-probability of an individual receptor being bound as a function of the ligand
-expression, $\gamma$, and the ligand gradient, $\mu$;
-\begin{equation}
-P(b=1|\gamma,\mu) = \frac{\gamma(1 + \mu r)}{1 + \gamma(1 + \mu r)},
-\end{equation}
-where $r$ is the distance of the receptor from the centre of a 1D growth
-cone. To determine the growth cone's estimate of the gradient, we assume
-evenly spaced receptors and sample from a uniform distribution for each
-location across the growth cone. Activated receptors contribute
+For axon-axon interactions, attraction causes a movement along a vector
+between two axons causing them to move closer together; repulsion cause the
+opposite effect.
 
@@
 \end{document}
+
+%% For the noise based paper:
+%% Instead, we treat the receptor binding at the growth cone with greater
+%% care. Based on a Michaelis-Menten model applied to
+%% receptor binding, \citet{mortimer_bayesian_2009} give an expression for the
+%% probability of an individual receptor being bound as a function of the ligand
+%% expression, $\gamma$, and the ligand gradient, $\mu$;
+%% \begin{equation}
+%% P(b=1|\gamma,\mu) = \frac{\gamma(1 + \mu r)}{1 + \gamma(1 + \mu r)},
+%% \end{equation}
+%% where $r$ is the distance of the receptor from the centre of a 1D growth
+%% cone. To determine the growth cone's estimate of the gradient, we assume
+%% evenly spaced receptors and sample from a uniform distribution for each
+%% location across the growth cone. Activated receptors contribute

### 2021-07-15 paper/paper.tex
WIP on paper methods and a simplification of axon-axon interactions

@@
 %
-For axon-tectum interactions, we implement attraction as a climbing of the
+For axon-tectum interactions, we implemented attraction as a climbing of the
 tectal ligand expression gradient and repulsion as gradient descent.
 %
-To determine the strength of the effect we assume a purely linear receptor
-binding model, and set the movement of the growth cone to be $\sum_i \pm
-r_i \times \grad l_i$.
+To determine the strength of the effect we assumed a purely linear receptor
+binding model, and set the chemotactic movement vector of the growth cone to
+be
+%
+\begin{equation}
+\mathbf{G} = \sum_i^N D_i r_i \nabla l_i
+\end{equation}
+%
+where $r_i$ is the receptor expression for ligand-receptor pair $i$,
+$\nabla l_i$ is the gradient of the ligand expression and $D_i$ denotes the
+direction of the interaction induced when a molecule of $l_i$ binds to
+$r_i$. $D_i$ takes the value -1 for a repulsive interaction or 1 for an
+attractive interaction.
 
-For axon-axon interactions, attraction causes a movement along a vector
-between two axons causing them to move closer together; repulsion cause the
-opposite effect.
+For axon-axon interactions, attraction causes a movement along a vector,
+$\mathbf{x}_{b_0}$, between two axons causing them to move closer together;
+repulsion causes the opposite effect. The interaction, $\mathbf{I}$, is given
+by
+%
+\begin{equation}
+\mathbf{I} = \sum_b \mathbf{x}_{b_0} Q(d_{b0})
+\end{equation}
+%
+where $Q(d_{b0})$ is the signalling strength between two growth cones, given by
+%
+\begin{equation}
+Q(d_{b0}) = \begin{cases}
+     \sum_i^N D_i l_i r_i    & d_{b0} < 2r \\
+     0 & d_{b0} \geq 2r
+     \end{cases}
+\end{equation}
 

### 2021-07-15 sim/agent/branch.h

@@ -96,8 +96,12 @@ struct branch : public branch_base<T,N>
         // here, sum up the unit vectors kb.
         morph::Vector<T, 2> kb = this->current - kp->current;
         T d = kb.length();
-        T W = d <= this->two_r ? (T{1} - d/this->two_r) : T{0};
+
+        // Q collects the overall signal transmitted by ligands binding to
+        // receptors. Repulsive interactions add to Q; attractive interactions make Q
+        // more negative.
         T Q = T{0};
+
         if constexpr (N == 4) {
             // Forward signalling is activation of the receptor by the ligand. Treat as a multiplicative signal.
             Q = kp->lgnd[0] * this->rcpt[0] * (source_tissue->forward_interactions[0] == interaction::repulsion ? 1 : -1)
@@ -112,12 +116,16 @@ struct branch : public branch_base<T,N>
 
         kb.renormalize(); // vector bk is a unit vector
 
-        I += kb * Q * W;
+        // I is the axon-axon interaction. Q is the signal. Interaction occurs if distance is <= two_r
+        I += kb * Q * (d <= this->two_r ? T{1} : T{0});
 
-        C += kb * W; // Generic, distance based competition (unused, mostly)
+        // W is a distance-dependent weight, which is 0 outside a distance of two_r and
+        // linearly increases to 1 when d=0.
+        T W = d <= this->two_r ? (T{1} - d/this->two_r) : T{0};
+        C += kb * W; // Generic, distance based competition
 
         // In client code, should we add to n_k or not?
-        return (W > T{0} ? T{1} : T{0});
+        return (d <= this->two_r ? T{1} : T{0});
     }
 
     // A subroutine of compute_next

### 2021-07-15 paper/paper.tex
WIP on paper; have a draft model section now

--- 2021-07-14 paper/paper.tex
@@
 correct locations to recreate the topography of the retinal cells. The model
-of Simpson & Goodhill (2011) provides an account of the basic topographic
+of \citet{simpson_simple_2011} provides an account of the basic topographic
 arrangement of cells on the tectum, as well as reproducing well known surgical
@@
 specified destination. In addition to the globally supervised chemoaffinity
-term, Simpson & Goodhill introduced a competitive interaction for space
-between growth cone agents and a receptor-ligand axon-axon interaction in
-order to account for the full set of experimental manipulations. Here, we
+term, \citet{simpson_simple_2011} introduced a competitive interaction for
+space between growth cone agents and a receptor-ligand axon-axon interaction
+in order to account for the full set of experimental manipulations. Here, we
 propose the replacement of the chemoaffinity term with a gradient following
@@
 reproduced. Furthermore, we show that the observation that competition is not
-and essential requirement for axons to find their way (Gosse et al., 2008) is
-also accounted for by the model, due to the opposing influences of signalling
-gradient pairs. Finally, we demonstrate that, assuming exponentially varying
-receptor expression in the retina, ligand expression should either be
-exponential if the receptor-ligand signal induces repulsion (i.e. gradient
-descent) or logarithmic if the signal induces attraction (gradient
-ascent). Thus, we find that a model analogous to the one we presented in James
-et al (2020) that accounts for murine barrel patterning is also a candidate
-mechanism for the arrangement of the more continuous retinotectal system.}
+and essential requirement for axons to find their
+way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
+the opposing influences of signalling gradient pairs. Finally, we demonstrate
+that, assuming exponentially varying receptor expression in the retina, ligand
+expression should either be exponential if the receptor-ligand signal induces
+repulsion (i.e. gradient descent) or logarithmic if the signal induces
+attraction (gradient ascent). Thus, we find that a model analogous to the one
+we presented in \citet{james_modelling_2020} that accounts for murine barrel
+patterning is also a candidate mechanism for the arrangement of the more
+continuous retinotectal system.}
 
@@
 There have been many approaches to modelling the organisation of the
-retino-tectal projection. Many of these models are wholly or partially
+retinotectal projection. Many of these models are wholly or partially
 phenomenological.
@@
 
-Gradient based models: \citet{nakomoto_topographically_1996}
+Gradient based models: \citet{nakamoto_topographically_1996}
 
@@
 %
-\citet{simpson_simple_2011} models three contributions to the movement of the growth
+\citet{simpson_simple_2011} model three contributions to the movement of the growth
 cones; chemoaffinity, competition for space and axon-axon interactions. The
@@
 vector towards the target. Competition is implemented as a movement of
-constant distance along the vector between two agents if two agents are within
-a threshold distance of one another. Interaction is signal-induced repulsion,
-which is triggered if two agents are both within the threshold distance of
-each other and have compatible EphA receptor expression.
+constant distance along the vector between two agents if those two agents are
+within a threshold distance of one another. Interaction is signal-induced
+repulsion, which is triggered if two agents are both within the threshold
+distance of each other and have compatible EphA receptor expression.
 
@@
 
-Each retinal ganglion cell projects $n=8$ growth cone agents which carry a set
-of 4 receptors, $r$, and 4 ligands, $l$, expressed at levels determined by the
-cell's location on the retina. The expression levels of the receptors and
-ligands vary with respect to the cell's position on the retinal surface. Each
-receptor or ligand varies with respect only to one dimension.  Receptor
-expression gradients are arranged in orthogonal pairs with the gradient of
-$r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to
-$r_0$, is orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs,
-with $l_0$ opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$
-orthogonal to $l_0$ and opposing $r_1$.
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expressions_fig.png}
+\caption{Simulated retinal and tectal signalling molecule expression. Four
+receptor types are expressed, and four corresponding ligand types. Receptors
+of type $i$ are activated by ligands of type $i$.
+%
+\textbf{A} Retinal receptor expression, $r_i$. Four receptors are expressed and
+displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
+$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential
+function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$
+and $r_3$ have the opposite sense. Retinal receptor expression is modelled
+with exponential functions because there is ample evidence to support this
+form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
+%
+\textbf{B} Retinal ligand expression, $l_i$, is expressed in complimentary
+gradients \citep{hornberger_modulation_1999}.
+%
+\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown in a linear form,
+as there is less evidence to determine whether ligand expression is
+exponential, linear or logarithmic on the tectum.
+$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
+}
+\label{f:1}
+\end{figure}
+
+Each retinal ganglion cell projects $n=8$ growth cone agents (referred to
+as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands, $l_i$,
+expressed at levels determined by the cell's location on the retina. The
+expression levels of the receptors and ligands vary with respect to the cell's
+position on the retinal surface. Each receptor or ligand varies with respect
+only to one dimension.  Receptor expression gradients are arranged in
+orthogonal pairs with the gradient of $r_0$ being orthogonal to that of
+$r_1$. $r_2$, whose gradient is opposite to $r_0$, is orthogonal to
+$r_3$. Ligands are also arranged in orthogonal pairs, with $l_0$ opposing
+$r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$ orthogonal to $l_0$ and
+opposing $r_1$.
 
@@
 %
-The tectum expresses the same sets of receptors and ligands, also in
-orthogonal pairs of gradients. In this work, we did not consider the effect of
-reverse signalling from tectal receptors to RGC ligands so we only modelled
-tectal ligand expression. Although several studies model tectal ligand
+The simulated tectum expresses the same types of receptors and ligands, also
+in orthogonal pairs of gradients. In this work, we did not consider the effect
+of reverse signalling from tectal receptors to RGC ligands so we only modelled
+tectal ligand expression, $L$. Although several studies model tectal ligand
 expression with exponential functions \citep{koulakov_stochastic_2004}, the
 experimental evidence for ligand expression is more ambiguous (Supp. Fig
-X). Correspondingly, we leave open the possibility that ligand expression may
-be modelled by exponential, linear or logarithmic functions.
+X). Correspondingly, we leave open the possibility that tectal ligand
+expression may be modelled by exponential, linear or logarithmic functions.
 
@@
 EphB, Ryk \citep{schmitt_wntryk_2006} and
-Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $l$ includes the
-ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
+Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $l$ and $L$
+include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
 RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
-role in retino-tectal map formation.
+role in retinotectal map formation.
 
@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Retinal and tectal gene expression.
-%
-\textbf{A} Retinal receptor expression. Four receptors are expressed and
-displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and
-green;$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an
-exponential function of $x$; $r_1$ increases with an exponential function of
-$y$. $r_2$ and $r_3$ have the opposite sense. The expression is modelled with
-exponential functions because there is ample evidence that shows that retinal
-receptor expressions follow this form.
-%
-\textbf{B} Retinal ligand expression. May not be included.
-%
-\textbf{C} Tectal receptor expression. May not be included.
-%
-\textbf{D} Tectal ligand expression. $l_0$ and $l_1$ are red and green; $l_2$
-and $l_3$ are cyan and magenta. These expressions are shown in a linear form,
-as there is less evidence to determine whether tectal ligand expression is
-exponential, linear or logarithmic in form.
-}
-\label{f:1}
-\end{figure}
-
-\subsection{Signalling}
+\subsection{Chemotaxis}
 
 In this model, the signal transmitted when a ligand binds to a receptor on a
-growth cone can lead to one of two effects; the signal may induce an
-attraction towards the ligand-expression region or a repulsion away from it.
+growth cone (forward signalling) can lead to one of two effects; the
+signal may induce an attraction towards the ligand-expressing region or a
+repulsion away from it.
 %
-For axon-tectum interactions, we implement attraction as a climbing of the
+For axon-tectum interactions, we implemented attraction as a climbing of the
 tectal ligand expression gradient and repulsion as gradient descent.
 %
-To determine the strength of the effect we assume a purely linear receptor
-binding model, and set the movement of the growth cone to be $\sum_i \pm
-r_i \times \grad l_i$.
+To determine the strength of the effect we assumed a purely linear receptor
+binding model, and set the chemotactic movement vector of the growth cone $b$ at
+location $\mathbf{x}_b$ on the tectum to be
+%
+\begin{equation}
+\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
+\end{equation}
+%
+where $r_{i,b}$ is the receptor expression on axon branch $b$ for
+ligand-receptor pair $i$, $\nabla L_i$ is the gradient of expression of ligand
+$i$ on the tectum and $F_i$ denotes the direction of the interaction induced
+when a molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes
+the value -1 for a repulsive interaction or 1 for an attractive interaction.
 
-For axon-axon interactions, attraction causes a movement along a vector
-between two axons causing them to move closer together; repulsion cause the
-opposite effect.
+\subsection{Axon-axon interactions}
 
-%This is where I really want to say ``the only interactions in the model are
-%via forward (ligand-activates-receptor) and reverse
-%(receptor-activates-ligand) signalling''. Something like:
+For axon-axon interactions, repulsion causes a movement of branch $b$ along a
+unit vector, $\hat{\mathbf{x}}_{kb}$, from branch $k$ to branch $b$ causing
+them to move further apart; attraction causes the opposite movement. The
+interaction, $\mathbf{I}_b$, acting on axon branch $b$, is given by
+%
+\begin{equation}
+\mathbf{I}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
+\end{equation}
+%
+where $Q(d_{kb})$ is the signalling strength between two growth cones of
+radius $r$ a distance $d_{kb}$ from one another, given by
+%
+\begin{equation}
+Q(d_{kb}) = \begin{cases}
+     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r \\
+     0 & d_{kb} > 2r
+     \end{cases}
+\end{equation}
+%
+The sign of $Q$ (which is dependent on the values of $F_i$) determines whether
+the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
+expression of ligand $i$ on branch $k$.
 
-%Interactions in the model are by receptor-ligand interactions only. Forward
-%interactions are those in which the ligand activates the receptor, delivering
-%a signal into the receptor-expressing cell. Reverse signals occur when a
-%membrane-bound ligand is activated~\citep{wu_role_2019}.
+\subsection{Movement}
+
+At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, of each
+branch, $b$, is updated according to
+%
+\begin{equation}
+\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
+\end{equation}
+%
+where
+%
+\begin{equation}
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b
+\end{equation}
+%
+evaluated at time $t$.
 

### 2021-07-16 paper/paper.tex
WIP, esp on interactions - added Ihist

@@
 experimental evidence for ligand expression is more ambiguous (Supp. Fig
-X). Correspondingly, we leave open the possibility that tectal ligand
+\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
 expression may be modelled by exponential, linear or logarithmic functions.
@@
 In this model, the signal transmitted when a ligand binds to a receptor on a
-growth cone (forward signalling) can lead to one of two effects; the
-signal may induce an attraction towards the ligand-expressing region or a
-repulsion away from it.
+branch (forward signalling) can lead to one of two effects; the signal may
+induce an attraction towards the ligand-expressing region or a repulsion away
+from it.
 %
@@
 To determine the strength of the effect we assumed a purely linear receptor
-binding model, and set the chemotactic movement vector of the growth cone $b$ at
+binding model, and set the chemotactic movement vector of the branch $b$ at
 location $\mathbf{x}_b$ on the tectum to be
@@
 %
-where $r_{i,b}$ is the receptor expression on axon branch $b$ for
-ligand-receptor pair $i$, $\nabla L_i$ is the gradient of expression of ligand
-$i$ on the tectum and $F_i$ denotes the direction of the interaction induced
-when a molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes
-the value -1 for a repulsive interaction or 1 for an attractive interaction.
+where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
+pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
+tectum and $F_i$ denotes the direction of the interaction induced when a
+molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
+-1 for a repulsive interaction or 1 for an attractive interaction.
 
@@
 them to move further apart; attraction causes the opposite movement. The
-interaction, $\mathbf{I}_b$, acting on axon branch $b$, is given by
+interaction, $\mathbf{I}_b$, acting on branch $b$, is given by
 %

### 2021-07-16 paper/paper.sh

@@ -1,0 +1,6 @@
+#!/bin/sh
+
+xelatex paper.tex
+bibtex paper.aux
+xelatex paper.tex
+xelatex paper.tex

### 2021-07-16 sim/agent/branch_base.h

@@ -29,6 +29,8 @@ struct branch_base
     }
 #endif
 
+    virtual void init() = 0;
+
     // Place the next computed location for path in 'next' so that while computing, we
     // don't modify the numbers we're working from. After looping through all branches,
     // place current into next and move on to next time step.

### 2021-07-16 sim/agent/agent1.h
Friday WIP

@@ -571,6 +571,7 @@ struct Agent1
             // Set the branch's termination zone
             unsigned int ri = i/bpa; // retina index
             this->pending_branches[i].init();
+            this->pending_branches[i].setr (this->mconf->getFloat ("r", 0.05f));
             this->pending_branches[i].aid = (int)ri; // axon index
             if (conf->getBool ("singleaxon", false)) {
                 this->pending_branches[i].rcpt = this->ret->rcpt[singleaxon_idx]; // FIXME: Use seeaxons

### 2021-07-16 sim/agent/branch_geb.h

@@ -15,8 +15,11 @@ template<typename T, size_t N>
 struct branch_geb : public branch_base<T,N>
 {
     // The cone radius, hard coded
-    static constexpr T r = T{0.05};
+    void setr (T _r) { this->r = _r; }
+protected:
+    T r = T{0.05};
 
+public:
     // For random angle choice
     morph::RandUniform<T> rng;
 

### 2021-07-19 paper/paper.tex
Friday WIP
WIP, towards first results figure

@@
 and, in common with that work, consists of agents representing axonal growth
-cones, originating from a square retina and moving upon a square tectum.
+cones, originating from a square retina and moving upon a square tectum of
+side length 1 in arbitrary units.
 %
@@
 
-In our model, space-based competition is omitted and chemoaffinity and
-axon-axon interactions occur only via signal transmission between receptors
-and ligands, which are expressed in graded patterns on both retina and
-tectum. Signals induce attractive or repulsive interactions; for chemoaffinity
-the interaction is to ascend or descend the gradient of ligand expression on
-the tectum; for axon-axon interactions, the strength of short-range repulsion
-(or attraction) is governed by the receptor-ligand expression on pairs of axons.
+In our model, space-based competition was omitted and chemoaffinity and
+axon-axon interactions were remodelled so that all interactions were via
+signal transmission between receptors and ligands, which are expressed in
+graded patterns on both retina and tectum. Signals induce attractive or
+repulsive interactions; for chemoaffinity the interaction is to ascend or
+descend the gradient of ligand expression on the tectum; for axon-axon
+interactions, the strength of short-range repulsion (or attraction) is
+governed by the receptor-ligand expression on pairs of axons.
 
-\subsection{Receptor and ligand expression}
+\subsection*{Receptor and ligand expression}
 
@@
 Each retinal ganglion cell projects $n=8$ growth cone agents (referred to
-as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands, $l_i$,
+as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands,
+$l_i$, indexed by $i$ and
 expressed at levels determined by the cell's location on the retina. The
@@
 studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-%
+We adopt the same precise form for the retinal receptor expression
+as \citet{simpson_simple_2011} which is
+\begin{equation}
+r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
+\end{equation}
+
 The simulated tectum expresses the same types of receptors and ligands, also
@@
 of reverse signalling from tectal receptors to RGC ligands so we only modelled
-tectal ligand expression, $L$. Although several studies model tectal ligand
+tectal ligand expression, $L_i$. Although several studies model tectal ligand
 expression with exponential functions \citep{koulakov_stochastic_2004}, the
@@
 \textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
-expression may be modelled by exponential, linear or logarithmic functions.
+expression may be modelled by exponential, linear or logarithmic functions:
+\begin{equation}
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
+\end{equation}
+or
+\begin{equation}
+L_i^{\text{lin}}(x,y) = 1.31 + 2.33 x;
+\end{equation}
+or
+\begin{equation}
+L_i^{\text{log}}(x,y) = 2.32 + 1.29 \log (2.3(x+0.2))
+\end{equation}
+(each of which has approximately the same expression values at $x=0$ and $x=1$).
 
@@
 
-\subsection{Chemotaxis}
+\subsection*{Movement}
+
+At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, of each
+branch, $b$, is updated according to
+%
+\begin{equation}
+\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
+\end{equation}
+%
+where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
+effect, $\mathbf{G}_b$ and an axon-axon interaction, $\mathbf{I}_b$:
+%
+\begin{equation}
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b
+\end{equation}
+%
+evaluated at time $t$. $m_1$ and $m_2$ are scalar parameters.
+
+\subsection*{Chemoaffinity}
 
@@
 
-\subsection{Axon-axon interactions}
+\subsection*{Axon-axon interactions}
 
@@
 \begin{equation}
-\mathbf{I}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
+% No \frac{1}{|B_b|} in this one
+\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
 \end{equation}
@@
 
-\subsection{Movement}
+\subsection*{Initial conditions}
 
-At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, of each
-branch, $b$, is updated according to
-%
+Branches were randomly distributed in a stripe at the rostral side of the
+tectum at the start of each simulation. Each RGC axon was assigned a random
+initial position coordinate:
 \begin{equation}
-\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
+\mathbf{x}_{\text{axon},t=0} = (U(0,1), U(-0.2,0))
 \end{equation}
-%
-where
-%
-\begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b
-\end{equation}
-%
-evaluated at time $t$.
+where U(p,q) is a number selected from a random uniform distribution in the
+range $[p,q)$. Each of the $n$ branches per RGC axon was given its parent
+axon's initial position, plus a randomly generated offset with coordinates
+derived from a normal distribution of mean 0 and standard deviation 0.1.
 
 \section{Results}
+
+\subsection*{Chemoaffinity model}
+
+First, we describe the behaviour of the model with $m_2 = 0$, which has only
+the gradient-based chemoaffinity effect and no axon-axon interaction.
+
+\subsubsection*{Wildtype}
+\subsubsection*{Tectal ablation etc}
+
+\subsection*{Chemoaffinitiy with axon-axon interaction}
+
+\subsubsection*{Wildtype}
+\subsubsection*{Tectal ablation etc}
 

### 2021-07-19 sim/agent/branch.h

@@ -385,6 +385,10 @@ public:
             for (size_t i = 0; i<this->ihs; ++i) { I += this->Ihist[i]; }
         }
         // Collected non-border movement components
+        //std::cout << "Movement for branch " << this->id
+        //          << ": m[0]*G:" << (G*m[0])
+        //          << "+ m[1]*C:" << (C*m[1])
+        //          << "+ m[2]*I:" << (I*m[2]) << std::endl;
         morph::Vector<T, 2> nonB = G * m[0] + C * m[1] + I * m[2];
 
         // Option for how movement is dealt with near the border; how to get axons back inside domain

### 2021-07-19 sim/agent/tissue.h

@@ -60,7 +60,8 @@ enum class expression_form
     lin,  // A linear function
     quad, // A quadratic function
     exp,  // An exponential function
-    log   // A logarithmic function
+    log,  // A logarithmic function
+    exp2  // A slightly different exponential function, used to demonstrate dependence of chemo-only model on expressions
 };
 
 // Which sense is an expression pattern becoming stronger?
@@ -332,6 +333,7 @@ struct guidingtissue : public tissue<T>
     T linear_expression (const T& x) const { return T{1.31} + T{2.3333} * x; }
     T quadratic_expression (const T& x) const { return T{1.31} + T{2.3333} * x * x; }
     T exponential_expression (const T& x) const { return T{1.05} + T{0.26} * std::exp (T{2.3} * x); }
+    T exponential_expression2 (const T& x) const { return T{1.05} + T{0.26} * std::exp (T{1.2} * x); }
     T logarithmic_expression (const T& x) const { return T{2.32} + T{1.29} * std::log (T{2.3} * (x+T{0.2})); }
 
     T expression_function (const T& x, const expression_form& ef) const
@@ -349,6 +351,9 @@ struct guidingtissue : public tissue<T>
             break;
         case expression_form::log:
             rtn = this->logarithmic_expression (x);
+            break;
+        case expression_form::exp2:
+            rtn = this->exponential_expression2 (x);
             break;
         default:
             break;

### 2021-07-19 paper/paper.tex
WIP

@@
 }
-\label{f:1}
+\label{f:ex}
 \end{figure}
 
-Each retinal ganglion cell projects $n=8$ growth cone agents (referred to
+Each retinal ganglion cell projects $n$ growth cone agents (referred to
 as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands,
@@
 
-Figure~\ref{f:1} shows colour maps of receptor and ligand
-expression, in this case tectal ligand expression maps are linear.
+Figure~\ref{f:ex} shows colour maps of receptor and ligand
+expression, in this case tectal ligand expression maps are linear. Note that
+the expression is discretized, and in simulations, the gradient is computed
+numerically from the discretized expression. This makes it straightforward to
+perform tectal and retinal `graft' manipulations.
 
@@
 where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
-effect, $\mathbf{G}_b$ and an axon-axon interaction, $\mathbf{I}_b$:
+effect, $\mathbf{G}_b$, an axon-axon interaction, $\mathbf{I}_b$ and a `border effect',
+$\mathbf{B}_b$, which acts to retain branches within the tectal region:
 %
 \begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b + m_3 \mathbf{B}_b
 \end{equation}
 %
-evaluated at time $t$. $m_1$ and $m_2$ are scalar parameters.
+evaluated at time $t$. $m_1$, $m_2$ and $m_3$ are scalar parameters.
 
@@
 
+\subsection*{Border effect}
+
+We implement a border effect based on gradient following by assuming that
+there is some other molecular signal which acts on all branches near the
+boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
+given by:
+%
+\begin{equation}
+B_{b,x} = \begin{cases}
+        2r-x    & x<2r \\
+        1-2r-x    & x>1-2r
+\end{cases}
+B_{b,y} = \begin{cases}
+        2r-y    & y<2r \\
+        1-2r-y    & y>1-2r
+\end{cases}
+\end{equation}
+%
+Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
+molecule expressed around the border of the tectum, whose expression increases
+quadratically outside the tectum.
+
 \subsection*{Initial conditions}
@@
 
-First, we describe the behaviour of the model with $m_2 = 0$, which has only
-the gradient-based chemoaffinity effect and no axon-axon interaction.
+First, we describe the behaviour of a model which has only the gradient-based
+chemoaffinity effect and no axon-axon interaction or border effect
+(i.e.~$m_2=m_3=0$). This model is run with $n=8$ branches per RGC axon. All
+receptor-ligand signals induce repulsive interactions (i.e.~gradient
+descent). Receptor and ligand expressions are as shown in
+Fig.\,\ref{f:ex}. Fig.\,\ref{f:ch}A gives a colour map of retinal ganglion
+cell soma locations, with the colour red encoding $x$ position on the
+temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
+position on the dorso-ventral axis.
+%
+Fig.\,\ref{f:ch}A+ shows the expected final location of each RGC axon,
+according to experimental observations in a `fishnet' plot, which draws lines
+between adjacent axons.
+%
+Fig.\,\ref{f:ch}B shows the position of
+400 branches at $t=500$ (in arbitrary time units). Considering each
+coordinate axis separately, each branch interacts differentially with two
+opposing gradients. Branches which settle nearest the rostral edge of the
+tectum interact more strongly with the gradient $L_0$ than with $L_2$
+(Fig\,\ref{f:ex}C). A similar pair of gradients separate branches along the
+M-L tectal axis. The branches become positioned on a grid due to the
+discretization of the tectal ligand gradient (as the number of branches and
+tectal elements tends towards infinity, the arangement tends towards a
+continuous map---see Fig.\,S1 \textbf{FIXME}).
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_chemo.png}
+\caption{Chemoaffinity model}
+\label{f:ch}
+\end{figure}
+
+Fig.\,\ref{f:ch}C shows the RGC axon centroids; the mean location of $n$
+branches per axon. Because all $n$ branches for each RGC have identical
+receptor expression and they experience only the chemoaffinity effect, each
+branch attains almost the same final position and so the centroid graph
+strongly resembles the graph of all the branches. Fig.\,\ref{f:ch}C also shows
 

### 2021-07-19 sim/agent/agent1.h

@@ -552,8 +552,8 @@ struct Agent1
 
         // Axon initial positions x and y can be uniformly randomly selected...
         morph::RandUniform<T, std::mt19937> rng_x(T{0}, T{1.0});
-        //morph::RandUniform<T, std::mt19937> rng_y(T{-0.2}, T{0}); // S&G
-        morph::RandUniform<T, std::mt19937> rng_y(T{0.0}, T{0.2}); // S&G
+        morph::RandUniform<T, std::mt19937> rng_y(T{-0.2}, T{0}); // S&G
+        //morph::RandUniform<T, std::mt19937> rng_y(T{0.0001}, T{0.2}); // All within field
         // ...or set from the ideal position plus a random perturbation
         morph::RandNormal<T, std::mt19937> rng_p0(T{0}, T{0.1});
         // A normally distributed perturbation is added for each branch. SD=0.1.
@@ -867,7 +867,7 @@ struct Agent1
         this->bv->rcpt_scale.compute_autoscale (rcpt_min, rcpt_max);
         this->bv->target_scale.compute_autoscale (0, 1);
         this->bv->finalize();
-        this->bv->addLabel ("Growth cones", {0.0f, 1.1f, 0.0f});
+        this->bv->addLabel ("Branches", {0.0f, 1.1f, 0.0f});
         this->addOrientationLabels (this->bv, std::string("Tectal"));
         v->addVisualModel (this->bv);
 
@@ -881,7 +881,7 @@ struct Agent1
         this->cv->maxlen = this->conf->getDouble ("maxnetline", 1.0);
         this->cv->viewmode = netvisual_viewmode::actual;
         this->cv->finalize();
-        this->cv->addLabel ("axon centroids", {0.0f, 1.1f, 0.0f});
+        this->cv->addLabel ("Axon centroids", {0.0f, 1.1f, 0.0f});
         this->addOrientationLabels (this->cv, std::string("Tectal"));
         v->addVisualModel (this->cv);
 
@@ -890,7 +890,7 @@ struct Agent1
         this->tcv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
         this->tcv->viewmode = netvisual_viewmode::targetplus;
         this->tcv->finalize();
-        this->tcv->addLabel ("experiment suggests", {0.0f, 1.1f, 0.0f});
+        this->tcv->addLabel ("Experiment", {0.0f, 1.1f, 0.0f});
[…]

### 2021-07-19 sim/agent/netvisual.h

@@ -129,20 +129,22 @@ public:
                                this->locations->clr[i], this->locations->clr[i],
                                this->radiusFixed, 16);
 
-            // A line (cyl. tube) from target to actual position
-            this->computeTube (idx,
-                               this->locations->targ[i],
-                               this->locations->p[i]+actualpuckoffs,
-                               this->locations->clr[i], this->locations->clr[i],
-                               this->puckthick[2], 8);
+            if (this->draw_actual == true) {
+                // A line (cyl. tube) from target to actual position
+                this->computeTube (idx,
+                                   this->locations->targ[i],
+                                   this->locations->p[i]+actualpuckoffs,
+                                   this->locations->clr[i], this->locations->clr[i],
+                                   this->puckthick[2], 8);
 
-            // A slightly smaller puck for actual position
-            this->computeTube (idx,
-                               this->locations->p[i]+actualpuckoffs+puckthick,
-                               this->locations->p[i]+actualpuckoffs-puckthick,
-                               morph::Vector<float,3>({1,0,0}), morph::Vector<float,3>({0,1,0}),
-                               this->locations->clr[i], this->locations->clr[i],
-                               this->radiusFixed*0.667, 16);
+                // A slightly smaller puck for actual position
+                this->computeTube (idx,
+                                   this->locations->p[i]+actualpuckoffs+puckthick,
+                                   this->locations->p[i]+actualpuckoffs-puckthick,
+                                   morph::Vector<float,3>({1,0,0}), morph::Vector<float,3>({0,1,0}),
+                                   this->locations->clr[i], this->locations->clr[i],
+                                   this->radiusFixed*0.667, 16);
+            }
         }
         // Connections
         for (auto c : this->locations->c) {
@@ -186,6 +188,8 @@ public:
     net<Flt>* locations = (net<Flt>*)0;
     Flt radiusFixed = 0.01;
     Flt linewidth = 0.004;
+    //! If true, draw small pucks and lines showing actual position imposed on the 'expt suggests' view
+    bool draw_actual = false;
     //! The maximum length of a line betwen two vertices for it to be visualised
     Flt maxlen = 1e9;
     morph::Vector<float,3> puckthick = { 0, 0, 0.002 };

### 2021-07-19 paper/paper.tex
Monday WIP, figure generation from sims in a repeatable way

@@
 expression may be modelled by exponential, linear or logarithmic functions:
-\begin{equation}
+\begin{equation} \label{e:tecexp}
 L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
@@
 \end{equation}
-(each of which has approximately the same expression values at $x=0$ and $x=1$).
+Each function has approximately the same expression values at $x=0$ and
+$x=1$. Opposing functions are obtained by substituting $(1-x)$ for $x$.
 
@@
 
+\subsubsection*{Wildtype}
+
 First, we describe the behaviour of a model which has only the gradient-based
 chemoaffinity effect and no axon-axon interaction or border effect
-(i.e.~$m_2=m_3=0$). This model is run with $n=8$ branches per RGC axon. All
-receptor-ligand signals induce repulsive interactions (i.e.~gradient
-descent). Receptor and ligand expressions are as shown in
-Fig.\,\ref{f:ex}. Fig.\,\ref{f:ch}A gives a colour map of retinal ganglion
-cell soma locations, with the colour red encoding $x$ position on the
+(i.e.~$m_2=m_3=0$). This model was run with $n=8$ branches per RGC axon. All
+receptor-ligand signals induced repulsive interactions (i.e.~gradient
+descent). Receptor and ligand expressions were as shown in Fig.\,\ref{f:ex},
+noting that tectal ligand expression was of exponential form
+(Eq.\,\ref{e:tecexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
+ganglion cell soma locations, with the colour red encoding $x$ position on the
 temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
@@
 %
-Fig.\,\ref{f:ch}A+ shows the expected final location of each RGC axon,
-according to experimental observations in a `fishnet' plot, which draws lines
-between adjacent axons.
+Fig.\,\ref{f:ch}B shows the expected final location of each RGC axon,
+according to experimental observations, in a `fishnet' plot which draws lines
+between adjacent axons giving a visual indication of the topology of the
+arrangement.
 %
-Fig.\,\ref{f:ch}B shows the position of
+Fig.\,\ref{f:ch}C shows the position of
 400 branches at $t=500$ (in arbitrary time units). Considering each
@@
 
-Fig.\,\ref{f:ch}C shows the RGC axon centroids; the mean location of $n$
-branches per axon. Because all $n$ branches for each RGC have identical
-receptor expression and they experience only the chemoaffinity effect, each
-branch attains almost the same final position and so the centroid graph
-strongly resembles the graph of all the branches. Fig.\,\ref{f:ch}C also shows
+Fig.\,\ref{f:ch}D shows a fishnet plot of the RGC axon centroids; the mean
+location of $n$ branches per axon. Because all $n$ branches for each RGC have
+identical receptor expression and they experience only the chemoaffinity
+effect, each branch attains almost the same final position and so the centroid
+graph strongly resembles Fig.\,\ref{f:ch}B. Fig.\,\ref{f:ch}D indicates that
+the topology of the arrangement resembles that of the experimental result, but
+the area over which the branches arrange themselves is limited to about 50\%
+of the tectum. It is possible to tailor the tectal ligand expression, whcih
+adjusts the relative contributions of the opposing interactions, to tune the
+final map. This allows the chemoaffinity model to attain a good match to the
+experimental arrangement (Fig.\~S2 \textbf{fixme}). However, the match is
+highly dependent on the form of receptor/ligand expression in both retinal and
+tectal cells.
 
-\subsubsection*{Wildtype}
 \subsubsection*{Tectal ablation etc}

### 2021-07-19 scripts/fig_expression.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+./build/sim/agent/tissuevis ./configs/a1/m_ee_chemoonly.json

### 2021-07-19 sim/agent/agent1.cpp

@@ -94,19 +94,23 @@ int main (int argc, char **argv)
     if (num_guiders == 4) {
         if (branch_model == "gebhardt") {
             Agent1<float, 4, branch_geb<float, 4>> model (conf, mconf);
+            model.title = std::string("geb_") + m_id + std::string("_") + e_id;
             model.run();
             model.save (outfile);
         } else if (branch_model == "stochastic") {
             Agent1<float, 4, branch_stochastic<float, 4, 10>> model (conf, mconf);
+            model.title = std::string("stoc_") + m_id + std::string("_") + e_id;
             model.run();
             model.save (outfile);
         } else {
             Agent1<float, 4, branch<float, 4>> model (conf, mconf);
+            model.title = std::string("j4_") + m_id + std::string("_") + e_id;
             model.run();
             model.save (outfile);
         }
     } else if (num_guiders == 2) {
         Agent1<float, 2, branch<float, 2>> model (conf, mconf);
+        model.title = std::string("j2_") + m_id + std::string("_") + e_id;
         model.run();
         model.save (outfile);
     }

### 2021-07-19 sim/agent/agent1.h

@@ -34,6 +34,7 @@
 # include "branchvisual.h"
 # include "netvisual.h"
 # include "tissuevisual.h"
+enum class graph_layout { a, b };
 #endif
 
 template<typename T, size_t N, typename B=branch<T, N>>
@@ -123,6 +124,10 @@ struct Agent1
         }
         std::cout << "Done simulating\n";
 #ifdef VISUALISE
+        // Save final image based on config file names
+        std::stringstream nss;
+        nss << this->title << ".png";
+        this->v->saveImage (nss.str());
         this->vis(this->conf->getUInt ("steps", 1000));
         this->v->keepOpen();
 #endif
@@ -148,11 +153,13 @@ struct Agent1
         } else {
             glfwPollEvents();
         }
-        this->bv->reinit();
-        this->av->reinit();
-        this->cv->reinit();
-        this->tcv->reinit();
-        this->gv->append ((float)stepnum, this->ax_centroids.sos(), 0);
+        this->bv->reinit(); // Branches
+        this->cv->reinit(); // Centroids
+        this->tcv->reinit(); // Experiment
+        if (this->layout == graph_layout::a) {
+            this->av->reinit(); // Selected axons
+            this->gv->append ((float)stepnum, this->ax_centroids.sos(), 0);
+        }
         this->v->render();
         if (this->conf->getBool ("movie", false)) {
             std::stringstream frame;
@@ -764,7 +771,8 @@ struct Agent1
     {
 #ifdef VISUALISE
         const unsigned int ww = this->conf->getUInt ("win_width", 2000);
-        this->tvv = new morph::Visual (ww, ww, "Retinal and Tectal expression");
+        const unsigned int wh = this->conf->getUInt ("win_height", 2000);
+        this->tvv = new morph::Visual (ww, wh, "Retinal and Tectal expression");
         if constexpr (use_ortho_tvv) {
             float orthoside = 2.5;
             this->tvv->ptype = morph::perspective_type::orthographic;
@@ -845,9 +853,19 @@ struct Agent1
 #ifdef VISUALISE
         // morph::Visual init
         const unsigned int ww = this->conf->getUInt ("win_width", 1200);
-        unsigned int wh = static_cast<unsigned int>(widthtoheight * (float)ww);
-        std::cout << "New morph::Visual with width/height: " << ww << "/" << wh << std::endl;
-        this->v = new morph::Visual (ww, wh, "Agent based retinotectal model");
[…]

### 2021-07-20 paper/paper.tex
WIP. About to Change graph_layout::c again

@@
 
-\subsection*{Chemoaffinity model}
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_wt.png}
+\caption{Gradient following model with axon-axon interaction}
+\label{f:wt}
+\end{figure}
 
-\subsubsection*{Wildtype}
+We first present the result of running the model above, which assumes that
+axonal growth cones express graded levels of receptors which interact with
+ligands expressed as gradients on the tectum and on the surface of other
+growth cones to guide their movement. We make the following additional
+assumptions: i) All receptor-ligand interactions are repulsive, as for
+EphA-ephrinA coupling \citep{nakamoto_topographically_1996}; and ii) tectal
+ligand expressions vary exponentially along a given axis, following
+Eq.\,\ref{e:tecexp}.
 
-First, we describe the behaviour of a model which has only the gradient-based
+\subsection*{Chemoaffinity---Wildtype}
+
+Introductory sentence.
+
+We now describe the behaviour of a model which has only the gradient-based
 chemoaffinity effect and no axon-axon interaction or border effect
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_chemo.png}
+\includegraphics[width=\linewidth]{./images/j4_ee_chemoonly_wt_chemoonly.png}
 \caption{Chemoaffinity model}
@@
 tectal cells.
-
-\subsubsection*{Tectal ablation etc}
-
-\subsection*{Chemoaffinitiy with axon-axon interaction}
-
-\subsubsection*{Wildtype}
-\subsubsection*{Tectal ablation etc}
 

### 2021-07-20 sim/agent/agent1.h

@@ -69,6 +69,13 @@ struct Agent1
     void run()
     {
 #ifdef VISUALISE
+        if (this->layout == graph_layout::c) {
+            this->freeze_times[0] = 0;
+            this->freeze_times[1] = this->conf->getUInt ("freeze_time1", 20);
+            this->freeze_times[2] = this->conf->getUInt ("freeze_time2", 80);
+            this->freeze_times[3] = this->conf->getUInt ("freeze_time3", 120);
+        }
+
         this->visinit();
 #endif
         this->gradient_rng = new morph::RandNormal<T, std::mt19937>(1, this->conf->getDouble("gradient_rng_width", 0.0));
@@ -126,7 +133,7 @@ struct Agent1
 #ifdef VISUALISE
         // Save final image based on config file names
         std::stringstream nss;
-        nss << this->title << ".png";
+        nss << "./paper/images/" << this->title << ".png";
         this->v->saveImage (nss.str());
         this->vis(this->conf->getUInt ("steps", 1000));
         this->v->keepOpen();
@@ -158,11 +165,11 @@ struct Agent1
             this->bv->reinit(); // Branches
         }
 
-         // Centroids that should show up to a certain time
+        // Centroids that should show up to a certain time
         if (this->layout == graph_layout::c) {
-            if (stepnum < 20) { this->cv1->reinit(); }
-            if (stepnum < 80) { this->cv2->reinit(); }
-            //if (stepnum < 80) { this->cv3->reinit(); }
+            if (stepnum <= this->freeze_times[1]) { this->cv1->reinit(); }
+            if (stepnum <= this->freeze_times[2]) { this->cv2->reinit(); }
+            //if (stepnum <= this->freeze_times[3]) { this->cv3->reinit(); }
         }
         this->cv->reinit(); // Centroids to end
 
@@ -870,7 +877,7 @@ struct Agent1
         const unsigned int ww = this->conf->getUInt ("win_width", wdefault);
         unsigned int hdefault = 800;
         if (this->layout == graph_layout::b) { hdefault = 700; }
-        if (this->layout == graph_layout::c) { hdefault = 1130; }
+        if (this->layout == graph_layout::c) { hdefault = 1180; }
         const unsigned int wh = this->conf->getUInt ("win_height", hdefault);
 
[…]

### 2021-07-20 paper/paper.tex
Modified border effect to be effective from r, not 2r - neater and perfectly sensible

@@
 B_{b,x} = \begin{cases}
-        2r-x    & x<2r \\
-        1-2r-x    & x>1-2r
+        r-x      & x<r \\
+        1-r-x    & x>1-r
 \end{cases}
 B_{b,y} = \begin{cases}
-        2r-y    & y<2r \\
-        1-2r-y    & y>1-2r
+        r-y      & y<r \\
+        1-r-y    & y>1-r
 \end{cases}
@@
 molecule expressed around the border of the tectum, whose expression increases
-quadratically outside the tectum.
+quadratically outside the tectum and affects any branch touching (or outside)
+the boundary
 
@@
 \includegraphics[width=\linewidth]{./images/j4_ee_wt.png}
-\caption{Gradient following model with axon-axon interaction}
+\caption{Gradient following model with axon-axon interactions.}
 \label{f:wt}
@@
 
-We first present the result of running the model above, which assumes that
-axonal growth cones express graded levels of receptors which interact with
-ligands expressed as gradients on the tectum and on the surface of other
-growth cones to guide their movement. We make the following additional
-assumptions: i) All receptor-ligand interactions are repulsive, as for
-EphA-ephrinA coupling \citep{nakamoto_topographically_1996}; and ii) tectal
-ligand expressions vary exponentially along a given axis, following
-Eq.\,\ref{e:tecexp}.
+To enable a simulation it is necessary to make the assumptions, in addition to
+those implicit in the description of the model above, that i) All
+receptor-ligand interactions are repulsive, as for EphA-ephrinA
+coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
+all tectal ligand expressions vary exponentially along a given axis, following
+Eq.\,\ref{e:tecexp}. The result of running this system are presented in
+Fig.\,\ref{f:wt}.
+%
+Fig.\,\ref{f:wt}A shows a dual-colour map of retinal ganglion cell positions
+on the square retina, with red encoding position on the $x$ axis (temporal to
+nasal) and green encoding position on the $y$ axis (ventral to
+dorsal). Fig.\,\ref{f:wt}B presents a fishnet plot of the expected location on
+the tectal surface of each RGC at the end of the development process. Thus,
+the `red' RGC from the ventro-nasal corner should map to the caudo-medial
+corner of the tectum and the `green' RGC from the dorso-temporal corner should
+find its way to the rostro-lateral corner on the tectum. The lines between the
+dots of Fig.\,\ref{f:wt}B indicate the adjacency relationships between
+cells. Thus, colour in each panel indicates a branch's location of origin on
+the retina.
+%
+Fig.\,\ref{f:wt}C shows the initial state of the system, with branches
+stochastically arranged around the rostral edge of the tectum. Each dot
+represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}D-E
+show the development of the system with time, with Fig.\,\ref{f:wt}E showing
 
@@
 
+The model described above assumes that branches express graded levels of
+receptors which interact with ligands expressed as gradients on the tectum,
+the tectal border and on the surface of other branches, and that only these
+interactions guide their movement.
+
 % Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)

### 2021-07-20 sim/agent/agent1.h

@@ -70,6 +70,8 @@ struct Agent1
     {
 #ifdef VISUALISE
         if (this->layout == graph_layout::c) {
+            // Set up (from config file if necssary) the times at which the various
+            // graphs will be frozen (i.e. no longer updated from the sim)
             this->freeze_times[0] = 0;
             this->freeze_times[1] = this->conf->getUInt ("freeze_time1", 20);
             this->freeze_times[2] = this->conf->getUInt ("freeze_time2", 80);
@@ -868,6 +870,8 @@ struct Agent1
 #endif // VISUALISE
     }
 
+    // Set up the simulation visualisation scene. This depends on whether this->layout
+    // is graph_layout::a, ::b or ::c, etc.
     void visinit()
     {
 #ifdef VISUALISE

### 2021-07-20 paper/paper.tex
Final Tuesday updates/WIP

@@
 
-To enable a simulation it is necessary to make the assumptions, in addition to
-those implicit in the description of the model above, that i) All
+To enable a simulation it is necessary to make the assumptions (in addition to
+those implicit in the description of the model above) that i) All
 receptor-ligand interactions are repulsive, as for EphA-ephrinA
@@
 %
-Fig.\,\ref{f:wt}A shows a dual-colour map of retinal ganglion cell positions
+Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
 on the square retina, with red encoding position on the $x$ axis (temporal to
 nasal) and green encoding position on the $y$ axis (ventral to
-dorsal). Fig.\,\ref{f:wt}B presents a fishnet plot of the expected location on
+dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
 the tectal surface of each RGC at the end of the development process. Thus,
@@
 find its way to the rostro-lateral corner on the tectum. The lines between the
-dots of Fig.\,\ref{f:wt}B indicate the adjacency relationships between
-cells. Thus, colour in each panel indicates a branch's location of origin on
-the retina.
+dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
+cells. Thus, colour in each panel indicates a branch or axon's location of
+origin on the retina.
 %
-Fig.\,\ref{f:wt}C shows the initial state of the system, with branches
+Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
 stochastically arranged around the rostral edge of the tectum. Each dot
-represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}D-E
-show the development of the system with time, with Fig.\,\ref{f:wt}E showing
+represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
+show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
+well ordered arrangement, with a majority of adjacency relationships matching
+the experimental prediction (i.e. there are few crossed lines). The
+intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
+tectal surface, with final adjacency relationships developing over a longer
+timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
+shows \textbf{RMS error or some metric} of the axon centroids with respect to
+the experimental prediction. Fig.\,\ref{f:wt}F shows the final
+location of each branch in the simulation, which gives a more continuous view
+of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
+branches and their centroid path history for five selected RGCs, with path
+histories strongly resembling those presented by \citet{simpson_simple_2011}.
 
-\subsection*{Chemoaffinity---Wildtype}
+This result shows that a gradient following model based on receptor-ligand
+interactions and incorporating axon-axon repulsion is able to generate the
+basic mapping observed in the retinotectal projection.
+
+\subsection*{Manipulations}
+
+To determine if the model can explain the many surgical and genetic
+manipulations described in the literature
+\citep*[for a review, see][]{goodhill_retinotectal_1999} we performed
+simulations matching those in \citet{simpson_simple_2011} where we either
+rotated or translocated grafts of the tectum, ablated portions of the retina
+and/or the tectum or manipulated the receptor or ligand expressions in the
+model...
+
+\subsection*{Grating simulations}
+
+I'd like to do this if it's not too much work!
+
+\subsection*{Is chemoaffinity alone sufficient to explain the retinotectal projection?}
 

### 2021-07-20 scripts/fig_wt.sh

@@ -1,0 +1,2 @@
+#!/bin/sh
+./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_wt.json

### 2021-07-20 sim/agent/agent1.h

@@ -163,14 +163,12 @@ struct Agent1
             glfwPollEvents();
         }
 
-        if (this->layout != graph_layout::c) {
-            this->bv->reinit(); // Branches
-        }
+        this->bv->reinit(); // Branches
 
         // Centroids that should show up to a certain time
         if (this->layout == graph_layout::c) {
             if (stepnum <= this->freeze_times[1]) { this->cv1->reinit(); }
-            if (stepnum <= this->freeze_times[2]) { this->cv2->reinit(); }
+            //if (stepnum <= this->freeze_times[2]) { this->cv2->reinit(); }
             //if (stepnum <= this->freeze_times[3]) { this->cv3->reinit(); }
         }
         this->cv->reinit(); // Centroids to end
@@ -1000,45 +998,19 @@ struct Agent1
 
         } else  if (this->layout == graph_layout::c) { // Layout with diff. time end points
 
-            offset[1] -= 1.6f;
-
             // A Retinal cell positions
             v->addVisualModel (this->createTissueVisual (v->shaderprog, v->tshaderprog, offset, ret, "Retinal", expression_view::cell_positions, 0, 2));
 
+            offset[1] -= 1.6f;
+
             // B Experiment: Another NetVisual view showing the target locations
-            offset[0] += 1.3f;
             this->tcv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
             this->tcv->viewmode = netvisual_viewmode::targetplus;
             this->tcv->finalize();
             this->tcv->addLabel ("Experiment", {0.0f, 1.1f, 0.0f});
             v->addVisualModel (this->tcv);
 
-            // C Selected axons: This one gives an 'axon view'
             offset[0] += 1.3f;
-            this->av = new BranchVisual<T, N, B> (v->shaderprog, v->tshaderprog, offset, &this->branches, &this->ax_history);
-            this->av->axonview = true;
-            for (auto sa : this->seeaxons) { this->av->seeaxons.insert(sa); }
-            this->av->rcpt_scale.compute_autoscale (rcpt_min, rcpt_max);
-            this->av->target_scale.compute_autoscale (0, 1);
-            this->av->finalize();
-            this->av->addLabel ("Selected axons", {0.0f, 1.1f, 0.0f});
-            v->addVisualModel (this->av);
-
-            offset[0] += 1.3f;
-
[…]

### 2021-07-21 paper/paper.tex
Working on generating manipulations figures

@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_wt.png}
+\includegraphics[width=\linewidth]{./images/j4_ee_wt_fig.png}
 \caption{Gradient following model with axon-axon interactions.}
@@
 \end{figure}
+
+\subsection*{Normal development}
 
@@
 \citep*[for a review, see][]{goodhill_retinotectal_1999} we performed
-simulations matching those in \citet{simpson_simple_2011} where we either
-rotated or translocated grafts of the tectum, ablated portions of the retina
-and/or the tectum or manipulated the receptor or ligand expressions in the
-model...
+simulations matching those in \citet{simpson_simple_2011} in which we either
+rotated or translocated grafts of the tectum; ablated portions of the retina
+and/or the tectum; or manipulated the receptor or ligand expressions in the
+model.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
+\caption{Surgical manipulations}
+\label{f:manip}
+\end{figure}
 

### 2021-07-21 sim/agent/agent1.h

@@ -171,14 +171,12 @@ struct Agent1
         // Centroids that should show up to a certain time
         if (this->layout == graph_layout::c) {
             if (stepnum <= this->freeze_times[1]) { this->cv1->reinit(); }
-            //if (stepnum <= this->freeze_times[2]) { this->cv2->reinit(); }
-            //if (stepnum <= this->freeze_times[3]) { this->cv3->reinit(); }
         }
         this->cv->reinit(); // Centroids to end
 
         this->tcv->reinit(); // Experiment
+        this->av->reinit(); // Selected axons
         if (this->layout == graph_layout::a || this->layout == graph_layout::c) {
-            this->av->reinit(); // Selected axons
             this->gv->append ((float)stepnum, this->ax_centroids.sos(), 0);
         }
         this->v->render();
@@ -879,7 +877,7 @@ struct Agent1
         // morph::Visual init
         unsigned int wdefault = 1200;
         if (this->layout == graph_layout::a) { wdefault = 1920; }
-        if (this->layout == graph_layout::b) { wdefault = 2200; }
+        if (this->layout == graph_layout::b) { wdefault = 2650; }
         if (this->layout == graph_layout::c) { wdefault = 2200; }
         const unsigned int ww = this->conf->getUInt ("win_width", wdefault);
         unsigned int hdefault = 800;
@@ -895,7 +893,7 @@ struct Agent1
         if (this->layout == graph_layout::a) {
             this->v->setSceneTrans (-0.413126f, 0.6811f, -5.0f);
         } else if (this->layout == graph_layout::b) {
-            this->v->setSceneTrans (-0.326813f, -0.00970902f, -2.6f);
+            this->v->setSceneTrans (-0.890077f, -0.0236414f, -2.6f);
         } else if (this->layout == graph_layout::c) {
             this->v->setSceneTrans (-0.943763f, 0.800606f, -5.9f);
         }
@@ -995,12 +993,24 @@ struct Agent1
             // Axon centroids: Centroids of branches viewed with a NetVisual
             offset[0] += 1.3f;
             this->cv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
-            this->cv->maxlen = this->conf->getDouble ("maxnetline", 1.0);
-            this->cv->viewmode = netvisual_viewmode::actual;
+            //this->cv->maxlen = this->conf->getDouble ("maxnetline", 1.0);
[…]

### 2021-07-21 sim/agent/branchvisual.h

@@ -47,6 +47,20 @@ public:
     // Target location will have a different scale
     morph::Scale<Flt, Flt> target_scale;
 
+    void drawBoundary (VBOint& idx)
+    {
+        std::array<float, 3> gry = { 0.2, 0.2, 0.2 };
+        float w = 1, h = 1;
+        this->computeLine (idx, morph::Vector<Flt, 3>({0, 0, 0}), morph::Vector<Flt, 3>({w, 0, 0}),
+                           this->uz, gry, gry, this->blinewidth, this->blinewidth/4);
+        this->computeLine (idx, morph::Vector<Flt, 3>({w, 0, 0}), morph::Vector<Flt, 3>({w, h, 0}),
+                           this->uz, gry, gry, this->blinewidth, this->blinewidth/4);
+        this->computeLine (idx, morph::Vector<Flt, 3>({w, h, 0}), morph::Vector<Flt, 3>({0, h, 0}),
+                           this->uz, gry, gry, this->blinewidth, this->blinewidth/4);
+        this->computeLine (idx, morph::Vector<Flt, 3>({0, h, 0}), morph::Vector<Flt, 3>({0, 0, 0}),
+                           this->uz, gry, gry, this->blinewidth, this->blinewidth/4);
+    }
+
     void initializeVertices()
     {
         if (this->axonview) { return this->initializeAxonview(); }
@@ -96,6 +110,8 @@ public:
             this->computeSphere (idx, cur, clr, clr2, this->radiusFixed, 14, 12);
             this->computeFlatLineRnd (idx, (*this->ax_history)[b.aid].back(), cur, this->uz, clr, this->linewidth/2.0f, 0.0f, true, false);
         }
+
+        this->drawBoundary (idx);
     }
 
     //! Set this->radiusFixed, then re-compute vertices.
@@ -113,6 +129,7 @@ public:
     //! Change this to get larger or smaller spheres.
     Flt radiusFixed = 0.01;
     Flt linewidth = 0.008;
+    Flt blinewidth = 0.004;
     //! A normal vector, fixed as pointing up
     morph::Vector<float, 3> uz = {0,0,1};
     //! In axon view, show just a small selection of axons

### 2021-07-22 paper/paper.tex
Looking at the linear model, but it seems to perform almost as well as the exp-exp model, which is odd

@@
 those implicit in the description of the model above) that i) All
-receptor-ligand interactions are repulsive, as for EphA-ephrinA
+receptor-ligand signalled interactions are repulsive, as for EphA-ephrinA
 coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
@@
 
+Fig.\,\ref{f:manip} shows the result of simulating three surgical
+manipulations alongside the wildtype result, which is shown on Row A. Rows B
+and C show graft rotation simulations (90$\degree$ and 180$\degree$). In each
+case an 8x8 square on the tectum (each unit is 0.05x0.05) is rotated. The
+rotational sense can be observed clearly in the `Branches' column; for the
+90$\degree$ rotation, the top-right (caudal-lateral) corner of the rotated
+square is red, whereas the top-left corner of the wildtype tectum is red. \textbf{etc}.
+
 \begin{figure}
 \includegraphics[width=\linewidth]{./images/fig_manipulations.png}
-\caption{Surgical manipulations}
+\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
+rotation. Row C: 180 degree rotation. Row D: Graft swap.}
 \label{f:manip}
@@
 
-\subsection*{Grating simulations}
+\subsection*{ligand expression profiles}
 
-I'd like to do this if it's not too much work!
+A section to show the effect of assuming linear or logarithmic expression
+profiles. Especially that linear expression fails to reproduce the rotation
+and graft swap manipulations. Hmm, actually, it looks ok?!
+
+\subsection*{Attractive receptor-ligand interactions}
+
+Show that, if you switch one of the receptor-ligand  interactions in the
+system to be attractive instead of repulsive, then you have to make the ligand
+expression form logarithmic instead of exponential.
 
@@
 
+\subsection*{Grating simulations}
+
+I'd like to do this if it's not too much work!
+
 \section{Discussion}

### 2021-07-23 paper/paper.tex
Splurging out results figures into the draft paper

@@
 
-Fig.\,\ref{f:manip} shows the result of simulating three surgical
+Fig.\,\ref{f:surg1} shows the result of simulating three surgical
 manipulations alongside the wildtype result, which is shown on Row A. Rows B
@@
 rotation. Row C: 180 degree rotation. Row D: Graft swap.}
-\label{f:manip}
+\label{f:surg1}
+\end{figure}
+
+Fig.\,\ref{f:surg2} shows retinal ablation, tectal ablation, mismatch and
+compound eye experiments (Stuart: see Simpson and Goodhill Fig. 4).
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_manipulations2.png}
+\caption{Surgical manipulations 2. Row A: Retinal ablation, Row B: Tectal
+ablation, Row C: Mismatch (ablate retina and also the tectum that matches the
+left-behind retina), Row D: Compound eye (Two nasal retinal halves `glued
+together')}
+\label{f:surg2}
 \end{figure}
@@
 expression form logarithmic instead of exponential.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_ephb_wt.png}
+\caption{The effect of making one interaction attractive. This models
+switching one receptor/ligand pair to be an analogue of EphB, but retains an
+assumption that the ephrin-B ligand has exponential form. The change compared
+with the model of Fig.\,\ref{f:wt} is that the tectal ligand direction of
+$L_1$ is reversed and the interaction of $r_1$ and $L_1$/$l_1$ is made
+attractive. $L_1$ retains the assumption of exponential change. Note the
+`bunching up' of the field to the medial side of the tectal sheet.}
+\label{f:ee_oneattractive}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_eel_ephb_wt.png}
+\caption{The arrangement is restored by making the expression of $L_1$ on the
+tectal sheet change logarithmically.}
+\label{f:eel_oneattractive}
+\end{figure}
 
@@
 
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_chemo_manipulations.png}
+\caption{Surgical manipulations in the chemoaffinity-only model. Row A: wildtype. Row B: 90 degree
+rotation. Row C: 180 degree rotation. Row D: Graft swap}
+\label{f:chemosurg}
+\end{figure}
+
 \subsection*{Grating simulations}

### 2021-07-23 scripts/fig_chemo_surgical1.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+# Rotational wildtype, but with different selected axons:
+./build/sim/agent/agent1 ./configs/a1/m_ee_chemoonly.json ./configs/a1/e_wt_rotcmp.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee_chemoonly.json ./configs/a1/e_tecrot90.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee_chemoonly.json ./configs/a1/e_tecrot180.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee_chemoonly.json ./configs/a1/e_tecswap.json

### 2021-07-23 scripts/fig_surgical2.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+# Rotational wildtype, but with different selected axons:
+./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_retablate.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_tecablate.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_mismatch.json
+
+./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_compound.json

### 2021-08-02 paper/paper.tex
Looking at the genetic manipulation expts brown and reber

@@
 
-To enable a simulation it is necessary to make the assumptions (in addition to
-those implicit in the description of the model above) that i) All
-receptor-ligand signalled interactions are repulsive, as for EphA-ephrinA
+To enable a simulation it is necessary to make assumptions about the nature of
+the receptor-ligand signalled interactions (i.e. whether they are attractive
+or repulsive) and about the pattern of ligand expression on the tectal
+surface. We assume that i) All receptor-ligand signalled interactions are
+repulsive, as for EphA-ephrinA
 coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
 all tectal ligand expressions vary exponentially along a given axis, following
-Eq.\,\ref{e:tecexp}. The result of running this system are presented in
+Eq.\,\ref{e:tecexp}. The results of running such a system are presented in
 Fig.\,\ref{f:wt}.
@@
 
+\subsubsection*{Surgical Manipulations}
+
 Fig.\,\ref{f:surg1} shows the result of simulating three surgical
@@
 \end{figure}
+
+\subsubsection*{Genetic manipulations}
+
+Brown et al. EphA knock-in.
+
+Describe Reber and other gene knock-in and knock-out experiments.
 

### 2021-08-02 sim/agent/agent1.h

@@ -536,14 +536,24 @@ struct Agent1
         }
 
         // Knockin will increase expression for half of all cells. There's code in tissue.h to make this randomised or regular
+        bool genetic_manipulation = false;
         T affected = T{0.5};
-        T amount = T{1};
+        T amount = this->conf->getDouble ("amount", 1);
         if (this->conf->getBool ("reber", false)) {
             if (manipulated) { throw std::runtime_error ("Code is only tested for one manipulation at a time!"); }
             // Knockin/knockdown receptor 0:
             this->ret->receptor_knockin (0, affected, amount);
             this->ret->receptor_knockdown (0, amount);
             manipulated = true;
+            genetic_manipulation = true;
+        }
+
+        if (this->conf->getBool ("brown", false)) {
+            if (manipulated) { throw std::runtime_error ("Code is only tested for one manipulation at a time!"); }
+            // Knockin receptor 0 in half of RGCs (but with no knockdown):
+            this->ret->receptor_knockin (0, affected, amount);
+            manipulated = true;
+            genetic_manipulation = true;
         }
 
         std::cout << "Retina has " << this->ret->num() << " cells\n";
@@ -593,6 +603,9 @@ struct Agent1
         bool totally_random = this->mconf->getBool ("totally_random_init", true);
         std::string branch_model = this->mconf->getString ("branch_model", "james_agent");
 
+        std::array<float, 3> red = { 1.0f, 0.0f, 0.0f };
+        std::array<float, 3> blue = { 0.0f, 0.0f, 1.0f };
+
         // A loop to set up each branch object in pending_branches.
         for (unsigned int i = 0; i < this->pending_branches.size(); ++i) {
             // Set the branch's termination zone
@@ -633,6 +646,12 @@ struct Agent1
             }
 
             this->ax_centroids.p[ri] += initpos / static_cast<T>(bpa);
+
+            if (genetic_manipulation == true) { // genetic manipulation of retinal receptor 0
+                // Set colour red or blue depending on if receptor 0 in the retina was manipulated or not.
+
+                this->ax_centroids.clr[ri] = this->ret->rcpt_manipulated[ri][0] == true ? red : blue;
+            }
 
             // "experiment suggests": The target for axon centroids is defined by their
             // origin location on the retina. However, their target on the tectum is the
@@ -738,6 +757,11 @@ struct Agent1
         }
 
         if (this->conf->getBool ("reber", false)) {
[…]

### 2021-08-04 scripts/fig_surgical.sh
WIP

@@ -1,10 +1,10 @@
 #!/bin/sh
 
 # Rotational wildtype, but with different selected axons:
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_wt_rotcmp.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_wt_rotcmp.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_tecrot90.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_tecrot90.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_tecrot180.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_tecrot180.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_tecswap.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_tecswap.json

### 2021-08-04 scripts/fig_surgical2.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 
 # Rotational wildtype, but with different selected axons:
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_retablate.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_retablate.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_tecablate.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_tecablate.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_mismatch.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_mismatch.json
 
-./build/sim/agent/agent1 ./configs/a1/m_ee.json ./configs/a1/e_compound.json
+./build/sim/agent/agent1 ./configs/a1/m_ee_axonaxon.json ./configs/a1/e_compound.json

### 2021-08-04 sim/agent/agent1.h

@@ -37,7 +37,7 @@ static constexpr bool visualise = false;
 #endif
 
 // A selection of possible graph layouts to show when running the program
-enum class graph_layout { a, b, c, d };
+enum class graph_layout { a, b, c, d, e };
 
 template<typename T, size_t N, typename B=branch<T, N>>
 struct Agent1
@@ -141,23 +141,31 @@ struct Agent1
         std::cout << "Done simulating\n";
         if constexpr (visualise == true) {
 
-            // Update retinal TN position vs tectal RC position graph for the 'd' layout:
-            if (this->layout == graph_layout::d) {
-                morph::vVector<T>tn(this->ret->posn.size(), 0);
+            // Update retinal NT position vs tectal RC position graph for the 'd' layout:
+            if (this->layout == graph_layout::d || this->layout == graph_layout::e) {
+                morph::vVector<T>nt;
+                morph::vVector<T>rc;
+                morph::vVector<T>nt_m; // manipulated
+                morph::vVector<T>rc_m;
+
                 size_t ii = 0;
-                for (auto p : this->ret->posn) {
-                    tn[ii++] = p[0]; // Get x-component of retinal position into tn.
+                for (ii = 0; ii < this->ret->posn.size(); ++ii) {
+                    if (this->ret->rcpt_manipulated[ii][0]) {
+                        nt_m.push_back(1-this->ret->posn[ii][0]);
+                        rc_m.push_back(this->ax_centroids.p[ii][1]);
+                    } else {
+                        nt.push_back(1-this->ret->posn[ii][0]);
+                        rc.push_back(this->ax_centroids.p[ii][1]);
+                    }
                 }
-                morph::vVector<T>rc(this->ax_centroids.p.size(), 0);
-                ii = 0;
-                for (auto p : this->ax_centroids.p) {
-                    rc[ii++] = p[1]; // Get y-component of final tectal position into rc
-                }
-                if (rc.size() != tn.size()) {
-                    throw std::runtime_error ("rc and tn not same size");
-                }
-                this->gv->setdata (rc, tn);
-                this->gv->reinit(); // re-finalize?
+
+                this->gv->setdata (nt, rc, "wt");
+                T ki_amount = this->conf->getDouble ("knockin", 1);
+                T kd_amount = this->conf->getDouble ("knockdown", 0);
+                std::stringstream ss;
[…]

### 2021-08-04 sim/agent/tissue.h

@@ -658,7 +658,7 @@ struct guidingtissue : public tissue<T>
 
     // For receptor index idx, knock-in affected proportion (range [0,1]) of cells,
     // adding amount as a constant increase in receptor expression
-    static constexpr bool random_knockin = false;
+    static constexpr bool random_knockin = true;
     void receptor_knockin (size_t idx, T affected, T amount)
     {
         if (idx >= N) { throw std::runtime_error ("receptor index out of range"); }

### 2021-08-06 paper/paper.tex
WIP
WIP. Thinking about competition and axon interaction distances

@@
   in a noisy environment if axons compete \\
+  \emph{or} \\
   Time to stop; modelling chemoaffinity with competition \\
+  \emph{or} \\
+  Can the retinotectal projection be modelled exclusively by receptor-ligand interactions?
 }
@@
 
-In our model, space-based competition was omitted and chemoaffinity and
-axon-axon interactions were remodelled so that all interactions were via
-signal transmission between receptors and ligands, which are expressed in
-graded patterns on both retina and tectum. Signals induce attractive or
-repulsive interactions; for chemoaffinity the interaction is to ascend or
-descend the gradient of ligand expression on the tectum; for axon-axon
-interactions, the strength of short-range repulsion (or attraction) is
-governed by the receptor-ligand expression on pairs of axons.
+In our model, space-based competition was retained (though sometimes omitted)
+and chemoaffinity and axon-axon interactions were remodelled so that all
+interactions occurred via signal transmission between receptors and ligands,
+both of which are expressed in graded patterns on both retina and
+tectum. Signals induce attractive or repulsive interactions; for chemoaffinity
+the interaction is to ascend or descend the gradient of ligand expression on
+the tectum; for axon-axon interactions, the strength of short-range repulsion
+(or attraction) is governed by the receptor-ligand expression on pairs of
+axons.
 
@@
 
-Figure~\ref{f:ex} shows colour maps of receptor and ligand
-expression, in this case tectal ligand expression maps are linear. Note that
-the expression is discretized, and in simulations, the gradient is computed
-numerically from the discretized expression. This makes it straightforward to
-perform tectal and retinal `graft' manipulations.
+Figure~\ref{f:ex} shows colour maps of receptor and ligand expression, in this
+case tectal ligand expression maps are linear. Note that the expression is
+discretized, and in simulations, the gradient is computed numerically from the
+discretized expression. This makes it straightforward to perform tectal and
+retinal `graft' manipulations. Coordinates within the simulated retina and
+tectum are normalised, with 1 corresponding to 2 mm on the tectum and on the
+retina for mouse \citep{reber_relative_2004}. This means that growth cones of
+approximate radius 5 $\mu$m = 0.005 mm \citep{goodhill_molecular_2016} have a
+radius $r$ = 0.0025 in the arbitrary units of the simulation. For a realistic
+simulation based on mouse, the approximate RGC density in the retina is around
+44000 cells per retina \citep{jeon_major_1998}. This sets the side length of
+the retina to $\sqrt{44000} \approx 200$. This results in a simulation of
+prodigious computational size, which although tractable (10 s per step,
+requiring several hours, rather than days to run) is not convenient. To
+maintain the relationship between growth cone size and cell density, we permit
+a smaller number of growth cones ($400\times8=3200$) but keep the fractional
+area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
+per axon) or 6.9 (if there are 8 cones per axon).
 
@@
 
-At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, of each
-branch, $b$, is updated according to
+At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
+$t$ of each branch $b$ is updated according to
 %
@@
 where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
-effect, $\mathbf{G}_b$, an axon-axon interaction, $\mathbf{I}_b$ and a `border effect',
+effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
+receptor-ligand interaction, $\mathbf{I}_b$ and a `border effect',
 $\mathbf{B}_b$, which acts to retain branches within the tectal region:
@@
 \begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{I}_b + m_3 \mathbf{B}_b
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b + m_4 \mathbf{B}_b
 \end{equation}
 %
-evaluated at time $t$. $m_1$, $m_2$ and $m_3$ are scalar parameters.
+evaluated at time $t$. $m_1$--$m_4$ are scalar parameters.
 
@@
 -1 for a repulsive interaction or 1 for an attractive interaction.
+
+\subsection*{Axon-axon competition}
+
+Competition, in the context of this model, is a repulsive effect which acts
+between any two branches, whether they originate from different retinal
+ganglion cells or from the same cell. It assumes that there is some kind of
+self-other signal (\textbf{be sure to discuss possible origins}) which acts to
+discourage two cells from approaching each other.
+%
+\begin{equation}
+\mathbf{C}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb} W
+\end{equation}
+where the distance based weight $W$ is given by
+\begin{equation}
+W(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r \\
+     0 & d_{kb} > 2r
+     \end{cases}
+\end{equation}
+for two growth cones of radius $r$ separated by a distance $d_{kb}$. $B_b$ is
+the set of growth cones within a distance $2r$ of branch $b$.
 
@@
 
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_brown.png}
+\caption{Brown et al simulation. This is EphA3 knockin - it adds a constant to
+half of the RGC's $r_0$ (randomly selected). Compare with S\&G Fig.\,6.}
+\label{f:brown}
+\end{figure}
+
 Describe Reber and other gene knock-in and knock-out experiments.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_reber.png}
+\caption{Reber et al simulation. Follow on to Brown et al (same
+group). Combines the selective EphA3 knockin with a uniform EphA4 knockdown.}
+\label{f:reber}
+\end{figure}
 
@@
 Introductory sentence.
+
+\textbf{FIXME: Find the expression functions for the tectum which give a
+decent WT result. Perhaps I already did this? Experiment with these and
+axon-axon with smaller radius of interaction.}
 

### 2021-08-06 sim/agent/agent1.h

@@ -642,12 +642,17 @@ struct Agent1
         std::array<float, 3> red = { 1.0f, 0.0f, 0.0f };
         std::array<float, 3> blue = { 0.0f, 0.0f, 1.0f };
 
+        float r_conf = this->mconf->getFloat ("r", 0.05f);
+        float rc_conf = this->mconf->getFloat ("rc", r_conf);
+        float rrl_conf = this->mconf->getFloat ("rrl", r_conf);
         // A loop to set up each branch object in pending_branches.
         for (unsigned int i = 0; i < this->pending_branches.size(); ++i) {
             // Set the branch's termination zone
             unsigned int ri = i/bpa; // retina index
             this->pending_branches[i].init();
-            this->pending_branches[i].setr (this->mconf->getFloat ("r", 0.05f));
+            this->pending_branches[i].setr (r_conf);
+            this->pending_branches[i].setrc (rc_conf);
+            this->pending_branches[i].setrrl (rrl_conf);
             this->pending_branches[i].aid = (int)ri; // axon index
             if (conf->getBool ("singleaxon", false)) {
                 this->pending_branches[i].rcpt = this->ret->rcpt[singleaxon_idx]; // FIXME: Use seeaxons

### 2021-08-06 sim/agent/branch_geb.h

@@ -16,8 +16,15 @@ struct branch_geb : public branch_base<T,N>
 {
     // The cone radius, hard coded
     void setr (T _r) { this->r = _r; }
+    T getr() { return this->r; }
+    T getrc() { return this->rc; }
+    T getrrl() { return this->rrl; }
+    void setrc(T _r) {}
+    void setrrl(T _r) {}
 protected:
     T r = T{0.05};
+    T rc = T{0.0};
+    T rrl = T{0.0};
 
 public:
     // For random angle choice

### 2021-09-07 paper/paper.tex
Work in progress to add receptor-ligand AND receptor-receptor interactions

@@
 \title {
-  Competition can account for stopping in a gradient-following model
-  of the retinotectal projection \\
-  \emph{or} \\
-  Competition provides stopping for the chemoaffinity
-  theory which is robust to noise \\
-  \emph{or} \\
-  Self-organisation of topographically ordered axon connections can take place
-  in a noisy environment if axons compete \\
-  \emph{or} \\
-  Time to stop; modelling chemoaffinity with competition \\
-  \emph{or} \\
-  Can the retinotectal projection be modelled exclusively by receptor-ligand interactions?
+%  Competition can account for stopping in a gradient-following model
+%  of the retinotectal projection \\
+%  \emph{or} \\
+%  Competition provides stopping for the chemoaffinity
+%  theory which is robust to noise \\
+%  \emph{or} \\
+%  Self-organisation of topographically ordered axon connections can take place
+%  in a noisy environment if axons compete \\
+%  \emph{or} \\
+  Time to stop; agent-based modelling of chemoaffinity with competition
 }

### 2021-09-07 sim/agent/branch.h

@@ -30,15 +30,19 @@ struct branch : public branch_base<T,N>
     T getr() { return this->r; }
     T getrc() { return this->rc; }
     T getrrl() { return this->rrl; }
+    T getrrr() { return this->rrr; }
     void setrc (T _r) { this->rc = _r; this->two_rc = _r*T{2}; }
     void setrrl (T _r) { this->rrl = _r; this->two_rrl = _r*T{2}; }
+    void setrrr (T _r) { this->rrr = _r; this->two_rrr = _r*T{2}; }
 protected:
-    T r = T{0.04};  // Note - arrangement is quite strongly dependent on this interaction radius
+    T r = T{0.04};   // Actual radius of a growth cone
     T two_r = T{2}*r;
-    T rc = T{0.04};
+    T rc = T{0.04};  // competition interaction distance (arrangement is quite strongly dependent on this interaction radius)
     T two_rc = T{2}*r;
-    T rrl = T{0.04};
+    T rrl = T{0.04}; // receptor-ligand interaction distance
     T two_rrl = T{2}*r;
+    T rrr = T{0.04}; // receptor-receptor interaction distance
+    T two_rrr = T{2}*r;
 public:
     // Signalling ratio parameter for S&G-type interaction (but on 4 receptors, not 1)
     static constexpr bool s_g_interaction = false;
@@ -108,12 +112,13 @@ public:
 
     // A subroutine of compute_next
     //
-    // Within a set distance of two_r, consider that receptors and ligands on
-    // the two growth cones interact and produce an attractive (and/or repulsive)
+    // Within a set distance of two_r, consider that receptors and ligands on the two
+    // growth cones interact and produce an attractive (and/or repulsive)
     // interaction. Use the strength of this interaction to weight a vector between the
-    // two cones.
+    // two cones. I is the receptor-receptor interaction of S&G, J is my receptor-ligand
+    // axon-axon interaction.
     T compute_for_branch (const guidingtissue<T, N>* source_tissue, branch_base<T, N>* kp,
-                          morph::Vector<T, 2>& C, morph::Vector<T, 2>& I)
+                          morph::Vector<T, 2>& C, morph::Vector<T, 2>& I, morph::Vector<T, 2>& J)
     {
         // Paper deals with U_C(b,k) - the vector from branch b to branch k - and
         // sums these. However, that gives a competition term with a sign error. So
@@ -122,6 +127,9 @@ public:
         T d = kb.length();
[…]

### 2021-09-07 sim/agent/branch_geb.h

@@ -41,7 +41,7 @@ public:
     void compute_next (const std::vector<branch_geb<T, N>>& branches,
                        const guidingtissue<T, N>* source_tissue,
                        const guidingtissue<T, N>* tissue,
-                       const morph::Vector<T, 4>& m,
+                       const morph::Vector<T, 5>& m,
                        const morph::Vector<T, 2*N>& rns)
     {
         // Sums

### 2021-09-09 paper/paper.tex
WIP. Seem to have broken something...

@@
 effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
-receptor-ligand interaction, $\mathbf{I}_b$ and a `border effect',
+receptor-receptor interaction, $\mathbf{I}_b$, and axon-axon receptor-ligand
+interaction, $\mathbf{J}_b$, and a `border effect',
 $\mathbf{B}_b$, which acts to retain branches within the tectal region:
@@
 \begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b + m_4 \mathbf{B}_b
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b +
+m_4 \mathbf{J}_b + m_5 \mathbf{B}_b
 \end{equation}
 %
-evaluated at time $t$. $m_1$--$m_4$ are scalar parameters.
+evaluated at time $t$. $m_1$--$m_4$ are scalar parameters. $m_2$
 

### 2021-09-09 configs/a1/make_models.py

@@ -9,44 +9,51 @@ def make_pobj():
         "totally_random_init" : True,
         "intro_every" : 0,
 
-        "desc_arrays" : "Each array has num_guiders entries, in pairs. the order is [ x1, y1, x2, y2, etc ] where x1 means the 1st receptor/ligand that varies in the x direction",
-        "desc_forms" : "0: linear  1: quadratic  2: exponential  3: logarithmic  4: alt exponential",
+        #"desc_arrays" : "Each array has num_guiders entries, in pairs. the order is [ x1, y1, x2, y2, etc ] where x1 means the 1st receptor/ligand that varies in the x direction",
+        #"desc_forms" : "0: linear  1: quadratic  2: exponential  3: logarithmic  4: alt exponential",
         "ret_receptor_forms"    : [2,2,2,2],
         "ret_ligand_forms"      : [2,2,2,2],
         "tectum_ligand_forms"   : [4,4,4,4],
-        "desc_unused_arrays" : "The forms for tectal receptors are just used when setting up guidingtissue objects",
+        #"desc_unused_arrays" : "The forms for tectal receptors are just used when setting up guidingtissue objects",
         "tectum_receptor_forms" : [2,2,2,2],
 
-        "desc_directions" : "+x: increases with increasing x, etc",
+        #"desc_directions" : "+x: increases with increasing x, etc",
         "ret_receptor_directions"    : ["-x", "-y", "+x", "+y"],
         "ret_ligand_directions"      : ["+x", "+y", "-x", "-y"],
         "tectum_ligand_directions"   : ["+y", "+x", "-y", "-x"],
 
-        "desc_interaction" : "retina forward interaction gives the behaviour of retinal receptor expressing projections when the receptors are activated. May be attraction (1), no effect (0), or repulsion (-1)",
+        #"desc_interaction" : "retina forward interaction gives the behaviour of retinal receptor expressing projections when the receptors are activated. May be attraction (1), no effect (0), or repulsion (-1)",
         "ret_forward_interactions" : [-1,-1,-1,-1],
[…]

### 2021-09-09 scripts/fig_expression.sh

@@ -1,2 +1,2 @@
 #!/bin/sh
-./build/sim/agent/tissuevis ./configs/a1/m_ee_chemoonly.json
+./build/sim/agent/tissuevis ./configs/a1/m_eE_G.json

### 2021-09-09 sim/agent/branch.h

@@ -8,15 +8,6 @@
 #include <morph/MathAlgo.h>
 #include "tissue.h"
 
-// A choice of schemes to deal with axon branches that try to move outside the tissue domain
-enum class border_effect
-{
-    original, // The original scheme proposed by Simpson & Goodhill
-    gradients, // A scheme which acts like there are gradients pushing axon branchs back inside tissue
-    penned,     // Axon branches are 'penned in' - once inside the tissue, they can't move outside the tissue region
-    penned_pushed // Like 'penned in' but also pushed away from the border with 1/r component as in original
-};
-
 // The branch which does chemoaffinity + comp + axon-axon interaction
 //
 // For receptor/ligand interactions we define:
@@ -39,9 +30,9 @@ protected:
     T two_r = T{2}*r;
     T rc = T{0.04};  // competition interaction distance (arrangement is quite strongly dependent on this interaction radius)
     T two_rc = T{2}*r;
-    T rrl = T{0.04}; // receptor-ligand interaction distance
+    T rrl = T{0.04}; // receptor-ligand interaction distance for axon-axon interactions
     T two_rrl = T{2}*r;
-    T rrr = T{0.04}; // receptor-receptor interaction distance
+    T rrr = T{0.04}; // receptor-receptor interaction distance for axon-axon interactions
     T two_rrr = T{2}*r;
 public:
     // Signalling ratio parameter for S&G-type interaction (but on 4 receptors, not 1)
@@ -127,9 +118,6 @@ public:
         T d = kb.length();
         kb.renormalize(); // vector kb is a unit vector
 
-        //if constexpr (debug_compute_branch == true) {
-        //    std::cout << "d=" << d << ", 2rrl=" << this->two_rrl << ", 2rrr=" << this->two_rrr <<  std::endl;
-        //}
         // Space-based competition, C
         //
         // W is a distance-dependent weight, which is 0 outside a distance of two_r and
@@ -156,10 +144,7 @@ public:
         }
         //QI /= N;
         morph::Vector<T, 2> nullvec = {0, 0};
-        //if constexpr (debug_compute_branch == true) {
-        //    std::cout << "QI=" << QI << ", this->s = " << this->s << std::endl;
-        //}
-        I += QI > this->s ? kb * W : nullvec;
+        I += QI > this->s ? kb * W : nullvec; // FIXME FIXME: Need to count up number of interactions and divide by this number (like for comp)
 #else
         T QI = T{0};
         if constexpr (N == 4) {
[…]

### 2021-09-09 sim/agent/branch_geb.h

@@ -19,12 +19,15 @@ struct branch_geb : public branch_base<T,N>
     T getr() { return this->r; }
     T getrc() { return this->rc; }
     T getrrl() { return this->rrl; }
+    T getrrr() { return this->rrr; }
     void setrc(T _r) {}
     void setrrl(T _r) {}
+    void setrrr(T _r) {}
 protected:
     T r = T{0.05};
     T rc = T{0.0};
     T rrl = T{0.0};
+    T rrr = T{0.0};
 
 public:
     // For random angle choice

### 2021-09-10 paper/paper.tex
Add division by |B| to I and J

@@
 
+\subsection*{Observations/Notes}
+
+J has a repulsive effect, which is generally disordering. If competition has
+`a lot of work to do' (i.e. in the m\_ee... models) then a strong J will push
+the axons across the whole tissue, but will upset the local ordering
+relationships. Perhaps J also needs a self-self interaction (i.e. for members
+of the same axon?).
+
 \begin{figure}

### 2021-09-10 configs/a1/make_models.py

@@ -37,22 +37,22 @@ def make_pobj():
         #"desc_r" : "Radius of branches/cones (for visualisation)",
         "r" : 0.015, # 0.005 accurate ish, 0.015 good for vis
         #"desc_rc" : "Interaction radius via competition interaction",
-        "rc" : 0.05,
+        "rc" : 0.04,
         #"desc_rrr" : "Interaction radius via rcpt-rcpt interaction",
-        "rrr" : 0.015, # 0.015 ok. 0.05 S&G
+        "rrr" : 0.04, # 0.015 ok. 0.05 S&G
         #"desc_rrl" : "Interaction radius via rcpt-lgnd interaction",
-        "rrl" : 0.015,
+        "rrl" : 0.04,
         #"desc_s" : "Rcpt-rcpt signalling threshold value",
         "s" : 1.1,
-        #"desc_m1" : "chemoaffinity axon-tectum rcpt-lgnd",
+        #"desc_m1" : "chemoaffinity axon-tectum rcpt-lgnd (G)",
         "m1" : -1,
-        #"desc_m2" : "the axon-axon rcpt-lgnd interactions",
+        #"desc_m2" : "the axon-axon rcpt-lgnd interactions (J)",
         "m2" : -1,
-        #"desc_m3" : "the axon-axon rcpt-rcpt interactions",
+        #"desc_m3" : "the axon-axon rcpt-rcpt interactions (I)",
         "m3" : -1,
         #"desc_mborder" : "border effect",
         "mborder" : 0.5,
-        #"desc_c1" : "axon-axon competition (leave out of paper)",
+        #"desc_c1" : "axon-axon competition (C)",
         "c1" : -1
     }
     return pobj
@@ -122,11 +122,11 @@ def save_model_set (modeltag, pobj):
         json.dump(pobj, outfile, indent=4)
     print ('Wrote {0}'.format (fpath))
 
-# Set params once here
+# Set params once here. Sensitive to ee vs eE, I think
 m1 = 0.002  # G
-m2 = 0.001 # J
-m3 = 0.002  # I
-c1 = 0.2  # C
+m2 = 0.001  # J
+m3 = 0.04   # I
+c1 = 0.2    # C
 
 # These are all the model combos I might have:
 modeltag = 'eE'

### 2021-09-10 sim/agent/branch_stochastic.h

@@ -82,36 +82,43 @@ struct branch_stochastic : public branch<T,N>
         // Chemoaffinity, graded by origin position (i.e termination zone) of each retinal axon
         morph::Vector<T, 2> G = this->compute_chemo (source_tissue, tissue);
 
-        // Competition, C, and Axon-axon interactions, I, computed during the same loop
+        // Competition, C, and Axon-axon interactions, I&J, computed during the same loop
         // over the other branches
         morph::Vector<T, 2> C = {0, 0};
         morph::Vector<T, 2> I = {0, 0};
         morph::Vector<T, 2> J = {0, 0};
 
-        // Other branches are called k, making a set B_b, with a number of members that I call n_k
+        // Other branches are called k, making a set (3 sets) B_b, with a number of members that I call n_k etc
         T n_k = T{0};
+        T n_ki = T{0};
+        T n_kj = T{0};
         for (auto k : branches) {
             if (k.id == this->id) { continue; } // Don't interact with self
-            n_k += this->compute_for_branch (source_tissue, &k, C, I, J);
+            std::bitset<3> cij_added = this->compute_for_branch (source_tissue, (&k), C, I, J);
+            n_k += cij_added[0] ? T{1} : T{0};
+            n_ki += cij_added[1] ? T{1} : T{0};
+            n_kj += cij_added[2] ? T{1} : T{0};
         }
-
-        // Do the 1/|B_b| multiplication
-        if (n_k > T{0}) {
-            C = C/n_k;
-            I = I/n_k;
-        } // else C and I will be {0,0} still
+        // Do the 1/|B_b| multiplication to normalize C and I(!!)
+        if (n_k > T{0}) { C = C/n_k; } // else C will be {0,0} still
+        I = n_ki > T{0} ? I/n_ki : I;
+        J = n_kj > T{0} ? I/n_kj : J;
+        //std::cout << "C=" << C << std::endl;
 
         // Collected non-border movement components
-        morph::Vector<T, 2> nonB = G * m[0] + C * m[3] + I * m[2];
+#if 0
+        morph::Vector<T, 2> R = {0, 0}; // A little random movement, too
+        brng::i()->get(R);
+#endif
+        morph::Vector<T, 2> nonB = G * m[0] + J * m[1] + I * m[2]  + C * m[3]; // + R * m[5];
 
         // Border effect. A 'force' to move agents back inside the tissue boundary
         morph::Vector<T, 2> B = this->apply_border_effect (tissue, nonB);
[…]

### 2021-09-10 paper/paper.tex
WIP, switch to laptop

@@
 \usepackage[outercaption]{sidecap}
+\usepackage{subfigure}
 
@@
 %
-evaluated at time $t$. $m_1$--$m_4$ are scalar parameters. $m_2$
+evaluated at time $t$. $m_1$--$m_5$ are scalar parameters.
 
@@
 
-\subsection*{Axon-axon competition}
+\subsection*{Axon-axon interactions}
+
+\subsection*{Space based (or self-other) competition}
 
@@
 \begin{equation}
-\mathbf{C}_b = \frac{1}{|B_b|} \sum_k \hat{\mathbf{x}}_{kb} W
+\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
 \end{equation}
@@
 W(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r \\
-     0 & d_{kb} > 2r
+      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r_c \\
+     0 & d_{kb} > 2r_c
      \end{cases}
 \end{equation}
-for two growth cones of radius $r$ separated by a distance $d_{kb}$. $B_b$ is
-the set of growth cones within a distance $2r$ of branch $b$.
+for two growth cones of interaction radius $r_c$ separated by a distance
+$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
+branch $b$.
 
-\subsection*{Axon-axon interactions}
+\subsection*{Receptor-receptor axon-axon interactions}
 
-For axon-axon interactions, repulsion causes a movement of branch $b$ along a
-unit vector, $\hat{\mathbf{x}}_{kb}$, from branch $k$ to branch $b$ causing
-them to move further apart; attraction causes the opposite movement. The
-interaction, $\mathbf{I}_b$, acting on branch $b$, is given by
+We model interactions between axon growth cones due to signalling between
+receptors of the same type. For these interactions, repulsion causes a
+movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
+branch $k$ to branch $b$ causing them to move further apart; attraction causes
+the opposite movement. The interaction, $\mathbf{I}_b$, acting on branch $b$,
+is given by
+
+\textbf{FIXME} In the code, I'm mostly using the Reber-like ratio based
+interaction, as described in Simpson and Goodhill.
+
 %
 \begin{equation}
-% No \frac{1}{|B_b|} in this one
+% No \frac{1}{|B_b|} in this one. Really?
 \mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
@@
 %
-where $Q(d_{kb})$ is the signalling strength between two growth cones of
-radius $r$ a distance $d_{kb}$ from one another, given by
+where $Q_I(d_{kb})$ is the signalling strength between two growth cones of
+receptor-receptor interaction radius $r_i$ a distance $d_{kb}$ from one
+another, given by
 %
 \begin{equation}
-Q(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r \\
-     0 & d_{kb} > 2r
+Q_I(d_{kb}) = \begin{cases}
+     \sum_i^N E_i\,r_{i,b}\,r_{i,k}    & d_{kb} \leq 2r_i \\
+     0 & d_{kb} > 2r_i
      \end{cases}
@@
 %
-The sign of $Q$ (which is dependent on the values of $F_i$) determines whether
+The sign of $Q_I$ is dependent on the values of $E_i$, which determines
+whether the receptor-recetpro interaction, $\mathbf{I}_b$, is repulsive or
+attractive. $r_{i,k}$ is the expression of receptor $i$ on branch $k$.
+
+\subsection*{Receptor-ligand axon-axon interactions}
+
+%
+\begin{equation}
+% No \frac{1}{|B_b|} in this one. Really?
+\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
+\end{equation}
+%
+where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
+receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
+given by
+%
+\begin{equation}
+Q_J(d_{kb}) = \begin{cases}
+     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
+     0 & d_{kb} > 2r_j
+     \end{cases}
+\end{equation}
+%
+The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
 the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_wt_fig.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt.png}
 \caption{Gradient following model with axon-axon interactions.}
@@
 
-This result shows that a gradient following model based on receptor-ligand
-interactions and incorporating axon-axon repulsion is able to generate the
-basic mapping observed in the retinotectal projection.
+This result shows that a gradient-following model based on receptor-ligand
+interactions (both for axon-to-tectum and axon-to-axon interactions) is
+able to generate the basic mapping observed in the retinotectal projection.
 
@@
 
+%\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
+
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GC_wt_rotcmp.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot90.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot180.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecswap.png}
+\end{subfigure}
 \caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree

### 2021-09-10 scripts/fig_surgical_GC.sh

@@ -1,0 +1,7 @@
+#!/bin/sh
+
+# Rotational wildtype, but with different selected axons:
+./build/sim/agent/agent1 ./configs/a1/m_eE_GC.json ./configs/a1/e_wt_rotcmp.json
+./build/sim/agent/agent1 ./configs/a1/m_eE_GC.json ./configs/a1/e_tecrot90.json
+./build/sim/agent/agent1 ./configs/a1/m_eE_GC.json ./configs/a1/e_tecrot180.json
+./build/sim/agent/agent1 ./configs/a1/m_eE_GC.json ./configs/a1/e_tecswap.json

### 2021-09-10 scripts/fig_surgical_GJ.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+# Rotational wildtype, but with different selected axons:
+./build/sim/agent/agent1 ./configs/a1/m_eE_GJ.json ./configs/a1/e_wt_rotcmp.json
+
+./build/sim/agent/agent1 ./configs/a1/m_eE_GJ.json ./configs/a1/e_tecrot90.json
+
+./build/sim/agent/agent1 ./configs/a1/m_eE_GJ.json ./configs/a1/e_tecrot180.json
+
+./build/sim/agent/agent1 ./configs/a1/m_eE_GJ.json ./configs/a1/e_tecswap.json

### 2021-09-10 sim/agent/agent1.h

@@ -109,7 +109,9 @@ struct Agent1
             std::cout << "RMS error of axon centroids: " << this->ax_centroids.rms() << std::endl;
             if constexpr (visualise == true) {
                 this->vis(1);
-                this->v->keepOpen();
+                if (this->immediate_exit == false) {
+                    this->v->keepOpen();
+                }
             }
             return;
         }
@@ -178,7 +180,9 @@ struct Agent1
             nss << "./paper/images/" << this->title << ".png";
             this->v->saveImage (nss.str());
             this->vis(this->conf->getUInt ("steps", 1000));
-            this->v->keepOpen();
+            if (this->immediate_exit == false) {
+                this->v->keepOpen();
+            }
         }
     }
 
@@ -877,6 +881,7 @@ struct Agent1
 
         // Finally, set any additional parameters that will be needed with calling Agent1::run
         this->goslow = this->conf->getBool ("goslow", false);
+        this->immediate_exit = this->conf->getBool ("exit", false);
     }
 
 #ifdef VISUALISE
@@ -1287,6 +1292,8 @@ struct Agent1
     unsigned int rgcside = 20;
     // If true, then slow things down a bit in the visualization
     bool goslow = false;
+    // Exit or keep showing graphics?
+    bool immediate_exit = false;
     // How many steps to store history. Note, we might choose not to show all of these
     // in a visualisation?
     static constexpr size_t history = 2000;

### 2021-09-10 sim/agent/branch.h

@@ -156,20 +156,15 @@ public:
 #else
         T QI = T{0};
         if constexpr (N == 4) {
-            T QI0 = T{0};
-            T QI1 = T{0};
-            T QI2 = T{0};
-            T QI3 = T{0};
             // Receptor-receptor activation according to rcptrcpt_interactions
-            QI0 = kp->rcpt[0] * this->rcpt[0] * (source_tissue->rcptrcpt_interactions[0] == interaction::repulsion ? 1 :
-                                                 (source_tissue->rcptrcpt_interactions[0] == interaction::attraction ? -1 : 0));
-            QI1 = kp->rcpt[1] * this->rcpt[1] * (source_tissue->rcptrcpt_interactions[1] == interaction::repulsion ? 1 :
-                                                 (source_tissue->rcptrcpt_interactions[1] == interaction::attraction ? -1 : 0));
-            QI2 = kp->rcpt[2] * this->rcpt[2] * (source_tissue->rcptrcpt_interactions[2] == interaction::repulsion ? 1 :
-                                                 (source_tissue->rcptrcpt_interactions[2] == interaction::attraction ? -1 : 0));
-            QI3 = kp->rcpt[3] * this->rcpt[3] * (source_tissue->rcptrcpt_interactions[3] == interaction::repulsion ? 1 :
-                                                 (source_tissue->rcptrcpt_interactions[3] == interaction::attraction ? -1 : 0));
-            QI = QI0 + QI1 + QI2 + QI3;
+            QI = kp->rcpt[0] * this->rcpt[0] * (source_tissue->rcptrcpt_interactions[0] == interaction::repulsion ? 1 :
+                                                (source_tissue->rcptrcpt_interactions[0] == interaction::attraction ? -1 : 0))
+            + kp->rcpt[1] * this->rcpt[1] * (source_tissue->rcptrcpt_interactions[1] == interaction::repulsion ? 1 :
+                                             (source_tissue->rcptrcpt_interactions[1] == interaction::attraction ? -1 : 0))
+            + kp->rcpt[2] * this->rcpt[2] * (source_tissue->rcptrcpt_interactions[2] == interaction::repulsion ? 1 :
+                                             (source_tissue->rcptrcpt_interactions[2] == interaction::attraction ? -1 : 0))
+            + kp->rcpt[3] * this->rcpt[3] * (source_tissue->rcptrcpt_interactions[3] == interaction::repulsion ? 1 :
[…]

### 2021-09-14 paper/paper.tex
Monday's WIP

@@
 
+The I interaction is absent when the topographic ordering is good. This means
+that it does not fix the retablate condition. Like the other competition
+effects, it is strongly affected by the radius of interaction.
+
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_fig.png}
 \caption{Gradient following model with axon-axon interactions.}
@@
 \begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_wt_rotcmp.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_rotcmp.png}
 \end{subfigure}
 \begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot90.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecrot90.png}
 \end{subfigure}
 \begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot180.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecrot180.png}
 \end{subfigure}
 \begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecswap.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecswap.png}
 \end{subfigure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_manipulations2.png}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_retablate.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecablate.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_mismatch.png}
+\end{subfigure}
+\begin{subfigure}{\linewidth}
+\includegraphics[width=\linewidth]{./images/j4_eE_GJ_compound.png}
+\end{subfigure}
 \caption{Surgical manipulations 2. Row A: Retinal ablation, Row B: Tectal

### 2021-09-14 sim/agent/agent1.h

@@ -10,6 +10,8 @@
 
 // If true, turns parallel execution OFF and branch interaction debugging ON.
 static constexpr bool debug_compute_branch = false;
+// If true, then compile code to collect minimum and maximum interactions from branches
+static constexpr bool branch_min_maxes = false;
 
 #include <iostream>
 #include <string>
@@ -86,6 +88,9 @@ struct Agent1
                 this->freeze_times[2] = this->conf->getUInt ("freeze_time2", 80);
                 this->freeze_times[3] = this->conf->getUInt ("freeze_time3", 120);
             }
+
+            // How early to start showing the crossings metric?
+            this->crosscount_from = this->conf->getUInt ("crosscount_from", 1000);
 
             this->visinit();
         }
@@ -221,7 +226,7 @@ struct Agent1
         }
         if (this->layout == graph_layout::a || this->layout == graph_layout::c) {
             this->gv->append ((float)stepnum, this->ax_centroids.sos(), 0);
-            if (stepnum > 1000 && stepnum%50 == 0) {
+            if (stepnum > this->crosscount_from && stepnum%50 == 0) {
                 this->gv->append ((float)stepnum, this->ax_centroids.crosscount(), 1);
             }
         }
@@ -268,6 +273,14 @@ struct Agent1
     //! Perform one step of the simulation
     void step()
     {
+        if constexpr (branch_min_maxes == true) {
+            // Reset minses/maxes for branches to compute min/max in this step
+            for (auto& b : this->branches) {
+                b.minses.set_from (std::numeric_limits<T>::max());
+                b.maxes.set_from(std::numeric_limits<T>::lowest());
+            }
+        }
+
         // Compute the next position for each branch:
 #ifdef __OSX__
         // Mac compiler didn't like omp parallel for in front of a for(auto...
@@ -292,6 +305,22 @@ struct Agent1
                 b.compute_next (this->branches, this->ret, this->tectum, this->m, rns);
             }
 #endif
+        }
+
+        if constexpr (branch_min_maxes == true) {
+            // Check mins/maxes
+            morph::Vector<T,N> minses;
+            minses.set_from(std::numeric_limits<T>::max());
+            morph::Vector<T,N> maxes;
+            maxes.set_from(std::numeric_limits<T>::lowest());
+            for (auto b : this->branches) {
+                for (size_t i = 0; i < N; ++i) {
[…]

### 2021-09-14 sim/agent/branch_geb.h

@@ -32,7 +32,9 @@ protected:
 public:
     // For random angle choice
     morph::RandUniform<T> rng;
-    T s; // usused, dummy.
+    T s; // unused, dummy.
+    morph::Vector<T, N> minses; // unused, dummy.
+    morph::Vector<T, N> maxes; // unused, dummy.
 
     // No init for branch_geb
     virtual void init() {}

### 2021-11-15 paper/receptor_receptor_massaction.tex
A description of a mass-action receptor-receptor interaction, if it were to be used.

--- 2021-09-10 paper/paper.tex
@@
-% This is an example of using latex for a paper/report of specified
-% size/layout. It's useful if you want to provide a PDF that looks
-% like it was made in a normal word processor.
-
-% While writing, don't stop for errors
-\nonstopmode
-
-% Use the article doc class, with an 11 pt basic font size
-\documentclass[11pt, a4paper]{article}
-
-% Makes the main font Nimbus Roman, a Times New Roman lookalike:
-%\usepackage{mathptmx}% http://ctan.org/pkg/mathptmx
-% OR use this for proper Times New Roman (from msttcorefonts package
-% on Ubuntu). Use xelatex instead of pdflatex to compile:
-\usepackage{fontspec}
-\usepackage{xltxtra}
-\usepackage{xunicode}
-\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
-\setmainfont{Times New Roman}
-
-% Nice citations
-\usepackage{natbib}
-
-% Set margins
-\usepackage[margin=2.5cm]{geometry}
-
-% Multilingual support
-\usepackage[english]{babel}
-
-% Nice mathematics
-\usepackage{amsmath}
-
-% Left right harpoons for kinetic equations
-\usepackage{mathtools}
-
-% Control over maketitle
-\usepackage{titling}
-
-% Section styling
-\usepackage{titlesec}
-
-% Ability to use colour in text
-\usepackage[usenames]{color}
-
-% For the \degree symbol
-\usepackage{gensymb}
-
-% Allow includegraphics and nice wrapped figures
-\usepackage{graphicx}
-\usepackage{wrapfig}
-\usepackage[outercaption]{sidecap}
-\usepackage{subfigure}
-
-% Nice quotes
-\usepackage{csquotes}
-
-% Set formats using titlesec
-\titleformat*{\section}{\bfseries\rmfamily}
-\titleformat*{\subsection}{\bfseries\itshape\rmfamily}
-
-% thetitle is the number of the section. This sets the distance from
-% the number to the section text.
-\titlelabel{\thetitle.\hskip0.3em\relax}
-
-% Set title spacing with titlesec, too.  The first {1.0ex plus .2ex
-% minus .7ex} sets the spacing above the section title. The second
-% {-1.0ex plus 0.2ex} sets the spacing the section title to the
-% paragraph.
-\titlespacing{\section}{0pc}{1.0ex plus .2ex minus .7ex}{-1.1ex plus 0.2ex}
-
-%% Trick to define a language alias and permit language = {en} in the .bib file.
-% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
-\usepackage{letltxmacro}
-\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
-\makeatletter
-\DeclareRobustCommand{\selectlanguage}[1]{%
-  \@ifundefined{alias@\string#1}
-    {\ORIGselectlanguage{#1}}
-    {\begingroup\edef\x{\endgroup
-       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
-}
-\newcommand{\definelanguagealias}[2]{%
-  \@namedef{alias@#1}{#2}%
-}
-\makeatother
-\definelanguagealias{en}{english}
-\definelanguagealias{eng}{english}
-%% End language alias trick
-
-%% Any aliases here
-\newcommand{\mb}[1]{\mathbf{#1}} % this won't work?
-% Emphasis and bold.
-\newcommand{\e}{\emph}
-\newcommand{\code}[1]{\textsf{#1}}
-\newcommand{\dvrg}{\nabla\vcdot\nabla}
-%% END aliases
-
-% Custom font defs
-% fontsize is \fontsize{fontsize}{linespacesize}
-\def\authorListFont{\fontsize{11}{11} }
-\def\corrAuthorFont{\fontsize{10}{10} }
-\def\affiliationListFont{\fontsize{11}{11}\itshape }
-\def\titleFont{\fontsize{14}{11} \bfseries }
-\def\textFont{\fontsize{11}{11} }
-\def\sectionHdrFont{\fontsize{11}{11}\bfseries}
-\def\bibFont{\fontsize{10}{10} }
-\def\captionFont{\fontsize{10}{10} }
-
-% Caption font size to be small.
-\usepackage[font=small,labelfont=bf]{caption}
-
-% Make a dot for the dot product, call it vcdot for 'vector calculus
-% dot'. Bigger than \cdot, smaller than \bullet.
-\makeatletter
-\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
-\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
-\makeatother
-
-\def\firstAuthorLast{James}
-
-% Affiliations
-\def\Address{\\
-\affiliationListFont Adaptive Behaviour Research Group, Department of Psychology,
-  The University of Sheffield, Sheffield, UK \\
-}
-
-% The Corresponding Author should be marked with an asterisk. Provide
-% the exact contact address (this time including street name and city
-% zip code) and email of the corresponding author
-\def\corrAuthor{Seb James}
-\def\corrAddress{Department of Psychology, The University of Sheffield,
-  Western Bank, Sheffield, S10 2TP, UK}
-\def\corrEmail{seb.james@sheffield.ac.uk}
-
-% Figure out the font for the author list..
-\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address \\
-  \corrAuthorFont $^{*}$ Correspondence: \corrEmail}
-
-% No page numbering please
-\pagenumbering{gobble}
-
-% A trick to get the bibliography to show up with 1. 2. etc in place
-% of [1], [2] etc.:
-\makeatletter
-\renewcommand\@biblabel[1]{#1.}
-\makeatother
-
-% reduce separation between bibliography items if not using natbib:
-\let\OLDthebibliography\thebibliography
-\renewcommand\thebibliography[1]{
-  \OLDthebibliography{#1}
-  \setlength{\parskip}{0pt}
-  \setlength{\itemsep}{0pt plus 0.3ex}
-}
-
-% Set correct font for bibliography (doesn't work yet)
-%\renewcommand*{\bibfont}{\bibFont}
-
-% No paragraph indenting to match the VPH format
-\setlength{\parindent}{0pt}
-
-% Skip a line after paragraphs
-\setlength{\parskip}{0.5\baselineskip}
-\onecolumn
-
-% titling definitions
-\pretitle{\begin{center}\titleFont}
-\posttitle{\par\end{center}\vskip 0em}
-\preauthor{ % Fonts are set within \Authors
-        \vspace{-1.1cm} % Bring authors up towards title
-        \begin{center}
-        \begin{tabular}[t]{c}
-}
-\postauthor{\end{tabular}\par\end{center}}
-
-% Define title, empty date and authors
-\title {
-%  Competition can account for stopping in a gradient-following model
-%  of the retinotectal projection \\
-%  \emph{or} \\
-%  Competition provides stopping for the chemoaffinity
-%  theory which is robust to noise \\
-%  \emph{or} \\
-%  Self-organisation of topographically ordered axon connections can take place
-%  in a noisy environment if axons compete \\
-%  \emph{or} \\
-Time to stop; agent-based modelling of chemoaffinity with competition
-}
-\date{} % No date please
-\author{\Authors}
-
-% Simplified chemoaffinity explains retinotectal mapping via local axon-axon
-% interactions
-%
-% Focus on the chemoaffinity idea for this model. Show that local interactions
-% can provide an account for growth cones finding their targets.
-
-%% END OF PREAMBLE
-
-\begin{document}
-
-\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
-
-\vspace{-1.8cm} % HACK bring the introduction up towards the title. It
-                % would be better to do this with titling in \maketitle
-
-% Note this is the SfN draft abstract, and will need rewriting.
-\emph{In the classic Chemoaffinity theory, the retinotectal axon projection is
-thought to use pairs of orthogonal signalling gradients in the retina to
-specify the eventual location of synapses made on the surface of the
-tectum/superior colliculus. Similar orthogonal gradients in the tectum provide
-a coordinate system which allows the axons to match their prespecified
-destination with the correct location. Although the Ephrins have been shown to
-guide axons toward their destination, there has yet to emerge a complete
-account of the local interactions which halt the axonal growth cones in the
-correct locations to recreate the topography of the retinal cells. The model
-of \citet{simpson_simple_2011} provides an account of the basic topographic
-arrangement of cells on the tectum, as well as reproducing well known surgical
-and genetic manipulation experiments. However, it suffers from the absence of
-a local chemotactic guidance mechanism. Instead, each agent in their model is
-given instantaneous knowledge of the vector that would move it toward its pre
-specified destination. In addition to the globally supervised chemoaffinity
-term, \citet{simpson_simple_2011} introduced a competitive interaction for
-space between growth cone agents and a receptor-ligand axon-axon interaction
-in order to account for the full set of experimental manipulations. Here, we
-propose the replacement of the chemoaffinity term with a gradient following
-model consisting of axonal growth cone agents which carry receptor molecule
-expression determined by their soma's location of origin on the retina. Growth
-cones move on the simulated tectum guided by two pairs of opposing, orthogonal
-signalling molecules representing the Ephrin ligands. We show that with only
-the chemoaffinity term and a receptor-ligand based axon-axon interaction term
-(meaning that all growth cone interactions are by receptor-ligand signalling),
-a full range of experimental manipulations to the retinotectal system can be
-reproduced. Furthermore, we show that the observation that competition is not
-and essential requirement for axons to find their
-way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
-the opposing influences of signalling gradient pairs. Finally, we demonstrate
-that, assuming exponentially varying receptor expression in the retina, ligand
-expression should either be exponential if the receptor-ligand signal induces
-repulsion (i.e. gradient descent) or logarithmic if the signal induces
-attraction (gradient ascent). Thus, we find that a model analogous to the one
-we presented in \citet{james_modelling_2020} that accounts for murine barrel
-patterning is also a candidate mechanism for the arrangement of the more
-continuous retinotectal system.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-The retinotectal projection has proved to be a deep mine of information for
-the study of how the cells of the central nervous system are accurately
-connected together into functional networks. This projection connects the
-light-gathering cells in the retina to movement-related cells in the optic
-tectum (known as the superior colliculus in mammals). Light sources
-originating close to each other in the environment tend to activate retinal
-cells situated close together in the eye so that an image of the environment
-is formed on the retinal surface. It has been discovered that the topography
-of the retina is preserved within brain regions that process this information
-such that cells which are adjacent within the retina primarily excite cells
-adjacent in the tectum. This indicates that during development there must
-exist a mechanism which ensures the correct arrangement of the axons which
-leave the retina and connect to cells in the tectum.
-
-One reason for the success of the study of the retinotectal projection is its
-capacity to be experimentally manipulated. In some non-mammalian species, both
-the retina and the tectum can be partially ablated, or even physically
-reorganised in-vivo, after which axons regrow to restore the order and
-function of the system for the individual animal. This manipulability was
-exploited in influential work by R. W. Sperry and co-workers during the
-mid-twentieth century, leading to Sperry's 1963 summary of
-the \emph{chemoaffinity theory} \citep{sperry_chemoaffinity_1963} which
-proposes the existence of morphogenetic gradients that guide axons to their
-destination. The chemoaffinity theory was given robust support by the
-discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995}
-which have been shown to form into graded expression fields in the
-retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other
-sensory systems, such as the somatosensory
-system \citep{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
-effect on axonal outgrowth, as shown in in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
-in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
-%
-% A-P:
-%
-% EphA and ephrinAs: Anteroposterior axis of retiontectal projection
-% (Nasal-Temporal on retina; Anteroposterior on tectum).
-%
-% RGM (tectum) Neogenin (retina)
-%
-% D-V:
-%
-% EphB and ephrin B: Dorsoventral on retina; dorsoventral on tectum.
-%
-% En-2: Repels temporal growth cones of xenopus ret-tec axons and attracts
-% nasal ones. Brunet et all, Nature, 2005
-%
-% Ryk/Fz (retina) and Wnt3 (tectum) (Flanahan, 2006)
-%
-
-It is tempting to consider the retinotectal projection well understood. With a
-comprehensive theory supported by a biochemical mechanism, is there anything
-left to understand? That question can be answered by reviewing retinotectal
-modelling papers.
-
-Mini-review here, which contrasts some of the modelling papers. The upshot is
-that a central problem with models is not how the axons know how to get closer
-to their destination, but how they know they have \emph{arrived} at their true
-destination. Also make the point that many of the models are phenomenological
-in nature and do not elucidate the mechanisms behind the arrangement.
-
-In a recent paper, we proposed a self-organising mechanism, based on
-morphogenetic signalling gradients, which can arrange
-axons growing from the thalamus to the somatosensory cortex into the well
-known murine barrel cortex pattern~\citep{james_modelling_2020}. In
-characterising that system, we explored the effect of various types of
-noise. We found that the mechanism was robust to noise in the expression of
-the signalling molecules over a wide range of amplitudes and length-scales,
-and that noise in the interaction parameters, which are obtained by sampling
-the signalling molecules in the source tissue (the thalamic barreloid field)
-could cause topological defects. The question of noise in axon guidance has
-been explored by \citet{goodhill_can_2016}.
-
-\textbf{New introduction:} Simpson and Goodhill presented a model in which the
-chemoaffinity mechanism is non-biological. We offer a model in which a
-competition mechanism counteracts the drive of axons along the ephrin
-gradient. Will we need the axon-axon interaction mechanism based on
-non-same-Eph repulsion?
-
-\textbf{Another approach:} We present a model of gradient-following based on graded
-interactions between the growth cones of retinal ganglion cell axons and the
-tectal surface. We show that a combination of exponentially graded expression
-in the retina \citep{reber_relative_2004} and exponentially (for repulsively
-interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
-interacting receptor/ligands) graded expression on the tectum permits a purely
-local model to reproduce the majority of reported phenomena associated with
-retinotectal organisation.
-
-There have been many approaches to modelling the organisation of the
-retinotectal projection. Many of these models are wholly or partially
-phenomenological.
-
-A model which carefully deals with receptor
-binding is given by \citet{naoki_revisiting_2017} \citep[see also][]{mortimer_bayesian_2009}.
-
-Gradient based models: \citet{nakamoto_topographically_1996}
-
-\section{Model}
-
-Our model was inspired by the agent-based model of \cite{simpson_simple_2011}
-and, in common with that work, consists of agents representing axonal growth
-cones, originating from a square retina and moving upon a square tectum of
-side length 1 in arbitrary units.
-%
-\citet{simpson_simple_2011} model three contributions to the movement of the growth
-cones; chemoaffinity, competition for space and axon-axon interactions. The
-first two of these are modelled without reference to a particular mechanism;
-each agent is provided with globally-supervised information about its
-instantaneous location and its target location, allowing it to move along a
-vector towards the target. Competition is implemented as a movement of
-constant distance along the vector between two agents if those two agents are
-within a threshold distance of one another. Interaction is signal-induced
-repulsion, which is triggered if two agents are both within the threshold
-distance of each other and have compatible EphA receptor expression.
-
-In our model, space-based competition was retained (though sometimes omitted)
-and chemoaffinity and axon-axon interactions were remodelled so that all
-interactions occurred via signal transmission between receptors and ligands,
-both of which are expressed in graded patterns on both retina and
-tectum. Signals induce attractive or repulsive interactions; for chemoaffinity
-the interaction is to ascend or descend the gradient of ligand expression on
-the tectum; for axon-axon interactions, the strength of short-range repulsion
-(or attraction) is governed by the receptor-ligand expression on pairs of
-axons.
-
-\subsection*{Receptor and ligand expression}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Simulated retinal and tectal signalling molecule expression. Four
-receptor types are expressed, and four corresponding ligand types. Receptors
-of type $i$ are activated by ligands of type $i$.
-%
-\textbf{A} Retinal receptor expression, $r_i$. Four receptors are expressed and
-displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
-$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential
-function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$
-and $r_3$ have the opposite sense. Retinal receptor expression is modelled
-with exponential functions because there is ample evidence to support this
-form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
-%
-\textbf{B} Retinal ligand expression, $l_i$, is expressed in complimentary
-gradients \citep{hornberger_modulation_1999}.
-%
-\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown in a linear form,
-as there is less evidence to determine whether ligand expression is
-exponential, linear or logarithmic on the tectum.
-$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
-}
-\label{f:ex}
-\end{figure}
-
-Each retinal ganglion cell projects $n$ growth cone agents (referred to
-as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands,
-$l_i$, indexed by $i$ and
-expressed at levels determined by the cell's location on the retina. The
-expression levels of the receptors and ligands vary with respect to the cell's
-position on the retinal surface. Each receptor or ligand varies with respect
-only to one dimension.  Receptor expression gradients are arranged in
-orthogonal pairs with the gradient of $r_0$ being orthogonal to that of
-$r_1$. $r_2$, whose gradient is opposite to $r_0$, is orthogonal to
-$r_3$. Ligands are also arranged in orthogonal pairs, with $l_0$ opposing
-$r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$ orthogonal to $l_0$ and
-opposing $r_1$.
-
-Because there is convincing evidence that EphA and EphB receptors are
-expressed in exponentially increasing
-patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
-we use an exponential form for retinal receptor expressions, in common with
-other modelling
-studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression
-as \citet{simpson_simple_2011} which is
-\begin{equation}
-r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
-\end{equation}
-
-The simulated tectum expresses the same types of receptors and ligands, also
-in orthogonal pairs of gradients. In this work, we did not consider the effect
-of reverse signalling from tectal receptors to RGC ligands so we only modelled
-tectal ligand expression, $L_i$. Although several studies model tectal ligand
-expression with exponential functions \citep{koulakov_stochastic_2004}, the
-experimental evidence for ligand expression is more ambiguous (Supp. Fig
-\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
-expression may be modelled by exponential, linear or logarithmic functions:
-\begin{equation} \label{e:tecexp}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
-\end{equation}
-or
-\begin{equation}
-L_i^{\text{lin}}(x,y) = 1.31 + 2.33 x;
-\end{equation}
-or
-\begin{equation}
-L_i^{\text{log}}(x,y) = 2.32 + 1.29 \log (2.3(x+0.2))
-\end{equation}
-Each function has approximately the same expression values at $x=0$ and
-$x=1$. Opposing functions are obtained by substituting $(1-x)$ for $x$.
-
-While we do not explicitly name $r_0$, $r_1$,
-etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
-EphB, Ryk \citep{schmitt_wntryk_2006} and
-Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $l$ and $L$
-include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
-RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
-role in retinotectal map formation.
-
-Figure~\ref{f:ex} shows colour maps of receptor and ligand expression, in this
-case tectal ligand expression maps are linear. Note that the expression is
-discretized, and in simulations, the gradient is computed numerically from the
-discretized expression. This makes it straightforward to perform tectal and
-retinal `graft' manipulations. Coordinates within the simulated retina and
-tectum are normalised, with 1 corresponding to 2 mm on the tectum and on the
-retina for mouse \citep{reber_relative_2004}. This means that growth cones of
-approximate radius 5 $\mu$m = 0.005 mm \citep{goodhill_molecular_2016} have a
-radius $r$ = 0.0025 in the arbitrary units of the simulation. For a realistic
-simulation based on mouse, the approximate RGC density in the retina is around
-44000 cells per retina \citep{jeon_major_1998}. This sets the side length of
-the retina to $\sqrt{44000} \approx 200$. This results in a simulation of
-prodigious computational size, which although tractable (10 s per step,
-requiring several hours, rather than days to run) is not convenient. To
-maintain the relationship between growth cone size and cell density, we permit
-a smaller number of growth cones ($400\times8=3200$) but keep the fractional
-area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
-per axon) or 6.9 (if there are 8 cones per axon).
-
-\subsection*{Movement}
-
-At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
-$t$ of each branch $b$ is updated according to
-%
-\begin{equation}
-\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
-\end{equation}
-%
-where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
-effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
-receptor-receptor interaction, $\mathbf{I}_b$, and axon-axon receptor-ligand
-interaction, $\mathbf{J}_b$, and a `border effect',
-$\mathbf{B}_b$, which acts to retain branches within the tectal region:
-%
-\begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b +
-m_4 \mathbf{J}_b + m_5 \mathbf{B}_b
-\end{equation}
-%
-evaluated at time $t$. $m_1$--$m_5$ are scalar parameters.
-
-\subsection*{Chemoaffinity}
-
-In this model, the signal transmitted when a ligand binds to a receptor on a
-branch (forward signalling) can lead to one of two effects; the signal may
-induce an attraction towards the ligand-expressing region or a repulsion away
-from it.
-%
-For axon-tectum interactions, we implemented attraction as a climbing of the
-tectal ligand expression gradient and repulsion as gradient descent.
-%
-To determine the strength of the effect we assumed a purely linear receptor
-binding model, and set the chemotactic movement vector of the branch $b$ at
-location $\mathbf{x}_b$ on the tectum to be
-%
-\begin{equation}
-\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
-pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
-tectum and $F_i$ denotes the direction of the interaction induced when a
-molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
--1 for a repulsive interaction or 1 for an attractive interaction.
-
-\subsection*{Axon-axon interactions}
-
-\subsection*{Space based (or self-other) competition}
-
-Competition, in the context of this model, is a repulsive effect which acts
-between any two branches, whether they originate from different retinal
-ganglion cells or from the same cell. It assumes that there is some kind of
-self-other signal (\textbf{be sure to discuss possible origins}) which acts to
-discourage two cells from approaching each other.
-%
-\begin{equation}
-\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
-\end{equation}
-where the distance based weight $W$ is given by
-\begin{equation}
-W(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r_c \\
-     0 & d_{kb} > 2r_c
-     \end{cases}
-\end{equation}
-for two growth cones of interaction radius $r_c$ separated by a distance
-$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
-branch $b$.
-
 \subsection*{Receptor-receptor axon-axon interactions}
@@
 % No \frac{1}{|B_b|} in this one. Really?
-\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
+\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_I(d_{kb})
 \end{equation}
@@
 The sign of $Q_I$ is dependent on the values of $E_i$, which determines
-whether the receptor-recetpro interaction, $\mathbf{I}_b$, is repulsive or
+whether the receptor-receptor interaction, $\mathbf{I}_b$, is repulsive or
 attractive. $r_{i,k}$ is the expression of receptor $i$ on branch $k$.
-
-\subsection*{Receptor-ligand axon-axon interactions}
-
-%
-\begin{equation}
-% No \frac{1}{|B_b|} in this one. Really?
-\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
-\end{equation}
-%
-where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
-receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
-given by
-%
-\begin{equation}
-Q_J(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
-     0 & d_{kb} > 2r_j
-     \end{cases}
-\end{equation}
-%
-The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
-the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
-expression of ligand $i$ on branch $k$.
-
-\subsection*{Border effect}
-
-We implement a border effect based on gradient following by assuming that
-there is some other molecular signal which acts on all branches near the
-boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
-given by:
-%
-\begin{equation}
-B_{b,x} = \begin{cases}
-        r-x      & x<r \\
-        1-r-x    & x>1-r
-\end{cases}
-B_{b,y} = \begin{cases}
-        r-y      & y<r \\
-        1-r-y    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
-molecule expressed around the border of the tectum, whose expression increases
-quadratically outside the tectum and affects any branch touching (or outside)
-the boundary
-
-\subsection*{Initial conditions}
-
-Branches were randomly distributed in a stripe at the rostral side of the
-tectum at the start of each simulation. Each RGC axon was assigned a random
-initial position coordinate:
-\begin{equation}
-\mathbf{x}_{\text{axon},t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the
-range $[p,q)$. Each of the $n$ branches per RGC axon was given its parent
-axon's initial position, plus a randomly generated offset with coordinates
-derived from a normal distribution of mean 0 and standard deviation 0.1.
-
-\section{Results}
-
-\subsection*{Observations/Notes}
-
-J has a repulsive effect, which is generally disordering. If competition has
-`a lot of work to do' (i.e. in the m\_ee... models) then a strong J will push
-the axons across the whole tissue, but will upset the local ordering
-relationships. Perhaps J also needs a self-self interaction (i.e. for members
-of the same axon?).
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt.png}
-\caption{Gradient following model with axon-axon interactions.}
-\label{f:wt}
-\end{figure}
-
-\subsection*{Normal development}
-
-To enable a simulation it is necessary to make assumptions about the nature of
-the receptor-ligand signalled interactions (i.e. whether they are attractive
-or repulsive) and about the pattern of ligand expression on the tectal
-surface. We assume that i) All receptor-ligand signalled interactions are
-repulsive, as for EphA-ephrinA
-coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
-all tectal ligand expressions vary exponentially along a given axis, following
-Eq.\,\ref{e:tecexp}. The results of running such a system are presented in
-Fig.\,\ref{f:wt}.
-%
-Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
-on the square retina, with red encoding position on the $x$ axis (temporal to
-nasal) and green encoding position on the $y$ axis (ventral to
-dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
-the tectal surface of each RGC at the end of the development process. Thus,
-the `red' RGC from the ventro-nasal corner should map to the caudo-medial
-corner of the tectum and the `green' RGC from the dorso-temporal corner should
-find its way to the rostro-lateral corner on the tectum. The lines between the
-dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
-cells. Thus, colour in each panel indicates a branch or axon's location of
-origin on the retina.
-%
-Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
-stochastically arranged around the rostral edge of the tectum. Each dot
-represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
-show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
-well ordered arrangement, with a majority of adjacency relationships matching
-the experimental prediction (i.e. there are few crossed lines). The
-intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
-tectal surface, with final adjacency relationships developing over a longer
-timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
-shows \textbf{RMS error or some metric} of the axon centroids with respect to
-the experimental prediction. Fig.\,\ref{f:wt}F shows the final
-location of each branch in the simulation, which gives a more continuous view
-of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
-branches and their centroid path history for five selected RGCs, with path
-histories strongly resembling those presented by \citet{simpson_simple_2011}.
-
-This result shows that a gradient-following model based on receptor-ligand
-interactions (both for axon-to-tectum and axon-to-axon interactions) is
-able to generate the basic mapping observed in the retinotectal projection.
-
-\subsection*{Manipulations}
-
-To determine if the model can explain the many surgical and genetic
-manipulations described in the literature
-\citep*[for a review, see][]{goodhill_retinotectal_1999} we performed
-simulations matching those in \citet{simpson_simple_2011} in which we either
-rotated or translocated grafts of the tectum; ablated portions of the retina
-and/or the tectum; or manipulated the receptor or ligand expressions in the
-model.
-
-\subsubsection*{Surgical Manipulations}
-
-Fig.\,\ref{f:surg1} shows the result of simulating three surgical
-manipulations alongside the wildtype result, which is shown on Row A. Rows B
-and C show graft rotation simulations (90$\degree$ and 180$\degree$). In each
-case an 8x8 square on the tectum (each unit is 0.05x0.05) is rotated. The
-rotational sense can be observed clearly in the `Branches' column; for the
-90$\degree$ rotation, the top-right (caudal-lateral) corner of the rotated
-square is red, whereas the top-left corner of the wildtype tectum is red. \textbf{etc}.
-
-%\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
-
-\begin{figure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_wt_rotcmp.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot90.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecrot180.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GC_tecswap.png}
-\end{subfigure}
-\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
-rotation. Row C: 180 degree rotation. Row D: Graft swap.}
-\label{f:surg1}
-\end{figure}
-
-Fig.\,\ref{f:surg2} shows retinal ablation, tectal ablation, mismatch and
-compound eye experiments (Stuart: see Simpson and Goodhill Fig. 4).
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_manipulations2.png}
-\caption{Surgical manipulations 2. Row A: Retinal ablation, Row B: Tectal
-ablation, Row C: Mismatch (ablate retina and also the tectum that matches the
-left-behind retina), Row D: Compound eye (Two nasal retinal halves `glued
-together')}
-\label{f:surg2}
-\end{figure}
-
-\subsubsection*{Genetic manipulations}
-
-Brown et al. EphA knock-in.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_brown.png}
-\caption{Brown et al simulation. This is EphA3 knockin - it adds a constant to
-half of the RGC's $r_0$ (randomly selected). Compare with S\&G Fig.\,6.}
-\label{f:brown}
-\end{figure}
-
-Describe Reber and other gene knock-in and knock-out experiments.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_reber.png}
-\caption{Reber et al simulation. Follow on to Brown et al (same
-group). Combines the selective EphA3 knockin with a uniform EphA4 knockdown.}
-\label{f:reber}
-\end{figure}
-
-\subsection*{ligand expression profiles}
-
-A section to show the effect of assuming linear or logarithmic expression
-profiles. Especially that linear expression fails to reproduce the rotation
-and graft swap manipulations. Hmm, actually, it looks ok?!
-
-\subsection*{Attractive receptor-ligand interactions}
-
-Show that, if you switch one of the receptor-ligand  interactions in the
-system to be attractive instead of repulsive, then you have to make the ligand
-expression form logarithmic instead of exponential.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_ephb_wt.png}
-\caption{The effect of making one interaction attractive. This models
-switching one receptor/ligand pair to be an analogue of EphB, but retains an
-assumption that the ephrin-B ligand has exponential form. The change compared
-with the model of Fig.\,\ref{f:wt} is that the tectal ligand direction of
-$L_1$ is reversed and the interaction of $r_1$ and $L_1$/$l_1$ is made
-attractive. $L_1$ retains the assumption of exponential change. Note the
-`bunching up' of the field to the medial side of the tectal sheet.}
-\label{f:ee_oneattractive}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eel_ephb_wt.png}
-\caption{The arrangement is restored by making the expression of $L_1$ on the
-tectal sheet change logarithmically.}
-\label{f:eel_oneattractive}
-\end{figure}
-
-\subsection*{Is chemoaffinity alone sufficient to explain the retinotectal projection?}
-
-Introductory sentence.
-
-\textbf{FIXME: Find the expression functions for the tectum which give a
-decent WT result. Perhaps I already did this? Experiment with these and
-axon-axon with smaller radius of interaction.}
-
-We now describe the behaviour of a model which has only the gradient-based
-chemoaffinity effect and no axon-axon interaction or border effect
-(i.e.~$m_2=m_3=0$). This model was run with $n=8$ branches per RGC axon. All
-receptor-ligand signals induced repulsive interactions (i.e.~gradient
-descent). Receptor and ligand expressions were as shown in Fig.\,\ref{f:ex},
-noting that tectal ligand expression was of exponential form
-(Eq.\,\ref{e:tecexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
-ganglion cell soma locations, with the colour red encoding $x$ position on the
-temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
-position on the dorso-ventral axis.
-%
-Fig.\,\ref{f:ch}B shows the expected final location of each RGC axon,
-according to experimental observations, in a `fishnet' plot which draws lines
-between adjacent axons giving a visual indication of the topology of the
-arrangement.
-%
-Fig.\,\ref{f:ch}C shows the position of
-400 branches at $t=500$ (in arbitrary time units). Considering each
-coordinate axis separately, each branch interacts differentially with two
-opposing gradients. Branches which settle nearest the rostral edge of the
-tectum interact more strongly with the gradient $L_0$ than with $L_2$
-(Fig\,\ref{f:ex}C). A similar pair of gradients separate branches along the
-M-L tectal axis. The branches become positioned on a grid due to the
-discretization of the tectal ligand gradient (as the number of branches and
-tectal elements tends towards infinity, the arangement tends towards a
-continuous map---see Fig.\,S1 \textbf{FIXME}).
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_chemoonly_wt_chemoonly.png}
-\caption{Chemoaffinity model}
-\label{f:ch}
-\end{figure}
-
-Fig.\,\ref{f:ch}D shows a fishnet plot of the RGC axon centroids; the mean
-location of $n$ branches per axon. Because all $n$ branches for each RGC have
-identical receptor expression and they experience only the chemoaffinity
-effect, each branch attains almost the same final position and so the centroid
-graph strongly resembles Fig.\,\ref{f:ch}B. Fig.\,\ref{f:ch}D indicates that
-the topology of the arrangement resembles that of the experimental result, but
-the area over which the branches arrange themselves is limited to about 50\%
-of the tectum. It is possible to tailor the tectal ligand expression, whcih
-adjusts the relative contributions of the opposing interactions, to tune the
-final map. This allows the chemoaffinity model to attain a good match to the
-experimental arrangement (Fig.\~S2 \textbf{fixme}). However, the match is
-highly dependent on the form of receptor/ligand expression in both retinal and
-tectal cells.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_chemo_manipulations.png}
-\caption{Surgical manipulations in the chemoaffinity-only model. Row A: wildtype. Row B: 90 degree
-rotation. Row C: 180 degree rotation. Row D: Graft swap}
-\label{f:chemosurg}
-\end{figure}
-
-\subsection*{Grating simulations}
-
-I'd like to do this if it's not too much work!
-
-\section{Discussion}
-
-The model described above assumes that branches express graded levels of
-receptors which interact with ligands expressed as gradients on the tectum,
-the tectal border and on the surface of other branches, and that only these
-interactions guide their movement.
-
-% Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)
-% This relates to the work \cite{naoki_revisiting_2017}
-
-%
-% BIBLIOGRAPHY
-%
-\selectlanguage{English}
-\bibliographystyle{apalike}
-\bibliography{RetinoTectal}
-
-\end{document}
-
-%% For the noise based paper:
-%% Instead, we treat the receptor binding at the growth cone with greater
-%% care. Based on a Michaelis-Menten model applied to
-%% receptor binding, \citet{mortimer_bayesian_2009} give an expression for the
-%% probability of an individual receptor being bound as a function of the ligand
-%% expression, $\gamma$, and the ligand gradient, $\mu$;
-%% \begin{equation}
-%% P(b=1|\gamma,\mu) = \frac{\gamma(1 + \mu r)}{1 + \gamma(1 + \mu r)},
-%% \end{equation}
-%% where $r$ is the distance of the receptor from the centre of a 1D growth
-%% cone. To determine the growth cone's estimate of the gradient, we assume
-%% evenly spaced receptors and sample from a uniform distribution for each
-%% location across the growth cone. Activated receptors contribute

### 2021-11-15 paper/paper.tex
Minor changes to paper

@@
 \begin{equation}
-W(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r}   & d_{kb} \leq 2r_c \\
+W_c(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
      0 & d_{kb} > 2r_c
@@
 is given by
-
-\textbf{FIXME} In the code, I'm mostly using the Reber-like ratio based
-interaction, as described in Simpson and Goodhill.
-
 %
@@
 % No \frac{1}{|B_b|} in this one. Really?
-\mathbf{I}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q(d_{kb})
+\mathbf{I}_b
+= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_I \qquad \mathrm{if}~r_{i,b}
+/ r_{i,k} < 1.1
 \end{equation}
 %
-where $Q_I(d_{kb})$ is the signalling strength between two growth cones of
-receptor-receptor interaction radius $r_i$ a distance $d_{kb}$ from one
-another, given by
+the weight here is defined as a function of $r_I$:
 %
 \begin{equation}
-Q_I(d_{kb}) = \begin{cases}
-     \sum_i^N E_i\,r_{i,b}\,r_{i,k}    & d_{kb} \leq 2r_i \\
-     0 & d_{kb} > 2r_i
+W_I(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_I}   & d_{kb} \leq 2r_I \\
+     0 & d_{kb} > 2r_I
      \end{cases}
 \end{equation}
-%
-The sign of $Q_I$ is dependent on the values of $E_i$, which determines
-whether the receptor-recetpro interaction, $\mathbf{I}_b$, is repulsive or
-attractive. $r_{i,k}$ is the expression of receptor $i$ on branch $k$.
+$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
 

### 2021-11-15 paper/paper.tex
WIP on paper

@@
 and genetic manipulation experiments. However, it suffers from the absence of
-a local chemotactic guidance mechanism. Instead, each agent in their model is
-given instantaneous knowledge of the vector that would move it toward its pre
-specified destination. In addition to the globally supervised chemoaffinity
-term, \citet{simpson_simple_2011} introduced a competitive interaction for
-space between growth cone agents and a receptor-ligand axon-axon interaction
-in order to account for the full set of experimental manipulations. Here, we
-propose the replacement of the chemoaffinity term with a gradient following
-model consisting of axonal growth cone agents which carry receptor molecule
-expression determined by their soma's location of origin on the retina. Growth
-cones move on the simulated tectum guided by two pairs of opposing, orthogonal
-signalling molecules representing the Ephrin ligands. We show that with only
-the chemoaffinity term and a receptor-ligand based axon-axon interaction term
-(meaning that all growth cone interactions are by receptor-ligand signalling),
-a full range of experimental manipulations to the retinotectal system can be
+a local chemotactic guidance mechanism. Instead, each agent in that model is
+given instantaneous knowledge of the vector that would move it toward its
+pre-specified destination. In addition to the globally supervised
+chemoaffinity term, \citet{simpson_simple_2011} introduced a competitive
+interaction for space between growth cone agents and a receptor-receptor
+axon-axon interaction in order to account for the full set of experimental
+manipulations. Here, we propose the replacement of the chemoaffinity term with
+a gradient following model consisting of axonal growth cone agents which carry
+receptor molecule expression determined by their soma's location of origin on
+the retina. Growth cones move on the simulated tectum guided by two pairs of
+opposing, orthogonal signalling molecules representing the Ephrin ligands. We
+show that with only the receptor-ligand mediated chemoaffinity term and a
+receptor-receptor based axon-axon interaction term (meaning that all growth
+cone interactions are communicated by receptor signalling), a full range of
+experimental manipulations to the retinotectal system can be
 reproduced. Furthermore, we show that the observation that competition is not
-and essential requirement for axons to find their
+an essential requirement for axons to find their
 way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
 the opposing influences of signalling gradient pairs. Finally, we demonstrate
-that, assuming exponentially varying receptor expression in the retina, ligand
-expression should either be exponential if the receptor-ligand signal induces
-repulsion (i.e. gradient descent) or logarithmic if the signal induces
+that, assuming exponentially varying receptor expression in the retina, tectal
+ligand expression should either be exponential if the receptor-ligand signal
+induces repulsion (i.e. gradient descent) or logarithmic if the signal induces
 attraction (gradient ascent). Thus, we find that a model analogous to the one
@@
 
-The retinotectal projection has proved to be a deep mine of information for
-the study of how the cells of the central nervous system are accurately
-connected together into functional networks. This projection connects the
-light-gathering cells in the retina to movement-related cells in the optic
-tectum (known as the superior colliculus in mammals). Light sources
-originating close to each other in the environment tend to activate retinal
-cells situated close together in the eye so that an image of the environment
-is formed on the retinal surface. It has been discovered that the topography
-of the retina is preserved within brain regions that process this information
-such that cells which are adjacent within the retina primarily excite cells
-adjacent in the tectum. This indicates that during development there must
-exist a mechanism which ensures the correct arrangement of the axons which
-leave the retina and connect to cells in the tectum.
+The retinotectal projection has been an important model system in the study of
+how the cells of the central nervous system are accurately connected together
+into functional networks. This projection connects the light-gathering cells
+in the retina to movement-related cells in the optic tectum (known as the
+superior colliculus in mammals). Light sources originating close to each other
+in the environment tend to activate retinal ganglion cells (either
+directly \cite{iRGC_citation} or indirectly via rods and cones) situated close
+together in the eye so that an image of the environment is formed on the
+retinal surface. It has been discovered that the topography of the retina is
+preserved within brain regions that process this information such that cells
+which are adjacent within the retina primarily excite cells adjacent in the
+tectum. This indicates that during development there must exist a mechanism
+which ensures the correct arrangement of the axons which leave the retina and
+connect to cells in the tectum.
 
@@
 
-In our model, space-based competition was retained (though sometimes omitted)
-and chemoaffinity and axon-axon interactions were remodelled so that all
-interactions occurred via signal transmission between receptors and ligands,
-both of which are expressed in graded patterns on both retina and
-tectum. Signals induce attractive or repulsive interactions; for chemoaffinity
-the interaction is to ascend or descend the gradient of ligand expression on
-the tectum; for axon-axon interactions, the strength of short-range repulsion
-(or attraction) is governed by the receptor-ligand expression on pairs of
-axons.
+We sought to replace the globally-supervised mechanisms by local receptor and
+ligand based interactions. First, chemoaffinity was remodelled so that all
+interactions occurred via signal transmission between the receptors expressed
+on the growth cones of retinal ganglion cell axons and the ligands expressed
+across the tectum. Signals induced either an attractive or repulsive
+interaction to cause growth cones to ascend or descend the gradient of ligand
+expression on the tectum.
+
+We then explored the behaviour of models incorporating this chemoaffinity
+mechanism alone and in conjunction with several competition mechanisms. We
+found parameters for the best models of each type (chemo-only [G],
+chemo-competition [GC], chemo-receptor-receptor [GI] and chemo-receptor-ligand
+[GJ]) by an optimisation approach. We then compared each model's ability to
+reproduce the surgical and genetic manipulation results in a manner similar to
+that taken by \citet{simpson_simple_2011} to find the simplest model that
+would fit the data.
 
@@
 as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands,
-$l_i$, indexed by $i$ and
-expressed at levels determined by the cell's location on the retina. The
-expression levels of the receptors and ligands vary with respect to the cell's
-position on the retinal surface. Each receptor or ligand varies with respect
-only to one dimension.  Receptor expression gradients are arranged in
-orthogonal pairs with the gradient of $r_0$ being orthogonal to that of
-$r_1$. $r_2$, whose gradient is opposite to $r_0$, is orthogonal to
-$r_3$. Ligands are also arranged in orthogonal pairs, with $l_0$ opposing
-$r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$ orthogonal to $l_0$ and
-opposing $r_1$.
+$l_i$, indexed by $i$ and expressed at levels determined by the cell's
+location on the retina. The expression levels of the receptors and ligands
+vary with respect to the cell's position on the retinal surface. Each receptor
+or ligand varies only with respect to one dimension.  Receptor expression
+gradients are arranged in orthogonal pairs with the gradient of $r_0$ being
+orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, is
+orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs, with $l_0$
+opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$ orthogonal to
+$l_0$ and opposing $r_1$.
 
@@
 retina for mouse \citep{reber_relative_2004}. This means that growth cones of
-approximate radius 5 $\mu$m = 0.005 mm \citep{goodhill_molecular_2016} have a
+approximate radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a
 radius $r$ = 0.0025 in the arbitrary units of the simulation. For a realistic
@@
 %
-evaluated at time $t$. $m_1$--$m_5$ are scalar parameters.
+evaluated at time $t$. $m_1$--$m_5$ are scalar parameters. We typically
+considered models for which some of $m_2$ to $m_4$ were fixed to zero. A model
+for which $m_3$ and $m_4$ were zero was named a `$\mathbf{GC}$' model as it
+combined the chemoaffinity effect and space-based competition; a model for
+which $m_2$ was zero was a `$\mathbf{GIJ}$' model and so on.
 
@@
 
+We investigated three forms of competition that could be combined with the
+chemoaffinity effect. We investigated space-based competition
+(`$\mathbf{C}$'), as described in \citet{simpson_simple_2011}, a
+receptor-receptor signalling mechanism (`$\mathbf{I}$'), similar to that
+described in \citet{simpson_simple_2011}, and a receptor-ligand mass-action
+signalling model (`$\mathbf{J}$'). Our aim was to determine whether either of
+the receptor/ligand based signalling mechanisms could replace the space-based
+competition.
+
 \subsection*{Space based (or self-other) competition}
 
-Competition, in the context of this model, is a repulsive effect which acts
-between any two branches, whether they originate from different retinal
-ganglion cells or from the same cell. It assumes that there is some kind of
-self-other signal (\textbf{be sure to discuss possible origins}) which acts to
-discourage two cells from approaching each other.
+Space based competition is a repulsive effect which acts between any two
+branches, whether they originate from different retinal ganglion cells or from
+the same cell. It assumes that there is some kind of self-other signal which
+acts to discourage two cells from approaching each other.
 %
@@
 
-We model interactions between axon growth cones due to signalling between
-receptors of the same type. For these interactions, repulsion causes a
+We modelled interactions between axon growth cones due to signalling between
+receptors of the same type. For these interactions, repulsion caused a
 movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
-branch $k$ to branch $b$ causing them to move further apart; attraction causes
-the opposite movement. The interaction, $\mathbf{I}_b$, acting on branch $b$,
+branch $k$ to branch $b$ causing them to move further apart; attraction caused
+the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
 is given by
@@
 \begin{equation}
-% No \frac{1}{|B_b|} in this one. Really?
 \mathbf{I}_b
@@
 
-We implement a border effect based on gradient following by assuming that
+We implemented a border effect based on gradient following by assuming that
 there is some other molecular signal which acts on all branches near the
@@
 
+\subsection*{Simulation}
+
+% All the extra bits of info required to specify the simulation
+To enable a simulation of the model described above it is necessary to make
+some assumptions about the nature of the receptor-ligand signalled
+interactions (i.e. the values of $F_i$ that determine whether interactions are
+attractive or repulsive) and about the pattern of ligand expression on the
+tectal surface. Unless stated otherwise, we assumed that \textbf{i)} All
+receptor-ligand signalled interactions are repulsive, as for EphA-ephrinA
+coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996};
+and \textbf{ii)} all tectal ligand expressions vary exponentially along a
+given axis, following Eq.\,\ref{e:tecexp}.
+
+In all simulations, the border effect, $B_b$, had the same magnitude with
+$m_5=0.5$. The number of branches per axon was 4 unless otherwise stated.
+
 \section{Results}
 
-\subsection*{Observations/Notes}
+The first question we tried to answer was whether chemoaffinity alone,
+implemented as gradient following, is sufficient to pattern the tectum. We set
+$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300 steps.
 
-J has a repulsive effect, which is generally disordering. If competition has
-`a lot of work to do' (i.e. in the m\_ee... models) then a strong J will push
-the axons across the whole tissue, but will upset the local ordering
-relationships. Perhaps J also needs a self-self interaction (i.e. for members
-of the same axon?).
-
-The I interaction is absent when the topographic ordering is good. This means
-that it does not fix the retablate condition. Like the other competition
-effects, it is strongly affected by the radius of interaction.
+% ./build/sim/agent/agent1 configs/a1/m_ee_G.json  configs/a1/e_wt.json 300
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_G_wt.png}
+\caption{The pure gradient-following model.}
+\label{f:chemo}
+\end{figure}
 
@@
 
+\subsection*{Results Observations/Notes}
+
+J has a repulsive effect, which is generally disordering. If competition has
+`a lot of work to do' (i.e. in the m\_ee... models) then a strong J will push
+the axons across the whole tissue, but will upset the local ordering
+relationships. Perhaps J also needs a self-self interaction (i.e. for members
+of the same axon?).
+
+The I interaction is absent when the topographic ordering is good. This means
+that it does not fix the retablate condition. Like the other competition
+effects, it is strongly affected by the radius of interaction.
+
 % Discuss the Hill Equation! https://en.wikipedia.org/wiki/Hill_equation_(biochemistry)

### 2021-11-15 sim/agent/agent1.h

@@ -152,6 +152,7 @@ struct Agent1
 
                 // How early to start showing the crossings metric?
                 this->crosscount_from = this->conf->getUInt ("crosscount_from", 1000);
+                this->crosscount_every = this->conf->getUInt ("crosscount_every", 50);
 
                 this->visinit();
             }
@@ -306,7 +307,7 @@ struct Agent1
         }
         if (this->layout == graph_layout::a || this->layout == graph_layout::c) {
             this->gv->append ((float)stepnum, this->ax_centroids.sos(), 0);
-            if (stepnum > this->crosscount_from && stepnum%50 == 0) {
+            if (stepnum > this->crosscount_from && stepnum%this->crosscount_every == 0) {
                 this->gv->append ((float)stepnum, this->ax_centroids.crosscount(), 1);
             }
         }
@@ -1626,8 +1627,10 @@ struct Agent1
     NetVisual<T>* cv3 = nullptr;
     // Simulation times to stop updating graphs (see graph_layout::c)
     morph::Vector<size_t, 4> freeze_times;
-    // crosscount_from - how early to start showing the crossings count metric?
+    // How early to start showing the crossings count metric.
     unsigned int crosscount_from = 1000;
+    // How often to show the crossings count metric.
+    unsigned int crosscount_every = 50;
     // Centroid visual for targets
     NetVisual<T>* tcv = nullptr;
     // A graph for the SOS metric

### 2021-11-16 paper/paper.tex
The first results section on the chemo-only model

@@
 as \citet{simpson_simple_2011} which is
-\begin{equation}
+\begin{equation} \label{e:retrcpt}
 r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
@@
 expression may be modelled by exponential, linear or logarithmic functions:
-\begin{equation} \label{e:tecexp}
+\begin{equation} \label{e:tecligexp}
 L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
@@
 and \textbf{ii)} all tectal ligand expressions vary exponentially along a
-given axis, following Eq.\,\ref{e:tecexp}.
+given axis, following Eq.\,\ref{e:tecligexp}.
 
@@
 
+% ./build/sim/agent/agent1 configs/simpler/m_ee_G.json  configs/simpler/e_wt_fig1.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2.png}
+\caption{The pure gradient-following model.}
+\label{f:chemo}
+\end{figure}
+
 The first question we tried to answer was whether chemoaffinity alone,
 implemented as gradient following, is sufficient to pattern the tectum. We set
-$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300 steps.
+$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300
+steps. Fig.\,\ref{f:chemo}A indicates the colour code used for retinal cell
+position; dorsal retinal cells are indicated by green, nasal cells by
+red. Lightness gives position along the relevant axis so that cells at the
+dorsal-nasal corner are yellow and those at the ventral-transverse corner are
+black. Fig.\,\ref{f:chemo}B shows the expected layout of retinal ganglion cell
+axons on the tectum. The axons retain the colour of their originating cell on
+the retina, as shown in Fig.\,\ref{f:chemo}A. Nasal cell axons should be
+arranged caudally; dorsal axons laterally; axons should be evenly spaced and
+the pattern should span the entire tectum. Figs.\,\ref{f:chemo}C
+and \ref{f:chemo}D show, respectively, the positions of branches (i.e.~growth
+cones) and axon centroids at $t=300$. Topographic ordering is
+correct. Opposing gradients have sorted the axons along the two axes to form a
+grid of cells. However, the arrangement does not span the full
+tectum. Fig.\,\ref{f:chemo}E shows selected axons (one from the centre and one
+from each corner of the retina) and their position history.
+Fig.\,\ref{f:chemo}F shows two pattern metrics; the sum of squared distance
+errors of the axon centroids and the number of crossings in the 'fish net'
+plot of Fig.\,\ref{f:chemo}D which gives an indication of the disorder in the
+arrangement. The expected number of crossings is 0, as in
+Fig.\,\ref{f:chemo}B.
 
-% ./build/sim/agent/agent1 configs/a1/m_ee_G.json  configs/a1/e_wt.json 300
+The form of the retinal receptor expression (Eq.\,\ref{e:retrcpt}) is based on
+empirical measurement, but the tectal ligand expression
+(Eq.\,\ref{e:tecligexp}) was chosen arbitrarily. To expand the pattern in
+Fig.\,\ref{f:chemo}D so that it covers the entire tectum, we needed only to
+adjust the tectal ligand expression, substituting the alternative exponential
+expression:
+%
+\begin{equation} \label{e:tecligexp2}
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.2 x)
+\end{equation}
+
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig3.json
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_G_wt.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig2.png}
 \caption{The pure gradient-following model.}
@@
 all tectal ligand expressions vary exponentially along a given axis, following
-Eq.\,\ref{e:tecexp}. The results of running such a system are presented in
+Eq.\,\ref{e:tecligexp}. The results of running such a system are presented in
 Fig.\,\ref{f:wt}.
@@
 noting that tectal ligand expression was of exponential form
-(Eq.\,\ref{e:tecexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
+(Eq.\,\ref{e:tecligexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
 ganglion cell soma locations, with the colour red encoding $x$ position on the

### 2021-11-16 sim/agent/agent1.h
Graphing updates

@@ -302,7 +302,9 @@ struct Agent1
         this->cv->reinit(); // Centroids to end
 
         if (this->layout != graph_layout::e && this->layout != graph_layout::f) {
-            this->tcv->reinit(); // Experiment
+            if (this->layout != graph_layout::b) {
+                this->tcv->reinit(); // Experiment
+            }
             this->av->reinit(); // Selected axons
         }
         if (this->layout == graph_layout::a || this->layout == graph_layout::c) {
@@ -1151,6 +1153,13 @@ struct Agent1
 
     // Set up the simulation visualisation scene. This depends on whether this->layout
     // is graph_layout::a, ::b or ::c, etc.
+    //
+    // a: 2x3
+    // b: 1x4 (or 1x3?)
+    // c:
+    // d:
+    // e:
+    // f:
     void visinit()
     {
         // morph::Visual init
@@ -1196,7 +1205,6 @@ struct Agent1
 
         // Offset for visuals
         morph::Vector<float> offset = { -1.5f, -0.5f, 0.0f };
-        morph::Vector<float> offset0 = { -1.5f, -0.5f, 0.0f };
 
         // Adding to the Main Visual second
         this->v->setCurrent();
@@ -1204,7 +1212,6 @@ struct Agent1
         if (this->layout == graph_layout::a // 2 rows, 3 cols. Standard layout for investigations
             || this->layout == graph_layout::d) { // Standard layout tweaked with graphs like in Brown et al
 
-            offset = offset0;
             // Top left
             v->addVisualModel (this->createTissueVisual (v->shaderprog, v->tshaderprog, offset, ret, "Retinal", expression_view::cell_positions, 0, 2));
 
@@ -1322,8 +1329,8 @@ struct Agent1
 
             v->addVisualModel (jtvm);
 
-        } else if (this->layout == graph_layout::b) { // A pared down layout with just 3 graphs
-
+        } else if (this->layout == graph_layout::b) { // A pared down layout with 3 or 4 graphs (expt, branches, centroids, selected) (but decide 3 or 4)
+#if 0
             // Experiment: Another NetVisual view showing the target locations
             this->tcv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
             this->tcv->viewmode = netvisual_viewmode::targetplus;
@@ -1332,6 +1339,7 @@ struct Agent1
             v->addVisualModel (this->tcv);
 
[…]

### 2021-11-16 paper/paper.tex
A better alternative fn for tectal ligands
Graphing updates
More WIP

@@
 attractive or repulsive) and about the pattern of ligand expression on the
-tectal surface. Unless stated otherwise, we assumed that \textbf{i)} All
-receptor-ligand signalled interactions are repulsive, as for EphA-ephrinA
+tectal surface. Unless stated otherwise, we assumed that: \textbf{i)} all
+receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
+for EphA-ephrinA
 coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996};
@@
 
-% ./build/sim/agent/agent1 configs/simpler/m_ee_G.json  configs/simpler/e_wt_fig1.json
+The first question we tried to answer was whether chemoaffinity alone,
+implemented as gradient following, is sufficient to pattern the tectum. We set
+$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300
+steps.
+
+Fig.\,\ref{f:ch}A indicates the colour code used for retinal cell position;
+dorsal retinal cells are indicated by green, nasal cells by red. Lightness
+gives position along the relevant axis so that cells at the dorsal-nasal
+corner are yellow and those at the ventral-transverse corner are
+black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
+axons on the tectum. The axons retain the colour of their originating cell on
+the retina, as shown in Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged
+caudally; dorsal axons laterally; axons should be evenly spaced and the
+pattern should span the entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net'
+plots of the centroid of the branches of each axon at three timepoints. In
+these plots, lines are drawn between axons that are expected to arrange
+themselves adjacently to one another. The number of line crossings in this
+plot gives an indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G
+shows the positions of the individual branches (i.e.~growth cones) at $t=300$
+that contribute to the the plot in Fig.\,\ref{f:ch}D. By this time, the
+topological order matches that shown in Fig.\,\ref{f:ch}E. Opposing gradients
+have sorted the axons along the two axes to form a grid of cells. However, the
+arrangement does not span the full tectum. Fig.\,\ref{f:ch}F shows two pattern
+metrics; the sum of squared distance errors of the axon centroids and the
+number of crossings in the fish net plot of Fig.\,\ref{f:ch}D. The expected
+number of crossings is 0, as in Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows
+selected axons (one from the centre and one from each corner of the retina)
+and their position history.
+
+% ./build/sim/agent/agent1 configs/simpler/m_ee_G.json configs/simpler/e_wt_fig2.json
 \begin{figure}
@@
 \caption{The pure gradient-following model.}
-\label{f:chemo}
+\label{f:ch}
 \end{figure}
-
-The first question we tried to answer was whether chemoaffinity alone,
-implemented as gradient following, is sufficient to pattern the tectum. We set
-$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300
-steps. Fig.\,\ref{f:chemo}A indicates the colour code used for retinal cell
-position; dorsal retinal cells are indicated by green, nasal cells by
-red. Lightness gives position along the relevant axis so that cells at the
-dorsal-nasal corner are yellow and those at the ventral-transverse corner are
-black. Fig.\,\ref{f:chemo}B shows the expected layout of retinal ganglion cell
-axons on the tectum. The axons retain the colour of their originating cell on
-the retina, as shown in Fig.\,\ref{f:chemo}A. Nasal cell axons should be
-arranged caudally; dorsal axons laterally; axons should be evenly spaced and
-the pattern should span the entire tectum. Figs.\,\ref{f:chemo}C
-and \ref{f:chemo}D show, respectively, the positions of branches (i.e.~growth
-cones) and axon centroids at $t=300$. Topographic ordering is
-correct. Opposing gradients have sorted the axons along the two axes to form a
-grid of cells. However, the arrangement does not span the full
-tectum. Fig.\,\ref{f:chemo}E shows selected axons (one from the centre and one
-from each corner of the retina) and their position history.
-Fig.\,\ref{f:chemo}F shows two pattern metrics; the sum of squared distance
-errors of the axon centroids and the number of crossings in the 'fish net'
-plot of Fig.\,\ref{f:chemo}D which gives an indication of the disorder in the
-arrangement. The expected number of crossings is 0, as in
-Fig.\,\ref{f:chemo}B.
 
@@
 (Eq.\,\ref{e:tecligexp}) was chosen arbitrarily. To expand the pattern in
-Fig.\,\ref{f:chemo}D so that it covers the entire tectum, we needed only to
+Fig.\,\ref{f:ch}D so that it covers the entire tectum, we needed only to
 adjust the tectal ligand expression, substituting the alternative exponential
@@
 \begin{equation} \label{e:tecligexp2}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.2 x)
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.1 x)
 \end{equation}
 
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig3.json
+The result at $t=300$ is shown in Fig.\,\ref{f:chalt}A and B. The axon
+centroids are arranged across the full tectal surface.  The arrangement is
+slightly disordered around the middle of the tectum. This is caused by
+numerical artefacts from the discretization of the ligand expression and the
+fact that in the centre, the difference between the opposing receptor-ligand
+signals is very small.  Fig .\,\ref{f:chalt}C shows the error metric, which is
+closer to 0 with the new ligand expression.
+
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig2.json
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig2.png}
-\caption{The pure gradient-following model.}
-\label{f:chemo}
+\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig3.png}
+\caption{The pure gradient-following model (alt exponential).}
+\label{f:chalt}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_fig.png}
-\caption{Gradient following model with axon-axon interactions.}
-\label{f:wt}
-\end{figure}
+%\begin{figure}
+%\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_fig.png}
+%\caption{Gradient following model with axon-axon interactions.}
+%\label{f:wt}
+%\end{figure}
 
@@
 
-Introductory sentence.
-
-\textbf{FIXME: Find the expression functions for the tectum which give a
-decent WT result. Perhaps I already did this? Experiment with these and
-axon-axon with smaller radius of interaction.}
-
-We now describe the behaviour of a model which has only the gradient-based
-chemoaffinity effect and no axon-axon interaction or border effect
-(i.e.~$m_2=m_3=0$). This model was run with $n=8$ branches per RGC axon. All
-receptor-ligand signals induced repulsive interactions (i.e.~gradient
-descent). Receptor and ligand expressions were as shown in Fig.\,\ref{f:ex},
-noting that tectal ligand expression was of exponential form
-(Eq.\,\ref{e:tecligexp}). Fig.\,\ref{f:ch}A gives a colour map of retinal
-ganglion cell soma locations, with the colour red encoding $x$ position on the
-temporal-nasal axis (with red signifying nasal positions). Green encodes $y$
-position on the dorso-ventral axis.
-%
-Fig.\,\ref{f:ch}B shows the expected final location of each RGC axon,
-according to experimental observations, in a `fishnet' plot which draws lines
-between adjacent axons giving a visual indication of the topology of the
-arrangement.
-%
-Fig.\,\ref{f:ch}C shows the position of
-400 branches at $t=500$ (in arbitrary time units). Considering each
-coordinate axis separately, each branch interacts differentially with two
-opposing gradients. Branches which settle nearest the rostral edge of the
-tectum interact more strongly with the gradient $L_0$ than with $L_2$
-(Fig\,\ref{f:ex}C). A similar pair of gradients separate branches along the
-M-L tectal axis. The branches become positioned on a grid due to the
-discretization of the tectal ligand gradient (as the number of branches and
-tectal elements tends towards infinity, the arangement tends towards a
-continuous map---see Fig.\,S1 \textbf{FIXME}).
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_chemoonly_wt_chemoonly.png}
-\caption{Chemoaffinity model}
-\label{f:ch}
-\end{figure}
-
-Fig.\,\ref{f:ch}D shows a fishnet plot of the RGC axon centroids; the mean
-location of $n$ branches per axon. Because all $n$ branches for each RGC have
-identical receptor expression and they experience only the chemoaffinity
-effect, each branch attains almost the same final position and so the centroid
-graph strongly resembles Fig.\,\ref{f:ch}B. Fig.\,\ref{f:ch}D indicates that
-the topology of the arrangement resembles that of the experimental result, but
-the area over which the branches arrange themselves is limited to about 50\%
-of the tectum. It is possible to tailor the tectal ligand expression, whcih
-adjusts the relative contributions of the opposing interactions, to tune the
-final map. This allows the chemoaffinity model to attain a good match to the
-experimental arrangement (Fig.\~S2 \textbf{fixme}). However, the match is
-highly dependent on the form of receptor/ligand expression in both retinal and
-tectal cells.
+This section now at start of results
 
@@
 rotation. Row C: 180 degree rotation. Row D: Graft swap}
-\label{f:chemosurg}
+\label{f:chsurg}
 \end{figure}

### 2021-11-16 sim/agent/agent1.h

@@ -1154,12 +1154,12 @@ struct Agent1
     // Set up the simulation visualisation scene. This depends on whether this->layout
     // is graph_layout::a, ::b or ::c, etc.
     //
-    // a: 2x3
-    // b: 1x4 (or 1x3?)
-    // c:
-    // d:
-    // e:
-    // f:
+    // 0 a: 2x3
+    // 1 b: 1x4 (or 1x3?)
+    // 2 c: 2x4 layout with staggered time "snapshots"
+    // 3 d: 2x3 like standard but with final DV-NT graph for e.g. Reber result
+    // 4 e: 1x3 Branches, centroids and the DV-NT graph
+    // 5 f: 1x1 with inset (just centroids and expt inset)
     void visinit()
     {
         // morph::Visual init

### 2021-11-17 paper/paper.tex
Adds letters to Fig 3

@@
 
-\subsection*{Simulation}
+\subsection*{Assumptions}
 
@@
 
+\subsection*{Chemoaffinity alone}
+
 The first question we tried to answer was whether chemoaffinity alone,
@@
 
-Fig.\,\ref{f:ch}A indicates the colour code used for retinal cell position;
-dorsal retinal cells are indicated by green, nasal cells by red. Lightness
-gives position along the relevant axis so that cells at the dorsal-nasal
-corner are yellow and those at the ventral-transverse corner are
-black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
-axons on the tectum. The axons retain the colour of their originating cell on
-the retina, as shown in Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged
-caudally; dorsal axons laterally; axons should be evenly spaced and the
-pattern should span the entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net'
-plots of the centroid of the branches of each axon at three timepoints. In
-these plots, lines are drawn between axons that are expected to arrange
-themselves adjacently to one another. The number of line crossings in this
-plot gives an indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G
-shows the positions of the individual branches (i.e.~growth cones) at $t=300$
-that contribute to the the plot in Fig.\,\ref{f:ch}D. By this time, the
-topological order matches that shown in Fig.\,\ref{f:ch}E. Opposing gradients
-have sorted the axons along the two axes to form a grid of cells. However, the
-arrangement does not span the full tectum. Fig.\,\ref{f:ch}F shows two pattern
-metrics; the sum of squared distance errors of the axon centroids and the
-number of crossings in the fish net plot of Fig.\,\ref{f:ch}D. The expected
-number of crossings is 0, as in Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows
-selected axons (one from the centre and one from each corner of the retina)
-and their position history.
+Results are shown in Fig.\,\ref{f:ch}. Fig.\,\ref{f:ch}A indicates the colour
+code used for retinal cell position; dorsal retinal cells are indicated by
+green, nasal cells by red. Lightness gives position along the relevant axis so
+that cells at the dorsal-nasal corner are yellow and those at the
+ventral-transverse corner are black. Fig.\,\ref{f:ch}E shows the expected
+layout of retinal ganglion cell axons on the tectum. The axons retain the
+colour of their originating cell on the retina, as shown in
+Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged caudally; dorsal axons
+laterally; axons should be evenly spaced and the pattern should span the
+entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net' plots of the axon
+centroids at three timepoints. An axon's centroid is the average location of
+its 4 branches. In these plots, lines are drawn between axons that are
+expected to arrange themselves adjacently to one another by the end of the
+simulation. The number of line crossings in this plot gives an indication of
+the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the positions of the
+individual branches (i.e.~growth cones) at $t=300$ that contribute to the the
+plot in Fig.\,\ref{f:ch}D. By this time, the topological order matches that
+shown in Fig.\,\ref{f:ch}E. Opposing gradients have sorted the axons along the
+two axes to form a grid of cells. However, the arrangement does not span the
+full tectum. Fig.\,\ref{f:ch}F shows two pattern metrics; the sum of squared
+distance errors of the axon centroids and the number of crossings in the fish
+net plot of Fig.\,\ref{f:ch}D. The expected number of crossings is 0, as in
+Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows selected axons (one from the centre
+and one from each corner of the retina) and their position history.
 
@@
 
+for the tectal ligand expression.
+
 The result at $t=300$ is shown in Fig.\,\ref{f:chalt}A and B. The axon
@@
 fact that in the centre, the difference between the opposing receptor-ligand
-signals is very small.  Fig .\,\ref{f:chalt}C shows the error metric, which is
-closer to 0 with the new ligand expression.
+signals is very small. Nevertheless, it is demonstrated that a modified form
+for the tectal ligand expression (which was found easily by hand tuning)
+allows a well ordered map to form for the wildtype system. Fig
+.\,\ref{f:chalt}C shows the error metric, which is closer to 0 with the new
+ligand expression.
 
@@
 \end{figure}
+
+The `chemoaffinity only' model begins to encounter problems when we introduce
+simulations of experimental manipulations.
 

### 2021-11-17 sim/agent/agent1.h

@@ -1186,7 +1186,7 @@ struct Agent1
         if (this->layout == graph_layout::a || this->layout == graph_layout::d) {
             this->v->setSceneTrans (-0.3762f, 0.6880f, -5.3f);
         } else if (this->layout == graph_layout::b) {
-            this->v->setSceneTrans (-0.948019743f, -0.00213310868f, -2.69999981f);
+            this->v->setSceneTrans (-0.2375f, -0.03f, -2.8f);
         } else if (this->layout == graph_layout::e) {
             this->v->setSceneTrans (-0.890077f, -0.0236414f, -2.6f);
         } else if (this->layout == graph_layout::c) {
@@ -1329,52 +1329,8 @@ struct Agent1
 
             v->addVisualModel (jtvm);
 
-        } else if (this->layout == graph_layout::b) { // A pared down layout with 3 or 4 graphs (expt, branches, centroids, selected) (but decide 3 or 4)
-#if 0
-            // Experiment: Another NetVisual view showing the target locations
-            this->tcv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
-            this->tcv->viewmode = netvisual_viewmode::targetplus;
-            this->tcv->finalize();
-            this->tcv->addLabel ("Experiment", {0.0f, 1.1f, 0.0f});
-            v->addVisualModel (this->tcv);
-
-            offset[0] += 1.3f;
-#endif
-            // Branches: Visualise the branches with a custom VisualModel
-            this->bv = new BranchVisual<T, N, B> (v->shaderprog, v->tshaderprog, offset, &this->branches, &this->ax_history);
-            this->bv->rcpt_scale.compute_autoscale (rcpt_min, rcpt_max);
-            this->bv->target_scale.compute_autoscale (0, 1);
-            this->bv->view = branchvisual_view::discs;
-            this->bv->finalize();
-            this->bv->addLabel ("Branches", {0.0f, 1.1f, 0.0f});
-            this->addOrientationLabels (this->bv, std::string("Tectal"));
-            v->addVisualModel (this->bv);
-
-            // Axon centroids: Centroids of branches viewed with a NetVisual
-            offset[0] += 1.3f;
-            this->cv = new NetVisual<T> (v->shaderprog, v->tshaderprog, offset, &this->ax_centroids);
-            //this->cv->maxlen = this->conf->getDouble ("maxnetline", 1.0);
[…]

### 2021-11-18 paper/paper.tex
WIP. Next job, recreate the chemo-only figure with the eE model
More examples in chemo manips

@@
 
-\subsection*{Chemoaffinity alone}
+\subsection*{Chemoaffinity as the sole patterning mechanism}
 
@@
 \end{equation}
-
+%
 for the tectal ligand expression.
@@
 
-%\begin{figure}
-%\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_fig.png}
-%\caption{Gradient following model with axon-axon interactions.}
-%\label{f:wt}
-%\end{figure}
+% The chemo manipulations figure is made up from 4 figures generated with
+% graph_layout::g
+%
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot90_fig4row1.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot180_fig4row2.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecswap_fig4row3.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row4.json
+%
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{./images/fig_chemo_manipulations.png}
+\caption{Surgical manipulations in the chemoaffinity-only model. Row A: 90 degree
+rotation. Row B: 180 degree rotation. Row C: Graft swap. Row D: mismatch ablation}
+\label{f:chsurg}
+\end{figure}
 
@@
 
-\subsection*{Is chemoaffinity alone sufficient to explain the retinotectal projection?}
-
-This section now at start of results
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_chemo_manipulations.png}
-\caption{Surgical manipulations in the chemoaffinity-only model. Row A: wildtype. Row B: 90 degree
-rotation. Row C: 180 degree rotation. Row D: Graft swap}
-\label{f:chsurg}
-\end{figure}
-
 \subsection*{Grating simulations}

### 2021-11-18 scripts/fig_G_surgical.sh
Changes to automate generation of the chemo manipulations figure

@@ -1,0 +1,8 @@
+#!/bin/sh
+
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot90_fig4row1.json  1000
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot180_fig4row2.json  1000
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecswap_fig4row3.json  1000
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_retablat_fig4row4.json  1000
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecablat_fig4row5.json  1000
+./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row6.json  1000

### 2021-11-18 sim/agent/agent1.h

@@ -1652,9 +1652,9 @@ struct Agent1
         if (!startletter.empty()) { sl = startletter[0]; }
         morph::VisualModel* jtvm = new morph::VisualModel (v->shaderprog, v->tshaderprog, ozero);
         jtvm->addLabel (std::string({sl}), g_A, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
-        jtvm->addLabel (std::string({sl+1}), g_B, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
-        jtvm->addLabel (std::string({sl+2}), g_C, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
-        jtvm->addLabel (std::string({sl+3}), g_D, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
+        //jtvm->addLabel (std::string({sl+1}), g_B, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
+        //jtvm->addLabel (std::string({sl+2}), g_C, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
+        //jtvm->addLabel (std::string({sl+3}), g_D, morph::colour::black, morph::VisualFont::VeraBold, lfs, lpts);
         this->v->addVisualModel (jtvm);
     }
 

### 2021-11-18 paper/paper.tex
Changes to automate generation of the chemo manipulations figure
Incorporates tecrot90 gradients figure

@@
 The `chemoaffinity only' model begins to encounter problems when we introduce
-simulations of experimental manipulations.
+simulations of experimental manipulations. Fig\,\ref{f:chsurg} shows the
+results of 6 simulated surgical manipulations. In row A, a patch of tectal
+tissue has been cut out and rotated by a quarter turn. This is achieved in
+simulation by rotating a square region of the tectal region
+
+% Visualisation of ligand expression and gradients on the tectum for the
+% graft-rotate 90 degrees simulated surgical manipulation
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/Tissuevisb.png}
+\caption{Visualisation of ligand expression and gradients on the tectum for the
+graft-rotate 90 degrees simulated surgical manipulation}
+\label{f:tvrot90}
+\end{figure}
 
@@
 % ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecswap_fig4row3.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row4.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_retablat_fig4row4.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecablat_fig4row5.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row6.json
 %

### 2021-11-18 paper/paper.tex
More tissue visualisation. Next up: figure of all surgical manips, but only expression 0.1

@@
 The `chemoaffinity only' model begins to encounter problems when we introduce
-simulations of experimental manipulations. Fig\,\ref{f:chsurg} shows the
-results of 6 simulated surgical manipulations. In row A, a patch of tectal
-tissue has been cut out and rotated by a quarter turn. This is achieved in
-simulation by rotating a square region of the tectal region
+simulations of experimental surgical manipulations. Simulated surgery is
+performed by manipulating the maps shown in
+Fig.\,\ref{f:ex}. Fig.\,\ref{f:trot90} shows the expression maps and the
+resulting gradients that result from the surgical manipulation in which a
+graft of the tectum is rotated by a quarter turn and replaced. We rotate an
+8x8 square of the tectum and then recompute the gradients. As the expression
+is discretized, we also discretize the gradient, with the gradient in
+direction $x$ computed as the difference between the expression of the two
+neighbouring elements divided by twice the element distance. As there are four
+ligands expressed on the tectum, there are four pairs of gradient maps, all of
+which are shown in Fig.\,\ref{f:trot90}.
 
@@
 % graft-rotate 90 degrees simulated surgical manipulation
+% ./build/sim/agent/tissuevis configs/simpler/m_eE_G.json configs/simpler/e_tecrot90_fig4row1.json
+%
 \begin{figure}
@@
 graft-rotate 90 degrees simulated surgical manipulation}
-\label{f:tvrot90}
+\label{f:trot90}
 \end{figure}
+
+Fig.\,\ref{f:tmanip} shows the tectal ligand expression for ligands $L_1$ and
+$L_2$ for the five other surgical manipulations applied in this work. These
+are a 180\degree rotation (Fig\,\ref{f:tmanip}A), a graft-swap manipulation
+(Fig\,\ref{f:tmanip}B), ablation of half the retina  (Fig\,\ref{f:tmanip}C),
+ablation of half the tectum (Fig\,\ref{f:tmanip}D) and a mismatch ablation in
+which half of the retina is removed along with the half of the tectum to which
+the remaining retina would project (Fig\,\ref{f:tmanip}E).
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/TissueManips.png}
+\caption{Visualisation of ligand expression and gradients on the tectum for the
+other simulated surgical manipulations}
+\label{f:tmanip}
+\end{figure}
+
+Fig\,\ref{f:chsurg} shows the
+results of 6 simulated surgical manipulations. In row A, a patch of tectal
+tissue has been cut out and rotated by a quarter turn. This is achieved in
+simulation by rotating a square region of the tectal region.
 

### 2021-11-18 sim/agent/tissuevisual.h

@@ -57,6 +57,9 @@ public:
         // Note: VisualModel::finalize() should be called before rendering
     }
 
+    //! If true, then for ligand_grad_x/y_single, plot depth.
+    static constexpr bool plot_depth = false;
+
     //! Initialize as a rectangle made of 4 triangles for each rect, with z position
     //! of each of the 4 outer edges of the triangles interpolated, but a single colour
     //! for each rectangle. Gives a smooth surface in which you can see the pixels.
@@ -146,6 +149,13 @@ public:
 
             std::array<float, 3> clr = this->setColour (ri);
 
+            if constexpr (plot_depth == true) {
+                if (this->view == expression_view::ligand_grad_x_single
+                    || this->view == expression_view::ligand_grad_y_single) {
+                    // Optionally show z as depth
+                    z = this->dcolour[ri] / 5.0f;
+                }
+            }
             // First push the 5 positions of the triangle vertices, starting with the centre. All at z=0
             this->vertex_push (this->gtissue->posn[ri][0], this->gtissue->posn[ri][1], z, this->vertexPositions);
             // Use the centre position as the first location for finding the normal vector

### 2021-11-18 scripts/fig_G_surgical.sh
Adds a figure showing expressions for the different manipulations

@@ -6,3 +6,10 @@
 ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_retablat_fig4row4.json  1000
 ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecablat_fig4row5.json  1000
 ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row6.json  1000
+
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_tecrot90_fig4row1.json  1000
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_tecrot180_fig4row2.json  1000
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_tecswap_fig4row3.json  1000
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_retablat_fig4row4.json  1000
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_tecablat_fig4row5.json  1000
+./build/sim/agent/tissuevis configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row6.json  1000

### 2021-11-19 paper/paper.tex
Adds a figure showing expressions for the different manipulations
Ready to move on to the competition section in the 1st draft

--- 2021-11-18 paper/paper.tex
@@
 two axes to form a grid of cells. However, the arrangement does not span the
-full tectum. Fig.\,\ref{f:ch}F shows two pattern metrics; the sum of squared
-distance errors of the axon centroids and the number of crossings in the fish
-net plot of Fig.\,\ref{f:ch}D. The expected number of crossings is 0, as in
-Fig.\,\ref{f:ch}E. Fig.\,\ref{f:ch}H shows selected axons (one from the centre
-and one from each corner of the retina) and their position history.
+full tectum. Fig.\,\ref{f:ch}F shows two pattern metrics. The first metric,
+$\epsilon$, is the sum of squared distance errors of the axon centroids, with
+$\mathbf{x}_{k}$, the location of the centroid of the branches of axon $k$ and
+$\mathbf{x}'_{k}$ the target tectal location for axon $k$:
+%
+\begin{equation}\label{e:eps}
+\epsilon = \sum_k |\mathbf{x}_{k} - \mathbf{x}'_{k}|^2.
+\end{equation}
+%
+$\epsilon$ tends to 0 as the pattern approaches the target layout. However,
+two patterns may have the same non-zero value for $\epsilon$, despite one
+being qualitatively more disordered than the other. For this reason, we also
+use the number of crossings in the fish net plot of Fig.\,\ref{f:ch}D as a
+second metric, called $\eta$. The expected number of crossings in the wildtype
+pattern is 0, as in Fig.\,\ref{f:ch}E, although note that in some patterns,
+the expected number of crossings is non-zero. Fig.\,\ref{f:ch}H shows selected
+axons (one from the centre and one from each corner of the retina) and their
+position history.
 
@@
 
-The result at $t=300$ is shown in Fig.\,\ref{f:chalt}A and B. The axon
-centroids are arranged across the full tectal surface.  The arrangement is
-slightly disordered around the middle of the tectum. This is caused by
+The result at $t=300$ is shown in Figs.\,\ref{f:chalt}A and \ref{f:chalt}B. The
+axon centroids are arranged across the full tectal surface.  The arrangement
+is slightly disordered around the middle of the tectum. This is caused by
 numerical artefacts from the discretization of the ligand expression and the
 fact that in the centre, the difference between the opposing receptor-ligand
-signals is very small. Nevertheless, it is demonstrated that a modified form
-for the tectal ligand expression (which was found easily by hand tuning)
+signals is very small. Nevertheless, the result demonstrates that a modified
+form for the tectal ligand expression (which was found easily by hand tuning)
 allows a well ordered map to form for the wildtype system. Fig
-.\,\ref{f:chalt}C shows the error metric, which is closer to 0 with the new
-ligand expression.
+.\,\ref{f:chalt}C shows the error metric (\textbf{FIXIMAGE}), $\epsilon$,
+which is much closer to 0 with the new ligand expression.
 
@@
 
-The `chemoaffinity only' model begins to encounter problems when we introduce
-simulations of experimental manipulations. Fig\,\ref{f:chsurg} shows the
-results of 6 simulated surgical manipulations. In row A, a patch of tectal
-tissue has been cut out and rotated by a quarter turn. This is achieved in
-simulation by rotating a square region of the tectal region
+This `chemoaffinity only' model begins to encounter problems when we introduce
+simulations of experimental surgical manipulations. Simulated surgery is
+performed by manipulating the maps shown in
+Fig.\,\ref{f:ex}. Fig.\,\ref{f:trot90} shows the expression maps and the
+resulting gradients that result from the surgical manipulation in which a
+graft of the tectum is rotated by a quarter turn and replaced. We rotate an
+8x8 square of the tectum and then recompute the gradients. As the ligand
+expression is discretize rather than continuous, we compute the gradient
+numerically, with the gradient in direction $x$ computed as the difference
+between the expression of the two neighbouring elements divided by twice the
+element distance. As there are four ligands expressed on the tectum, there are
+four pairs of gradient maps, each of which is shown in Fig.\,\ref{f:trot90}.
 
@@
 graft-rotate 90 degrees simulated surgical manipulation}
-\label{f:tvrot90}
-\end{figure}
+\label{f:trot90}
+\end{figure}
+
+Fig.\,\ref{f:tmanip} shows the tectal ligand expression for ligands $L_1$ and
+$L_2$ for the five other surgical manipulations applied in this work. These
+are a 180\degree~rotation (Fig\,\ref{f:tmanip}A), a graft-swap manipulation
+(Fig\,\ref{f:tmanip}B), ablation of half the retina (Fig\,\ref{f:tmanip}C),
+and ablation of half the tectum (Fig\,\ref{f:tmanip}D). A mismatch
+manipulation was also performed in which which half of the retina is removed
+along with the half of the tectum to which the remaining retina would project.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expressions_manipulations.png}
+\caption{Visualisation of ligand expression and gradients on the tectum for the
+other simulated surgical manipulations}
+\label{f:tmanip}
+\end{figure}
+
+Fig\,\ref{f:chsurg} shows the results of 6 simulated surgical
+manipulations. In row A, a patch of tectal tissue has been cut out and rotated
+by a quarter turn, as in Fig.\,\ref{f:trot90}. From left to right, graphs show
+i) the expected final layout of the tectum, according to the results of
+experimental work, with the colour code giving the origin location of the
+axon's soma on the retina, ii) The locations of individual axon branches at
+simulation end, iii) the centroids of each axons' branches and iv) a rather
+useless selected axons box \textbf{FIXME}. Simulation time was 1000 time steps. The final
+locations of branches and axon centroids indicates that a qualitative rotation
+of the graft patch has occurred, with red branches inside the patch being
+located at the lateral-caudal corner rather than the medial-caudal corner. A
+portion of the graft region (the medial side, after rotation) is free of
+branches, indicating that the pattern is partial. A similar result is seen for
+the 180\degree~example in row B, in whcih the caudal half of the rotated patch
+is free of branches, but the rostral half contains branches with the correctly
+rotated orientations.
+
+Fig.\,\ref{f:chsurg}C shows the result of the patch swap manipulation, in
+which 12x4 element grafts, one from the caudal side of the tectum and one from
+the rostral side are swapped. This simulation is unsuccessful in patterning
+the caudal half of the tectum. The large gradients, both positive and
+negative, caused by the discontinuities along the long edges of the grafts
+attracts almost all the branches (whose initial position is the rostral edge)
+and very few pass through to the caudal side.
+
+The retinal ablation shown in Fig.\,\ref{f:chsurg}D is also unsuccessful, with
+the axons from the surviving nasal retina (the transverse half was removed)
+failing to arrange themselves across the entire tectum, due to the pure
+`position specification' which the dual gradient model achieves. Similarly,
+the tectal ablation simulation fails to reproduce experimentally observed
+results in which the retinal map `squashes up' within the remaining tectal
+region. Instead, axons from the transverse retina arrange with normal spacing
+and axons from the nasal retina are pushd to the caudal-most side of the
+remaining tectum (Fig.\,\ref{f:chsurg}E). A similar result occurs in the
+mismatch experiment (Fig.\,\ref{f:chsurg}F), but in this case there are no
+axons from teh transverse retina, which was been ablated and only the axons
+from the nasal retina are present, arranged in a line.
 
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{./images/fig_chemo_manipulations.png}
-\caption{Surgical manipulations in the chemoaffinity-only model. Row A: 90 degree
-rotation. Row B: 180 degree rotation. Row C: Graft swap. Row D: mismatch ablation}
+\includegraphics[width=0.95\linewidth]{./images/fig_chemo_manipulations.png}
+\caption{Surgical manipulations in the chemoaffinity-only model. \textbf{A} 90 degree
+rotation. textbf{B} 180 degree rotation. \textbf{C} Graft swap. \textbf{D}
+Retinal ablation \textbf{E} Tectal ablation \textbf{F} mismatch ablation}
 \label{f:chsurg}
@@
 
+Together, these results show that, while a pure gradient-following
+implementation of Sperry's chemoaffinity theory is able to reproduce the
+wildtype tectal arrangement of retinal ganglion cell axons, it cannot account
+for the results of surgical manipulation experiments.
+
+\subsection*{Competition}
+
+\section{ZEROTH DRAFT/NOTES}
+
 \subsection*{Normal development}
 
-To enable a simulation it is necessary to make assumptions about the nature of
-the receptor-ligand signalled interactions (i.e. whether they are attractive
-or repulsive) and about the pattern of ligand expression on the tectal
-surface. We assume that i) All receptor-ligand signalled interactions are
-repulsive, as for EphA-ephrinA
-coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}; and ii)
-all tectal ligand expressions vary exponentially along a given axis, following
-Eq.\,\ref{e:tecligexp}. The results of running such a system are presented in
-Fig.\,\ref{f:wt}.
 %

### 2021-11-22 paper/paper.tex
A few changes, switching to laptop

@@
 \includegraphics[width=\linewidth]{./images/Tissuevisb.png}
-\caption{Visualisation of ligand expression and gradients on the tectum for the
+\caption{Ligand expression and gradients on the tectum for the
 graft-rotate 90 degrees simulated surgical manipulation}
@@
 \includegraphics[width=\linewidth]{./images/expressions_manipulations.png}
-\caption{Visualisation of ligand expression and gradients on the tectum for the
+\caption{Ligand expression and gradients on the tectum for the
 other simulated surgical manipulations}
@@
 
-Together, these results show that, while a pure gradient-following
+Together, these results confirm that, while a pure gradient-following
 implementation of Sperry's chemoaffinity theory is able to reproduce the
@@
 \subsection*{Competition}
+
+Previous work has shown that competition can contribute to the correct layout
+of the retinotectal system \citep{stuff} and to the patterning of other
+systems including the murine cortical whisker
+barrels\citep{james_modelling_2020} \textbf{(FIXME owt else?)}.
+
+\citet{simpson_simple_2011} showed that repulsive interactions combined with a
+chemoaffinity-like effect could account for the surgical manipulations shown
+in Figs.\,\ref{f:trot90}--\ref{f:chsurg}. To confirm that a competitive
+mechanism could operate successfully in combination with the
+gradient-following chemoaffinity mechanism described above, we introduced
+three competition mechanisms and investigated their behaviour.
+
+These were: space-based competition (`$\mathbf{C}$'), as described
+in \citet{simpson_simple_2011}; a receptor-receptor signalling mechanism
+(`$\mathbf{I}$'), similar to that described in \citet{simpson_simple_2011},
+and a receptor-ligand mass-action signalling model
+(`$\mathbf{J}$'). \citet{simpson_simple_2011} found that both a space based
+competition and a receptor-receptor signalling mechanism were required to
+reproduce the full range of surgical and genetic manipulations explored in
+that work. The first of these, space-based competition, offers no mechanism by
+which axon growth cones are able to communicate and the second mechanism
+explored by \citet{simpson_simple_2011} was implemented only for a single axis
+in the model (EphA receptors, expressed in a gradient along the XX retinal
+axis). We wondered whether a single signalling mechanism could provide a
+suitable competitive, repulsive effect. To investigate, we compared both of
+the \citet{simpson_simple_2011} effects along with a third possible signalling
+mechanism involving the same receptor-ligand interactions employed by the
+gradient following mechansim, because the ligands expressed on the tectum are
+also expressed by the retinal ganglion cells.
+
+\subsection*{Space based (or self-other) competition}
+
+Space based competition is a repulsive effect which acts between any two
+branches, whether they originate from different retinal ganglion cells or from
+the same cell. It assumes that there is some kind of self-other signal which
+acts to discourage two cells from approaching each other.
+%
+\begin{equation}
+\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
+\end{equation}
+where the distance-dependent weight $W_c$ is given by
+\begin{equation}
+W_c(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
+     0 & d_{kb} > 2r_c
+     \end{cases}
+\end{equation}
+for two growth cones of interaction radius $r_c$ separated by a distance
+$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
+branch $b$.
+
+\subsection*{Receptor-receptor axon-axon interactions}
+
+We modelled interactions between axon growth cones due to signalling between
+receptors of the same type. For these interactions, repulsion caused a
+movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
+branch $k$ to branch $b$ causing them to move further apart; attraction caused
+the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
+is given by
+%
+\begin{equation}
+\mathbf{I}_b
+= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_I \qquad \mathrm{if}~r_{i,b}
+/ r_{i,k} < 1.1
+\end{equation}
+%
+the weight here is defined as a function of $r_I$:
+%
+\begin{equation}
+W_I(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_I}   & d_{kb} \leq 2r_I \\
+     0 & d_{kb} > 2r_I
+     \end{cases}
+\end{equation}
+$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
+
+\subsection*{Receptor-ligand axon-axon interactions}
+
+%
+\begin{equation}
+% No \frac{1}{|B_b|} in this one. Really?
+\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
+\end{equation}
+%
+where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
+receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
+given by
+%
+\begin{equation}
+Q_J(d_{kb}) = \begin{cases}
+     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
+     0 & d_{kb} > 2r_j
+     \end{cases}
+\end{equation}
+%
+The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
+the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
+expression of ligand $i$ on branch $k$.
 

### 2021-11-22 paper/paper.tex
Brings the description of the model into the chemoaffinity section of the results

@@
 \usepackage[usenames]{color}
+\definecolor{grey}{rgb}{0.6, 0.6, 0.63}
+\definecolor{black}{rgb}{0, 0, 0}
 
@@
 
+\color{grey}
 \textbf{New introduction:} Simpson and Goodhill presented a model in which the
@@
 \subsection*{Receptor and ligand expression}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Simulated retinal and tectal signalling molecule expression. Four
-receptor types are expressed, and four corresponding ligand types. Receptors
-of type $i$ are activated by ligands of type $i$.
-%
-\textbf{A} Retinal receptor expression, $r_i$. Four receptors are expressed and
-displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
-$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential
-function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$
-and $r_3$ have the opposite sense. Retinal receptor expression is modelled
-with exponential functions because there is ample evidence to support this
-form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
-%
-\textbf{B} Retinal ligand expression, $l_i$, is expressed in complimentary
-gradients \citep{hornberger_modulation_1999}.
-%
-\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown in a linear form,
-as there is less evidence to determine whether ligand expression is
-exponential, linear or logarithmic on the tectum.
-$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
-}
-\label{f:ex}
-\end{figure}
 
@@
 
+\color{black}
 \section{Results}
@@
 
-The first question we tried to answer was whether chemoaffinity alone,
-implemented as gradient following, is sufficient to pattern the tectum. We set
-$m_1=0.002$ and $m_2=m_3=m_4=0$ and ran the model for 300
-steps.
-
-Results are shown in Fig.\,\ref{f:ch}. Fig.\,\ref{f:ch}A indicates the colour
-code used for retinal cell position; dorsal retinal cells are indicated by
-green, nasal cells by red. Lightness gives position along the relevant axis so
-that cells at the dorsal-nasal corner are yellow and those at the
-ventral-transverse corner are black. Fig.\,\ref{f:ch}E shows the expected
-layout of retinal ganglion cell axons on the tectum. The axons retain the
-colour of their originating cell on the retina, as shown in
-Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged caudally; dorsal axons
-laterally; axons should be evenly spaced and the pattern should span the
-entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net' plots of the axon
-centroids at three timepoints. An axon's centroid is the average location of
-its 4 branches. In these plots, lines are drawn between axons that are
-expected to arrange themselves adjacently to one another by the end of the
-simulation. The number of line crossings in this plot gives an indication of
-the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the positions of the
-individual branches (i.e.~growth cones) at $t=300$ that contribute to the the
-plot in Fig.\,\ref{f:ch}D. By this time, the topological order matches that
-shown in Fig.\,\ref{f:ch}E. Opposing gradients have sorted the axons along the
-two axes to form a grid of cells. However, the arrangement does not span the
-full tectum. Fig.\,\ref{f:ch}F shows two pattern metrics. The first metric,
-$\epsilon$, is the sum of squared distance errors of the axon centroids, with
-$\mathbf{x}_{k}$, the location of the centroid of the branches of axon $k$ and
-$\mathbf{x}'_{k}$ the target tectal location for axon $k$:
+The first question we tried to answer was whether a gradient following
+mechansim, acting alone, is sufficient to pattern the tectum. To do so, we
+simulated the retina and tectum as unit square regions and defined retinal
+receptor expression and tectal ligand expression across these squares
+(Fig.\,\ref{f:ex}).
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expressions_fig.png}
+\caption{Simulated retinal and tectal signalling molecule expression. Four
+receptor types are expressed, and four corresponding ligand types. Receptors
+of type $i$ are activated by ligands of type $i$.
+%
+\textbf{A} Retinal receptor expression, $r_i$. Four receptors are expressed and
+displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
+$r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential
+function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$
+and $r_3$ have the opposite sense. Retinal receptor expression is modelled
+with exponential functions because there is ample evidence to support this
+form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
+%
+\textbf{B} Retinal ligand expression, $l_i$, is expressed in complimentary
+gradients \citep{hornberger_modulation_1999}.
+%
+\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown in a linear form,
+as there is less evidence to determine whether ligand expression is
+exponential, linear or logarithmic on the tectum.
+$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
+}
+\label{f:ex}
+\end{figure}
+
+Each retinal ganglion cell projects growth cones (referred to
+as \emph{branches}) which carry a set of 4 receptors, $r_i$ indexed by $i$ and
+expressed at levels determined by the cell soma's location on the
+retina. (Four ligands, $l_i$, are also expressed by RGCs, but these are not
+used by the chemoaffinity simulation.) The expression levels of the receptors
+vary with respect to the soma's position on the retinal surface. Each receptor
+varies only with respect to one dimension.  Receptor expression gradients are
+arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to
+that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, is orthogonal to
+$r_3$.
+
+Because there is convincing evidence that EphA and EphB receptors are
+expressed in exponentially increasing
+patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
+we use an exponential form for retinal receptor expressions, in common with
+other modelling
+studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
+We adopt the same precise form for the retinal receptor expression
+as \citet{simpson_simple_2011} which is
+\begin{equation} \label{e:retrcpt}
+r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
+\end{equation}
+for even values of $i$. The opposing expressions (for even values of $i$)
+are given by substituting $(1-x)$ for $x$.
+
+The tectum expresses ligands, $L_i$, for the retinal receptors, also
+in orthogonal pairs of gradients. Although several studies model tectal ligand
+expression with exponential functions \citep{koulakov_stochastic_2004}, the
+experimental evidence for ligand expression is more ambiguous than that for
+retinal receptor expression (Supp.~Fig
+\textbf{XXXXX}). We kept in mind the possibility that tectal ligand
+expression may be modelled by exponential, linear or logarithmic functions,
+but initially, we set it to the function
+\begin{equation} \label{e:tecligexp}
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
+\end{equation}
+for odd values of $i$ and using the same substitution of $(1-x)$ for $x$ to
+obtain the opposing ligand expressions for $i$ even.
+
+While we do not explicitly name $r_0$, $r_1$,
+etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
+EphB, Ryk \citep{schmitt_wntryk_2006} and
+Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ includes the
+ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
+RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
+r\^ole in retinotectal map formation.
+
+Figure~\ref{f:ex} shows colour maps of receptor and ligand
+expression. Coordinates within the simulated retina and tectum are normalised,
+with 1 corresponding to 2 mm on the tectum and on the retina for
+mouse \citep{reber_relative_2004}.
+
+\color{grey}
+This means that growth cones of approximate
+radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a radius $r$ =
+0.0025 in the arbitrary units of the simulation. For a realistic simulation
+based on mouse, the approximate RGC density in the retina is around 44000
+cells per retina \citep{jeon_major_1998}. This sets the side length of the
+retina to $\sqrt{44000} \approx 200$. This results in a simulation of
+prodigious computational size, which although tractable (10 s per step,
+requiring several hours, rather than days to run) is not convenient. To
+maintain the relationship between growth cone size and cell density, we permit
+a smaller number of growth cones ($400\times8=3200$) but keep the fractional
+area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
+per axon) or 6.9 (if there are 8 cones per axon).
+
+\color{black}
+Each discrete element in Fig.\,\ref{f:ex}B projects one RGC axon with 4
+branches to the tectum, carrying the receptor expression at the retinal
+element's location.
+At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
+$t$ of each branch $b$ is updated according to
+%
+\begin{equation}
+\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
+\end{equation}
+%
+where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
+effect and a `border effect', $\mathbf{B}_b$, which acts to retain branches
+within the tectal region:
+%
+\begin{equation}
+\mathbf{M}_{b} = m_1 \mathbf{G}_b + 0.5 \mathbf{B}_b
+\end{equation}
+%
+evaluated at time $t$.
+
+% G - chemoaffinity description
+The chemoaffinity component, $\mathbf{G}_b$, is governed by receptor-ligand
+signalling. The signal transmitted when a tectal ligand binds to a RGC
+receptor on a branch (forward signalling) can lead to one of two effects; the
+signal may induce an attraction towards the ligand-expressing region or a
+repulsion away from it.
+%
+For axon-tectum interactions, we implemented attraction as a climbing of the
+tectal ligand expression gradient and repulsion as gradient descent.
+%
+To determine the strength of the effect we assumed a purely linear receptor
+binding model, and set the chemotactic movement vector of the branch $b$ at
+location $\mathbf{x}_b$ on the tectum to be
+%
+\begin{equation}
+\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
+\end{equation}
+%
+where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
+pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
+tectum and $F_i$ denotes the direction of the interaction induced when a
+molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
+-1 for a repulsive interaction or 1 for an attractive interaction.
+%
+We assumed that all
+receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
+for EphA-ephrinA
+coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
+%
+$m_1$ is a
+scalar parameter which controls how much movement is generated for a given
+receptor-ligand signal size.
+% end chemoaffinity description
+
+% border effect description
+The border effect is also based on gradient following by assuming that
+there is some other molecular signal which acts on all branches near the
+boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
+given by:
+%
+\begin{equation}
+B_{b,x} = \begin{cases}
+        r-x      & x<r \\
+        1-r-x    & x>1-r
+\end{cases}
+B_{b,y} = \begin{cases}
+        r-y      & y<r \\
+        1-r-y    & y>1-r
+\end{cases}
+\end{equation}
+%
+Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
+molecule expressed around the border of the tectum, whose expression increases
+quadratically outside the tectum and affects any branch touching (or outside)
+the boundary.
+% end border effect description
+
+% Initial conditions
+Branches were initially randomly distributed in a stripe at the rostral side
+of the tectum at the start of each simulation. Each RGC axon was assigned a
+random initial position coordinate:
+\begin{equation}
+\mathbf{x}_{\text{axon},t=0} = (U(0,1), U(-0.2,0))
+\end{equation}
+where U(p,q) is a number selected from a random uniform distribution in the
+range $[p,q)$. Each of the 4 branches per RGC axon was given its parent
+axon's initial position, plus a randomly generated offset with coordinates
+derived from a normal distribution of mean 0 and standard deviation 0.1.
+% end initial conditions
+
+The results of running this simulation for 300 time steps are shown in
+Fig.\,\ref{f:ch}. Fig.\,\ref{f:ch}A indicates the colour code used for retinal
+cell position; dorsal retinal cells are indicated by green, nasal cells by
+red. Lightness gives position along the relevant axis so that cells at the
+dorsal-nasal corner are yellow and those at the ventral-transverse corner are
+black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
+axons on the tectum. The axons retain the colour of their originating cell on
+the retina, as shown in Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged
+caudally; dorsal axons laterally; axons should be evenly spaced and the
+pattern should span the entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net'
+plots of the axon centroids at three timepoints. An axon's centroid is the
+average location of its 4 branches. In these plots, lines are drawn between
+axons that are expected to arrange themselves adjacently to one another by the
+end of the simulation. The number of line crossings in this plot gives an
+indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the
+positions of the individual branches (i.e.~growth cones) at $t=300$ that
+contribute to the the plot in Fig.\,\ref{f:ch}D. By this time, the topological
+order matches that shown in Fig.\,\ref{f:ch}E. Opposing gradients have sorted
+the axons along the two axes to form a grid of cells. However, the arrangement
+does not span the full tectum. Fig.\,\ref{f:ch}F shows two pattern
+metrics. The first metric, $\epsilon$, is the sum of squared distance errors
+of the axon centroids, with $\mathbf{x}_{k}$, the location of the centroid of
+the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for
+axon $k$:
 %
@@
 
+\section{Discussion}
+Discussion goes here.
+
+\color{grey}
 \section{ZEROTH DRAFT/NOTES}

### 2021-11-22 paper/paper.tex
Got to a point where I need to create some more figures

@@
 
-\section{Model}
-
-Our model was inspired by the agent-based model of \cite{simpson_simple_2011}
-and, in common with that work, consists of agents representing axonal growth
-cones, originating from a square retina and moving upon a square tectum of
-side length 1 in arbitrary units.
-%
-\citet{simpson_simple_2011} model three contributions to the movement of the growth
-cones; chemoaffinity, competition for space and axon-axon interactions. The
-first two of these are modelled without reference to a particular mechanism;
-each agent is provided with globally-supervised information about its
-instantaneous location and its target location, allowing it to move along a
-vector towards the target. Competition is implemented as a movement of
-constant distance along the vector between two agents if those two agents are
-within a threshold distance of one another. Interaction is signal-induced
-repulsion, which is triggered if two agents are both within the threshold
-distance of each other and have compatible EphA receptor expression.
-
-We sought to replace the globally-supervised mechanisms by local receptor and
-ligand based interactions. First, chemoaffinity was remodelled so that all
-interactions occurred via signal transmission between the receptors expressed
-on the growth cones of retinal ganglion cell axons and the ligands expressed
-across the tectum. Signals induced either an attractive or repulsive
-interaction to cause growth cones to ascend or descend the gradient of ligand
-expression on the tectum.
-
-We then explored the behaviour of models incorporating this chemoaffinity
-mechanism alone and in conjunction with several competition mechanisms. We
-found parameters for the best models of each type (chemo-only [G],
-chemo-competition [GC], chemo-receptor-receptor [GI] and chemo-receptor-ligand
-[GJ]) by an optimisation approach. We then compared each model's ability to
-reproduce the surgical and genetic manipulation results in a manner similar to
-that taken by \citet{simpson_simple_2011} to find the simplest model that
-would fit the data.
-
-\subsection*{Receptor and ligand expression}
-
-Each retinal ganglion cell projects $n$ growth cone agents (referred to
-as \emph{branches}) which carry a set of 4 receptors, $r_i$, and 4 ligands,
-$l_i$, indexed by $i$ and expressed at levels determined by the cell's
-location on the retina. The expression levels of the receptors and ligands
-vary with respect to the cell's position on the retinal surface. Each receptor
-or ligand varies only with respect to one dimension.  Receptor expression
-gradients are arranged in orthogonal pairs with the gradient of $r_0$ being
-orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, is
-orthogonal to $r_3$. Ligands are also arranged in orthogonal pairs, with $l_0$
-opposing $r_0$ (as $r_0$ increases, $l_0$ decreases) and $l_1$ orthogonal to
-$l_0$ and opposing $r_1$.
-
-Because there is convincing evidence that EphA and EphB receptors are
-expressed in exponentially increasing
-patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004},
-we use an exponential form for retinal receptor expressions, in common with
-other modelling
-studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression
-as \citet{simpson_simple_2011} which is
-\begin{equation} \label{e:retrcpt}
-r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
-\end{equation}
-
-The simulated tectum expresses the same types of receptors and ligands, also
-in orthogonal pairs of gradients. In this work, we did not consider the effect
-of reverse signalling from tectal receptors to RGC ligands so we only modelled
-tectal ligand expression, $L_i$. Although several studies model tectal ligand
-expression with exponential functions \citep{koulakov_stochastic_2004}, the
-experimental evidence for ligand expression is more ambiguous (Supp. Fig
-\textbf{X}). Correspondingly, we leave open the possibility that tectal ligand
-expression may be modelled by exponential, linear or logarithmic functions:
-\begin{equation} \label{e:tecligexp}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
-\end{equation}
-or
-\begin{equation}
-L_i^{\text{lin}}(x,y) = 1.31 + 2.33 x;
-\end{equation}
-or
-\begin{equation}
-L_i^{\text{log}}(x,y) = 2.32 + 1.29 \log (2.3(x+0.2))
-\end{equation}
-Each function has approximately the same expression values at $x=0$ and
-$x=1$. Opposing functions are obtained by substituting $(1-x)$ for $x$.
-
-While we do not explicitly name $r_0$, $r_1$,
-etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA,
-EphB, Ryk \citep{schmitt_wntryk_2006} and
-Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $l$ and $L$
-include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and
-RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
-role in retinotectal map formation.
-
-Figure~\ref{f:ex} shows colour maps of receptor and ligand expression, in this
-case tectal ligand expression maps are linear. Note that the expression is
-discretized, and in simulations, the gradient is computed numerically from the
-discretized expression. This makes it straightforward to perform tectal and
-retinal `graft' manipulations. Coordinates within the simulated retina and
-tectum are normalised, with 1 corresponding to 2 mm on the tectum and on the
-retina for mouse \citep{reber_relative_2004}. This means that growth cones of
-approximate radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a
-radius $r$ = 0.0025 in the arbitrary units of the simulation. For a realistic
-simulation based on mouse, the approximate RGC density in the retina is around
-44000 cells per retina \citep{jeon_major_1998}. This sets the side length of
-the retina to $\sqrt{44000} \approx 200$. This results in a simulation of
-prodigious computational size, which although tractable (10 s per step,
-requiring several hours, rather than days to run) is not convenient. To
-maintain the relationship between growth cone size and cell density, we permit
-a smaller number of growth cones ($400\times8=3200$) but keep the fractional
-area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
-per axon) or 6.9 (if there are 8 cones per axon).
-
-\subsection*{Movement}
-
-At each timestep in a simulation, the position, $\mathbf{x}_{b,t}$, at time
-$t$ of each branch $b$ is updated according to
-%
-\begin{equation}
-\mathbf{x}_{b,t+1} = \mathbf{x}_{b,t} + \mathbf{M}_{b}
-\end{equation}
-%
-where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
-effect, $\mathbf{G}_b$, axon-axon competition, $\mathbf{C}_b$, an axon-axon
-receptor-receptor interaction, $\mathbf{I}_b$, and axon-axon receptor-ligand
-interaction, $\mathbf{J}_b$, and a `border effect',
-$\mathbf{B}_b$, which acts to retain branches within the tectal region:
-%
-\begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + m_2 \mathbf{C}_b + m_3 \mathbf{I}_b +
-m_4 \mathbf{J}_b + m_5 \mathbf{B}_b
-\end{equation}
-%
-evaluated at time $t$. $m_1$--$m_5$ are scalar parameters. We typically
-considered models for which some of $m_2$ to $m_4$ were fixed to zero. A model
-for which $m_3$ and $m_4$ were zero was named a `$\mathbf{GC}$' model as it
-combined the chemoaffinity effect and space-based competition; a model for
-which $m_2$ was zero was a `$\mathbf{GIJ}$' model and so on.
-
-\subsection*{Chemoaffinity}
-
-In this model, the signal transmitted when a ligand binds to a receptor on a
-branch (forward signalling) can lead to one of two effects; the signal may
-induce an attraction towards the ligand-expressing region or a repulsion away
-from it.
-%
-For axon-tectum interactions, we implemented attraction as a climbing of the
-tectal ligand expression gradient and repulsion as gradient descent.
-%
-To determine the strength of the effect we assumed a purely linear receptor
-binding model, and set the chemotactic movement vector of the branch $b$ at
-location $\mathbf{x}_b$ on the tectum to be
-%
-\begin{equation}
-\mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor
-pair $i$, $\nabla L_i$ is the gradient of expression of ligand $i$ on the
-tectum and $F_i$ denotes the direction of the interaction induced when a
-molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
--1 for a repulsive interaction or 1 for an attractive interaction.
-
-\subsection*{Axon-axon interactions}
-
-We investigated three forms of competition that could be combined with the
-chemoaffinity effect. We investigated space-based competition
-(`$\mathbf{C}$'), as described in \citet{simpson_simple_2011}, a
-receptor-receptor signalling mechanism (`$\mathbf{I}$'), similar to that
-described in \citet{simpson_simple_2011}, and a receptor-ligand mass-action
-signalling model (`$\mathbf{J}$'). Our aim was to determine whether either of
-the receptor/ligand based signalling mechanisms could replace the space-based
-competition.
-
-\subsection*{Space based (or self-other) competition}
-
-Space based competition is a repulsive effect which acts between any two
-branches, whether they originate from different retinal ganglion cells or from
-the same cell. It assumes that there is some kind of self-other signal which
-acts to discourage two cells from approaching each other.
-%
-\begin{equation}
-\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
-\end{equation}
-where the distance based weight $W$ is given by
-\begin{equation}
-W_c(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r_c}   & d_{kb} \leq 2r_c \\
-     0 & d_{kb} > 2r_c
-     \end{cases}
-\end{equation}
-for two growth cones of interaction radius $r_c$ separated by a distance
-$d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
-branch $b$.
-
-\subsection*{Receptor-receptor axon-axon interactions}
-
-We modelled interactions between axon growth cones due to signalling between
-receptors of the same type. For these interactions, repulsion caused a
-movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
-branch $k$ to branch $b$ causing them to move further apart; attraction caused
-the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
-is given by
-%
-\begin{equation}
-\mathbf{I}_b
-= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_I \qquad \mathrm{if}~r_{i,b}
-/ r_{i,k} < 1.1
-\end{equation}
-%
-the weight here is defined as a function of $r_I$:
-%
-\begin{equation}
-W_I(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r_I}   & d_{kb} \leq 2r_I \\
-     0 & d_{kb} > 2r_I
-     \end{cases}
-\end{equation}
-$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
-
-\subsection*{Receptor-ligand axon-axon interactions}
-
-%
-\begin{equation}
-% No \frac{1}{|B_b|} in this one. Really?
-\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
-\end{equation}
-%
-where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
-receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
-given by
-%
-\begin{equation}
-Q_J(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
-     0 & d_{kb} > 2r_j
-     \end{cases}
-\end{equation}
-%
-The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
-the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
-expression of ligand $i$ on branch $k$.
-
-\subsection*{Border effect}
-
-We implemented a border effect based on gradient following by assuming that
-there is some other molecular signal which acts on all branches near the
-boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}_b$ is
-given by:
-%
-\begin{equation}
-B_{b,x} = \begin{cases}
-        r-x      & x<r \\
-        1-r-x    & x>1-r
-\end{cases}
-B_{b,y} = \begin{cases}
-        r-y      & y<r \\
-        1-r-y    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}_b$ is equivalent to the action of a repulsive signalling
-molecule expressed around the border of the tectum, whose expression increases
-quadratically outside the tectum and affects any branch touching (or outside)
-the boundary
-
-\subsection*{Initial conditions}
-
-Branches were randomly distributed in a stripe at the rostral side of the
-tectum at the start of each simulation. Each RGC axon was assigned a random
-initial position coordinate:
-\begin{equation}
-\mathbf{x}_{\text{axon},t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the
-range $[p,q)$. Each of the $n$ branches per RGC axon was given its parent
-axon's initial position, plus a randomly generated offset with coordinates
-derived from a normal distribution of mean 0 and standard deviation 0.1.
-
-\subsection*{Assumptions}
-
-% All the extra bits of info required to specify the simulation
-To enable a simulation of the model described above it is necessary to make
-some assumptions about the nature of the receptor-ligand signalled
-interactions (i.e. the values of $F_i$ that determine whether interactions are
-attractive or repulsive) and about the pattern of ligand expression on the
-tectal surface. Unless stated otherwise, we assumed that: \textbf{i)} all
-receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as
-for EphA-ephrinA
-coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996};
-and \textbf{ii)} all tectal ligand expressions vary exponentially along a
-given axis, following Eq.\,\ref{e:tecligexp}.
-
-In all simulations, the border effect, $B_b$, had the same magnitude with
-$m_5=0.5$. The number of branches per axon was 4 unless otherwise stated.
-
 \color{black}
@@
 The first question we tried to answer was whether a gradient following
-mechansim, acting alone, is sufficient to pattern the tectum. To do so, we
+mechanism, acting alone, is sufficient to pattern the tectum. To do so, we
 simulated the retina and tectum as unit square regions and defined retinal
@@
 %
-\begin{equation}
-\mathbf{M}_{b} = m_1 \mathbf{G}_b + 0.5 \mathbf{B}_b
+\begin{equation} \label{e:mv}
+\mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + 0.5 \mathbf{B}_b
 \end{equation}
@@
 % G - chemoaffinity description
+%\subsection{Chemoaffinity}
 The chemoaffinity component, $\mathbf{G}_b$, is governed by receptor-ligand
@@
 %
-$m_1$ is a
+$m_{\!_G}$ is a
 scalar parameter which controls how much movement is generated for a given
@@
 
-% border effect description
+% \subsection{Border effect} % border effect description
 The border effect is also based on gradient following by assuming that
@@
 
-% Initial conditions
+% \subsection*{Initial conditions}
 Branches were initially randomly distributed in a stripe at the rostral side
@@
 branches, indicating that the pattern is partial. A similar result is seen for
-the 180\degree~example in row B, in whcih the caudal half of the rotated patch
+the 180\degree~example in row B, in which the caudal half of the rotated patch
 is free of branches, but the rostral half contains branches with the correctly
@@
 systems including the murine cortical whisker
-barrels\citep{james_modelling_2020} \textbf{(FIXME owt else?)}.
+barrels \citep{james_modelling_2020} \textbf{(FIXME owt else?)}.
 
@@
 and a receptor-ligand mass-action signalling model
-(`$\mathbf{J}$'). \citet{simpson_simple_2011} found that both a space based
-competition and a receptor-receptor signalling mechanism were required to
-reproduce the full range of surgical and genetic manipulations explored in
-that work. The first of these, space-based competition, offers no mechanism by
-which axon growth cones are able to communicate and the second mechanism
-explored by \citet{simpson_simple_2011} was implemented only for a single axis
-in the model (EphA receptors, expressed in a gradient along the XX retinal
-axis). We wondered whether a single signalling mechanism could provide a
-suitable competitive, repulsive effect. To investigate, we compared both of
-the \citet{simpson_simple_2011} effects along with a third possible signalling
-mechanism involving the same receptor-ligand interactions employed by the
-gradient following mechansim, because the ligands expressed on the tectum are
-also expressed by the retinal ganglion cells.
-
-\subsection*{Space based (or self-other) competition}
+(`$\mathbf{J}$'). The movement vector (Eq.\,\ref{e:mv}) becomes
+%
+\begin{equation} \label{e:mv2}
+\mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b +  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+\end{equation}
+%
+
+Although two of these competitive mechanisms were investigated
+by \citet{simpson_simple_2011}, the first, space-based competition, offers no
+mechanism by which axon growth cones are able to communicate and the second
+mechanism was implemented only for a single axis in the model (EphA receptors,
+expressed in a gradient along the XX retinal axis). We wondered whether a
+\emph{single} signalling mechanism working alongside the `untuned'
+chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
+successfully reproduce the wildtype and manipulation results. To investigate,
+we studied each competitive mechanism alongside chemoaffinity in turn, setting
+$m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model, $m_c=m_{\!_J}=0$ to
+give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a `$\mathbf{GJ}$'
+model.
+
+\subsection*{Chemoaffinity plus space-based competition (GC)}
 
@@
 
-\subsection*{Receptor-receptor axon-axon interactions}
+The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
+\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} with $m_{\!_I}=m_{\!_J}=0$ has three
+free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
+these parameters, we reasoned that evolution would optimize the arrangement of
+the wildtype pattern. We performed a simulated annealing optimization to
+minimize the error metric, $\epsilon$, for the wildtype pattern
+(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were $m_c$
+= 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
+0.39918. Fig.\,\ref{f:nonexistent_yet} shows a simulation with these parameters.
+
+\subsection*{Receptor-receptor axon-axon interactions (GI)}
 
@@
 \mathbf{I}_b
-= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_I \qquad \mathrm{if}~r_{i,b}
+= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
 / r_{i,k} < 1.1
@@
 %
-the weight here is defined as a function of $r_I$:
-%
-\begin{equation}
-W_I(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r_I}   & d_{kb} \leq 2r_I \\
-     0 & d_{kb} > 2r_I
+the weight here is defined as a function of $r_{\!_I}$:
+%
+\begin{equation}
+W_{\!_I}(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
+     0 & d_{kb} > 2r_{\!_I}
      \end{cases}
@@
 
-\subsection*{Receptor-ligand axon-axon interactions}
+\subsection*{Receptor-ligand axon-axon interactions (GJ)}
 
@@
 % No \frac{1}{|B_b|} in this one. Really?
-\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_J(d_{kb})
-\end{equation}
-%
-where $Q_J(d_{kb})$ is the signalling strength between two growth cones of
-receptor-ligand interaction radius $r_j$ a distance $d_{kb}$ from one another,
+\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
+\end{equation}
+%
+where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
+receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
 given by
@@
 \begin{equation}
-Q_J(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_j \\
-     0 & d_{kb} > 2r_j
+Q_{\!_J}(d_{kb}) = \begin{cases}
+     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
+     0 & d_{kb} > 2r_{\!_J}
      \end{cases}
@@
 Discussion goes here.
-
-\color{grey}
-\section{ZEROTH DRAFT/NOTES}
-
-\subsection*{Normal development}
-
-%
-Fig.\,\ref{f:wt}Ai shows a dual-colour map of retinal ganglion cell positions
-on the square retina, with red encoding position on the $x$ axis (temporal to
-nasal) and green encoding position on the $y$ axis (ventral to
-dorsal). Fig.\,\ref{f:wt}Aii presents a fishnet plot of the expected location on
-the tectal surface of each RGC at the end of the development process. Thus,
-the `red' RGC from the ventro-nasal corner should map to the caudo-medial
-corner of the tectum and the `green' RGC from the dorso-temporal corner should
-find its way to the rostro-lateral corner on the tectum. The lines between the
-dots of Fig.\,\ref{f:wt}Aii indicate the adjacency relationships between
-cells. Thus, colour in each panel indicates a branch or axon's location of
-origin on the retina.
-%
-Fig.\,\ref{f:wt}B shows the initial state of the system, with branches
-stochastically arranged around the rostral edge of the tectum. Each dot
-represents the centroid of $n=8$ branches per RGC axon. Figs.\,\ref{f:wt}C and \ref{f:wt}D
-show the development of the system with time, with Fig.\,\ref{f:wt}D showing a
-well ordered arrangement, with a majority of adjacency relationships matching
-the experimental prediction (i.e. there are few crossed lines). The
-intermediate timepoint at $t=40$ indicates that branches quickly cover the simulated
-tectal surface, with final adjacency relationships developing over a longer
-timescale. This dynamic behaviour is also evident in Fig.\,\ref{f:wt}E, which
-shows \textbf{RMS error or some metric} of the axon centroids with respect to
-the experimental prediction. Fig.\,\ref{f:wt}F shows the final
-location of each branch in the simulation, which gives a more continuous view
-of the order present in Fig.\,\ref{f:wt}D. Finally, Fig.\,\ref{f:wt}G shows the
-branches and their centroid path history for five selected RGCs, with path
-histories strongly resembling those presented by \citet{simpson_simple_2011}.
-
-This result shows that a gradient-following model based on receptor-ligand
-interactions (both for axon-to-tectum and axon-to-axon interactions) is
-able to generate the basic mapping observed in the retinotectal projection.
-
-\subsection*{Manipulations}
-
-To determine if the model can explain the many surgical and genetic
-manipulations described in the literature
-\citep*[for a review, see][]{goodhill_retinotectal_1999} we performed
-simulations matching those in \citet{simpson_simple_2011} in which we either
-rotated or translocated grafts of the tectum; ablated portions of the retina
-and/or the tectum; or manipulated the receptor or ligand expressions in the
-model.
-
-\subsubsection*{Surgical Manipulations}
-
-Fig.\,\ref{f:surg1} shows the result of simulating three surgical
-manipulations alongside the wildtype result, which is shown on Row A. Rows B
-and C show graft rotation simulations (90$\degree$ and 180$\degree$). In each
-case an 8x8 square on the tectum (each unit is 0.05x0.05) is rotated. The
-rotational sense can be observed clearly in the `Branches' column; for the
-90$\degree$ rotation, the top-right (caudal-lateral) corner of the rotated
-square is red, whereas the top-left corner of the wildtype tectum is red. \textbf{etc}.
-
-%\includegraphics[width=\linewidth]{./images/fig_manipulations.png}
-
-\begin{figure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_wt_rotcmp.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecrot90.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecrot180.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecswap.png}
-\end{subfigure}
-\caption{Surgical manipulations. Row A: wildtype. Row B: 90 degree
-rotation. Row C: 180 degree rotation. Row D: Graft swap.}
-\label{f:surg1}
-\end{figure}
-
-Fig.\,\ref{f:surg2} shows retinal ablation, tectal ablation, mismatch and
-compound eye experiments (Stuart: see Simpson and Goodhill Fig. 4).
-
-\begin{figure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_retablate.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_tecablate.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_mismatch.png}
-\end{subfigure}
-\begin{subfigure}{\linewidth}
-\includegraphics[width=\linewidth]{./images/j4_eE_GJ_compound.png}
-\end{subfigure}
-\caption{Surgical manipulations 2. Row A: Retinal ablation, Row B: Tectal
-ablation, Row C: Mismatch (ablate retina and also the tectum that matches the
-left-behind retina), Row D: Compound eye (Two nasal retinal halves `glued
-together')}
-\label{f:surg2}
-\end{figure}
-
-\subsubsection*{Genetic manipulations}
-
-Brown et al. EphA knock-in.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_brown.png}
-\caption{Brown et al simulation. This is EphA3 knockin - it adds a constant to
-half of the RGC's $r_0$ (randomly selected). Compare with S\&G Fig.\,6.}
-\label{f:brown}
-\end{figure}
-
-Describe Reber and other gene knock-in and knock-out experiments.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_reber.png}
-\caption{Reber et al simulation. Follow on to Brown et al (same
-group). Combines the selective EphA3 knockin with a uniform EphA4 knockdown.}
-\label{f:reber}
-\end{figure}
-
-\subsection*{ligand expression profiles}
-
-A section to show the effect of assuming linear or logarithmic expression
-profiles. Especially that linear expression fails to reproduce the rotation
-and graft swap manipulations. Hmm, actually, it looks ok?!
-
-\subsection*{Attractive receptor-ligand interactions}
-
-Show that, if you switch one of the receptor-ligand  interactions in the
-system to be attractive instead of repulsive, then you have to make the ligand
-expression form logarithmic instead of exponential.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_ephb_wt.png}
-\caption{The effect of making one interaction attractive. This models
-switching one receptor/ligand pair to be an analogue of EphB, but retains an
-assumption that the ephrin-B ligand has exponential form. The change compared
-with the model of Fig.\,\ref{f:wt} is that the tectal ligand direction of
-$L_1$ is reversed and the interaction of $r_1$ and $L_1$/$l_1$ is made
-attractive. $L_1$ retains the assumption of exponential change. Note the
-`bunching up' of the field to the medial side of the tectal sheet.}
-\label{f:ee_oneattractive}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eel_ephb_wt.png}
-\caption{The arrangement is restored by making the expression of $L_1$ on the
-tectal sheet change logarithmically.}
-\label{f:eel_oneattractive}
-\end{figure}
-
-\subsection*{Grating simulations}
-
-I'd like to do this if it's not too much work!
-
-\section{Discussion}
-
-The model described above assumes that branches express graded levels of
-receptors which interact with ligands expressed as gradients on the tectum,
-the tectal border and on the surface of other branches, and that only these
-interactions guide their movement.
-
-\subsection*{Results Observations/Notes}
-
-J has a repulsive effect, which is generally disordering. If competition has
-`a lot of work to do' (i.e. in the m\_ee... models) then a strong J will push
-the axons across the whole tissue, but will upset the local ordering
-relationships. Perhaps J also needs a self-self interaction (i.e. for members
-of the same axon?).
-
-The I interaction is absent when the topographic ordering is good. This means
-that it does not fix the retablate condition. Like the other competition
-effects, it is strongly affected by the radius of interaction.
 

### 2021-11-23 paper/paper.tex
A few tweaks to paper text before generating some more figures

@@
 
-\subsection*{Competition}
+\subsection*{Combining chemoaffinity with competition}
 
@@
 chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
-successfully reproduce the wildtype and manipulation results. To investigate,
-we studied each competitive mechanism alongside chemoaffinity in turn, setting
-$m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model, $m_c=m_{\!_J}=0$ to
-give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a `$\mathbf{GJ}$'
-model.
+successfully reproduce the wildtype and manipulation results. Alongside the
+same chemoaffinity mechanism, we studied each competitive mechanism in turn,
+setting $m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model,
+$m_c=m_{\!_J}=0$ to give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a
+`$\mathbf{GJ}$' model.
 
@@
 
-Space based competition is a repulsive effect which acts between any two
-branches, whether they originate from different retinal ganglion cells or from
-the same cell. It assumes that there is some kind of self-other signal which
-acts to discourage two cells from approaching each other.
+Space based competition was implemented as a repulsive effect acting between
+any two branches, whether they originated from different retinal ganglion
+cells or from the same cell. An assumption is that there is some kind of
+self-other signal which acts to discourage two cells from approaching each
+other. A competition movement vector, $\mathbf{C}_b$, acting on branch $b$,
+which is a proximity-weighted sum of the repulsion due to all nearby branches
+is defined as:
 %
 \begin{equation}
-\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W
+\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W_c
 \end{equation}
-where the distance-dependent weight $W_c$ is given by
+where $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to branch
+$b$. The distance-dependent weight $W_c$ is given by
 \begin{equation}
@@
 minimize the error metric, $\epsilon$, for the wildtype pattern
-(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were $m_c$
-= 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
-0.39918. Fig.\,\ref{f:nonexistent_yet} shows a simulation with these parameters.
+(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were
+$m_c$ = 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
+0.39918. Fig.\,\ref{f:nonexistent_yet} shows the $\mathbf{GC}$ simulation with
+these parameters.
 

### 2021-11-23 paper/paper.tex
Show GC model at t=1000
WIP before switching to PC
WIP

@@
 %
+% (see scripts/fig_G_surgical.sh)
 \begin{figure}
@@
 
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GC_best_1.json configs/simpler/e_wt_figcomp1.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_wt_figcomp1.png}
+\caption{$\mathbf{GC}$ model}
+\label{f:GC}
+\end{figure}
+
 The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
@@
 (Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were
-$m_c$ = 0.09959, $m_{\!_G}$ = 0.003841 and $r_c$ =
-0.39918. Fig.\,\ref{f:nonexistent_yet} shows the $\mathbf{GC}$ simulation with
-these parameters.
+$m_{\!_G}$ = 0.003841, $m_c$ = 0.09959 and $r_c$ = 0.39918. Fig.\,\ref{f:GC}
+shows the $\mathbf{GC}$ simulation with these parameters.
 
@@
 
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2.png}
+\caption{$\mathbf{GI}$ model}
+\label{f:GI}
+\end{figure}
+
 \subsection*{Receptor-ligand axon-axon interactions (GJ)}
@@
 expression of ligand $i$ on branch $k$.
+
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3.png}
+\caption{$\mathbf{GJ}$ model}
+\label{f:GJ}
+\end{figure}
 

### 2021-11-23 paper/paper.tex
And another layout, replacing the selected axons from the NT vs RC graph

@@
 region. Instead, axons from the transverse retina arrange with normal spacing
-and axons from the nasal retina are pushd to the caudal-most side of the
+and axons from the nasal retina are pushed to the caudal-most side of the
 remaining tectum (Fig.\,\ref{f:chsurg}E). A similar result occurs in the

### 2021-11-23 sim/agent/agent1.h

@@ -248,6 +248,24 @@ struct Agent1
                 this->gv->setdata (nt_m, rc_m, ss.str());
                 this->gv->reinit();
                 this->v->render();
+
+            } else if (this->layout == graph_layout::h) {
+                // Graph layout h shows the nt vs rc (for the mid band of the retina/central column of the tectum?).
+                morph::vVector<T>nt;
+                morph::vVector<T>rc;
+                morph::vVector<T>rc_expt;
+                size_t ii = 0;
+                for (ii = 0; ii < this->ret->posn.size(); ++ii) {
+                    nt.push_back(1-this->ret->posn[ii][0]);
+                    rc.push_back(this->ax_centroids.p[ii][1]);
+                    rc_expt.push_back(this->ax_centroids.targ[ii][1]);
+                }
+                // Also show expt. version.
+
+                this->gv->setdata (nt, rc_expt, "Expt");
+                this->gv->setdata (nt, rc, "Sim");
+                this->gv->reinit();
+                this->v->render();
             }
 
             // Save final image based on config file names
@@ -306,7 +324,7 @@ struct Agent1
         }
         this->cv->reinit(); // Centroids to end
 
-        if (this->layout != graph_layout::e && this->layout != graph_layout::f) {
+        if (this->layout != graph_layout::e && this->layout != graph_layout::f && this->layout != graph_layout::h) {
             if (this->layout != graph_layout::b) {
                 this->tcv->reinit(); // Experiment
             }
@@ -1302,14 +1320,14 @@ struct Agent1
         unsigned int wdefault = 1200;
         if (this->layout == graph_layout::a || this->layout == graph_layout::d) { wdefault = 1920; }
         if (this->layout == graph_layout::b || this->layout == graph_layout::e
-            || this->layout == graph_layout::g) { wdefault = 2480; }
+            || this->layout == graph_layout::g || this->layout == graph_layout::h) { wdefault = 2480; }
         if (this->layout == graph_layout::c) { wdefault = 2200; }
         if (this->layout == graph_layout::f) { wdefault = 675; }
         const unsigned int ww = this->conf->getUInt ("win_width", wdefault);
         unsigned int hdefault = 800;
[…]

### 2021-11-24 sim/agent/agent1.h
Adds dashed borders to figures

@@ -99,7 +99,7 @@ std::ostream& operator<< (std::ostream& os, const AgentMetrics<T>& am)
 }
 
 // A selection of possible graph layouts to show when running the program
-enum class graph_layout { a, b, c, d, e, f, g, h };
+enum class graph_layout { a, b, c, d, e, f, g, h, i };
 
 // Agent1 coordinates an agent based simulation of axon branches of type B. This class
 // also incorporates the visualisation code that displays the state of the simulation as
@@ -146,14 +146,12 @@ struct Agent1
                 // Get layout from config file. Default to 0 or 'a'
                 this->layout = (graph_layout)this->conf->getUInt ("graph_layout", 0);
 
-                if (this->layout == graph_layout::c) {
-                    // Set up (from config file if necssary) the times at which the various
-                    // graphs will be frozen (i.e. no longer updated from the sim)
-                    this->freeze_times[0] = 0;
-                    this->freeze_times[1] = this->conf->getUInt ("freeze_time1", 20);
-                    this->freeze_times[2] = this->conf->getUInt ("freeze_time2", 80);
-                    this->freeze_times[3] = this->conf->getUInt ("freeze_time3", 120);
-                }
+                // Set up (from config file if necssary) the times at which the various
+                // graphs will be frozen (i.e. no longer updated from the sim)
+                this->freeze_times[0] = 0;
+                this->freeze_times[1] = this->conf->getUInt ("freeze_time1", 20);
+                this->freeze_times[2] = this->conf->getUInt ("freeze_time2", 80);
+                this->freeze_times[3] = this->conf->getUInt ("freeze_time3", 120);
 
                 // How early to start showing the crossings metric?
                 this->crosscount_from = this->conf->getUInt ("crosscount_from", 1000);
@@ -389,6 +387,17 @@ struct Agent1
             // Show RMS error
             AgentMetrics<T> am = this->get_metrics();
             this->sos_txt->setupText (std::to_string(am.rms[0]));
+            break;
+        }
+        case graph_layout::i:
+        {
+            if (stepnum <= this->freeze_times[1]) { this->cv1->reinit(); } // Time freeze branches
+            this->cv->reinit();  // Centroids
[…]

### 2021-11-24 sim/agent/netvisual.h

@@ -163,6 +163,7 @@ public:
         this->drawBoundary (idx);
     }
 
+    // The tissue boundary, indicated by dashed lines
     void drawBoundary (VBOint& idx)
     {
         std::array<float, 3> gry = { 0.2, 0.2, 0.2 };
@@ -172,14 +173,29 @@ public:
             h *= (this->locations->domain_h-1) * this->locations->dx[1];
         }
         float _z = puckthick[2]*float{0.5001}; // Ensure boundary is visible above rest of drawing
-        this->computeLine (idx, morph::Vector<Flt, 3>({0, 0, _z}), morph::Vector<Flt, 3>({w, 0, _z}),
-                           this->uz, gry, gry, this->linewidth*zoom, this->linewidth/4*zoom);
-        this->computeLine (idx, morph::Vector<Flt, 3>({w, 0, _z}), morph::Vector<Flt, 3>({w, h, _z}),
-                           this->uz, gry, gry, this->linewidth*zoom, this->linewidth/4*zoom);
-        this->computeLine (idx, morph::Vector<Flt, 3>({w, h, _z}), morph::Vector<Flt, 3>({0, h, _z}),
-                           this->uz, gry, gry, this->linewidth*zoom, this->linewidth/4*zoom);
-        this->computeLine (idx, morph::Vector<Flt, 3>({0, h, _z}), morph::Vector<Flt, 3>({0, 0, _z}),
-                           this->uz, gry, gry, this->linewidth*zoom, this->linewidth/4*zoom);
+
+        this->computeFlatDashedLine (idx,
+                                     morph::Vector<Flt, 3>({0, 0, _z}),
+                                     morph::Vector<Flt, 3>({w, 0, _z}),
+                                     this->uz,
+                                     gry,
+                                     this->linewidth*zoom, 0.0f,
+                                     this->linewidth*5.0f, 0.4f);
+        this->computeFlatDashedLine (idx, morph::Vector<Flt, 3>({w, 0, _z}), morph::Vector<Flt, 3>({w, h, _z}),
+                                     this->uz,
+                                     gry,
+                                     this->linewidth*zoom, 0.0f,
+                                     this->linewidth*5.0f, 0.4f);
+        this->computeFlatDashedLine (idx, morph::Vector<Flt, 3>({w, h, _z}), morph::Vector<Flt, 3>({0, h, _z}),
+                                     this->uz,
+                                     gry,
+                                     this->linewidth*zoom, 0.0f,
[…]

### 2022-01-31 paper/paper.tex
Adds dashed borders to figures
Incorporate new figures
Work in progress

@@
 scalar parameter which controls how much movement is generated for a given
-receptor-ligand signal size.
+receptor-ligand signal.
 % end chemoaffinity description
@@
 \includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2.png}
-\caption{The pure gradient-following model.}
+\caption{The pure gradient-following model. \textbf{A} The origin position
+on the tectum gives axons a colour according to a dual-colour map. Axons in
+panels B-H that are red are those which originated in the most naso-ventral
+corner of the square simulated retina. \textbf{B}-\textbf{D} Axon centroid
+positions on the simulated tectum (dashed line) at time 0, 40 and 300
+units. Lines show adjacency relationships between the axons' origin positions
+on the retina. \textbf{E} Layout of axon centroids on the tectum observed in
+experimental results. \textbf{F} The sum-of-squared difference error,
+$\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon
+arrangement plotted with respect to simulation time. \textbf{G} The
+arrangement of axon branches at t=300. \textbf{H} Selected axon locations at
+t=300, with path histories in grey.  \emph{Abbreviations: D dorsal, V ventral, T
+temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
 \label{f:ch}
@@
 
-The result at $t=300$ is shown in Figs.\,\ref{f:chalt}A and \ref{f:chalt}B. The
-axon centroids are arranged across the full tectal surface.  The arrangement
-is slightly disordered around the middle of the tectum. This is caused by
-numerical artefacts from the discretization of the ligand expression and the
-fact that in the centre, the difference between the opposing receptor-ligand
-signals is very small. Nevertheless, the result demonstrates that a modified
-form for the tectal ligand expression (which was found easily by hand tuning)
-allows a well ordered map to form for the wildtype system. Fig
-.\,\ref{f:chalt}C shows the error metric (\textbf{FIXIMAGE}), $\epsilon$,
-which is much closer to 0 with the new ligand expression.
+The result at $t=300$ is shown in Figs.\,\ref{f:chalt}A
+and \ref{f:chalt}B. The axon centroids are arranged across the full tectal
+surface.  The arrangement is slightly disordered around the middle of the
+tectum. This is caused by numerical artefacts from the discretization of the
+ligand expression and the fact that in the centre, the difference between the
+opposing receptor-ligand signals is very small. Nevertheless, the result
+demonstrates that a modified form for the tectal ligand expression (which was
+found easily by hand tuning) allows a well ordered map to form for the
+wildtype system. Fig.\,\ref{f:chalt}C shows the error metric
+(\textbf{FIXIMAGE}), $\epsilon$, which is much closer to 0 with the new ligand
+expression.
 
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig2.json
+% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_wt_fig3.json
 \begin{figure}
@@
 mismatch experiment (Fig.\,\ref{f:chsurg}F), but in this case there are no
-axons from teh transverse retina, which was been ablated and only the axons
+axons from the transverse retina, which was been ablated and only the axons
 from the nasal retina are present, arranged in a line.
@@
 \caption{Surgical manipulations in the chemoaffinity-only model. \textbf{A} 90 degree
-rotation. textbf{B} 180 degree rotation. \textbf{C} Graft swap. \textbf{D}
+rotation. \textbf{B} 180 degree rotation. \textbf{C} Graft swap. \textbf{D}
 Retinal ablation \textbf{E} Tectal ablation \textbf{F} mismatch ablation}
@@
 Although two of these competitive mechanisms were investigated
-by \citet{simpson_simple_2011}, the first, space-based competition, offers no
+by \citet{simpson_simple_2011}, the first (space-based competition) offers no
 mechanism by which axon growth cones are able to communicate and the second
@@
 self-other signal which acts to discourage two cells from approaching each
-other. A competition movement vector, $\mathbf{C}_b$, acting on branch $b$,
-which is a proximity-weighted sum of the repulsion due to all nearby branches
-is defined as:
+other. A competition movement vector, $\mathbf{C}_b$, acting on branch $b$, is
+defined as a proximity-weighted sum of the repulsion due to all nearby branches:
 %
@@
 $d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
-branch $b$.
+branch $b$. This description of the competition matches that
+in \citet{simpson_simple_2011}.
 
@@
 shows the $\mathbf{GC}$ simulation with these parameters.
+
+% ./scripts/fig_GC_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GC_surgical.png}
+\caption{$\mathbf{GC}$ model, surgical manipulations}
+\label{f:GC}
+\end{figure}
 
@@
 
+% ./scripts/fig_GI_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GI_surgical.png}
+\caption{$\mathbf{GI}$ model, surgical manipulations}
+\label{f:GI}
+\end{figure}
+
 \subsection*{Receptor-ligand axon-axon interactions (GJ)}
@@
 
+% ./scripts/fig_GJ_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
+\caption{$\mathbf{GJ}$ model, surgical manipulations}
+\label{f:GJ}
+\end{figure}
+
+\subsection*{Single axon behaviour}
+
+Acknowledge the result of \citet{gosse_retinotopic_2008} and show the
+behaviour of the GC model (and by extension the other models) if a single axon
+is allowed to develop.
+
 \section{Discussion}

### 2022-02-01 paper/paper.tex
Adds caption

@@
 \includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig3.png}
-\caption{The pure gradient-following model (alt exponential).}
+\caption{By tuning the tectal ligand expression (replacing Eq.\,\ref{e:tecligexp} with
+ Eq.\,\ref{e:tecligexp2}), the pure gradient-following model can achieve a
+ successful layout in which the axons are arranged evenly across the full
+ tectum. At the centre of the tectum, where neither retinal receptor
+ expression nor tectal ligand expression is high, the ordering is weaker than
+ towards the edge. \textbf{A} Axon branches \textbf{B} axon
+ centroids \textbf{C} Five selected axons.}
 \label{f:chalt}

### 2022-02-02 paper/paper.tex
Remaking the graph for chemo-only surgical manipulations

@@
 
+\color{grey}
 % Note this is the SfN draft abstract, and will need rewriting.
@@
 continuous retinotectal system.}
+\color{black}
+
+\emph{The development of the axonal projection from the retina to the tectum
+is an important model for the study of how the brain organises its
+connectivity during development. In fish, amphibians and mammals, retinal
+ganglion cells grow their axons into the brain and make topographically
+ordered connections on the tectum.
+%
+Computational models seeking to provide a plausible mechanism by which this
+system can order itself have existed since the 1970s. However, most of these
+have been phenomenological, some only possible with the invokation of a global
+supervising process that provides the state of the entire system to each
+axonal agent.
+%
+Here, we investigated whether a purely self-organising model based on receptor
+and ligand interactions could be found, that would successfully reproduce both
+the normal development of the projection and its behaviour under the many
+experimental manipulations carried out by neurobiologists.
+%
+We show that a model based on gradient following (via receptor-ligand
+interactions) combined with receptor-receptor mediated axonal repulsion can
+account for a broad range of experimental results, with information only being
+transferred by nearest-neighbour interactions.
+}
 
@@
 in the environment tend to activate retinal ganglion cells (either
-directly \cite{iRGC_citation} or indirectly via rods and cones) situated close
-together in the eye so that an image of the environment is formed on the
-retinal surface. It has been discovered that the topography of the retina is
-preserved within brain regions that process this information such that cells
-which are adjacent within the retina primarily excite cells adjacent in the
-tectum. This indicates that during development there must exist a mechanism
-which ensures the correct arrangement of the axons which leave the retina and
-connect to cells in the tectum.
+directly \cite{iRGC_citation} or indirectly via rods and
+cones \cite{rods_cones_citations}) situated close together in the eye so that
+an image of the environment is formed on the retinal surface. It has been
+discovered that the topography of the retina is preserved within brain regions
+that process this information such that cells which are adjacent within the
+retina primarily excite cells adjacent in the tectum. This indicates that
+during development there must exist a mechanism which ensures the correct
+arrangement of the axons which leave the retina and connect to cells in the
+tectum.
 
@@
 reorganised in-vivo, after which axons regrow to restore the order and
-function of the system for the individual animal. This manipulability was
-exploited in influential work by R. W. Sperry and co-workers during the
-mid-twentieth century, leading to Sperry's 1963 summary of
-the \emph{chemoaffinity theory} \citep{sperry_chemoaffinity_1963} which
-proposes the existence of morphogenetic gradients that guide axons to their
-destination. The chemoaffinity theory was given robust support by the
-discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995}
-which have been shown to form into graded expression fields in the
-retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other
+function of the system for the individual animal. Influential work by
+R. W. Sperry and co-workers during the mid-twentieth century exploited this
+manipulability, leading to Sperry's 1963 summary of the \emph{chemoaffinity
+theory} \citep{sperry_chemoaffinity_1963} which proposes the existence of
+morphogenetic gradients that guide axons to their destination. The
+chemoaffinity theory was given robust support by the discovery of the ephrin
+ligands and their
+receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been
+shown to form into graded expression fields in the
+retina \citep{braisted_graded_1997},
+tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other
 sensory systems, such as the somatosensory
 system \citep{vanderhaeghen_mapping_2000}. The Ephrin ligands have a clear
-effect on axonal outgrowth, as shown in in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and
-in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
+effect on axonal outgrowth, as shown
+in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004}
+and by
+in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}
+studies. However, a gradient fo
 %
@@
 
-It is tempting to consider the retinotectal projection well understood. With a
-comprehensive theory supported by a biochemical mechanism, is there anything
-left to understand? That question can be answered by reviewing retinotectal
-modelling papers.
+%It is tempting to consider the retinotectal projection well understood. With a
+%comprehensive theory supported by a biochemical mechanism, is there anything
+%left to understand? That question can be answered by reviewing retinotectal
+%modelling papers.
 
-Mini-review here, which contrasts some of the modelling papers. The upshot is
-that a central problem with models is not how the axons know how to get closer
-to their destination, but how they know they have \emph{arrived} at their true
-destination. Also make the point that many of the models are phenomenological
-in nature and do not elucidate the mechanisms behind the arrangement.
+%Mini-review here, which contrasts some of the modelling papers. The upshot is
+%that a central problem with models is not how the axons know how to get closer
+%to their destination, but how they know they have \emph{arrived} at their true
+%destination. Also make the point that many of the models are phenomenological
+%in nature and do not elucidate the mechanisms behind the arrangement.%
 
@@
 % Visualisation of ligand expression and gradients on the tectum for the
-% graft-rotate 90 degrees simulated surgical manipulation
+% graft-rotated 90 degrees simulated surgical manipulation
 % ./build/sim/agent/tissuevis configs/simpler/m_eE_G.json configs/simpler/e_tecrot90_fig4row1.json
@@
 along with the half of the tectum to which the remaining retina would project.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_manipulations.png}
-\caption{Ligand expression and gradients on the tectum for the
-other simulated surgical manipulations}
-\label{f:tmanip}
-\end{figure}
 
@@
 \caption{$\mathbf{GC}$ model, surgical manipulations}
-\label{f:GC}
+\label{f:GCsurg}
 \end{figure}
@@
 
+See Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.
+
 % ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json
@@
 \caption{$\mathbf{GI}$ model, surgical manipulations}
-\label{f:GI}
+\label{f:GIsurg}
 \end{figure}
@@
 
-% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3.png}
-\caption{$\mathbf{GJ}$ model}
-\label{f:GJ}
-\end{figure}
-
-% ./scripts/fig_GJ_surgical.sh
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
-\caption{$\mathbf{GJ}$ model, surgical manipulations}
-\label{f:GJ}
-\end{figure}
+See Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
 
@@
 
+%
+% Supplementary figures in an appendix
+%
+\appendix
+
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+\section*{Supplementary figures}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expressions_manipulations.png}
+\caption{Ligand expression and gradients on the tectum for the
+other simulated surgical manipulations}
+\label{f:tmanip}
+\end{figure}
+
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3.png}
+\caption{$\mathbf{GJ}$ model}
+\label{f:GJ}
+\end{figure}
+
+% ./scripts/fig_GJ_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
+\caption{$\mathbf{GJ}$ model, surgical manipulations}
+\label{f:GJsurg}
+\end{figure}
+
 \end{document}

### 2022-02-02 sim/agent/agent1.h

@@ -226,6 +226,7 @@ struct Agent1
 
             // Update retinal NT position vs tectal RC position graph for the 'd' layout:
             if (this->layout == graph_layout::d || this->layout == graph_layout::e) {
+
                 morph::vVector<T>nt;
                 morph::vVector<T>rc;
                 morph::vVector<T>nt_m; // manipulated
@@ -251,7 +252,8 @@ struct Agent1
                 this->gv->reinit();
                 this->v->render();
 
-            } else if (this->layout == graph_layout::h || this->layout == graph_layout::k) {
+            } else if (this->layout == graph_layout::h || this->layout == graph_layout::k ||
+                       (this->layout == graph_layout::f && this->conf->getBool("rc_vs_nt", false) == true)) {
                 // Graph layout h/k shows the nt vs rc (for the mid band of the retina/central column of the tectum?).
                 morph::vVector<T>nt;
                 morph::vVector<T>rc;
@@ -368,17 +370,28 @@ struct Agent1
             this->cv->reinit(); // Centroids
             this->sim_time_txt->setupText (std::to_string(stepnum)); // Text
             AgentMetrics<T> am = this->get_metrics();
-            this->emetric_txt->setupText (am.rms.str());
-            this->crossings_txt->setupText (am.crosscount.str());
+            std::stringstream amss, ccss;
+            amss.precision(4);
+            ccss.precision(4);
+            amss << am.rms[0];
+            ccss << am.crosscount[0];
+            this->emetric_txt->setupText (amss.str());
+            this->crossings_txt->setupText (ccss.str());
+            if (this->conf->getBool ("rc_vs_nt", false) == true) {
+            } else {
+                this->gv->append ((float)stepnum, this->ax_centroids.rms(), 0); // RMS (whole)
+                this->gv->append ((float)stepnum, this->ax_centroids.rms_outside({0.25,0.25,0},{0.75,0.75,0}), 1); // RMS (surround)
+                this->gv->append ((float)stepnum, this->ax_centroids.rms_inside({0.25,0.25,0},{0.75,0.75,0}), 2); // RMS (patch)
+            }
             break;
         }
         case graph_layout::g:
         {
             this->bv->reinit(); // Branches
             this->cv->reinit(); // Centroids
[…]

### 2022-02-03 paper/paper.tex
WIP

@@
 
-\color{grey}
-% Note this is the SfN draft abstract, and will need rewriting.
-\emph{In the classic Chemoaffinity theory, the retinotectal axon projection is
-thought to use pairs of orthogonal signalling gradients in the retina to
-specify the eventual location of synapses made on the surface of the
-tectum/superior colliculus. Similar orthogonal gradients in the tectum provide
-a coordinate system which allows the axons to match their prespecified
-destination with the correct location. Although the Ephrins have been shown to
-guide axons toward their destination, there has yet to emerge a complete
-account of the local interactions which halt the axonal growth cones in the
-correct locations to recreate the topography of the retinal cells. The model
-of \citet{simpson_simple_2011} provides an account of the basic topographic
-arrangement of cells on the tectum, as well as reproducing well known surgical
-and genetic manipulation experiments. However, it suffers from the absence of
-a local chemotactic guidance mechanism. Instead, each agent in that model is
-given instantaneous knowledge of the vector that would move it toward its
-pre-specified destination. In addition to the globally supervised
-chemoaffinity term, \citet{simpson_simple_2011} introduced a competitive
-interaction for space between growth cone agents and a receptor-receptor
-axon-axon interaction in order to account for the full set of experimental
-manipulations. Here, we propose the replacement of the chemoaffinity term with
-a gradient following model consisting of axonal growth cone agents which carry
-receptor molecule expression determined by their soma's location of origin on
-the retina. Growth cones move on the simulated tectum guided by two pairs of
-opposing, orthogonal signalling molecules representing the Ephrin ligands. We
-show that with only the receptor-ligand mediated chemoaffinity term and a
-receptor-receptor based axon-axon interaction term (meaning that all growth
-cone interactions are communicated by receptor signalling), a full range of
-experimental manipulations to the retinotectal system can be
-reproduced. Furthermore, we show that the observation that competition is not
-an essential requirement for axons to find their
-way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
-the opposing influences of signalling gradient pairs. Finally, we demonstrate
-that, assuming exponentially varying receptor expression in the retina, tectal
-ligand expression should either be exponential if the receptor-ligand signal
-induces repulsion (i.e. gradient descent) or logarithmic if the signal induces
-attraction (gradient ascent). Thus, we find that a model analogous to the one
-we presented in \citet{james_modelling_2020} that accounts for murine barrel
-patterning is also a candidate mechanism for the arrangement of the more
-continuous retinotectal system.}
-\color{black}
+%% \color{grey}
+%% % Note this is the SfN draft abstract, and will need rewriting
+%% % COULD go into the introduction, with some modification.
+%% \emph{In the classic Chemoaffinity theory, the retinotectal axon projection is
+%% thought to use pairs of orthogonal signalling gradients in the retina to
+%% specify the eventual location of synapses made on the surface of the
+%% tectum/superior colliculus. Similar orthogonal gradients in the tectum provide
+%% a coordinate system which allows the axons to match their prespecified
+%% destination with the correct location. Although the Ephrins have been shown to
+%% guide axons toward their destination, there has yet to emerge a complete
+%% account of the local interactions which halt the axonal growth cones in the
+%% correct locations to recreate the topography of the retinal cells. The model
+%% of \citet{simpson_simple_2011} provides an account of the basic topographic
+%% arrangement of cells on the tectum, as well as reproducing well known surgical
+%% and genetic manipulation experiments. However, it suffers from the absence of
+%% a local chemotactic guidance mechanism. Instead, each agent in that model is
+%% given instantaneous knowledge of the vector that would move it toward its
+%% pre-specified destination. In addition to the globally supervised
+%% chemoaffinity term, \citet{simpson_simple_2011} introduced a competitive
+%% interaction for space between growth cone agents and a receptor-receptor
+%% axon-axon interaction in order to account for the full set of experimental
+%% manipulations. Here, we propose the replacement of the chemoaffinity term with
+%% a gradient following model consisting of axonal growth cone agents which carry
+%% receptor molecule expression determined by their soma's location of origin on
+%% the retina. Growth cones move on the simulated tectum guided by two pairs of
+%% opposing, orthogonal signalling molecules representing the Ephrin ligands. We
+%% show that with only the receptor-ligand mediated chemoaffinity term and a
+%% receptor-receptor based axon-axon interaction term (meaning that all growth
+%% cone interactions are communicated by receptor signalling), a full range of
+%% experimental manipulations to the retinotectal system can be
+%% reproduced. Furthermore, we show that the observation that competition is not
+%% an essential requirement for axons to find their
+%% way \citep{gosse_retinotopic_2008} is also accounted for by the model, due to
+%% the opposing influences of signalling gradient pairs. Finally, we demonstrate
+%% that, assuming exponentially varying receptor expression in the retina, tectal
+%% ligand expression should either be exponential if the receptor-ligand signal
+%% induces repulsion (i.e. gradient descent) or logarithmic if the signal induces
+%% attraction (gradient ascent). Thus, we find that a model analogous to the one
+%% we presented in \citet{james_modelling_2020} that accounts for murine barrel
+%% patterning is also a candidate mechanism for the arrangement of the more
+%% continuous retinotectal system.}
+%% \color{black}
 
@@
 %
-Here, we investigated whether a purely self-organising model based on receptor
-and ligand interactions could be found, that would successfully reproduce both
-the normal development of the projection and its behaviour under the many
+Here, we investigated whether, using only receptor and ligand interactions, a
+purely self-organising model could be found that would successfully reproduce
+both the normal development of the projection and its behaviour under the many
 experimental manipulations carried out by neurobiologists.
@@
 We show that a model based on gradient following (via receptor-ligand
-interactions) combined with receptor-receptor mediated axonal repulsion can
-account for a broad range of experimental results, with information only being
-transferred by nearest-neighbour interactions.
+interactions) combined with receptor-receptor mediated axonal repulsion, in
+which information is only transferred by nearest-neighbour interactions, can
+account for a broad range of experimental results.
 }
@@
 that process this information such that cells which are adjacent within the
-retina primarily excite cells adjacent in the tectum. This indicates that
-during development there must exist a mechanism which ensures the correct
-arrangement of the axons which leave the retina and connect to cells in the
-tectum.
+retina primarily excite cells adjacent in the tectum. Of the retinal axons
+that project directly to the tectum, those originating in the temporal retina
+make connections in the rostral tectum; thoes originating in dorsal retina
+connect to lateral tectum (Fig.\,\ref{}A).  This pattern of connectivity
+establishes itself before activity \citep{something}, indicating that during
+development there must exist a mechanism which ensures the correct arrangement
+of the axons which leave the retina and connect to cells in the tectum.
 
@@
 function of the system for the individual animal. Influential work by
-R. W. Sperry and co-workers during the mid-twentieth century exploited this
+R.~W.~Sperry and co-workers during the mid-twentieth century exploited this
 manipulability, leading to Sperry's 1963 summary of the \emph{chemoaffinity
@@
 in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}
-studies. However, a gradient fo
+studies. \emph{Does this para need finishing?}
 %
@@
 receptor expression and tectal ligand expression across these squares
-(Fig.\,\ref{f:ex}).
+(Fig.\,\ref{f:ex}B-D). The simulated retina possessed $n$=400 (20$\times$20) retinal
+ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square
+elements, with defined levels of expression for each ligand. One spatial unit
+corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
 
@@
 %
-\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown in a linear form,
-as there is less evidence to determine whether ligand expression is
-exponential, linear or logarithmic on the tectum.
+\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown
+in an exponential form,
+although there is less evidence to determine whether tectal expression is
+exponential, linear or logarithmic.
 $L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
@@
 
-Each retinal ganglion cell projects growth cones (referred to
+Each retinal ganglion cell (RGC) projects growth cones (referred to
 as \emph{branches}) which carry a set of 4 receptors, $r_i$ indexed by $i$ and
 expressed at levels determined by the cell soma's location on the
-retina. (Four ligands, $l_i$, are also expressed by RGCs, but these are not
-used by the chemoaffinity simulation.) The expression levels of the receptors
-vary with respect to the soma's position on the retinal surface. Each receptor
-varies only with respect to one dimension.  Receptor expression gradients are
+retinal grid. (Four ligands, $l_i$, are also expressed by RGCs, but these are not
+used by the chemoaffinity simulation.) The expression of each receptor
+varies only with respect to one spatial dimension.  Receptor expression gradients are
 arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to
@@
 \begin{equation} \label{e:retrcpt}
-r_i(x,y) = 1.05 + 0.26 \exp(2.3 x)
+r_i(x,y) = 1.05 + 0.26 \exp(2.3 u)
 \end{equation}
-for even values of $i$. The opposing expressions (for even values of $i$)
-are given by substituting $(1-x)$ for $x$.
+Where $u$ gives the direction with which the expression increases and is
+substituted by $1-x$ for $r_0$, $1-y$ for $r_1$, $x$ for $r_2$ and by $y$ for
+$r_3$ (Fig.\,\ref{1}B).
 
-The tectum expresses ligands, $L_i$, for the retinal receptors, also
-in orthogonal pairs of gradients. Although several studies model tectal ligand
+The tectum expresses ligands, $L_i$, for the retinal receptors, also in
+orthogonal pairs of gradients. Although several studies model tectal ligand
 expression with exponential functions \citep{koulakov_stochastic_2004}, the
-experimental evidence for ligand expression is more ambiguous than that for
-retinal receptor expression (Supp.~Fig
-\textbf{XXXXX}). We kept in mind the possibility that tectal ligand
-expression may be modelled by exponential, linear or logarithmic functions,
-but initially, we set it to the function
+experimental evidence for the form of tectal ligand expression is more ambiguous than that for
+retinal receptor expression (Fig.\,\ref{f:expressions} summarises results from
+the literature). We kept in mind the
+possibility that tectal ligand expression may be modelled by exponential,
+linear or logarithmic functions, but initially, we set it to the exponential function
 \begin{equation} \label{e:tecligexp}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 x)
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 u)
 \end{equation}
-for odd values of $i$ and using the same substitution of $(1-x)$ for $x$ to
-obtain the opposing ligand expressions for $i$ even.
+and use the substitutions $y$, $x$, $(1-y)$ or $1-x$ in place of $u$ to obtain
+the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$
+(Fig.\,\ref{1}D).
 
@@
 RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a
-r\^ole in retinotectal map formation.
+r\^ole in retinotectal map formation. The fact that we have chosen opposing
+gradients is also guided by experimental observations that these exist in the
+brain \citep{opposing_gradients}. Without a system of opposing gradients, it
+would not be possible for axons to stop without an additional signal
+instructing them to do so.
 
-Figure~\ref{f:ex} shows colour maps of receptor and ligand
-expression. Coordinates within the simulated retina and tectum are normalised,
-with 1 corresponding to 2 mm on the tectum and on the retina for
-mouse \citep{reber_relative_2004}.
+%already said that:
+%Figure~\ref{f:ex} shows colour maps of receptor and ligand
+%expression. Coordinates within the simulated retina and tectum are normalised,
+%with 1 corresponding to 2 mm on the tectum and on the retina for
+%mouse \citep{reber_relative_2004}.
 
@@
 \color{black}
-Each discrete element in Fig.\,\ref{f:ex}B projects one RGC axon with 4
+Each discrete, square element in Fig.\,\ref{f:ex}B projects one RGC axon with 4
 branches to the tectum, carrying the receptor expression at the retinal
@@
 %
-where $\mathbf{M}_{b}$, the movement vector, is the sum of a chemoaffinity
+where $\mathbf{M}_{b}$, the branch movement vector, is the sum of a chemoaffinity
 effect and a `border effect', $\mathbf{B}_b$, which acts to retain branches
@@
 molecule of ligand $i$ binds to a receptor $i$ molecule. $F_i$ takes the value
--1 for a repulsive interaction or 1 for an attractive interaction.
+$-1$ for a repulsive interaction or $1$ for an attractive interaction.
 %
@@
 %
-$m_{\!_G}$ is a
+$m_{\!_G}$ in Eq.\,\ref{e:mv} is a
 scalar parameter which controls how much movement is generated for a given
-receptor-ligand signal.
+chemotactic movement vector, $\mathbf{G}_b$.
 % end chemoaffinity description
@@
 Branches were initially randomly distributed in a stripe at the rostral side
-of the tectum at the start of each simulation. Each RGC axon was assigned a
+of the tectum at the start of each simulation. Each RGC axon $k$ was assigned a
 random initial position coordinate:
-\begin{equation}
-\mathbf{x}_{\text{axon},t=0} = (U(0,1), U(-0.2,0))
+\begin{equation}\label{e:ic}
+\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
 \end{equation}
@@
 red. Lightness gives position along the relevant axis so that cells at the
-dorsal-nasal corner are yellow and those at the ventral-transverse corner are
+dorsal-nasal corner are yellow and those at the ventral-temporal corner are
 black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
 axons on the tectum. The axons retain the colour of their originating cell on
-the retina, as shown in Fig.\,\ref{f:ch}A. Nasal cell axons should be arranged
+the retina. Nasal cell axons should be arranged
 caudally; dorsal axons laterally; axons should be evenly spaced and the
-pattern should span the entire tectum.  Fig.\,\ref{f:ch}B--D show `fish net'
-plots of the axon centroids at three timepoints. An axon's centroid is the
-average location of its 4 branches. In these plots, lines are drawn between
+pattern should span the entire tectum.  Fig.\,\ref{f:ch}, panels B--D show
+`fish net' plots of the average location of the 4 branches of each axon (the
+axon's centroid) at three timepoints. Fig.\,\ref{f:ch}B indicates the
+randomised placement of axon branches at the rostral end of the tectum as
+specified by the initial conditions (Eq.\,\ref{e:ic}). In these plots, lines are drawn between
 axons that are expected to arrange themselves adjacently to one another by the
 end of the simulation. The number of line crossings in this plot gives an
-indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the
-positions of the individual branches (i.e.~growth cones) at $t=300$ that
-contribute to the the plot in Fig.\,\ref{f:ch}D. By this time, the topological
-order matches that shown in Fig.\,\ref{f:ch}E. Opposing gradients have sorted
-the axons along the two axes to form a grid of cells. However, the arrangement
-does not span the full tectum. Fig.\,\ref{f:ch}F shows two pattern
-metrics. The first metric, $\epsilon$, is the sum of squared distance errors
-of the axon centroids, with $\mathbf{x}_{k}$, the location of the centroid of
-the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for
-axon $k$:
+indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the positions of the individual branches
+(i.e.~growth cones) at $t=300$ that contribute to the plot in
+Fig.\,\ref{f:ch}D. By this time, the topological order matches that shown in
+Fig.\,\ref{f:ch}E. Opposing gradients have sorted the axons along the two axes
+to form a grid of cells. However, the arrangement does not span the full
+tectum. Fig.\,\ref{f:ch}F shows two pattern metrics; $\epsilon$ is the mean
+distance error of the axon centroids, given by:
 %
 \begin{equation}\label{e:eps}
-\epsilon = \sum_k |\mathbf{x}_{k} - \mathbf{x}'_{k}|^2.
+\epsilon = \sqrt{\frac{\sum_k (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
 \end{equation}
+%
+where $\mathbf{x}_{k}$ is the location of the centroid of the
+branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon
+$k$. $n$ is the number of RGC axons in the system.
 %
@@
 use the number of crossings in the fish net plot of Fig.\,\ref{f:ch}D as a
-second metric, called $\eta$. The expected number of crossings in the wildtype
-pattern is 0, as in Fig.\,\ref{f:ch}E, although note that in some patterns,
-the expected number of crossings is non-zero. Fig.\,\ref{f:ch}H shows selected
+second metric, called $\eta$, which tends towards an expected target number of
+crossings, $\eta_t$ as the pattern approaches the target pattern.  For the wildtype
+pattern, $\eta_t=0$, as in Fig.\,\ref{f:ch}E, although note that in some patterns,
+$\eta_t$ is non-zero. Fig.\,\ref{f:ch}H shows selected
 axons (one from the centre and one from each corner of the retina) and their
@@
 \begin{equation} \label{e:tecligexp2}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.1 x)
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.1 u)
 \end{equation}
 %
-for the tectal ligand expression.
+for Eq.\,\ref{e:tecligexp}.
 
@@
 found easily by hand tuning) allows a well ordered map to form for the
-wildtype system. Fig.\,\ref{f:chalt}C shows the error metric
-(\textbf{FIXIMAGE}), $\epsilon$, which is much closer to 0 with the new ligand
-expression.
+wildtype system. Fig.\,\ref{f:chalt}C shows the error metric, $\epsilon$,
+which is much closer to 0 with the new ligand expression \textbf{Fix graph}.
 
@@
 graft of the tectum is rotated by a quarter turn and replaced. We rotate an
-8x8 square of the tectum and then recompute the gradients. As the ligand
-expression is discretize rather than continuous, we compute the gradient
+8$\times$8 square of the tectum and then recompute the gradients. As the ligand
+expression is discretized rather than continuous, we compute the gradient
 numerically, with the gradient in direction $x$ computed as the difference
@@
 Fig\,\ref{f:chsurg} shows the results of 6 simulated surgical
-manipulations. In row A, a patch of tectal tissue has been cut out and rotated
-by a quarter turn, as in Fig.\,\ref{f:trot90}. From left to right, graphs show
-i) the expected final layout of the tectum, according to the results of
-experimental work, with the colour code giving the origin location of the
-axon's soma on the retina, ii) The locations of individual axon branches at
-simulation end, iii) the centroids of each axons' branches and iv) a rather
-useless selected axons box \textbf{FIXME}. Simulation time was 1000 time steps. The final
-locations of branches and axon centroids indicates that a qualitative rotation
-of the graft patch has occurred, with red branches inside the patch being
-located at the lateral-caudal corner rather than the medial-caudal corner. A
-portion of the graft region (the medial side, after rotation) is free of
-branches, indicating that the pattern is partial. A similar result is seen for
-the 180\degree~example in row B, in which the caudal half of the rotated patch
-is free of branches, but the rostral half contains branches with the correctly
-rotated orientations.
+manipulations. In A, a patch of tectal tissue has been cut out and rotated by
+a quarter turn (the manipulation illustrated in Fig.\,\ref{f:trot90}). For
+this simulation, the figure shows i) The locations of the axon centroids at
+simulation end, ii) (inset) the expected final layout of the tectum, according
+to the results of experimental work, with the colour code giving the origin
+location of the axon's soma on the retina and iii) the error metric,
+$\epsilon$ plotted versus simulation time. Simulation time was 1000 time
+steps. The final locations of branches and axon centroids indicates that a
+qualitative rotation of the graft patch has occurred, with red branches inside
+the patch being located at the lateral-caudal corner rather than the
+medial-caudal corner. A portion of the graft region (the medial side, after
+rotation) is free of branches, indicating that the pattern is partial. A
+similar result is shown for the 180\degree~example in B, in which the caudal
+half of the rotated patch is free of branches, but the rostral half contains
+branches with the correctly rotated orientations.
 
 Fig.\,\ref{f:chsurg}C shows the result of the patch swap manipulation, in
-which 12x4 element grafts, one from the caudal side of the tectum and one from
+which 12$\times$4 element grafts, one from the caudal side of the tectum and one from
 the rostral side are swapped. This simulation is unsuccessful in patterning
@@
 The retinal ablation shown in Fig.\,\ref{f:chsurg}D is also unsuccessful, with
-the axons from the surviving nasal retina (the transverse half was removed)
+the axons from the surviving nasal retina (the temporal half was removed)
 failing to arrange themselves across the entire tectum, due to the pure
-`position specification' which the dual gradient model achieves. Similarly,
+`position specification' which the dual gradient model achieves. Here, the
+average axon position on the rostral-caudal axis of the tectum is plotted versus the average
+nasal-temporal origin of the axon on the retina. Similarly,
 the tectal ablation simulation fails to reproduce experimentally observed
 results in which the retinal map `squashes up' within the remaining tectal
-region. Instead, axons from the transverse retina arrange with normal spacing
+region. Instead, axons from the temporal retina arrange with normal spacing
 and axons from the nasal retina are pushed to the caudal-most side of the
@@
 mismatch experiment (Fig.\,\ref{f:chsurg}F), but in this case there are no
-axons from the transverse retina, which was been ablated and only the axons
+axons from the temporal retina, which was been ablated and only the axons
 from the nasal retina are present, arranged in a line.
@@
 \begin{figure}
+\includegraphics[width=\linewidth]{../misc/expression_profiling/expression_profiles.png}
+\caption{Expression profiles from the literature.}
+\label{f:expressions}
+\end{figure}
+
+\begin{figure}
 \includegraphics[width=\linewidth]{./images/expressions_manipulations.png}

### 2022-02-04 paper/paper.tex
WIP switch to other laptop

@@
 Each retinal ganglion cell (RGC) projects growth cones (referred to
-as \emph{branches}) which carry a set of 4 receptors, $r_i$ indexed by $i$ and
+as \emph{branches}) which carry a set of $m$ receptors, $r_i$ indexed by $i$ and
 expressed at levels determined by the cell soma's location on the
-retinal grid. (Four ligands, $l_i$, are also expressed by RGCs, but these are not
+retinal grid. ($m$ ligands, $l_i$, are also expressed by RGCs, but these are not
 used by the chemoaffinity simulation.) The expression of each receptor
-varies only with respect to one spatial dimension.  Receptor expression gradients are
-arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to
-that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, is orthogonal to
-$r_3$.
+varies only with respect to one spatial dimension.
+
+The minimum number of gradients required to provide a coordinate system on the
+surface of the tectum within which axons could find a target location is
+$m=2$, arranged in orthogonal directions. However if there are only two
+gradients, imagining that one guides axons laterally and the other guides
+axons in the caudal direction, then it is necessary for there to be an
+additional `stopping mechanism' to prevent all of the axons eventually being
+guided to the most lateral/caudal position. To avoid the need for a different
+stopping mechanism, we chose to model \emph{opposing pairs of orthogonal
+gradients} with $m=4$. With opposing gradients, an axon which interacts most
+strongly with a gradient in the caudal direction would be drawn to the caudal
+side of the tectum, whereas an axon whose strongest interaction was to climb a
+gradient in the rostral direction would be drawn to the rostral side. There is
+evidence that multiple gene expression gradients are found in the brain, with
+examples of opposing pairs including XX and YY.
+
+We defined 4 receptor expression gradients arranged in orthogonal pairs
+with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose
+gradient is opposite to $r_0$, was orthogonal to $r_3$.
 

### 2022-02-08 paper/paper.tex
WIP

@@
 %
-\begin{equation}
+\begin{equation}\label{e:G}
 \mathbf{G}_b = \sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x}_b)
@@
 
+\subsection*{Robustness to noise}
+
+\emph{This section might fit most naturally after the Chemoaffinity only section.}
+
+\begin{figure}
+\includegraphics[width=0.3\linewidth]{./images/GI_vs_noise_tweaked.png}
+\caption{The mean value of $\eta$ (number of crossings) and $\epsilon$
+(mean distance error), averaged over the period from simulation time $t=500$ to $t=1000$, plotted with respect to the
+noise gain in the $\mathbf{GI}$ model and the wildtype condition (compare with
+Fig.\,\ref{f:GI}). Both metrics indicate that a moderate
+amount of gradient sampling noise is \emph{beneficial} to the establishment of
+an ordered tectum.}
+\label{f:noise}
+\end{figure}
+
+One problem with axon gradient-following is the ability of the axon to
+successfully detect the direction of the gradient given the small distances
+over which the axon can sample signalling molecules and the size of receptor
+binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
+system to self-organise in the presence of noise, by introducing a stochastic
+element to each branch's sampling of the gradient.
+
+% See branch::estimate_ligand_gradient() in sim/agent/branch.h for the code
+Noise is added to the estimation of the ligand gradient by multiplying the
+true gradient by a random vector, $\mathbf{U}$, whose coordinates are drawn
+from a random uniform distribution centred about 0 with magnitude $\nu$, denoted:
+$\mathbf{U}(-\nu, \nu)$. Eq.\,\ref{e:G} becomes
+
+\begin{equation}\label{e:Gnu}
+\mathbf{G}_b = \sum_i^N F_i\,r_{i,b}\, \mathbf{U}(-\nu, \nu) \nabla L_i(\mathbf{x}_b)
+\end{equation}
+
+Fig.\,\ref{f:noise} shows the response of the $\mathbf{GI}$ model to the
+addition of ligand gradient estimation noise.
+
+What about retinal expression noise? This would have a disordering effect on
+the pattern.
+
 \subsection*{Single axon behaviour}

### 2022-02-08 paper/GI_vs_noise.py

@@ -38,15 +38,17 @@ with open ('GI_vs_noise.txt') as csvfile:
 
     fig, ax1 = plt.subplots()
 
-    color = clr1
+    color = clr2
     ax1.set_xlabel('Noise gain')
-    ax1.set_ylabel(r'$\bar{\eta}$', rotation=0,  color=blk)
+    # The bar reproduces well in matplotlib, but not in svg/inkscape
+    #ax1.set_ylabel(r'$\bar{\eta}$', rotation=0,  color=blk)
+    ax1.set_ylabel(r'$\eta$', rotation=0,  color=blk)
     ax1.plot(noise_gain, mean_eta, color=color)
     ax1.tick_params(axis='y', labelcolor=color)
 
     ax2 = ax1.twinx()  # instantiate a second axes that shares the same x-axis
-    color = clr2
-    ax2.set_ylabel(r'$\bar{\epsilon}$', rotation=0, color=blk)  # we already handled the x-label with ax1
+    color = clr1
+    ax2.set_ylabel(r'$\epsilon$', rotation=0, color=blk)  # we already handled the x-label with ax1
     ax2.plot(noise_gain, mean_eps, color=color)
     ax2.tick_params(axis='y', labelcolor=color)
 

### 2022-02-09 paper/paper.tex
Wednesday WIP

@@
 % Use the article doc class, with an 11 pt basic font size
-\documentclass[11pt, a4paper]{article}
+\documentclass[11pt, a4paper, draft]{article}
+%\documentclass[11pt, a4paper]{article}
 
@@
 Each retinal ganglion cell (RGC) projects growth cones (referred to
-as \emph{branches}) which carry a set of $m$ receptors, $r_i$ indexed by $i$ and
+as \emph{branches}) which carry a set of $N$ receptors, $r_i$ indexed by $i$ and
 expressed at levels determined by the cell soma's location on the
-retinal grid. ($m$ ligands, $l_i$, are also expressed by RGCs, but these are not
+retinal grid. ($N$ ligands, $l_i$, are also expressed by RGCs, but these are not
 used by the chemoaffinity simulation.) The expression of each receptor
@@
 surface of the tectum within which axons could find a target location is
-$m=2$, arranged in orthogonal directions. However if there are only two
+$N=2$, arranged in orthogonal directions. However if there are only two
 gradients, imagining that one guides axons laterally and the other guides
@@
 stopping mechanism, we chose to model \emph{opposing pairs of orthogonal
-gradients} with $m=4$. With opposing gradients, an axon which interacts most
+gradients} with $N=4$. With opposing gradients, an axon which interacts most
 strongly with a gradient in the caudal direction would be drawn to the caudal
@@
 Where $u$ gives the direction with which the expression increases and is
-substituted by $1-x$ for $r_0$, $1-y$ for $r_1$, $x$ for $r_2$ and by $y$ for
-$r_3$ (Fig.\,\ref{1}B).
-
-The tectum expresses ligands, $L_i$, for the retinal receptors, also in
+substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for
+$r_3$ (Fig.\,\ref{f:ex}B).
+
+Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in
 orthogonal pairs of gradients. Although several studies model tectal ligand
@@
 possibility that tectal ligand expression may be modelled by exponential,
-linear or logarithmic functions, but initially, we set it to the exponential function
+linear or logarithmic functions, but initially, we set it to the same
+exponential function used for retinal expression,
 \begin{equation} \label{e:tecligexp}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 u)
-\end{equation}
-and use the substitutions $y$, $x$, $(1-y)$ or $1-x$ in place of $u$ to obtain
+L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 u),
+\end{equation}
+and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain
 the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$
-(Fig.\,\ref{1}D).
+(Fig.\,\ref{f:ex}D).
 
@@
 \color{black}
-Each discrete, square element in Fig.\,\ref{f:ex}B projects one RGC axon with 4
+Each discrete, square element in Fig.\,\ref{f:ex}B projects one RGC axon with $M=4$
 branches to the tectum, carrying the receptor expression at the retinal
@@
 The results of running this simulation for 300 time steps are shown in
-Fig.\,\ref{f:ch}. Fig.\,\ref{f:ch}A indicates the colour code used for retinal
+Fig.\,\ref{f:G}. Fig.\,\ref{f:G}A indicates the colour code used for retinal
 cell position; dorsal retinal cells are indicated by green, nasal cells by
@@
 dorsal-nasal corner are yellow and those at the ventral-temporal corner are
-black. Fig.\,\ref{f:ch}E shows the expected layout of retinal ganglion cell
+black. Fig.\,\ref{f:G}E shows the expected layout of retinal ganglion cell
 axons on the tectum. The axons retain the colour of their originating cell on
@@
 caudally; dorsal axons laterally; axons should be evenly spaced and the
-pattern should span the entire tectum.  Fig.\,\ref{f:ch}, panels B--D show
+pattern should span the entire tectum.  Fig.\,\ref{f:G}, panels B--D show
 `fish net' plots of the average location of the 4 branches of each axon (the
-axon's centroid) at three timepoints. Fig.\,\ref{f:ch}B indicates the
+axon's centroid) at three timepoints. Fig.\,\ref{f:G}B indicates the
 randomised placement of axon branches at the rostral end of the tectum as
@@
 end of the simulation. The number of line crossings in this plot gives an
-indication of the disorder in the arrangement. Fig.\,\ref{f:ch}G shows the positions of the individual branches
+indication of the disorder in the arrangement. Fig.\,\ref{f:G}G shows the positions of the individual branches
 (i.e.~growth cones) at $t=300$ that contribute to the plot in
-Fig.\,\ref{f:ch}D. By this time, the topological order matches that shown in
-Fig.\,\ref{f:ch}E. Opposing gradients have sorted the axons along the two axes
+Fig.\,\ref{f:G}D. By this time, the topological order matches that shown in
+Fig.\,\ref{f:G}E. Opposing gradients have sorted the axons along the two axes
 to form a grid of cells. However, the arrangement does not span the full
-tectum. Fig.\,\ref{f:ch}F shows two pattern metrics; $\epsilon$ is the mean
+tectum. Fig.\,\ref{f:G}F shows two pattern metrics; $\epsilon$ is the mean
 distance error of the axon centroids, given by:
@@
 \begin{equation}\label{e:eps}
-\epsilon = \sqrt{\frac{\sum_k (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
+\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
 \end{equation}
@@
 being qualitatively more disordered than the other. For this reason, we also
-use the number of crossings in the fish net plot of Fig.\,\ref{f:ch}D as a
+use the number of crossings in the fish net plot of Fig.\,\ref{f:G}D as a
 second metric, called $\eta$, which tends towards an expected target number of
 crossings, $\eta_t$ as the pattern approaches the target pattern.  For the wildtype
-pattern, $\eta_t=0$, as in Fig.\,\ref{f:ch}E, although note that in some patterns,
-$\eta_t$ is non-zero. Fig.\,\ref{f:ch}H shows selected
+pattern, $\eta_t=0$, as in Fig.\,\ref{f:G}E, although note that in some patterns,
+$\eta_t$ is non-zero. Fig.\,\ref{f:G}H shows selected
 axons (one from the centre and one from each corner of the retina) and their
@@
 temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-\label{f:ch}
+\label{f:G}
 \end{figure}
@@
 (Eq.\,\ref{e:tecligexp}) was chosen arbitrarily. To expand the pattern in
-Fig.\,\ref{f:ch}D so that it covers the entire tectum, we needed only to
+Fig.\,\ref{f:G}D so that it covers the entire tectum, we needed only to
 adjust the tectal ligand expression, substituting the alternative exponential
@@
 
-The result at $t=300$ is shown in Figs.\,\ref{f:chalt}A
-and \ref{f:chalt}B. The axon centroids are arranged across the full tectal
+The result at $t=300$ is shown in Figs.\,\ref{f:G0}A
+and \ref{f:G0}B. The axon centroids are arranged across the full tectal
 surface.  The arrangement is slightly disordered around the middle of the
@@
 found easily by hand tuning) allows a well ordered map to form for the
-wildtype system. Fig.\,\ref{f:chalt}C shows the error metric, $\epsilon$,
-which is much closer to 0 with the new ligand expression \textbf{Fix graph}.
+wildtype system. Fig.\,\ref{f:G0}C shows the error metric, $\epsilon$,
+which is much closer to 0 with the new ligand expression.
 
@@
  centroids \textbf{C} Five selected axons.}
-\label{f:chalt}
+\label{f:G0}
 \end{figure}
@@
 
-Fig\,\ref{f:chsurg} shows the results of 6 simulated surgical
+Fig\,\ref{f:Gsurg} shows the results of 6 simulated surgical
 manipulations. In A, a patch of tectal tissue has been cut out and rotated by
@@
 
-Fig.\,\ref{f:chsurg}C shows the result of the patch swap manipulation, in
+Fig.\,\ref{f:Gsurg}C shows the result of the patch swap manipulation, in
 which 12$\times$4 element grafts, one from the caudal side of the tectum and one from
@@
 
-The retinal ablation shown in Fig.\,\ref{f:chsurg}D is also unsuccessful, with
+The retinal ablation shown in Fig.\,\ref{f:Gsurg}D is also unsuccessful, with
 the axons from the surviving nasal retina (the temporal half was removed)
@@
 and axons from the nasal retina are pushed to the caudal-most side of the
-remaining tectum (Fig.\,\ref{f:chsurg}E). A similar result occurs in the
-mismatch experiment (Fig.\,\ref{f:chsurg}F), but in this case there are no
+remaining tectum (Fig.\,\ref{f:Gsurg}E). A similar result occurs in the
+mismatch experiment (Fig.\,\ref{f:Gsurg}F), but in this case there are no
 axons from the temporal retina, which was been ablated and only the axons
@@
 Retinal ablation \textbf{E} Tectal ablation \textbf{F} mismatch ablation}
-\label{f:chsurg}
+\label{f:Gsurg}
 \end{figure}
@@
 chemoaffinity-like effect could account for the surgical manipulations shown
-in Figs.\,\ref{f:trot90}--\ref{f:chsurg}. To confirm that a competitive
+in Figs.\,\ref{f:trot90}--\ref{f:Gsurg}. To confirm that a competitive
 mechanism could operate successfully in combination with the
@@
 %
-\begin{equation}
-\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k \hat{\mathbf{x}}_{kb} W_c
+\begin{equation} \label{e:C}
+\mathbf{C}_b = \frac{1}{|B_{b,c}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb} W_c
 \end{equation}
@@
 $d_{kb}$. $B_{b,c}$ is the set of growth cones within a distance $2r_c$ of
-branch $b$. This description of the competition matches that
-in \citet{simpson_simple_2011}.
+branch $b$ and the total number of branches in the system is the number of RGC
+axons, $n$ multiplied by the number of branches per axon, $M$. This
+description of the competition matches that in \citet{simpson_simple_2011}.
 
@@
 minimize the error metric, $\epsilon$, for the wildtype pattern
-(Fig.\,\ref{f:ch}E) at simulation time 1000. The parameters returned were
+(Fig.\,\ref{f:G}E) at simulation time 1000. The parameters returned were
 $m_{\!_G}$ = 0.003841, $m_c$ = 0.09959 and $r_c$ = 0.39918. Fig.\,\ref{f:GC}
-shows the $\mathbf{GC}$ simulation with these parameters.
-
-% ./scripts/fig_GC_surgical.sh
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GC_surgical.png}
-\caption{$\mathbf{GC}$ model, surgical manipulations}
-\label{f:GCsurg}
-\end{figure}
+shows the $\mathbf{GC}$ simulation with these parameters. Fish net plots at
+$t=40$ and $t=1000$ show that the pattern develops quickly and compares well
+with the expected layout (Fig.\,\ref{f:G}E). We investigated the response of
+this model to the same set of surgical manipulations shown for the
+chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We found that the addition of
+competition leads to a better correspondence to the experimental results than
+for the chemoaffinity-only model (see supplementary Fig.\,\ref{f:GCsurg}).
+
+\subsection*{Receptor-ligand axon-axon interactions (GJ)}
+
+Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), requires a
+global supervisor to provide information to each branch about the location of
+all other branches, we looked for a receptor-ligand interaction that could
+signal between axon branches and enable them to adjust their movements away
+from one another. We investigated the idea that the same receptor-ligand
+interaction that is modelled between RGC axons and the tectum could also act
+as a signal between the axons, because retinal ganglion cells are known to
+express ephrin receptor ligands as well as Eph
+receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$, expressed
+by the RGCs in the model according to
+%
+\begin{equation} \label{e:retlig}
+l_i(x,y) = 1.05 + 0.26 \exp(2.3 u).
+\end{equation}
+%
+We used exactly the same form as the receptor expression given by
+Eq.\,\ref{e:retrcpt}, but with opposite gradient direction so here $u$ is
+substituted by $x$ for $l_0$, $y$ for $l_1$, $(1-x)$ for $l_2$ and by $(1-y)$
+for $l_3$ (see Fig.\,\ref{f:ex}C). We modelled a
+multiplicative (or mass-action?) interaction to give the movement
+vector, $\mathbf{J}_b$, as  follows:
+%
+\begin{equation}
+% No \frac{1}{|B_b|} in this one. Really?
+\mathbf{J}_b = \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
+\end{equation}
+%
+where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
+receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
+given by
+%
+\begin{equation}
+Q_{\!_J}(d_{kb}) = \begin{cases}
+     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
+     0 & d_{kb} > 2r_{\!_J}
+     \end{cases}
+\end{equation}
+%
+The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines whether
+the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$ is the
+expression of ligand $i$ on branch $k$.
+
+Considering just one receptor-ligand pair,
+for instance $r_0$/$l_0$, and defining their interaction as repulsive, an axon
+from the transverse side of the retina, which expresses a high level of
+receptor $r_0$ will interact most strongly with an axon from the nasal side of
+the retina, where $l_0$ is most highly expressed. The interaction with axons
+that originate nearby, on the transverse side will be minimal, as these
+express low levels of $l_0$. Axons originating from the nasal side of the
+retina interact weakly with axons originating nearby.
+
+See Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
 
@@
 \mathbf{I}_b
-= \frac{1}{|B_{b,I}|} \sum_k \sum_i \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
+= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
 / r_{i,k} < 1.1
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2.png}
+\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2_exit_true.png}
 \caption{$\mathbf{GI}$ model}
@@
 
-\subsection*{Receptor-ligand axon-axon interactions (GJ)}
-
-%
-\begin{equation}
-% No \frac{1}{|B_b|} in this one. Really?
-\mathbf{J}_b = \sum_k \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
-\end{equation}
-%
-where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
-receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
-given by
-%
-\begin{equation}
-Q_{\!_J}(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
-     0 & d_{kb} > 2r_{\!_J}
-     \end{cases}
-\end{equation}
-%
-The sign of $Q_J$ (which is dependent on the values of $F_i$) determines whether
-the interaction, $\mathbf{I}_b$, is repulsive or attractive. $l_{i,k}$ is the
-expression of ligand $i$ on branch $k$.
-
-See Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
-
 \subsection*{Robustness to noise}
@@
 \begin{figure}
-\includegraphics[width=0.3\linewidth]{./images/GI_vs_noise_tweaked.png}
-\caption{The mean value of $\eta$ (number of crossings) and $\epsilon$
-(mean distance error), averaged over the period from simulation time $t=500$ to $t=1000$, plotted with respect to the
-noise gain in the $\mathbf{GI}$ model and the wildtype condition (compare with
-Fig.\,\ref{f:GI}). Both metrics indicate that a moderate
+\centering
+\includegraphics[width=0.7\linewidth]{./images/model_vs_noise.png}
+\caption{ \textbf{A} Chemotaxis-only model. The mean value of $\eta$ (number of crossings) and $\epsilon$
+(mean distance error), averaged over the period from simulation time $t=750$ to $t=1000$, plotted with respect to the
+noise gain ($\nu$) in the best $\mathbf{G}$ model and the wildtype condition (compare with
+Fig.\,\ref{f:G0}). Noise for $\nu < 0.5$ is tolerated by the model, but
+disorder increases about this value.
+\textbf{B} Competitive interaction in the $\mathbf{GI}$ model extends the
+range of ligand gradient noise which can be tolerated by the system.
+Both metrics indicate that a moderate
 amount of gradient sampling noise is \emph{beneficial} to the establishment of
@@
 
+\subsection*{Attractive receptor-ligand interactions}
+
+If we switch a receptor-ligand pair to have an attractive interaction, the
+tectal ligand expression should change from exponential to logarithmic.
+
 \section{Discussion}
@@
 \label{f:tmanip}
+\end{figure}
+
+% ./scripts/fig_GC_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GC_surgical.png}
+\caption{$\mathbf{GC}$ model, surgical manipulations. Demonstrating that this
+model reproduces these manipulations.}
+\label{f:GCsurg}
 \end{figure}

### 2022-02-10 paper/paper.tex
More WIP

@@
 %
-The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines whether
-the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$ is the
-expression of ligand $i$ on branch $k$.
-
-Considering just one receptor-ligand pair,
-for instance $r_0$/$l_0$, and defining their interaction as repulsive, an axon
-from the transverse side of the retina, which expresses a high level of
-receptor $r_0$ will interact most strongly with an axon from the nasal side of
-the retina, where $l_0$ is most highly expressed. The interaction with axons
-that originate nearby, on the transverse side will be minimal, as these
-express low levels of $l_0$. Axons originating from the nasal side of the
-retina interact weakly with axons originating nearby.
+The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines
+whether the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$
+is the expression of ligand $i$ on branch $k$.
+
+Considering just one receptor-ligand pair, for instance $r_0$/$l_0$, and
+defining their interaction as repulsive, an axon from the transverse side of
+the retina, which expresses a high level of receptor $r_0$ will interact most
+strongly with an axon from the nasal side of the retina, where $l_0$ is most
+highly expressed. The interaction with axons that originate nearby, on the
+transverse side will be minimal, as these express low levels of $l_0$. Axons
+originating from the nasal side of the retina also interact most strongly with
+axons originating nearby and in general the $r_0$/$l_0$ interaction is weaker
+between a nasal axon and a second retinal axon. However, when the opposing
+receptor/ligand pair ($r_2$/$l_2$) interaction is summed with that of
+$r_0$/$l_0$, a more even interaction is found, with interactions between axons
+from extreme sides of the retina being the strongest. (Poss.~supp.~figure
+based on the Octave script J.m).
+
+We performed the same simulated annealing process as for the
+$\mathbf{GC}$-model to find the parameters best able to recover the wildtype
+pattern and found $m_{\!_G} = 0.00047$, $m_{\!_J} = 0.00089$ and $r_{\!_J} = 0.341$.
 

### 2022-02-10 misc/J.m

@@ -28,15 +28,16 @@ plot (x, _r0); % exp for receptors on retina
 hold on;
 plot (x, _l0); % ligands
 plot (x, _r0 .* _l0, 'color', colour.orchid2, 'linestyle', '--');
-plot (x(1:3), _r0(1) .* _l0(1:3), 'ro-');
-plot ([x(1),x(1)], [_r0(2),_r0(1).*_l0(1)], 'ro-');
+plot (x, _r0(1) .* _l0, 'r-');
+plot ([x(1),x(1)], [_r0(2),_r0(1).*_l0(1)], 'r-');
 plot (x(3:7), _r0(5) .* _l0(3:7), 'b');
 plot ([x(5),x(5)], [_r0(5),_r0(5).*_l0(5)], 'b-');
 plot (x(7:11), _r0(9) .* _l0(7:11), 'g');
 plot ([x(9),x(9)], [_r0(9),_r0(9).*_l0(9)], 'g-');
 plot (x(11:15), _r0(13) .* _l0(11:15), 'm');
 plot ([x(13),x(13)], [_r0(13),_r0(13).*_l0(13)], 'm-');
-plot (x(15:19), _r0(17) .* _l0(15:19), 'k');
+%%plot (x(15:19), _r0(17) .* _l0(15:19), 'k');
+plot (x, _r0(17) .* _l0, 'k');
 plot ([x(17),x(17)], [_r0(17),_r0(17).*_l0(17)], 'k-');
 lg = legend (['rcpt0';'lgnd0'],'Location','North');
 
@@ -47,8 +48,9 @@ plot (x, _r1);
 hold on;
 plot (x, _l1); % ligands
 plot (x, _r1 .* _l1, 'color', colour.orchid2, 'linestyle', '--');
-plot (x(1:3), _r1(1) .* _l1(1:3), 'ro-');
-plot ([x(1),x(1)], [_r1(2),_r1(1).*_l1(1)], 'ro-');
+%%plot (x(1:3), _r1(1) .* _l1(1:3), 'r-');
+plot (x, _r1(1) .* _l1, 'r-');
+plot ([x(1),x(1)], [_r1(2),_r1(1).*_l1(1)], 'r-');
 plot (x(3:7), _r1(5) .* _l1(3:7), 'b');
 plot ([x(5),x(5)], [_r1(5),_r1(5).*_l1(5)], 'b-');
[…]

### 2022-02-10 paper/paper.tex
WIP moving to laptop

@@
 
-\color{grey}
-\textbf{New introduction:} Simpson and Goodhill presented a model in which the
-chemoaffinity mechanism is non-biological. We offer a model in which a
-competition mechanism counteracts the drive of axons along the ephrin
-gradient. Will we need the axon-axon interaction mechanism based on
-non-same-Eph repulsion?
-
-\textbf{Another approach:} We present a model of gradient-following based on graded
-interactions between the growth cones of retinal ganglion cell axons and the
-tectal surface. We show that a combination of exponentially graded expression
-in the retina \citep{reber_relative_2004} and exponentially (for repulsively
-interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
-interacting receptor/ligands) graded expression on the tectum permits a purely
-local model to reproduce the majority of reported phenomena associated with
-retinotectal organisation.
-
-There have been many approaches to modelling the organisation of the
-retinotectal projection. Many of these models are wholly or partially
-phenomenological.
+%% \color{grey}
+%% \textbf{New introduction:} Simpson and Goodhill presented a model in which the
+%% chemoaffinity mechanism is non-biological. We offer a model in which a
+%% competition mechanism counteracts the drive of axons along the ephrin
+%% gradient. Will we need the axon-axon interaction mechanism based on
+%% non-same-Eph repulsion?
+
+%% \textbf{Another approach:} We present a model of gradient-following based on graded
+%% interactions between the growth cones of retinal ganglion cell axons and the
+%% tectal surface. We show that a combination of exponentially graded expression
+%% in the retina \citep{reber_relative_2004} and exponentially (for repulsively
+%% interacting receptor ligand pairs) \emph{or} logarithmically (for attractively
+%% interacting receptor/ligands) graded expression on the tectum permits a purely
+%% local model to reproduce the majority of reported phenomena associated with
+%% retinotectal organisation.
+
+%% There have been many approaches to modelling the organisation of the
+%% retinotectal projection. Many of these models are wholly or partially
+%% phenomenological.
 
@@
 
-\color{grey}
-This means that growth cones of approximate
-radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a radius $r$ =
-0.0025 in the arbitrary units of the simulation. For a realistic simulation
-based on mouse, the approximate RGC density in the retina is around 44000
-cells per retina \citep{jeon_major_1998}. This sets the side length of the
-retina to $\sqrt{44000} \approx 200$. This results in a simulation of
-prodigious computational size, which although tractable (10 s per step,
-requiring several hours, rather than days to run) is not convenient. To
-maintain the relationship between growth cone size and cell density, we permit
-a smaller number of growth cones ($400\times8=3200$) but keep the fractional
-area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
-per axon) or 6.9 (if there are 8 cones per axon).
-
-\color{black}
+%% This means that growth cones of approximate
+%% radius 5 $\mu$m = 0.005 mm \citep{goodhill_can_2016} have a radius $r$ =
+%% 0.0025 in the arbitrary units of the simulation. For a realistic simulation
+%% based on mouse, the approximate RGC density in the retina is around 44000
+%% cells per retina \citep{jeon_major_1998}. This sets the side length of the
+%% retina to $\sqrt{44000} \approx 200$. This results in a simulation of
+%% prodigious computational size, which although tractable (10 s per step,
+%% requiring several hours, rather than days to run) is not convenient. To
+%% maintain the relationship between growth cone size and cell density, we permit
+%% a smaller number of growth cones ($400\times8=3200$) but keep the fractional
+%% area of the branches at 0.86 ($44000 \times \pi r^2$) (if there is one cone
+%% per axon) or 6.9 (if there are 8 cones per axon).
+
 Each discrete, square element in Fig.\,\ref{f:ex}B projects one RGC axon with $M=4$
@@
 mechanism could operate successfully in combination with the
-gradient-following chemoaffinity mechanism described above, we introduced
-three competition mechanisms and investigated their behaviour.
-
-These were: space-based competition (`$\mathbf{C}$'), as described
-in \citet{simpson_simple_2011}; a receptor-receptor signalling mechanism
-(`$\mathbf{I}$'), similar to that described in \citet{simpson_simple_2011},
-and a receptor-ligand mass-action signalling model
-(`$\mathbf{J}$'). The movement vector (Eq.\,\ref{e:mv}) becomes
-%
+gradient-following chemoaffinity mechanism described above, we first
+implemented the `space based' competition of \citet{simpson_simple_2011},
+updating the movement vector with a term $m_c \mathbf{C}_b$:
 \begin{equation} \label{e:mv2}
-\mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b +  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
-\end{equation}
-%
-
-Although two of these competitive mechanisms were investigated
-by \citet{simpson_simple_2011}, the first (space-based competition) offers no
-mechanism by which axon growth cones are able to communicate and the second
-mechanism was implemented only for a single axis in the model (EphA receptors,
-expressed in a gradient along the XX retinal axis). We wondered whether a
-\emph{single} signalling mechanism working alongside the `untuned'
-chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
-successfully reproduce the wildtype and manipulation results. Alongside the
-same chemoaffinity mechanism, we studied each competitive mechanism in turn,
-setting $m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model,
-$m_c=m_{\!_J}=0$ to give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a
-`$\mathbf{GJ}$' model.
-
-\subsection*{Chemoaffinity plus space-based competition (GC)}
-
-Space based competition was implemented as a repulsive effect acting between
-any two branches, whether they originated from different retinal ganglion
-cells or from the same cell. An assumption is that there is some kind of
-self-other signal which acts to discourage two cells from approaching each
-other. A competition movement vector, $\mathbf{C}_b$, acting on branch $b$, is
-defined as a proximity-weighted sum of the repulsion due to all nearby branches:
+ \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b + 0.5 \mathbf{B}_b
+ %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+\end{equation}
+where $m_c$ is simply a scalar parameter.
+%
+%
+%% introduced
+%% three competition mechanisms and investigated their behaviour.
+%
+%% These were: space-based competition (`$\mathbf{C}$'), as described
+%% in \citet{simpson_simple_2011}; a receptor-receptor signalling mechanism
+%% (`$\mathbf{I}$'), similar to that described in \citet{simpson_simple_2011},
+%% and a receptor-ligand mass-action signalling model
+%% (`$\mathbf{J}$'). The movement vector (Eq.\,\ref{e:mv}) becomes
+%% %
+%% \begin{equation} \label{e:mv2}
+%% \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b +  m_c \mathbf{C}_b +  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+%% \end{equation}
+%% %
+%
+%% Although two of these competitive mechanisms were investigated
+%% by \citet{simpson_simple_2011}, the first (space-based competition) offers no
+%% mechanism by which axon growth cones are able to communicate and the second
+%% mechanism was implemented only for a single axis in the model (EphA receptors,
+%% expressed in a gradient along the XX retinal axis). We wondered whether a
+%% \emph{single} signalling mechanism working alongside the `untuned'
+%% chemoaffinity with ligand expression given by Eq.\,\ref{e:tecligexp} could
+%% successfully reproduce the wildtype and manipulation results. Alongside the
+%% same chemoaffinity mechanism, we studied each competitive mechanism in turn,
+%% setting $m_{\!_I}=m_{\!_J}=0$ to give a `$\mathbf{GC}$' model,
+%% $m_c=m_{\!_J}=0$ to give a `$\mathbf{GI}$' model or $m_c=m_{\!_I}=0$ to give a
+%% `$\mathbf{GJ}$' model.
+%
+%%\subsection*{Chemoaffinity plus space-based competition (GC)}
+%
+The competition movement vector, $\mathbf{C}_b$, introduces a repulsive effect
+acting between any two branches, whether they originated from different
+retinal ganglion cells or from the same cell. An assumption is that there is
+some kind of self-other signal which acts to discourage two cells from
+approaching each other. $\mathbf{C}_b$ is defined as a proximity-weighted sum
+of the repulsion due to all nearby branches:
 %
@@
 axons, $n$ multiplied by the number of branches per axon, $M$. This
-description of the competition matches that in \citet{simpson_simple_2011}.
+description of the competition matches that
+in \citet{simpson_simple_2011}. Because the model incorporates a chemoaffinity
+effect and a competition effect, we called it the `$\mathbf{GC}$ model'.
 
@@
 The model defined by the receptor and ligand expressions of Eqs.\,\ref{e:retrcpt}
-\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} with $m_{\!_I}=m_{\!_J}=0$ has three
+\& \ref{e:tecligexp} along with Eq.\,\ref{e:mv2} has three
 free parameters; $m_{\!_G}$, $m_c$ and $r_c$. To find the best combination of
@@
 
-\subsection*{Receptor-ligand axon-axon interactions (GJ)}
+\subsection*{Receptor-ligand axon-axon interactions}
 
@@
 from one another. We investigated the idea that the same receptor-ligand
-interaction that is modelled between RGC axons and the tectum could also act
-as a signal between the axons, because retinal ganglion cells are known to
-express ephrin receptor ligands as well as Eph
-receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$, expressed
-by the RGCs in the model according to
+interaction that leads to gradient-based movement of RGC axons on the tectum
+could also act as a signal between the axons. This is plausible, because
+retinal ganglion cells are known to express ephrin receptor ligands as well as
+Eph receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$,
+expressed by the RGCs in the model according to
 %
@@
 substituted by $x$ for $l_0$, $y$ for $l_1$, $(1-x)$ for $l_2$ and by $(1-y)$
-for $l_3$ (see Fig.\,\ref{f:ex}C). We modelled a
-multiplicative (or mass-action?) interaction to give the movement
-vector, $\mathbf{J}_b$, as  follows:
+for $l_3$ (see Fig.\,\ref{f:ex}C).
+%
+We combined chemotaxis with a multiplicative (or mass-action?) interaction,
+$\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the movement
+vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
+%
+\begin{equation} \label{e:mv3}
+ \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+ %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+\end{equation}
+%
+$\mathbf{J}_b$ is defined as
 %
@@
 We performed the same simulated annealing process as for the
-$\mathbf{GC}$-model to find the parameters best able to recover the wildtype
-pattern and found $m_{\!_G} = 0.00047$, $m_{\!_J} = 0.00089$ and $r_{\!_J} = 0.341$.
+$\mathbf{GC}$ model to find the parameters best able to recover the wildtype
+pattern and found $m_{\!_G} = 0.00047$, $m_{\!_J} = 0.00089$ and $r_{\!_J} =
+0.341$.
 
@@
 
-\subsection*{Receptor-receptor axon-axon interactions (GI)}
+\subsection*{Receptor-receptor axon-axon interactions}
 

### 2022-02-10 misc/J.m

@@ -21,59 +21,74 @@ _exp2_ki_kd = _exp2_ki - kd;
 _exp2_kd = _exp2 - kd;
 
 %% The J effect
-_r0 = _exp;
-_l0 = flip(_exp);
+_r0 = flip(_exp);
+_l0 = _exp;
 figure(1); clf;
-plot (x, _r0); % exp for receptors on retina
+plot (x, _r0, 'linestyle', ':'); % exp for receptors on retina
 hold on;
-plot (x, _l0); % ligands
-plot (x, _r0 .* _l0, 'color', colour.orchid2, 'linestyle', '--');
+plot (x, _l0, 'linestyle', ':'); % ligands
+%%plot (x, _r0 .* _l0, 'color', colour.orchid2, 'linestyle', ':');
 plot (x, _r0(1) .* _l0, 'r-');
-plot ([x(1),x(1)], [_r0(2),_r0(1).*_l0(1)], 'r-');
+plot ([x(1),x(1)], [_r0(2),_r0(1).*_l0(1)], 'ro-');
 plot (x(3:7), _r0(5) .* _l0(3:7), 'b');
-plot ([x(5),x(5)], [_r0(5),_r0(5).*_l0(5)], 'b-');
+plot ([x(5),x(5)], [_r0(5),_r0(5).*_l0(5)], 'bo-');
 plot (x(7:11), _r0(9) .* _l0(7:11), 'g');
-plot ([x(9),x(9)], [_r0(9),_r0(9).*_l0(9)], 'g-');
+plot ([x(9),x(9)], [_r0(9),_r0(9).*_l0(9)], 'go-');
 plot (x(11:15), _r0(13) .* _l0(11:15), 'm');
-plot ([x(13),x(13)], [_r0(13),_r0(13).*_l0(13)], 'm-');
-%%plot (x(15:19), _r0(17) .* _l0(15:19), 'k');
-plot (x, _r0(17) .* _l0, 'k');
-plot ([x(17),x(17)], [_r0(17),_r0(17).*_l0(17)], 'k-');
+plot ([x(13),x(13)], [_r0(13),_r0(13).*_l0(13)], 'mo-');
+plot (x(15:19), _r0(17) .* _l0(15:19), 'c');
+%%plot (x, _r0(17) .* _l0, 'c');
+plot ([x(17),x(17)], [_r0(17),_r0(17).*_l0(17)], 'co-');
[…]

### 2022-02-21 paper/paper.tex
Correction to comparison in paper
Work in progress

@@
 % Use the article doc class, with an 11 pt basic font size
-\documentclass[11pt, a4paper, draft]{article}
-%\documentclass[11pt, a4paper]{article}
+%\documentclass[11pt, a4paper, draft]{article}
+\documentclass[11pt, a4paper]{article}
 
@@
 
-\subsection*{Receptor-ligand axon-axon interactions}
+\subsection*{Axon-axon interactions}
 
@@
 signal between axon branches and enable them to adjust their movements away
-from one another. We investigated the idea that the same receptor-ligand
+from one another.
+
+% First, look at I, from Simpson and Goodhill.
+First, we took inspiration from Simpson and Goodhill's `I' interaction, which
+is defined as:
+
+%We modelled interactions between axon growth cones due to signalling between
+%receptors of the same type. For these interactions, repulsion caused a
+%movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
+%branch $k$ to branch $b$ causing them to move further apart; attraction caused
+%the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
+%is given by
+%
+\begin{equation}
+\mathbf{I}_b
+= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
+/ r_{i,k} > 1.1
+\end{equation}
+%
+the weight here is defined as a function of $r_{\!_I}$:
+%
+\begin{equation}
+W_{\!_I}(d_{kb}) = \begin{cases}
+      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
+     0 & d_{kb} > 2r_{\!_I}
+     \end{cases}
+\end{equation}
+$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
+
+ (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.
+
+% Next, replicate with receptor-ligand interaction.
+
+% OR
+
+% Look at receptor-ligand interaction, and just point out that the
+
+Next, we investigated the idea that the same receptor-ligand
 interaction that leads to gradient-based movement of RGC axons on the tectum
@@
 
-\subsection*{Receptor-receptor axon-axon interactions}
-
-We modelled interactions between axon growth cones due to signalling between
-receptors of the same type. For these interactions, repulsion caused a
-movement of branch $b$ along a unit vector, $\hat{\mathbf{x}}_{kb}$, from
-branch $k$ to branch $b$ causing them to move further apart; attraction caused
-the opposite movement. This interaction, $\mathbf{I}_b$, acting on branch $b$,
-is given by
-%
-\begin{equation}
-\mathbf{I}_b
-= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
-/ r_{i,k} < 1.1
-\end{equation}
-%
-the weight here is defined as a function of $r_{\!_I}$:
-%
-\begin{equation}
-W_{\!_I}(d_{kb}) = \begin{cases}
-      1 - \frac{d_{kb}}{2r_{\!_I}}   & d_{kb} \leq 2r_{\!_I} \\
-     0 & d_{kb} > 2r_{\!_I}
-     \end{cases}
-\end{equation}
-$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
-
-See Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.
+%%\subsection*{Receptor-receptor axon-axon interactions}
 

### 2022-02-21 misc/J.m

@@ -16,21 +16,6 @@ x = [0:0.05:1];
 
 _exp = 0.26 .* exp (2.3.*x) + 1.05;
 _deriv_exp = 0.26 .* 2.3 .* exp (2.3.*x);
-## _exp2 = 0.26 .* exp (1.1.*x) + 1.05;
-## _log = 2.32 + 1.29 .* log (2.3 .* (x+0.2));
-## _quad = 1.31 + 2.333 .* x .* x;
-## _lin = 1.31 + 2.333 .* x;
-
-ki = 1.1
-kd = 0.6
-
-_exp_ki = _exp + ki;
-_exp_ki_kd = _exp_ki - kd;
-_exp_kd = _exp - kd;
-
-_exp2_ki = _exp2 + ki;
-_exp2_ki_kd = _exp2_ki - kd;
-_exp2_kd = _exp2 - kd;
 
 %% The J effect
 _r0 = flip(_exp);
@@ -120,6 +105,55 @@ plot ([x(21),x(21)], [0,_r2(21).*_l2(21)+_r0(21).*_l0(21)], 'ko-');
 lg3 = legend (['r0[0]xl0 + r2[0]xl2 combined interaction (competitive)';'r0[.2]xl0 + r2[.2]xl2 combined interaction (competitive)';'etc'],'Location','North');
 strlbl = ['Temporal ------------------------> Nasal'];
 ##ylim([0,20])
+xlabel(strlbl)
+ylabel('Expression/Interaction')
+title('Summed mass-action interaction')
+
+subplot(2,4,4);
+hold on;
+###
+plot (x, _r0, 'linestyle', '--'); % exp for receptors on retina
+hold on;
+plot (x, _l0, 'linestyle', '--'); % ligands
+plot (x, _r0(1) .* _l0, 'r-');
+plot (x, _r0(5) .* _l0, 'b');
+plot (x(7:11), _r0(9) .* _l0(7:11), 'g');
+plot (x, _r0(13) .* _l0, 'm');
+plot (x(15:19), _r0(17) .* _l0(15:19), 'c');
+
+plot ([x(1),x(1)], [_r0(1),_r0(1).*_l0(1)], 'ro-');
+plot ([x(5),x(5)], [_r0(5),_r0(5).*_l0(5)], 'bo-');
+plot ([x(9),x(9)], [_r0(9),_r0(9).*_l0(9)], 'go-');
[…]

### 2022-02-21 sim/agent/anneal1.cpp

@@ -214,15 +214,15 @@ int main (int argc, char **argv)
         } else if (params.back() == "r_j") {
             param_ranges.push_back ({0.001, 0.5});
         } else if (params.back() == "s") {
-            param_ranges.push_back ({0.01, 0.99});
+            param_ranges.push_back ({0.01, 6});
         } else if (params.back() == "m_g") {
-            param_ranges.push_back ({0.0001, 0.01});
+            param_ranges.push_back ({0.0001, 0.004});
         } else if (params.back() == "m_c") {
             param_ranges.push_back ({0.01, 0.1});
         } else if (params.back() == "m_i") {
             param_ranges.push_back ({0.01, 0.8});
         } else if (params.back() == "m_j") {
-            param_ranges.push_back ({0.00001, 0.001});
+            param_ranges.push_back ({0.001, 0.4});
         }
     }
     morph::vVector<float> param_range_max(param_ranges.size(), float{0});

### 2022-02-21 sim/agent/branch.h

@@ -116,8 +116,7 @@ public:
         return G;
     }
 
-    static constexpr bool distance_based_weight_for_J = false;
-    // Summing up mass-action interactinos for J was the old way. Alternative is to
+    // Summing up mass-action interactions for J was the old way. Alternative is to
     // compete if any interaction is supra-threshold.
     static constexpr bool sum_up_mass_action_interactions = false;
 
@@ -182,8 +181,8 @@ public:
         // binding of receptors and ligands.
         //
         // One important difference is the lack of a distance based weighting (cf C and
-        // I effects). I have tried adding this, and as far as I can tell, it makes
-        // little difference to the competition.
+        // I effects). In early 2022, I tried adding this, and as far as I could tell,
+        // it makes little difference to the result.
         morph::Vector<T,N> QJar;
         for (size_t i = 0; i < N; ++i) {
             QJar[i] = (source_tissue->forward_interactions[i] == interaction::repulsion ? T{1} :
@@ -192,27 +191,19 @@ public:
         QJar *= kp->lgnd * this->rcpt;
 
         if constexpr (sum_up_mass_action_interactions == true) {
-            T QJ = QJar.sum()/N; // The key is not to SUM, but to have, as with I, a
-            // threshold, above which competition is triggered.
-
-            if constexpr (distance_based_weight_for_J == true) {
-                // Incorporate a distance based weight for each contribution [the 1 - d / (2 r_j)]
-                J += kb * QJ * (d <= this->two_r_j ? (T{1} - d/this->two_r_j) : T{0});
-            } else {
-                // Just add to J. This will require smaller m_j parameter value than with
-                // a weight-basis. I prefer this as how would a cell know the distance?
-                J += kb * QJ * (d <= this->two_r_j ? T{1} : T{0});
-            }
-
+            // Original coding of the 'J' competition
+            T QJ = QJar.sum()/N;
+            J += kb * QJ * (d <= this->two_r_j ? T{1} : T{0});
             rtn[2] = (d <= this->two_r_j && QJ > T{0}) ? true : false;
         } else {
             // Add to competition if any of the mass-action interactions is true.
             // Note: use <= operator which is true if EVERY member of the Vector is <= threshold
             bool subthreshold = false;
-            if ((subthreshold = QJar.abs() <= this->s) == false) {
-                T QJ = QJar.sum()/N; // Sum the interactions to modulate the competition strength?
[…]

### 2022-02-21 sim/agent/tissue.h

@@ -62,7 +62,8 @@ enum class expression_form
     quad, // A quadratic function
     exp,  // An exponential function
     log,  // A logarithmic function
-    exp2  // A slightly different exponential function, used to demonstrate dependence of chemo-only model on expressions
+    exp2,  // A slightly different exponential function, used to demonstrate dependence of chemo-only model on expressions
+    exp_inv // 1/exp
 };
 
 // Which sense is an expression pattern becoming stronger?
@@ -459,6 +460,8 @@ struct guidingtissue : public tissue<T>
         case expression_form::exp2:
             rtn = this->exponential_expression2 (x);
             break;
+        case expression_form::exp_inv:
+            rtn = T{1}/this->exponential_expression (x);
         default:
             break;
         }

### 2022-02-23 paper/paper.tex
Paper WIP

@@
 % Use the article doc class, with an 11 pt basic font size
-%\documentclass[11pt, a4paper, draft]{article}
-\documentclass[11pt, a4paper]{article}
+\documentclass[11pt, a4paper, draft]{article}
+%\documentclass[11pt, a4paper]{article}
 
@@
 this model to the same set of surgical manipulations shown for the
-chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We found that the addition of
-competition leads to a better correspondence to the experimental results than
-for the chemoaffinity-only model (see supplementary Fig.\,\ref{f:GCsurg}).
+chemoaffinity-only model in Fig.\,\ref{f:Gsurg}. We confirmed that the
+addition of competition leads to a better correspondence to the experimental
+results than for the chemoaffinity-only model (see supplementary
+Fig.\,\ref{f:GCsurg}).
 
@@
 % First, look at I, from Simpson and Goodhill.
-First, we took inspiration from Simpson and Goodhill's `I' interaction, which
-is defined as:
+First, we took inspiration from Simpson and Goodhill's model for the
+interaction between EphA receptors, which is defined (with receptor type $i$
+standing for EphA) as:
 
@@
 \mathbf{I}_b
-= \frac{1}{|B_{b,I}|} \sum_k^{nM} \sum_i^N \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
+= \frac{1}{|B_{b,I}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_I} \qquad \mathrm{if}~r_{i,b}
 / r_{i,k} > 1.1
@@
 %
-the weight here is defined as a function of $r_{\!_I}$:
+$r_{i,k}$ is the expression of receptor $i$ on branch $k$ and the weight here
+is defined as a function of the distance, $d_{kb}$, between branches:
 %
@@
 \end{equation}
-$r_{i,k}$ is the expression of receptor $i$ on branch $k$.
-
- (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.
+%
+$B_{b,I}$ is the set of growth cones within a distance $2r_{\!_I}$ of branch
+$b$. $\hat{\mathbf{x}}_{kb}$ is as defined in Eq.\,\ref{e:C}.
+
+Simpson and Goodhill introduced this interaction as an additive modification
+to the main, distance based competition mechanism of Eq.\,\ref{e:C} (adding a
+term $m_{\!_I} \mathbf{I}_b$ to the equation). They introduced an interaction
+varying along the temporal-nasal retinal axis and mediated by one receptor,
+EphA, for which there was experimental data supporting a repulsive
+effect. They noted that the use of a receptor ratio (comparing the ratio with
+a `discrimination limit', above which repulsion would be activated) is
+equivalent to a multiplicative interaction between EphA receptors and ephrin-A
+ligands with the strong assumption that ligand expression is the inverse of
+receptor expression, i.e. that $l_i = 1/r_i$.
+
+We asked i) whether this mechanism could \emph{replace} the non-biological
+competition, C, (making necessary its extension to the dorsal-ventral retinal
+axis) and ii) if cast in terms of receptor-ligand interactions, could the
+strong ligand expression assumption be relaxed?
+
+% (supp?) Fig.\,\ref{f:GI} and Fig.\,\ref{f:GIsurg}.
 
@@
 % Look at receptor-ligand interaction, and just point out that the
+
+In order to replace the distance-based competition, $\mathbf{C}_b$, we first
+suggest that each receptor ligand pair involved in the retinotectal system
+could mediate a repulsive interaction. DO WE NEED TO? Taken alone, the EphA
+interaction between one branch, $b$, and another, $k$, will cause a repulsion
+if $k$ originated more \textbf{nasally?} than $b$. What about a branch
+originating on the temporal side of $b$? This could be handled by the
+interaction between $k$ and $b$ and may not require an additional interaction.
 

### 2022-02-24 paper/paper.tex
Thursday work

@@
 \includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Simulated retinal and tectal signalling molecule expression. Four
-receptor types are expressed, and four corresponding ligand types. Receptors
-of type $i$ are activated by ligands of type $i$.
-%
-\textbf{A} Retinal receptor expression, $r_i$. Four receptors are expressed and
+\caption{Simulated retinal and tectal signalling molecule
+expression. \textbf{A} A schematic of retino-tectal mapping.
+%
+\textbf{B} Retinal receptor expression, $r_i$. Four receptors are expressed and
 displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue;
@@
 %
-\textbf{B} Retinal ligand expression, $l_i$, is expressed in complimentary
+\textbf{C} Retinal ligand expression, $l_i$, is expressed in complimentary
 gradients \citep{hornberger_modulation_1999}.
 %
-\textbf{C} Tectal ligand expression $L_i$. Tectal ligand expression is shown
+\textbf{D} Tectal ligand expression $L_i$. Tectal ligand expression is shown
 in an exponential form,
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2.png}
+\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2_exit_true.png}
 \caption{The pure gradient-following model. \textbf{A} The origin position
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig3.png}
+\includegraphics[width=\linewidth]{./images/j4_eE_G_wt_fig3_exit_true.png}
 \caption{By tuning the tectal ligand expression (replacing Eq.\,\ref{e:tecligexp} with
@@
 % graph_layout::g
-%
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot90_fig4row1.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecrot180_fig4row2.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecswap_fig4row3.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_retablat_fig4row4.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_tecablat_fig4row5.json
-% ./build/sim/agent/agent1 configs/simpler/m_eE_G.json  configs/simpler/e_mismatch_fig4row6.json
-%
 % (see scripts/fig_G_surgical.sh)
@@
 effect and a competition effect, we called it the `$\mathbf{GC}$ model'.
-
-% ./build/sim/agent/agent1 configs/simpler/m_ee_GC_best_1.json configs/simpler/e_wt_figcomp1.json
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_wt_figcomp1.png}
-\caption{$\mathbf{GC}$ model}
-\label{f:GC}
-\end{figure}
 
@@
 global supervisor to provide information to each branch about the location of
-all other branches, we looked for a receptor-ligand interaction that could
-signal between axon branches and enable them to adjust their movements away
-from one another.
+all other branches (But \textbf{fixme}, arguably, so does the GJ model now!),
+we looked for a receptor-ligand interaction that could signal between axon
+branches and enable them to adjust their movements away from one another.
 
@@
 First, we took inspiration from Simpson and Goodhill's model for the
-interaction between EphA receptors, which is defined (with receptor type $i$
-standing for EphA) as:
+interaction between EphA receptors, which is defined (with receptor $r_i$
+representing EphA) as:
 
@@
 In order to replace the distance-based competition, $\mathbf{C}_b$, we first
-suggest that each receptor ligand pair involved in the retinotectal system
-could mediate a repulsive interaction. DO WE NEED TO? Taken alone, the EphA
-interaction between one branch, $b$, and another, $k$, will cause a repulsion
-if $k$ originated more \textbf{nasally?} than $b$. What about a branch
-originating on the temporal side of $b$? This could be handled by the
-interaction between $k$ and $b$ and may not require an additional interaction.
-
+point out that many of the ligands, $l_i$, expressed on the tectum are also
+expressed by retinal ganglion cells and so where there is signalling between
+tectal and RG cells, there is also the possibility for inter-RGC
+signalling. Furthermore, if a receptor-ligand interaction on the tectum leads
+to growth cones moving away from a region of high ligand expression, then the
+equivalent axon-axon interaction would also be expected to be repulsive.
+
+Taken alone, the EphA forward interaction between one branch, $b$, and
+another, $k$, will cause a repulsion if $k$ originated more nasally than
+$b$. A branch originating on the \emph{temporal} side of $b$ will cause no
+repulsion of $b$. However, although $k$ does not induce repulsion in $b$, $b$
+should induce a repulsion in $k$ and thus, it could be argued that only a
+single ephrin-A ligand gradient should be required in the retina to ensure
+that repulsive signals are transmitted between axons whose cell bodies
+originte far apart in the retina. However, there is an asymmetry in the
+interaction. Consider a cluster of growth cones, $k$, originating from the
+nasal retina near to a single cone, $b$, from the temporal retina and assume
+that $r_b/r_k>1.1$. The single cone is affected by the sum of 3 interactions
+and has a movement of 3 units added to its movement vector (in the temporal
+direction). Now reverse the originating locations so that the cluster, $k$,
+originates temporally and the individual cone, $b$, comes from the nasal
+retina. Now $r_k/r_b>1.1$ and each member of the cluster experiences an
+addition of just one unit to its movement vector and the cones move apart at a
+speed that is only $1/3$ of that for the original origins.
+
+To avoid this asymmetry, we allow two receptor-ligand pairs, which are
+expressed in opposing directions, to contribute to axon-axon interactions,
+ensuring that interactions between single growth cones and clusters are not
+dependent on their origin locations. Fig.\,\ref{f:expressions}C shows ligand
+expressions representing an ephrin-A like expression ($l_0$) and an opposing
+expression ($l_2$). It remains the case (with a suitable choice of threshold,
+$s$) that branches originating from the same location on the temporal-nasal
+axis will not experience competitive interactions.
+
+Note that it would be possible to postulate two additional retinal ligand
+expression fields (for $l_1$ and $l_3$), but we found this not to be
+necessary; a single opposing pair of fields ensures that axons that originated
+close together (and branches of the same axon) do not repel one another,
+whereas axons that are close on the temporal-nasal axis, but far apart on the
+dorso-ventral axis are generally driven far apart by chemotaxis, so whether
+they compete or not when close together does not affect the behaviour of the
+system.
+
+% Now describe J model, and show the result.
+
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3_exit_true_steps_1500.png}
+\caption{$\mathbf{GJ}$ model}
+\label{f:GJ}
+\end{figure}
+
+% ./scripts/fig_GJ_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
+\caption{$\mathbf{GJ}$ model, surgical manipulations}
+\label{f:GJsurg}
+\end{figure}
+
+%% TO ADD: Figure for the Reber/Brown manipulation on the GJ model.
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_reber.png}
+\caption{$\mathbf{GJ}$ model, Reber et al.~manipulation. Look at THAT!}
+\label{f:GJreber}
+\end{figure}
+
+%% HEREish
 Next, we investigated the idea that the same receptor-ligand
@@
 %%\subsection*{Receptor-receptor axon-axon interactions}
-
-% ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2_exit_true.png}
-\caption{$\mathbf{GI}$ model}
-\label{f:GI}
-\end{figure}
-
-% ./scripts/fig_GI_surgical.sh
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GI_surgical.png}
-\caption{$\mathbf{GI}$ model, surgical manipulations}
-\label{f:GIsurg}
-\end{figure}
 
@@
 
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GC_best_1.json configs/simpler/e_wt_figcomp1_exit_true.json
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_wt_figcomp1_exit_true.png}
+\caption{$\mathbf{GC}$ model}
+\label{f:GC}
+\end{figure}
+
 % ./scripts/fig_GC_surgical.sh
@@
 
-% ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3.png}
-\caption{$\mathbf{GJ}$ model}
-\label{f:GJ}
-\end{figure}
-
-% ./scripts/fig_GJ_surgical.sh
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
-\caption{$\mathbf{GJ}$ model, surgical manipulations}
-\label{f:GJsurg}
-\end{figure}
+% No need to show the GI model
+% ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json
+%\begin{figure}
+%\includegraphics[width=\linewidth]{./images/j4_ee_GI_best_1_wt_figcomp2_exit_true.png}
+%\caption{$\mathbf{GI}$ model}
+%\label{f:GI}
+%\end{figure}
+
+% ./scripts/fig_GI_surgical.sh
+%\begin{figure}
+%\includegraphics[width=\linewidth]{./images/fig_GI_surgical.png}
+%\caption{$\mathbf{GI}$ model, surgical manipulations}
+%\label{f:GIsurg}
+%\end{figure}
 

### 2022-02-24 scripts/fig_GJ_noise.sh

@@ -1,0 +1,10 @@
+#!/bin/sh
+
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.0 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.2 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.4 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.6 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.8 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=1.0 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=1.5 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=2.0 -co:exit=true

### 2022-02-24 scripts/search_GC.sh

@@ -1,0 +1,9 @@
+#!/bin/sh
+
+#
+# Use anneal1 to search for the best parameters for the GC model.
+#
+# m_ee_GC.json contains the starting model paramters. s_GC.json
+# contains the (simulated annealing) search parameters
+#
+./build/sim/agent/anneal1c ./optimization/m_ee_GC.json ./optimization/s_GC.json

### 2022-02-24 scripts/search_GJ.sh

@@ -3,7 +3,7 @@
 #
 # Use anneal1 to search for the best parameters for the GJ model.
 #
-# GJ.json contains the starting model paramteres. s_GJ.json contains
-# the (simulated annealing) search parameters
+# m_ee_GJ.json contains the starting model paramters. s_GJ.json
+# contains the (simulated annealing) search parameters
 #
 ./build/sim/agent/anneal1c ./optimization/m_ee_GJ.json ./optimization/s_GJ.json

### 2022-02-24 sim/agent/anneal1.cpp

@@ -66,7 +66,10 @@ T objfn (Agent1<T, N, branch<T, N>>& model1,
         std::cout << params[i] << " = " << param_values[i] << ", ";
     }
     std::cout << " (Sim count: " << model_sim_count << ")\n";
-    T rtn = m1m.sos.back();
+
+    // Compute the metric.
+    T scaled_nc = T{1} + T{0.01} * m1m.crosscount.back();
+    T rtn = m1m.sos.back() * scaled_nc;
 
     return rtn;
 }
@@ -248,6 +251,8 @@ int main (int argc, char **argv)
         param_range_min[i] = param_ranges[i][0];
         param_range_max[i] = param_ranges[i][1];
     }
+    morph::vVector<float> param_range_diff = param_range_max - param_range_min;
+    morph::vVector<float> param_range_offs = param_range_min / param_range_diff;
 
     optimiser = new morph::Anneal<double>(param_values, param_ranges);
     // Anneal ASA params from sconf:
@@ -406,13 +411,16 @@ int main (int argc, char **argv)
             graph3->append ((float)optimiser->steps, optimiser->f_x_cand, 0);
 
             // Add parameter set to the scattervisual. Scale coords by param_maxes
-            morph::vVector<float> coord = one_over_param_maxes * optimiser->x_cand;
+            morph::vVector<float> x_cand_dbl (optimiser->x_cand.size());
+            std::copy (optimiser->x_cand.begin(), optimiser->x_cand.end(), x_cand_dbl.begin());
+            morph::vVector<float> coord = (x_cand_dbl / param_range_diff) - param_range_offs;
             // Deal with < 3 coords
             size_t sz = coord.size();
             if (sz < 3) {
                 coord.resize (3);
                 for (size_t i = sz; i < 3; ++i) { coord[i] = float{0}; }
             }
+            static constexpr float scatter_max_sz = 0.05f; // How big should the bad blobs be allowed to get?
             // If there are >3 coords, then have a 'start index' and show 3 of
             // them. Allow user to swich the 'start index' so they can show dimensions
             // 0,1,2 or 1,2,3, or 2,3,4 etc, looping back to 0 as necessary. If start
@@ -430,10 +438,10 @@ int main (int argc, char **argv)
                 std::cout << "param_hist with accepted size is " << param_hist.size() << std::endl;
                 // Now add all the points
                 for (size_t i = 0; i < param_hist.size(); ++i) {
[…]

### 2022-02-28 paper/paper.tex
WIP

@@
 %
-\begin{equation}
+\begin{equation} \label{e:WI}
 W_{\!_I}(d_{kb}) = \begin{cases}
@@
 
-Taken alone, the EphA forward interaction between one branch, $b$, and
+Taken alone, the EphA forward interaction between one RGC branch, $b$, and
 another, $k$, will cause a repulsion if $k$ originated more nasally than
@@
 % Now describe J model, and show the result.
+We defined a movement vector arising from a multiplicative receptor-ligand
+interaction, $\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the
+movement vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
+%
+\begin{equation} \label{e:mv3}
+ \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
+\end{equation}
+%
+% noting that $m_J J_b$ \emph{replaces} $m_c C_b$.
+%
+$\mathbf{J}_b$ is defined as
+%
+\begin{equation}
+\mathbf{J}_b = \frac{1}{|B_{b,J}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_J}\,Q_{\!_J}(d_{kb})
+\end{equation}
+%
+where $W_{\!_J}$ is distance based weighting analogous to $W_{\!_I}$,
+$B_{b,J}$ is analogous to $B_{b,I}$ and $Q_{\!_J}(d_{kb})$ is the signalling
+strength between two growth cones of receptor-ligand interaction radius
+$r_{\!_J}$ a distance $d_{kb}$ from one another, given by
+%
+\begin{equation}
+Q_{\!_J}(d_{kb}) = \begin{cases}
+                 F_i & \sum_i^N r_{i,b}\,l_{i,k} > s,~d_{kb} \leq 2r_{\!_J} \\
+                 0 & \mathrm{otherwise}
+     \end{cases}
+\end{equation}
+%
+The sign of $Q_{\!_J}$ determines whether the interaction, $\mathbf{J}_b$, is
+repulsive or attractive. $l_{i,k}$ is the expression of ligand $i$ on branch
+$k$.
+
+$\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ if there was a single
+$l_i$ expressed in the retina and its form was $1/r_i$. As well as allowing
+for two, opp
+
+With $m_{\!_G}$ fixed at the value 0.0012, and allowing the threshold
+parameter $s$ to vary along with $m_{\!_J}$ and $r_{\!_J}$ we performed a
+simulated annealing optimization. The parameters best able to recover the
+wildtype pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} =
+0.079$. The simulation of the system with these parameters is shown in
+Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
 
@@
 \includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_figcomp3_exit_true_steps_1500.png}
-\caption{$\mathbf{GJ}$ model}
+\caption{Chemoaffinity and receptor-ligand mediated competitive interactions
+($\mathbf{GJ}$ model). }
 \label{f:GJ}
@@
 
-%% TO ADD: Figure for the Reber/Brown manipulation on the GJ model.
+\subsection*{Knockin and Knockdown manipulations}
 
@@
 
-%% HEREish
-Next, we investigated the idea that the same receptor-ligand
-interaction that leads to gradient-based movement of RGC axons on the tectum
-could also act as a signal between the axons. This is plausible, because
-retinal ganglion cells are known to express ephrin receptor ligands as well as
-Eph receptors \citep{ephrins_on_RGCs}. We defined $N=4$ ligands, $l_i$,
-expressed by the RGCs in the model according to
-%
-\begin{equation} \label{e:retlig}
-l_i(x,y) = 1.05 + 0.26 \exp(2.3 u).
-\end{equation}
-%
-We used exactly the same form as the receptor expression given by
-Eq.\,\ref{e:retrcpt}, but with opposite gradient direction so here $u$ is
-substituted by $x$ for $l_0$, $y$ for $l_1$, $(1-x)$ for $l_2$ and by $(1-y)$
-for $l_3$ (see Fig.\,\ref{f:ex}C).
-%
-We combined chemotaxis with a multiplicative (or mass-action?) interaction,
-$\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the movement
-vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
-%
-\begin{equation} \label{e:mv3}
- \mathbf{M}_{b} = m_{\!_G} \mathbf{G}_b + m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
- %%+  m_{\!_I} \mathbf{I}_b +  m_{\!_J} \mathbf{J}_b + 0.5 \mathbf{B}_b
-\end{equation}
-%
-$\mathbf{J}_b$ is defined as
-%
-\begin{equation}
-% No \frac{1}{|B_b|} in this one. Really?
-\mathbf{J}_b = \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,Q_{\!_J}(d_{kb})
-\end{equation}
-%
-where $Q_{\!_J}(d_{kb})$ is the signalling strength between two growth cones of
-receptor-ligand interaction radius $r_{\!_J}$ a distance $d_{kb}$ from one another,
-given by
-%
-\begin{equation}
-Q_{\!_J}(d_{kb}) = \begin{cases}
-     \sum_i^N F_i\,r_{i,b}\,l_{i,k}    & d_{kb} \leq 2r_{\!_J} \\
-     0 & d_{kb} > 2r_{\!_J}
-     \end{cases}
-\end{equation}
-%
-The sign of $Q_{\!_J}$ (which is dependent on the values of $F_i$) determines
-whether the interaction, $\mathbf{J}_b$, is repulsive or attractive. $l_{i,k}$
-is the expression of ligand $i$ on branch $k$.
-
-Considering just one receptor-ligand pair, for instance $r_0$/$l_0$, and
-defining their interaction as repulsive, an axon from the transverse side of
-the retina, which expresses a high level of receptor $r_0$ will interact most
-strongly with an axon from the nasal side of the retina, where $l_0$ is most
-highly expressed. The interaction with axons that originate nearby, on the
-transverse side will be minimal, as these express low levels of $l_0$. Axons
-originating from the nasal side of the retina also interact most strongly with
-axons originating nearby and in general the $r_0$/$l_0$ interaction is weaker
-between a nasal axon and a second retinal axon. However, when the opposing
-receptor/ligand pair ($r_2$/$l_2$) interaction is summed with that of
-$r_0$/$l_0$, a more even interaction is found, with interactions between axons
-from extreme sides of the retina being the strongest. (Poss.~supp.~figure
-based on the Octave script J.m).
-
-We performed the same simulated annealing process as for the
-$\mathbf{GC}$ model to find the parameters best able to recover the wildtype
-pattern and found $m_{\!_G} = 0.00047$, $m_{\!_J} = 0.00089$ and $r_{\!_J} =
-0.341$.
-
-See Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
+%Considering just one receptor-ligand pair, for instance $r_0$/$l_0$, and
+%defining their interaction as repulsive, an axon from the transverse side of
+%the retina, which expresses a high level of receptor $r_0$ will interact most
+%strongly with an axon from the nasal side of the retina, where $l_0$ is most
+%highly expressed. The interaction with axons that originate nearby, on the
+%transverse side will be minimal, as these express low levels of $l_0$. Axons
+%originating from the nasal side of the retina also interact most strongly with
+%axons originating nearby and in general the $r_0$/$l_0$ interaction is weaker
+%between a nasal axon and a second retinal axon. However, when the opposing
+%receptor/ligand pair ($r_2$/$l_2$) interaction is summed with that of
+%$r_0$/$l_0$, a more even interaction is found, with interactions between axons
+%from extreme sides of the retina being the strongest. (Poss.~supp.~figure
+%based on the Octave script J.m).
 
@@
 Fig.\,\ref{f:G0}). Noise for $\nu < 0.5$ is tolerated by the model, but
-disorder increases about this value.
-\textbf{B} Competitive interaction in the $\mathbf{GI}$ model extends the
+disorder increases above this value.
+\textbf{B} Competitive interaction in the $\mathbf{GJ}$ model extends the
 range of ligand gradient noise which can be tolerated by the system.
@@
 
-Fig.\,\ref{f:noise} shows the response of the $\mathbf{GI}$ model to the
-addition of ligand gradient estimation noise.
+Fig.\,\ref{f:noise} shows the response of both the chemoaffinity-only model
+and the full $\mathbf{GJ}$ model to the addition of ligand gradient estimation
+noise. The results indicate that a moderate amount of noise is beneficial for
+the establishment of an ordered map and that the competitive model is much
+more robust to higher levels of gradient estimation noise.
 
@@
 Discussion goes here.
+
+Discuss the requirement for a distance based weight and the fact that without
+this, the system doesn't converge on a stable solution.
 

### 2022-02-28 scripts/paper_figures.sh

@@ -29,3 +29,6 @@
 ./build/sim/agent/agent1 configs/simpler/m_ee_GC_best_1.json configs/simpler/e_wt_figcomp1.json -co:exit=true
 ./build/sim/agent/agent1 configs/simpler/m_ee_GI_best_1.json configs/simpler/e_wt_figcomp2.json -co:exit=true
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp3.json -co:exit=true -co:steps=1500
+
+# The Reber manipulation
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_reber.json -co:exit=true -co:steps=1500

### 2022-02-28 sim/agent/agent1.h

@@ -710,6 +710,7 @@ struct Agent1
     }
 #endif // VISUALISE
 
+    // This number decides which axon will be shown in the single axon expt. It's an axon from a central RGC.
     static constexpr size_t singleaxon_idx = 210;
     // Set true to use orthographic projection
     static constexpr bool use_ortho = false;

### 2022-02-28 sim/agent/branch.h

@@ -185,17 +185,27 @@ public:
                        (source_tissue->forward_interactions[i] == interaction::attraction ? T{-1} : T{0}));
         }
         QJar *= kp->lgnd * this->rcpt;
+        //std::cout << "Rcpt * lgnd = " << QJar << " (s is " << this->s << ")\n";
 
         // Add to competition if any of the mass-action interactions is true.
         // Note: use <= operator which is true if EVERY member of the Vector is <= threshold
         bool subthreshold = false;
-        if (d <= this->two_r_j && (subthreshold = QJar.abs() <= this->s) == false) {
-            // W = QJar.sum()/N; // Sum the interactions to modulate the competition strength?
-            // W = QJar.longest(); // Use the maximum suprathreshold interaction element?
-            W = T{1} - d/this->two_r_j; // or have a distance based weight?
-            J += kb * W;
-            rtn[2] = true;
+        if (d <= this->two_r_j) {
+            //std::cout << "d < 2rj: " << d << " cf. two_r_j " << this->two_r_j << std::endl;
+            if ((subthreshold = QJar.abs() <= this->s) == false) {
+                //std::cout << "subthreshold false so OVER threshold for interaction...\n";
+                // W = QJar.sum()/N; // Sum the interactions to modulate the competition strength?
+                // W = QJar.longest(); // Use the maximum suprathreshold interaction element?
+                W = T{1} - d/this->two_r_j; // or have a distance based weight?
+                J += kb * W;
+                rtn[2] = true;
+            } else {
+                //std::cout << "subthreshold true (no interaction)...\n";
+                rtn[2] = false;
+            }
+            //std::cout << "subthreshold is " << (subthreshold == true ? "true" : "false") << std::endl;
         } else {
+            //std::cout << "d > 2rj (no interaction): " << d << " cf. two_r_j " << this->two_r_j << std::endl;
             //std::cout << "NO interaction for |QJar| = " << QJar.abs() << std::endl;
             rtn[2] = false;
         }

### 2022-03-01 paper/paper.tex
Adds single axon inkscape figure
Small changes to manuscript

@@
 results than for the chemoaffinity-only model (see supplementary
-Fig.\,\ref{f:GCsurg}).
+Fig.\,\ref{f:GCsurg}). We also confirmed the result
+from \citet{simpson_simple_2011} that this description of a competitive
+interaction is not able to reproduce the result of \citet{reber_relative_2004}
+(Fig.\,\ref{f:GCreber}) Hmm. DID we? That graph looks nearly as good as the GJ
+one...
 
@@
 
-Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), requires a
-global supervisor to provide information to each branch about the location of
-all other branches (But \textbf{fixme}, arguably, so does the GJ model now!),
-we looked for a receptor-ligand interaction that could signal between axon
-branches and enable them to adjust their movements away from one another.
+Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), does not
+describe a signalling mechanism and cannot, on its own, explain all of the
+experimental results, we looked for a receptor-ligand interaction that could
+signal between axon branches and enable them to adjust their movements away
+from one another.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{./images/single_axon.png}
+\caption{ \textbf{A} Central \textbf{B} Caudal (nasal RGC) \textbf{C} Rostral
+(temporal RGC)}
+\label{f:single}
+\end{figure}
+
 \subsection*{Attractive receptor-ligand interactions}
@@
 \label{f:GCsurg}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_reber_exit_true_steps_1500.png}
+\caption{$\mathbf{GC}$ model, Reber experiment.}
+\label{f:GCreber}
 \end{figure}

### 2022-03-03 paper/paper.tex
Working through the model description

@@
 with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose
-gradient is opposite to $r_0$, was orthogonal to $r_3$.
+gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:expressions}B).
 
@@
 \begin{equation} \label{e:tecligexp}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(2.3 u),
+L_i(x,y) = 1.05 + 0.26 \exp(2.3 u),
 \end{equation}
@@
 \begin{equation} \label{e:tecligexp2}
-L_i^{\text{exp}}(x,y) = 1.05 + 0.26 \exp(1.1 u)
+L_i(x,y) = 1.05 + 0.26 \exp(1.1 u)
 \end{equation}
@@
  towards the edge. \textbf{A} Axon branches \textbf{B} axon
- centroids \textbf{C} Five selected axons.}
+ centroids \textbf{C} The pattern metric $\eta$ is zero, as it was in
+ Fig.\,\ref{f:G}F and $\epsilon$ approaches a much lower value as $t \to \infty.}
 \label{f:G0}
@@
 mismatch experiment (Fig.\,\ref{f:Gsurg}F), but in this case there are no
-axons from the temporal retina, which was been ablated and only the axons
+axons from the temporal retina, which was ablated and only the axons
 from the nasal retina are present, arranged in a line.
@@
 results than for the chemoaffinity-only model (see supplementary
-Fig.\,\ref{f:GCsurg}). We also confirmed the result
-from \citet{simpson_simple_2011} that this description of a competitive
-interaction is not able to reproduce the result of \citet{reber_relative_2004}
-(Fig.\,\ref{f:GCreber}) Hmm. DID we? That graph looks nearly as good as the GJ
-one...
+Fig.\,\ref{f:GCsurg}).
 
@@
 Because the competition effect, $\mathbf{C}_b$ (Eq.\,\ref{e:C}), does not
-describe a signalling mechanism and cannot, on its own, explain all of the
-experimental results, we looked for a receptor-ligand interaction that could
-signal between axon branches and enable them to adjust their movements away
-from one another.
+describe a signalling mechanism, and because \citet{simpson_simple_2011}
+reported that an additional mechanism is required to explain the gene
+manipulation experiments of \citet{brown_topographic_2000}
+and \citet{reber_relative_2004}, we looked for a receptor-ligand interaction
+that could signal between axon branches and enable them to adjust their
+movements away from one another.
 
@@
 $B_{b,I}$ is the set of growth cones within a distance $2r_{\!_I}$ of branch
-$b$. $\hat{\mathbf{x}}_{kb}$ is as defined in Eq.\,\ref{e:C}.
+$b$. $\hat{\mathbf{x}}_{kb}$ is as defined in Eq.\,\ref{e:C}. This mechanism
+adds a weighted repulsion if the ratio of receptor expression on
+branch $b$ and the expression on branch $k$ exceeds the threshold of 1.1.
 
@@
 
-To avoid this asymmetry, we allow two receptor-ligand pairs, which are
+To avoid this asymmetry, we allowed two receptor-ligand pairs, which are
 expressed in opposing directions, to contribute to axon-axon interactions,
 ensuring that interactions between single growth cones and clusters are not
-dependent on their origin locations. Fig.\,\ref{f:expressions}C shows ligand
+dependent on their origin locations. Fig.\,\ref{f:ex}C shows ligand
 expressions representing an ephrin-A like expression ($l_0$) and an opposing
-expression ($l_2$). It remains the case (with a suitable choice of threshold,
-$s$) that branches originating from the same location on the temporal-nasal
-axis will not experience competitive interactions.
+expression ($l_2$). We adopted a form for the ligand expression such that
+$r_i \times l_i = 4.7727$, with the scalar constant chosen so that the retinal
+ligand expression, $l_i(x,y)$ has approximately the same start and end values
+as the exponential retinal receptor expression:
+%
+\begin{equation}
+ \label{e:retligexp}
+l_i(x,y) = \frac{4.7727}{1.05 + 0.26 \exp (2.3 u)}
+\end{equation}
+$u$ was substituted with $(1-x)$ for $l_0$ or $x$ for $l_2$ (Fig.\,\ref{f:ex}C).
+%
+%With a suitable choice of threshold, $s$, branches originating from the same
+%location on the temporal-nasal axis will not experience competitive
+%interactions.
 
@@
 expression fields (for $l_1$ and $l_3$), but we found this not to be
-necessary; a single opposing pair of fields ensures that axons that originated
-close together (and branches of the same axon) do not repel one another,
-whereas axons that are close on the temporal-nasal axis, but far apart on the
-dorso-ventral axis are generally driven far apart by chemotaxis, so whether
-they compete or not when close together does not affect the behaviour of the
-system.
+necessary; a single opposing pair of fields can ensure that most pairs of
+axons are able to signal the necessity for a competitive interaction. We set
+$l_1 = l_3 = 0$.
 
@@
 interaction, $\mathbf{J}_b$ (scaled by a parameter $m_{\!_J}$) to define the
-movement vector, $\mathbf{M}_b$, for the $\mathbf{GJ}$ model as:
+movement vector, $\mathbf{M}_b$, for the `$\mathbf{GJ}$ model' as:
 %
@@
 \begin{equation}
-\mathbf{J}_b = \frac{1}{|B_{b,J}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_J}\,Q_{\!_J}(d_{kb})
-\end{equation}
-%
-where $W_{\!_J}$ is distance based weighting analogous to $W_{\!_I}$,
-$B_{b,J}$ is analogous to $B_{b,I}$ and $Q_{\!_J}(d_{kb})$ is the signalling
-strength between two growth cones of receptor-ligand interaction radius
-$r_{\!_J}$ a distance $d_{kb}$ from one another, given by
+\mathbf{J}_b = \frac{1}{|B_{b,J}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W_{\!_J}\,Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k})
+\end{equation}
+%
+where $W_{\!_J}$ is distance based weighting analogous to $W_{\!_I}$ and
+$B_{b,J}$ is analogous to $B_{b,I}$. $Q_{\!_J}$ determines if the signals
+between a branch $b$ expressing receptors $\mathbf{r}_b$ and branch $k$
+expressing ligands $\mathbf{l}_k$, separated by a distance $d_{kb}$ will cause
+$b$ to move away from $k$. $Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k})$
+can be expressed as
 %
 \begin{equation}
-Q_{\!_J}(d_{kb}) = \begin{cases}
-                 F_i & \sum_i^N r_{i,b}\,l_{i,k} > s,~d_{kb} \leq 2r_{\!_J} \\
-                 0 & \mathrm{otherwise}
+Q_{\!_J}(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
+                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
+                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_J} \\
+                 1 & \mathrm{otherwise}
      \end{cases}
 \end{equation}
-%
-The sign of $Q_{\!_J}$ determines whether the interaction, $\mathbf{J}_b$, is
-repulsive or attractive. $l_{i,k}$ is the expression of ligand $i$ on branch
-$k$.
-
-$\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ if there was a single
-$l_i$ expressed in the retina and its form was $1/r_i$. As well as allowing
-for two, opp
-
-With $m_{\!_G}$ fixed at the value 0.0012, and allowing the threshold
-parameter $s$ to vary along with $m_{\!_J}$ and $r_{\!_J}$ we performed a
-simulated annealing optimization. The parameters best able to recover the
-wildtype pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} =
-0.079$. The simulation of the system with these parameters is shown in
-Fig.\,\ref{f:GJ} and Fig.\,\ref{f:GJsurg}.
+
+where $r_{\!_J}$ is the receptor-ligand interaction radius, $l_{i,k}$ (the
+expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of
+$\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of
+$\mathbf{r}_b$. $F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$
+interaction signals repulsion and +1 if it signals attraction. $s$ is a
+positive signal threshold whose value should be around 4.7727; if higher, then
+branches whose retinal origin is close will not experience a competitive
+interaction. If $s\leq 4.7727$, then \emph{any} pair of branches will compete as they
+approch each other. With the constraint $s>0$ comes the assumption that
+attractive interactions will have no effect on the competitive movement
+vector---a receptor-ligand pair that signals attraction cannot cause
+$Q_{\!_J}$ to take the value 1. $Q_{\!_J}$ is set to 1 only if at least one
+repulsive signal exceeds the threshold, $s$ and the distance from branch
+$b$ to branch $k$ is smaller than $2 r_{\!_J}$.
+
+Note that $\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ (with $s=1.1$) if there
+was a single $l_i$ expressed in the retina and its form was $1/r_i$.
+
+% Now the simulation.
+Adopting retinal expressions given in Eqs.\,\ref{e:retrcpt}
+and \ref{e:retligexp}, we performed a simulated annealing optimization to find
+the best parameters for the $\mathbf{GJ}$ model.  With $m_{\!_G}$ fixed at the
+value 0.0012, we allowed the threshold parameter $s$ to vary along with
+$m_{\!_J}$ and $r_{\!_J}$ . The parameters best able to recover the wildtype
+pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} = 0.079$. Note that
+the discovered value for $s$ is less than 4.7727 meaning that for \emph{any
+pair of branches}, either $r_0\times l_0$ or $r_2\times l_2$ will exceed $s$.
+
+The
+simulation of the system with these parameters is shown in Fig.\,\ref{f:GJ}
+and Fig.\,\ref{f:GJsurg}.
 
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_reber.png}
-\caption{$\mathbf{GJ}$ model, Reber et al.~manipulation. Look at THAT!}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
+\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
 \label{f:GJreber}

### 2022-03-03 optimization/search_GJ_fixs.sh

@@ -1,0 +1,9 @@
+#!/bin/sh
+
+#
+# Use anneal1 to search for the best parameters for the GJ model.
+#
+# m_ee_GJ.json contains the starting model paramters. s_GJ.json
+# contains the (simulated annealing) search parameters
+#
+./build/sim/agent/anneal1c ./optimization/m_ee_GJ.json ./optimization/s_GJ_fixs.json

### 2022-03-07 paper/paper.tex
WIP

@@
 We asked i) whether this mechanism could \emph{replace} the non-biological
-competition, C, (making necessary its extension to the dorsal-ventral retinal
-axis) and ii) if cast in terms of receptor-ligand interactions, could the
-strong ligand expression assumption be relaxed?
+competition, $\mathbf{C}_b$, and ii) if cast in terms of receptor-ligand
+interactions, could the strong ligand expression assumption be relaxed?
 
@@
 that repulsive signals are transmitted between axons whose cell bodies
-originte far apart in the retina. However, there is an asymmetry in the
+originate far apart in the retina. However, there is an asymmetry in the
 interaction. Consider a cluster of growth cones, $k$, originating from the
@@
 interaction. If $s\leq 4.7727$, then \emph{any} pair of branches will compete as they
-approch each other. With the constraint $s>0$ comes the assumption that
+approach each other. With the constraint $s>0$ comes the assumption that
 attractive interactions will have no effect on the competitive movement
@@
 value 0.0012, we allowed the threshold parameter $s$ to vary along with
-$m_{\!_J}$ and $r_{\!_J}$ . The parameters best able to recover the wildtype
-pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} = 0.079$. Note that
-the discovered value for $s$ is less than 4.7727 meaning that for \emph{any
-pair of branches}, either $r_0\times l_0$ or $r_2\times l_2$ will exceed $s$.
-
-The
-simulation of the system with these parameters is shown in Fig.\,\ref{f:GJ}
-and Fig.\,\ref{f:GJsurg}.
+$m_{\!_J}$ and $r_{\!_J}$. The parameters best able to recover the wildtype
+pattern were $s=4.732$, $m_{\!_J} = 0.247$ and $r_{\!_J} = 0.079$, with the
+wildtype result shown in Fig.\,\ref{f:GJ}. Note that the discovered value for
+$s$ is \emph{less} than 4.7727 meaning that for any pair of branches, either
+$r_0\times l_0$ or $r_2\times l_2$ will exceed $s$ and branches all
+compete. With $s < 4.7727$ the $\mathbf{GJ}$ model is equivalent to the
+$\mathbf{GC}$ model. However, the parameters found for the $\mathbf{GJ}$ model
+did lead to better results for surgical manipulations, as shown in
+Fig.\,\ref{f:GJsurg}. Values for the pattern metric, $\epsilon$ are
+consistently lower than those for the $\mathbf{GC}$ model shown in
+Fig.\,\ref{f:GCsurg}. Given that the \emph{systems} are equivalent, the
+differences must have arisen due to the optimization process, which must have
+found alternative `good' parameters. We verified this by copying the
+parameters found for the $\mathbf{GJ}$ model into the $\mathbf{GC}$ model,
+setting $m_G = 0.0012$, $r_c = 0.079$ and $m_c =
+0.247$. Fig.\,\ref{f:GClikeGJsurg} shows the result for surgical
+manipulations. Their equivalence to Figs.~\ref{f:GCsurg} is evident by
+inspection. Nevertheless, we retain the receptor-ligand signalling of the
+$\mathbf{GJ}$ model as this maintains the ability of the model to respond to
+the genetic manipulations described in the next section.
 
@@
 \subsection*{Knockin and Knockdown manipulations}
+
+In the early 2000s, a series of genetic experiments added to the set of
+manipulations that a model of the retinotectal system would need to
+explain. First, \citet{brown_topographic_2000} showed that a selective RGC
+knock-in of EphA3 receptors led to an interesting modulation of the maps found
+in the tectum. For the strongest knock-in (a homozygote manipulation) two maps
+were formed, with knock-in cells finding termination zones more rostral than
+their wildtype retinal neighbours. By enhancing the quantity of EphA receptors
+expressed by half of all retino-tectal axons, they demonstrated the importance
+of retinal receptor expression for the arrangement of the tectal map. In the
+weaker, heterozygote manipulation, the maps were still separated, but the
+separation `collapsed' towards the rostral end of the tectum
+(see \citet{brown_topographic_2000}, Fig\,5). This led to a second study
+showing that with an additional knock-down of EphA4 receptors in \emph{all}
+RGCs, the retinal origin of the collapse point was shifted temporally and all
+heterozygote EphA3 knock-in cells were pushed even further rostrally
+(\citet{reber_relative_2004}, their Fig.\,3e). Does the $\mathbf{GJ}$ model
+reproduce the results of Brown et al and Reber et al?
+
+An assumption made in these papers is that all EphA receptor subtypes
+contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which
+increases from a low level at the nasal side of the retina, to a high level at
+the temporal side. The precise contribution of each subtype to the overall
+level is assumed not to be important. In our model, receptor $r_0$ increases
+in the retina in a nasal to temporal direction (Fig\,\ref{f:expressions}B) and
+so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et
+al.~experiments. In half of the RGCs, we knock-in EphA3 by adding to the value
+of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote
+EphA3 ki/ki). To add the additional knock-down of EphA4
+in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all}
+RGCs.
 
@@
 \caption{$\mathbf{GC}$ model, surgical manipulations. Demonstrating that this
-model reproduces these manipulations.}
+model performs better than the chemoaffinity-only model.}
 \label{f:GCsurg}
@@
 
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GC_best_1_reber_exit_true_steps_1500.png}
-\caption{$\mathbf{GC}$ model, Reber experiment.}
-\label{f:GCreber}
+% ./scripts/fig_GClikeGJ_surgical.sh
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GClikeGJ_surgical.png}
+\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
+optimization ( $m_G = 0.0012$, $r_c = 0.079$ and $m_c = 0.247$). Results are
+equivalent to those for the $\mathbf{GJ}$ model in Fig.\,\ref{f:GJsurg}.}
+\label{f:GClikeGJsurg}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations_GClikeGJ.png}
+\caption{$\mathbf{GC}$ model with parameters found from \emph{GJ}
+optimization, genetic manipulations. As for the surgical manipulations, the
+results are similar to those of the $\mathbf{GJ}$ model with the equivalent
+parameters. However, there are subtle differences such as XXX.}
+\label{f:GClikeGJreber}
 \end{figure}

### 2022-03-07 scripts/fig_GClikeGJ_surgical.sh

@@ -1,0 +1,8 @@
+#!/bin/sh
+
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_tecrot90_figcomp.json  -co:steps=1000
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_tecrot180_figcomp.json -co:steps=1000
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_tecswap_figcomp.json   -co:steps=1000
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_retablat_figcomp.json  -co:steps=1000
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_tecablat_figcomp.json  -co:steps=1000
+./build/sim/agent/agent1 configs/simpler/m_ee_GClikeGJ.json  configs/simpler/e_mismatch_figcomp.json  -co:steps=1000

### 2022-03-07 paper/paper.tex
WIP on laptop

@@
 branches whose retinal origin is close will not experience a competitive
-interaction. If $s\leq 4.7727$, then \emph{any} pair of branches will compete as they
-approach each other. With the constraint $s>0$ comes the assumption that
-attractive interactions will have no effect on the competitive movement
+interaction. If $s\leq 4.7727$, then \emph{any} pair of branches will compete
+as they approach each other. With the constraint $s>0$ comes the assumption
+that attractive interactions will have no effect on the competitive movement
 vector---a receptor-ligand pair that signals attraction cannot cause
 $Q_{\!_J}$ to take the value 1. $Q_{\!_J}$ is set to 1 only if at least one
-repulsive signal exceeds the threshold, $s$ and the distance from branch
-$b$ to branch $k$ is smaller than $2 r_{\!_J}$.
-
-Note that $\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ (with $s=1.1$) if there
-was a single $l_i$ expressed in the retina and its form was $1/r_i$.
+repulsive signal exceeds the threshold, $s$ and the distance from branch $b$
+to branch $k$ is smaller than $2 r_{\!_J}$.
+
+Note that $\mathbf{J}_b$ would be equivalent to $\mathbf{I}_b$ (with $s=1.1$)
+if there was a single $l_i$ expressed in the retina and its form was $1/r_i$.
 
@@
 
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
+\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
+\label{f:GJeph}
+\end{figure}
+
 An assumption made in these papers is that all EphA receptor subtypes
@@
 in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all}
-RGCs.
-
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
-\label{f:GJreber}
-\end{figure}
+RGCs. Fig.\,\ref{f:GJeph}A-C shows, for the single EphA3 knock-in (EphA3
+ki/+), branch positions, axon centroids and final rostral-caudal position on
+the tectum plotted versus the originating nasal-temporal retinal
+coordinate. In A, branches are coloured in the red/green duochrome scheme
+according to their origin location on the retina. In B \& C, axon centroids of
+RGCs which were treated with the knock-in manipulation are coloured red,
+wildtype cells are blue. The map of treated cells on the tectum is shifted
+slightly rostrally, but the shift becomes statistically insignificant at about
+0.75 along the nasal-temporal axis. While the map collapse is less abrupt than
+in the experimental result of \citet{brown_topographic_2000}, it is comparable
+with the result from \citet{simpson_simple_2011}.  Fig.\,\ref{f:GJeph}D-F
+shows the same trio of graphs for the homozygote double knock-in, EphA3
+ki/ki. Here, the treated axons are pushed further towards the rostral end of
+the tectum and the untreated axons are pushed caudally. The maps are
+completely separated; as in the experiment, there is no `collapse point'.
+
+The rostral movement of treated axons occurs due to their increased expression
+of EphA ($r_0$), which increases the interaction with the tectal
+rostral-caudal ligand gradient. The opposing receptor expression ($r_2$) is
+unchanged by the manipulation and so those axons with increased $r_0$
+expression are pulled to a more rostral position than they would otherwise
+take. The caudal movement of wildtype axons, whose receptor expression in
+Figs.\,\ref{f:GJeph}A-F is unchanged, occurs due to competition with the
+caudally-drawn treated axons.
+%
+%Interestingly, the feature of a slight 'elbow' in the double knockin
+%graph (F) at about 0.7 on the nasal-temporal axis matches that seen in the
+%experimental result. Well, kinda.
+
+Figs.\,\ref{f:GJeph}G-I show the results for the Reber et al.~manipulation, in
+which the EphA expression is selectively knocked in in 50\% of cells and
+knocked down in all cells. The collapse point has moved from about 0.75 in
+Fig.\,\ref{f:GJeph}C to about 0.85 (\textbf{check}), in agreement with the
+experimental result. However, the significant rostral movement of the treated
+axons seen in \citet{reber_relative_2004}, where not a single treated axon was
+found more than 30\% along the rostral-caudal axis, was absent from our result.
 
@@
 \caption{ \textbf{A} Chemotaxis-only model. The mean value of $\eta$ (number of crossings) and $\epsilon$
-(mean distance error), averaged over the period from simulation time $t=750$ to $t=1000$, plotted with respect to the
-noise gain ($\nu$) in the best $\mathbf{G}$ model and the wildtype condition (compare with
+(mean distance error), averaged over the period from simulation time $t=750$
+to $t=1000$, plotted with respect to the noise gain ($\nu$) in the best
+$\mathbf{G}$ model and the wildtype condition (compare with
 Fig.\,\ref{f:G0}). Noise for $\nu < 0.5$ is tolerated by the model, but
@@
 binding noise \citep{goodhill_noise_2014}.  We investigated the ability of the
-system to self-organise in the presence of noise, by introducing a stochastic
-element to each branch's sampling of the gradient.
+chemoaffinity-only and the $\mathbf{GJ}$ model systems to self-organise in the
+presence of noise, by introducing a stochastic element to each branch's
+sampling of the gradient.
 
@@
 true gradient by a random vector, $\mathbf{U}$, whose coordinates are drawn
-from a random uniform distribution centred about 0 with magnitude $\nu$, denoted:
-$\mathbf{U}(-\nu, \nu)$. Eq.\,\ref{e:G} becomes
+from a random uniform distribution centred about 0 with magnitude $\nu$,
+denoted: $\mathbf{U}(-\nu, \nu)$. Eq.\,\ref{e:G} becomes
 
@@
 
+% What about the EphA manipulations with noise? Ans: No strong effect (just
+% add "noise_gain" : 0.5, to one of the expt jsons.
+
 \subsection*{Single axon behaviour}
@@
 parameters. However, there are subtle differences such as XXX.}
-\label{f:GClikeGJreber}
+\label{f:GClikeGJeph}
 \end{figure}

### 2022-03-28 paper/main.tex

% Use the article doc class, with an 11 pt basic font size
%\documentclass[11pt, a4paper, draft]{article}
\documentclass[11pt, a4paper]{article}
% Nice citations
\usepackage{natbib}
% Set margins
\usepackage[margin=2.5cm]{geometry}
% Multilingual support
\usepackage[english]{babel}
% Nice mathematics and Left-right harpoons for kinetic equations
\usepackage{amsmath,mathtools}
% Control over maketitle
\usepackage{titling,titlesec}
% Ability to use colour in text
\usepackage[usenames]{color}
\definecolor{grey}{rgb}{0.6, 0.6, 0.63}
\definecolor{black}{rgb}{0, 0, 0}
% For the \degree symbol
\usepackage{textcomp,gensymb}
% Allow includegraphics and wrapped figures
\usepackage{graphicx,wrapfig,subfigure}
\usepackage[outercaption]{sidecap}
\usepackage{csquotes}

%% Trick to define a language alias and permit language = {en} in the .bib file.
% From: http://tex.stackexchange.com/questions/199254/babel-define-language-synonym
\usepackage{letltxmacro}
\LetLtxMacro{\ORIGselectlanguage}{\selectlanguage}
\makeatletter
\DeclareRobustCommand{\selectlanguage}[1]{%
  \@ifundefined{alias@\string#1}
    {\ORIGselectlanguage{#1}}
    {\begingroup\edef\x{\endgroup
       \noexpand\ORIGselectlanguage{\@nameuse{alias@#1}}}\x}%
}
\newcommand{\definelanguagealias}[2]{%
  \@namedef{alias@#1}{#2}%
}
\makeatother
\definelanguagealias{en}{english}
\definelanguagealias{eng}{english}
%% End language alias trick

%% Any aliases here
\newcommand{\mb}[1]{\mathbf{#1}} % this won't work?
% Emphasis and bold.
\newcommand{\e}{\emph}
\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\dvrg}{\nabla\vcdot\nabla}
%% END aliases

% Custom font defs
% fontsize is \fontsize{fontsize}{linespacesize}
\def\authorListFont{\fontsize{11}{11} }
\def\corrAuthorFont{\fontsize{10}{10} }
\def\affiliationListFont{\fontsize{11}{11}\itshape }
\def\titleFont{\fontsize{14}{11} \bfseries }
\def\textFont{\fontsize{11}{11} }
\def\sectionHdrFont{\fontsize{11}{11}\bfseries}
\def\bibFont{\fontsize{10}{10} }
\def\captionFont{\fontsize{10}{10} }

% Caption font size to be small.
\usepackage[font=small,labelfont=bf]{caption}

% Make a dot for the dot product, call it vcdot for 'vector calculus dot'. Bigger than \cdot, smaller than \bullet.
\makeatletter
\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
\makeatother

\def\firstAuthorLast{James}
\def\Address{\\
\affiliationListFont Department of Psychology,\\
 \affiliationListFont  The University of Sheffield, Sheffield, UK \\
}
\def\corrAuthor{Seb James}
\def\corrAddress{Department of Psychology, The University of Sheffield,
  Western Bank, Sheffield, S10 2TP, UK}
\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address}
%  \corrAuthorFont $^{*}$ Correspondence: {seb.james@sheffield.ac.uk}}

% If no page numbering wanted:
%\pagenumbering{gobble}

% A trick to get the bibliography to show up with 1. 2. etc in place of [1], [2] etc.:
\makeatletter
\renewcommand\@biblabel[1]{#1.}
\makeatother

% reduce separation between bibliography items if not using natbib:
\let\OLDthebibliography\thebibliography
\renewcommand\thebibliography[1]{
  \OLDthebibliography{#1}
  \setlength{\parskip}{0pt}
  \setlength{\itemsep}{0pt plus 0.3ex}
}

% Define title, empty date and authors
\title {
Time to stop; agent-based modelling of chemoaffinity with competition
}
\date{} % No date please
\author{\Authors}

%% END OF PREAMBLE

\begin{document}
%\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
\maketitle

\emph{We show that a true chemoaffinity model based on gradient following (via receptor-ligand interactions) combined with receptor-ligand-mediated axonal repulsion, in which information is only transferred by nearest-neighbour interactions, can account for a broad range of experimental results.}

%\begin{itemize}
%    \item Sperry's chemospecificity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development.
%    \item A key component of the theory is that growing axons find their way into position by chemotaxis, i.e., by evaluating local information in gradients of gene expression that span the target tissue
%    \item Computational models have shown that in addition to chemotaxis, local competitive interactions between growing axons can account for a wide range of results from experiments in which the factors known to influence the growth of axons have been surgically or genetically manipulated
%    \item However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemospecificity model, that axons sample and evaluate positional information only \emph{locally}, is yet to be tested directly.
%    \item Consider, for example, an important model by Goodhill \& Simpson, which shows how [a certain set of assumptions about the form of the] competition between axons can account for patterns of innervation in the developing tectum that result from ablating/rotating sections of the retina.
%    \item The formulation of this model includes a term that at every time-step moves the simulated axons in the direction of a pre-specified target location. Thus the model implicitly assumes a global supervisor, capable of evaluating the current state of each axon, computing a direction vector to correct its trajectory, and communicating that information to each axon.
%    \item This model was constructed primarily to establish the effects of competition, and to this end is elegantly formulated.
%    \item Nevertheless, the key assumptions of the chemospecificity model, and specifically the proposal that axons find their way via evaluating only local information (chemotaxis), are yet to be investigated directly.
%    \item The aim of the current study is therefore to construct a model that represents the assumptions of the chemospecificity model explicitly.
%    \item One advantage of such a model is that it can be used to investigate the sensitivity and robustness of local interactions in establishing topographic maps to various sources of noise.
%    \item Indeed, our simulation results show that the local mechanisms of axonal information-processing that are assumed by the chemospecificity model provide a remarkably robust solution to the problem of topographic map development.
%\end{itemize}

\section{Introduction}

Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_chemoaffinity_1963}.
%
The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
%
This key idea does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
An additional component of Sperry's theory \citep{gaze?} suggested that the axon could determine this direction by evaluating local information in gradients of gene expression spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
%
Sperry's theory was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
%
The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.

Computational models have shown that in addition to chemotaxis, local competitive interactions between growing axons can account for a wide range of results from experiments in which the factors known to influence the growth of axons have been surgically or genetically manipulated \citep{prestige_role_1975,simpson_simple_2011,suetterlin_target-independent_2014}.

However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemoaffinity model, that axons sample and evaluate positional information only \emph{locally}, is yet to be tested directly.
%
Consider, for example, an important model by Goodhill \& Simpson, which shows how [a certain set of assumptions about the form of the] competition between axons can account for patterns of innervation in the developing tectum that result from ablating/rotating sections of the retina \citep{simpson_simple_2011}.
%
The formulation of this model includes a term that at every time-step moves the simulated axons in the direction of a pre-specified target location.
Thus the model implicitly assumes a global supervisor, capable of evaluating the current state of each axon, computing a direction vector to correct its trajectory, and communicating that information to each axon.
%
This model was constructed primarily to establish the effects of competition, and to this end is elegantly formulated.
%
Nevertheless, the key assumptions of the chemoaffinity model, and specifically the proposal that axons find their way via evaluating only local information (chemotaxis), are yet to be investigated directly.

The aim of the current study is therefore to construct a model that represents the assumptions of the chemoaffinity model explicitly.

One advantage of such a model is that it can be used to investigate the sensitivity and robustness of local interactions in establishing topographic maps to various sources of noise.

Indeed, our simulation results show that the local mechanisms of axonal information-processing that are assumed by the chemoaffinity model provide a remarkably robust solution to the problem of topographic map development.

\section{Results}
\subsection*{A model of chemotaxis and competition}

In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \mathbf{M}$ with the movement vector $\mathbf{M}$ given by
\begin{equation} \label{e:mv2}
 \mathbf{M} = \mathbf{G} +  \mathbf{X} + \mathbf{B}.
\end{equation}
$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{X}$ is the movement due to competitive interactions with other branches.
$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue.

\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{X}$ was studied.
Their model, which successfully accounts for a wide range of experimental manipulations, found an $\mathbf{X}$ comprised of two components; a simple distance-based competition between \emph{all} branches ($\mathbf{C}$), along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon ($\mathbf{I}$).
We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.

\subsection*{A gradient based model of chemotaxis}

We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be

\begin{equation}\label{e:G}
\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
\end{equation}
%
where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
%
We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
%
$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.

%%%%% Describing the simulation
We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares (Fig.\,\ref{f:ex}B-D).
The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.

Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$ indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
($N$ ligands, $l_i$, are also expressed by RGCs, but these are not used for chemotaxis.)
The expression of each receptor varies only with respect to one spatial dimension.
%
We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$. $r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
%
Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}B).

Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Fig.\,\ref{f:expressions} summarises results from the literature).
We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}D).

While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ includes the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.

\subsection*{A signalling model for competition}

 We modelled competition as distance-based competitive movement vector, $\mathbf{X}$, which is `switched on' via receptor-ligand signalling:
%
\begin{equation}
\mathbf{X} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
\end{equation}
%
$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$) and $B_{b}$ is the set of branches that are within interaction distance of $b$. $m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
%
$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
%
\begin{equation}
Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_X} \\
                 1 & \mathrm{otherwise.}
     \end{cases}
\end{equation}
%
$r_{\!_X}$ is the receptor-ligand interaction radius, $l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$.
$F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
$s$ is a positive signal threshold for interaction.
With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.

\subsection*{Simulation}

The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$. We fixed $m_{\!_G}$ to the value $0.0012$ and used the optimization to find the competition parameters that would balance this level of chemotaxis.
To find the best combination of the remaining three parameters, we reasoned that evolution would optimize the arrangement of the wildtype pattern.
We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000.

We used a combination of pattern metrics to define the objective function. $\epsilon$ is the mean distance error of the axon centroids, given by:
%
\begin{equation}\label{e:eps}
\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
\end{equation}
%
where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
$n$ is the number of RGC axons in the system.
%
$\epsilon$ tends to 0 as the pattern approaches the target layout.
However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other.
For this reason, we also use the number of crossings in the fish net plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout. The annealing objective function, $\xi$, combined these metrics with $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
The parameters returned by the simulated annealing were
$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.

Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps.
Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red. Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
The axons retain the colour of their originating cell on the retina.
Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
Fig.\,\ref{f:GJ}, panels B--D show `fish net' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints.
Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation.
The number of line crossings in this plot gives an indication of the disorder in the arrangement.
Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
Opposing gradients have sorted the axons along the two axes to form a grid of cells.
The arrangement spans the full tectum due to competition between the growth cones (\textbf{ref supp. figure of chemo-only model}).
Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E.
Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.

\subsection*{Surgical manipulations}

After finding an optimum set of parameters for the wildtype arrangement, we tested our model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
Figs.~\ref{f:GJsurg}A and \ref{f:GJsurg}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
%
To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column).
After the manipulation, we re-computed the tectal ligand expression gradients. Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
The manipulated gradients, interacting with unchanged axon branches, lead to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
Long connecting lines in the fish net show that a number of axons from near the edges of the tectum have been drawn into the graft region.
The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg}A \& B.
We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red).
The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04).
The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
%
Fig.\,\ref{f:GJsurg}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. \textbf{Fixme: which of the two grafts does the RH graph refer to?}

Fig.\,\ref{f:GJsurg}D shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset) \citep{attardi_preferential_1963,schmidt_expansion_1978}.
The fish-net plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal tectal axis plotted against its position on the nasal-temporal retinal axis in red triangles, with the experimental positions plotted in blue squares.
%
Fig.\,\ref{f:GJsurg}E shows the result of a tectal ablation. Here the retina is left un-modified, but the caudal half of the tectum is surgically removed. Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue. Our model clearly replicates this result, indicated by both fish-net and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
%
A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}. Our model replicates this manipulation (Fig.\,\ref{f:GJsurg}F) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg}E; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
%
The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}. Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished. These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg}G).

\subsection*{Genetic manipulations}

In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
explain.
First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum.
For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours.
By enhancing the quantity of EphA receptors expressed by half of all retino-tectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map.
In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5).
This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
Does our model reproduce the results of Brown et al and Reber et al?

An assumption made in these papers is that all EphA receptor subtypes contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
The precise contribution of each subtype to the overall level is assumed not to be important.
In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et al.~experiments.
In half of the RGCs, we knock-in EphA3 by adding to the value of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote EphA3 ki/ki).
To add the additional knock-down of EphA4 in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all} RGCs.
Fig.\,\ref{f:GJeph}A-C shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, axon centroids and final rostral-caudal position on the tectum plotted versus the originating nasal-temporal retinal coordinate.
In A, branches are coloured in the red/green duochrome scheme according to their origin location on the retina.
In B \& C, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue.
The map of treated cells on the tectum is shifted slightly rostrally, but the shift becomes statistically insignificant at about 0.75 along the nasal-temporal axis.
While the map collapse is less abrupt than in the experimental result of \citet{brown_topographic_2000}, it is comparable with the result from \citet{simpson_simple_2011}.
Fig.\,\ref{f:GJeph}D-F shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki. Here, the treated axons are pushed further towards the rostral end of the tectum and the untreated axons are pushed caudally.
The maps are completely separated; as in the experiment, there is no `collapse point'.

The rostral movement of treated axons occurs due to their increased expression of EphA ($r_0$), which increases the interaction with the tectal rostral-caudal ligand gradient.
The opposing receptor expression ($r_2$) is unchanged by the manipulation and so those axons with increased $r_0$ expression are pulled to a more rostral position than they would otherwise take.
The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-F is unchanged, occurs due to competition with the caudally-drawn treated axons.

Figs.\,\ref{f:GJeph}G-I show the results for the Reber et al.~manipulation, in which the EphA expression is selectively knocked in in 50\% of cells and knocked down in all cells.
The collapse point has moved from about 0.75 in Fig.\,\ref{f:GJeph}C to about 0.85 (\textbf{check}), in agreement with the
experimental result.
However, the significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from our result.

\section{Discussion}

\subsection*{On chemotaxis}

Gradient based model discussed, but not simulated by \citet{hope_arrow_1976}.

\citet{nakamoto_topographically_1996} is Gradient based model.

The minimum number of gradients required to provide a coordinate system on the surface of the tectum within which axons could find a target location is $N=2$, arranged in orthogonal directions.
However if there are only two gradients, imagining that one guides axons laterally and the other guides axons in the caudal direction, then it is necessary for there to be an additional `stopping mechanism' to prevent all of the axons eventually being guided to the most lateral/caudal position.
To avoid the need for a different stopping mechanism, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$.
With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
There is evidence that multiple gene expression gradients are found in the brain, with examples of opposing pairs including XX and YY.
Adopting a pair of opposing gradients is mathematically reminiscent of the models of \citet{gierer_model_1983} which proposed... and \citet{fraser_differential_1980}.

\subsection*{Chemotaxis in isolation}

\subsection*{On competition}

Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive. An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.

Note that $\mathbf{X}$ would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$, if there was a single $l_i$ expressed in the retina and its form was $1/r_i$.
However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}C). If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.

\subsection*{Single axon behaviour}

\subsection*{Attractive receptor-ligand interactions}

\subsection*{The value of $s$}

% Bibliography here
\selectlanguage{English}
\bibliographystyle{apalike}
\bibliography{RetinoTectal}

\appendix

\section*{Method details}

\subsection*{Border effect} % border effect description
The border effect is also based on gradient following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue. For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
%
\begin{equation}\label{e:B}
B_x = \begin{cases}
        \frac{1}{2} (r-x)     & x<r \\
        \frac{1}{2} (1-r-x)    & x>1-r
\end{cases}
B_y = \begin{cases}
        \frac{1}{2} (r-y)      & y<r \\
        \frac{1}{2} (1-r-y)    & y>1-r
\end{cases}
\end{equation}
%
Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
% end border effect description

\subsection*{Initial conditions}
Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation.
Each RGC axon $k$ was assigned a random initial position coordinate:
\begin{equation}\label{e:ic}
\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
\end{equation}
where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$. Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
% end initial conditions

%
% FIGURES
%

% FIGURE 1
\begin{figure}
\includegraphics[width=\linewidth]{./images/expressions_fig.png}
\caption{Simulated retinal and tectal signalling molecule expression.
\textbf{A} A schematic of retino-tectal mapping.
%
\textbf{B} Retinal receptor expression, $r_i$. Four receptors are expressed and displayed here in two dual-colour maps. $r_0$ and $r_1$ are red and blue; $r_2$ and $r_3$ are cyan and magenta. $r_0$ increases as an exponential function of $-x$; $r_1$ increases with an exponential function of $-y$. $r_2$ and $r_3$ have the opposite sense. Retinal receptor expression is modelled with exponential functions because there is ample evidence to support this form. $N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
%
\textbf{C} Retinal ligand expression, $l_i$, is expressed in complimentary gradients \citep{hornberger_modulation_1999}.
%
\textbf{D} Tectal ligand expression $L_i$. Tectal ligand expression is shown in an exponential form, although there is less evidence to determine whether tectal expression is exponential, linear or logarithmic.
$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
}
\label{f:ex}
\end{figure}

% FIGURE 2
\begin{figure}
\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
\caption{Receptor-ligand model of chemotaxis and competition.
\textbf{A} The origin position on the tectum gives axons a colour according to a dual-colour map. Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina.
\textbf{E} Layout of axon centroids on the tectum observed in experimental results.
\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time.
\textbf{G} The arrangement of axon branches at t=1500.
\textbf{H} Selected axon locations at t=1500, with path histories in grey.
\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
\label{f:GJ}
\end{figure}

% FIGURE 3 ./scripts/fig_GJ_surgical.sh
\begin{figure}
\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
\caption{$\mathbf{GJ}$ model, surgical manipulations}
\label{f:GJsurg}
\end{figure}

% FIGURE 4
\begin{figure}
\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
\label{f:GJeph}
\end{figure}

% ANY SUPPLEMENTARY FIGURES GO AFTER THIS
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}

\begin{figure}
\includegraphics[width=\linewidth]{./images/Tissuevisb.png}
\caption{Ligand expression and gradients on the tectum for the
graft-rotate 90 degrees simulated surgical manipulation}
\label{f:trot90}
\end{figure}

\end{document}

### 2022-08-11 paper/main.tex
Current state of paper from Overleaf

@@
 \usepackage{csquotes}
+% For full page figures:
+%\usepackage[CaptionAfterwards]{fitpage}
 
@@
 
-\emph{We show that a true chemoaffinity model based on gradient following (via receptor-ligand interactions) combined with receptor-ligand-mediated axonal repulsion, in which information is only transferred by nearest-neighbour interactions, can account for a broad range of experimental results.}
+\emph{We show that a true chemoaffinity model based on gradient-following (via receptor-ligand interactions) combined with receptor-ligand-mediated axonal repulsion, in which information is only transferred by nearest-neighbour interactions, can account for a broad range of experimental manipulations of the retinotectal system.}
 
@@
 
-Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_chemoaffinity_1963}.
+% Could use chemoaffinity as a coverall theory, comprising chemospecificity and chemotaxis.
+
+Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
 %
@@
 %
-This key idea does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{gaze?} suggested that the axon could determine this direction by evaluating local information in gradients of gene expression spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-%
-Sperry's theory was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
+This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
+An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of gene expression spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
+
+Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity (\textbf{find citation}).
+In this system a two-dimensional map of the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
+The retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
+This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
+
+In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
+% And sounds?
 %
 The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
+Two classes of ephrins have been identified, labelled `A' and `B'.
+Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
+Ephrin B acts on medial-lateral ordering with an attractive interaction.
+Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
+
+Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
+The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
+Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
+Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
+He presented gradient following as a special case of a general class of \emph{potential models}.
+In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
+The axon then grows or moves to minimize $p$, which causes it to find its target location.
+This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
+This model forms a mathematical expression of Sperry's theory.
+Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
+ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
+Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
+He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
+
+Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
+These models have also helped establish how competition between axons for space in the target tissue can additionally account for many of the known effects of surgical manipulation on map development.
+
+Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
+These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
+In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
+% Not sure about this:
+However as the number of terms involved in PDEs necessarily grows large in systems representing differential chemotaxis, these systems quickly become impossible to analyse, making this potential advantage of such modelling redundant.
+Additionally, the numerical simulation of a PDE system can become intractable, because each axonal projection must be represented by several state variables, each of which require solution across the entire spatial domain (which may consist of many thousands of elements).
+% in james_modelling we did 41 projections. I found that trying to build a retinotectal system with about 100 projections was already intractable.
+
+% But know we can motivate agent based modelling because *simulation* of PDEs for differential chemotaxis is intractable.
+An alternative simulation approach is to treat each of $n$ axons as an agent that interacts with its environment and with the other $n-1$ axons.
+The computation of such a system (see for example, \citet{simpson_simple_2011}) is dominated by the $O(n^2)$ axon-axon interactions and so a limit will be approached as $n\rightarrow \infty$.
+We found that agent based simulations of $O(10^3)$ interacting axons can be achieved on a standard desktop computer whereas an equivalent PDE system would fail with about 100 axon projections.
+
+\emph{PDEs too complex (can't do maths analysis), so need agent based. Can represent more biological detail and ask how that affects the systems level.}
+
+\emph{List the questions that emerge from the discussion above}
+
+\emph{Can Gierer model explain ALL the manipulations that S and G did?}
+
+\emph{Allows us to make a specific representation of Gierer's model in which the receptor-ligand mechanism is elaborated.}
+
+\textbf{xxxxxxxxxxxxxxx}
 
@@
 
-However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemoaffinity model, that axons sample and evaluate positional information only \emph{locally}, is yet to be tested directly.
+However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemoaffinity model, that axons sample and evaluate positional information only \emph{locally}, has yet to be tested directly.
 %
@@
 %
-Nevertheless, the key assumptions of the chemoaffinity model, and specifically the proposal that axons find their way via evaluating only local information (chemotaxis), are yet to be investigated directly.
+Nevertheless, the key assumptions of the chemoaffinity model, and specifically the proposal that axons find their way by evaluating only local information (chemotaxis), are yet to be investigated directly.
 
-The aim of the current study is therefore to construct a model that represents the assumptions of the chemoaffinity model explicitly.
+The aim of the current study is therefore to construct a model that represents the assumptions of the chemotaxis model explicitly and to investigate how this mechanism behaves when combined with competitive interactions.
 
@@
 
-Indeed, our simulation results show that the local mechanisms of axonal information-processing that are assumed by the chemoaffinity model provide a remarkably robust solution to the problem of topographic map development.
+Indeed, our results show that the local mechanisms of axonal information-processing that are assumed by the chemoaffinity model provide a remarkably robust solution to the problem of topographic map development.
 
@@
 $\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{X}$ is the movement due to competitive interactions with other branches.
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue.
+$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
 
@@
 
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$ indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
+Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
 ($N$ ligands, $l_i$, are also expressed by RGCs, but these are not used for chemotaxis.)
@@
 
- We modelled competition as distance-based competitive movement vector, $\mathbf{X}$, which is `switched on' via receptor-ligand signalling:
+ We modelled competition as a distance-based competitive movement vector, $\mathbf{X}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
 %
-\begin{equation}
+\begin{equation} \label{e:X}
 \mathbf{X} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
@@
 %
-$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$) and $B_{b}$ is the set of branches that are within interaction distance of $b$.
+$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance of $b$.
 $m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
@@
 The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used the optimization to find the competition parameters that would balance this level of chemotaxis.
-To find the best combination of the remaining three parameters, we reasoned that evolution would optimize the arrangement of the wildtype pattern.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000.
+We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
+We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
 
@@
 However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other.
-For this reason, we also use the number of crossings in the fish net plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics with $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
+For this reason, we also used the number of crossings in the fish net plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
+The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
 The parameters returned by the simulated annealing were
@@
 Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (\textbf{ref supp. figure of chemo-only model}).
+The arrangement spans the full tectum due to competition between the growth cones (compare with the chemotaxis-only model in Fig.\,\ref{f:G}).
 Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
@@
 %
-Fig.\,\ref{f:GJsurg}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. \textbf{Fixme: which of the two grafts does the RH graph refer to?}
+Fig.\,\ref{f:GJsurg}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped.
+There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
+% Note: the graph shows the pattern metric for a region that covers BOTH grafts and for the region that is not the two grafts.
 
@@
 According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset) \citep{attardi_preferential_1963,schmidt_expansion_1978}.
+The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}).
 The fish-net plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
@@
 This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-Does our model reproduce the results of Brown et al and Reber et al?
+We tested whether our model would reproduce these results.
 
@@
 The precise contribution of each subtype to the overall level is assumed not to be important.
-In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we will manipulate $r_0$ to recreate the Brown et al.~and Reber et al.~experiments.
-In half of the RGCs, we knock-in EphA3 by adding to the value of $r_0$ in these cells either 0.6 (heterzygote EphA3 ki/+) or 3.0 (homozygote EphA3 ki/ki).
-To add the additional knock-down of EphA4 in \citet{reber_relative_2004}, we subtract 0.35 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A-C shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, axon centroids and final rostral-caudal position on the tectum plotted versus the originating nasal-temporal retinal coordinate.
-In A, branches are coloured in the red/green duochrome scheme according to their origin location on the retina.
-In B \& C, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue.
-The map of treated cells on the tectum is shifted slightly rostrally, but the shift becomes statistically insignificant at about 0.75 along the nasal-temporal axis.
-While the map collapse is less abrupt than in the experimental result of \citet{brown_topographic_2000}, it is comparable with the result from \citet{simpson_simple_2011}.
-Fig.\,\ref{f:GJeph}D-F shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons are pushed further towards the rostral end of the tectum and the untreated axons are pushed caudally.
-The maps are completely separated; as in the experiment, there is no `collapse point'.
+In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we manipulated $r_0$ to recreate the Brown et al.~and Reber et al.~experiments.
+In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 1.0 (heterozygote EphA3 ki/+) or 4.0 (homozygote EphA3 ki/ki).
+To include the additional knock-down of EphA4 from \citet{reber_relative_2004}, we subtracted 0.7 from $r_0$ in \emph{all} RGCs.
+Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, axon centroids and final rostral-caudal position on the tectum plotted versus the originating nasal-temporal retinal coordinate for time step 1500.
+The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
+In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue.
+The map of treated cells on the tectum was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
+However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because a bootstrapped studentized t-test of the two sets of points indicates that they are drawn from statistically different distributions (p<0.0025).
+% While the convergence of the manipulated and wildtype lines was less abrupt than in the experimental result of \citet{brown_topographic_2000}, it was comparable with the result from \citet{simpson_simple_2011}.
+Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
+Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally.
+The maps were completely separated; as in the experiment, there was no collapse point.
 
-The rostral movement of treated axons occurs due to their increased expression of EphA ($r_0$), which increases the interaction with the tectal rostral-caudal ligand gradient.
-The opposing receptor expression ($r_2$) is unchanged by the manipulation and so those axons with increased $r_0$ expression are pulled to a more rostral position than they would otherwise take.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-F is unchanged, occurs due to competition with the caudally-drawn treated axons.
+The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the tectal rostral-caudal ligand gradient.
+The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
+The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
 
-Figs.\,\ref{f:GJeph}G-I show the results for the Reber et al.~manipulation, in which the EphA expression is selectively knocked in in 50\% of cells and knocked down in all cells.
-The collapse point has moved from about 0.75 in Fig.\,\ref{f:GJeph}C to about 0.85 (\textbf{check}), in agreement with the
-experimental result.
-However, the significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from our result.
+Figs.\,\ref{f:GJeph}C show the results for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
+As in Fig.\,\ref{f:GJeph}A, the red and blue lines converge, but there is no map collapse (p<0.008).
+The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from our result.
+In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A because while $r_0$ is increased by 1 from the knock-in, the knock-down of 0.7 reduces these cells' interaction with tectal ephrin-A ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
+It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged.
+The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests some interaction effect between EphA4 and other EphA receptors that is not captured by our model.
+We address this in the next section.
+
+\subsection*{Two new mechanisms to address genetic results: Cluster size and $r_2$ collapse}
+
+The previous approach assumed that all EphA receptor subtypes (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
+This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
+We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different.
+We redefined $r_{0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of free EphA4 receptors (Fig.\,\ref{f:clustermech}B).
+Here, we assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
+The number of cis-bound EphA4 receptors is taken as $r_{A4}^{\mathrm{cis}} = 0.15\,l_0$ and $r_{\!\scriptscriptstyle A4, b} = 3.5 - r_{A4}^{\mathrm{cis}}$ where the value 0.15 is a binding weight and 3.5 represents the spatially homogeneous retinal   expression of EphA4 (Fig.\,\ref{f:clustermech}C).
+
+It is known that EphA receptors form clusters whose presence is necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
+It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
+Additionally, EphA4 receptors make `side attachments' to clusters of EphA receptors, possibly regulating the cluster size \citep{nikolov_ephephrin_2013}.
+%
+We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$.
+We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4,b}$, the density of EphA4 receptors on branch $b$:
+\begin{equation}
+    c_0 = 1/r_{\!\scriptscriptstyle A4,b}
+\end{equation}
+%
+Eq.\,\ref{e:G} was modified to include the cluster size:
+%
+\begin{equation}\label{e:Gcs}
+\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
+\end{equation}
+
+$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
+Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).
+%(This is reminiscent of the divisive relationship suggested by the work of Brown, Reber and co-workers who showed that when the ratio of EphA in treated to EphA in untreated cells drops below a threshold, the mapping system directs the cells to the same tectal location.)
+Fig\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
+Fig\,\ref{f:clustermech}D shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
+We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
+In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_1$, the effectiveness of $r_2$ is reduced to one fifth of its normal value.
+
+\textbf{Re-write this para for Fig.\,\ref{f:r2collapse}.}
+Fig.\,\ref{f:r2collapse}A shows the behaviour of the model in the EphA3 knock-in condition.
+The additional repulsion away from the caudal tectum leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
+The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for transverse RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for transverse RGCs (although transverse RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
+Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the result in the un-modified model (Fig.\,\ref{GJeph}B).
+The goal of reproducing the significant caudal shift for EphA3 ki/EphA4 kd cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
+This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
+This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
+To reproduce this shift, the additional $r_2$ collapse mechanism, driven by increased EphA expression and knocked-down EphA4, is required.
+$r_2$ collapse occurs across most of the N--T axis in the knock-in/knock-down condition \textbf{Add a graph to Fig.\,\ref{f:clustermech} that shows the region affected}.
+The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
+Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
+This gives a largely unperturbed map with \textbf{just a couple of features for us to adequately explain}.
+
+%To more fully reproduce the experimental results of \citet{reber_relative_2004}, we refer to Fig.\,\ref{f:ephadiv}D.
+If the nasal-temporal signal curve for the knock-down condition was similar to the wildtype curve, and all curves tended towards a common signal strength for RGCs with a temporal origin, we would expect to retain the rostral shift in knock-in cells, an enhanced rostral shift for knock-in/knock-down cells and no shift for knock-down cells.
+
+%Now find the sums, starting from
+
+%\begin{equation}
+%    c_0 = 3*(0.2 + \frac{(r_{A4}^{cis}-0.8) %\exp(0.5(r_0-1.31))}{10\,(r_{\!\scriptscriptstyle A4})^3 + 2})
+%\end{equation}
+
+%There was a reason for the offset 0.2 which I need to remember/articulate. The 1.3 is an offset to ensure that the cluster size tends towards 0.2 for nasal RGCs.
 
@@
 
+% Summarize the findings...
+
+Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
+If not, does the addition of competition resolve the remaining problems?
+Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
+To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
+The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold.
+These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals.
+The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development.
+The model also accounted for some aspects of the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}, but distinct aspects of the results were not accounted for, suggesting that additional mechanisms may be in play.
+\textbf{So we added to the model.}
+
+%Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate.
+%Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations. Finish summarising results...
+
+% Place findings in context... (starting with the next section)
+
 \subsection*{On chemotaxis}
 
-Gradient based model discussed, but not simulated by \citet{hope_arrow_1976}.
+The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
+In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
+The axon then grows or moves to minimize $p$, which causes it to find its target location.
+This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
+Gierer recognised that tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
+He addressed this problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and noting that this was equivalent to a more direct fibre-fibre competition.
 
-\citet{nakamoto_topographically_1996} is Gradient based model.
+Because of the discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
+Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
+This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
 
-The minimum number of gradients required to provide a coordinate system on the surface of the tectum within which axons could find a target location is $N=2$, arranged in orthogonal directions.
-However if there are only two gradients, imagining that one guides axons laterally and the other guides axons in the caudal direction, then it is necessary for there to be an additional `stopping mechanism' to prevent all of the axons eventually being guided to the most lateral/caudal position.
-To avoid the need for a different stopping mechanism, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$.
+%\citet{nakamoto_topographically_1996} discussed gradient based models  stems in a large part from \citet{gierer_directional_1987}.
+Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
+Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
+Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
+Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$.
 With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain, with examples of opposing pairs including XX and YY.
-Adopting a pair of opposing gradients is mathematically reminiscent of the models of \citet{gierer_model_1983} which proposed... and \citet{fraser_differential_1980}.
+There is evidence that multiple gene expression gradients are found in the brain, with examples of opposing pairs including XX and YY \citep{cang_developmental_2013}.
+Adopting a pair of opposing gradients is mathematically reminiscent of/actually implements the models of \citet{gierer_model_1983} (which proposed the existence of potential functions) and \citet{fraser_differential_1980}.
+
+\emph{(Place this:)} The model required opposing pairs of graded signalling molecules for the chemotaxis component of the model.
+The competition model, in which a close-range repulsion was enabled by receptor-ligand binding between RGC axons required the participation of only a single pair of opposed signalling molecules in the retina.
+
+\subsection*{Not ephrins}
+
+Placeholder section.
+
+A number of other signalling molecules have been identified, which are be expressed in gradients that are similar to those of the ephrins.
+The expression of \emph{Wnt} and its ligand \emph{Ryk} matches that of Eph B and ephrin B \citep{schmitt_wntryk_2006}.
+Wnt is expressed at a high level towards the medial tectum; Ryk is expressed in a low to high gradient from the dorsal to ventral retina.
+Wnt-Ryk signalling induces repulsion away from areas of high Ryk expression, so RGCs from ventral retina should be repelled from the medial tectum.
+This acts in direct opposition to the effect of ephrin B signalling, which induces attraction of ventral RGCs towards the medial tectum \citep{hindges_ephb_2002}.
+% In discussion will go into why that fails to provide a counter gradient for EphB/ephrin B. See expression_profiles.svg.
+
+Neogenin, RGM (repulsive guidance molecule) and netrin are a further set of signalling molecules that have been shown to act in the retinotectal system.
+Neogenin is a receptor which is found in a graded distribution in the retina.
 
 \subsection*{Chemotaxis in isolation}
+
+Describe results, shown in supplementary figure of what happens with only chemotaxis.
 
@@
 
+Why does competition improve the surgical manipulation behaviour? Genetic manipulations?
+
 \subsection*{Single axon behaviour}
+
+The chemoaffinity part of this model ensures that single axons will be drawn to approximately the correct location.
+The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
+This is in agreement with \citet{gosse_retinotopic_2008}... \emph{Do a similar linear fit to our models match Gosse Fig 2d/e}.
 
@@
 
+Discuss the form of tectal ligand expression that is required by this model if receptor-ligand interactions are attractive (under the condition that retinal expression is exponential).
+In this case, the expression should follow a logarithmic increase.
+This ties in with the finding that only one pair of receptor-ligand interactions is required to mediate competition; this leaves the other axis free to be either repulsive, or attractive (as EphB).
+
+\subsection*{Genetic manipulations}
+
+ An assumption of the simulation of genetic manipulations was that there is no difference between the EphA receptor sub-types.
+
+\subsection*{Robustness to noise}
+
 \subsection*{The value of $s$}
+
+This is worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
+Esesentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
+
+% Any unexpected results?
+\subsection*{Other models}
+
+Have covered Gierer in introduction.
+Also, be sure to discuss \citet{naoki_revisiting_2017}, which is a gradient model with just two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level.
+
+% Limitations/weaknesses of research
+\subsection*{Limitations}
+
+The sensitivity of the model to the interaction radius - receptor-ligand binding really quite close range, but the interaction radius in this model is required to be of the order of the target axon-axon spacing.
+
+Low resolution of gradient (limited by size of grid).
+
+% Restate significant findings and implications
 
@@
 \subsection*{Border effect} % border effect description
-The border effect is also based on gradient following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
+The border effect is also based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
 For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
@@
 
+\subsection*{Bootstrap analyses of R-C tectal position vs.~N-T retinal origin}
+
+To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R-C positions obtained for the $n=20$ different dorso-ventral origins for each N-T location.
+The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
+Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
+The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
+%
+$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
+%
+where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
+
+A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
+The transformations are defined as:
+%
+$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
+%
+where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
+$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$.
+The Studentized difference of the means of the resamples is computed:
+
+$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
+
+If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
+If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$.
+We made 1000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL).
+The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
+
+The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
+%% End bootstrap desc
+
 %
@@
 % FIGURE 1
+% Expand panel A to be able to lose B-D.
 \begin{figure}
@@
 \caption{Receptor-ligand model of chemotaxis and competition.
-\textbf{A} The origin position on the tectum gives axons a colour according to a dual-colour map.
+\textbf{A} The origin position on the retina gives axons a colour according to a dual-colour map.
 Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
 \textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina.
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results.
+\textbf{E} Layout of axon centroids on the tectum observed in experimental results \textbf{expectation or prediction}.
 \textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time.
+\textbf{two subgraphs better}
 \textbf{G} The arrangement of axon branches at t=1500.
@@
 % FIGURE 3 ./scripts/fig_GJ_surgical.sh
+% Could split into retina and tectum or
+% a: rotation/graft (as hand waved by Gierer)
+% b: ablation (as simulated in 1D by Gierer)
+% c: mismatch/compound (as extensions of Gierer)
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GJ_surgical.png}
-\caption{$\mathbf{GJ}$ model, surgical manipulations}
+{\centering
+\includegraphics[width=0.75\linewidth]{./images/fig_GJ_surgical.png}
+\par }
+\caption{Performance of the model with surgical manipulations.
+Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
+\textbf{A.} Graft rotation.
+An 8$\times$8 square patch of tectum has been rotated by 90$\degree$, altering the tectal ligand expression pattern.
+High gradients form around the new discontinuities at the edge of the patch (see Fig\,\ref{f:trot90}), driving axons to follow the rotation of the patch.
+The left graph shows the axon centroid positions after 1500 time steps, the smaller inset graph shows the (idealised) anticipated layout seen in experiments \citep{levine_deployment_1974}.
+The right-hand graph shows the time-course of the pattern quality metric, $\epsilon$ (i.e.~how closely the centroid pattern matches the inset) for the entire tectum (blue), the graft region (yellow) and the surround (red).
+The axons within the graft are slightly less well ordered than those outside.
+\textbf{B} Shows the same type of graft rotation as \textbf{A} for a 180$\degree$ rotation \citep{yoon_retention_1973}.
+\textbf{C.} A simulated graft swap experiment.
+Two 12$\times$4 rectangular regions of tectum are cut out and grafted back in swapped positions \citep{hope_arrow_1976,gaze_visuotectal_1983}.
+While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
+Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.
+\textbf{D.} The temporal half of the retina is removed in this manipulation.
+Axons from the nasal half grow onto an unmodified tectum.
+This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
+In this work, it is the competition that achieves this effect.
+In panels D-G, the axon centroid graphs are plotted as in A-C, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
+\textbf{E}.
+Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
+As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
+\textbf{F.} A `mismatch' experiment \citep{horder_retention_1971}.
+The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
+As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
+\textbf{G} Compound eye.
+If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
+The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
 \label{f:GJsurg}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{$\mathbf{GJ}$ model, Brown et al./Reber et al.~manipulations.}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations.png}
+\caption{Three genetic manipulations simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
+In each manipulation, the left graph shows axon branch agents at the end of a simulation ($t=1500$), the centre graph shows axon centroids (also at simulation end), with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
+In the right hand graph, the final rostro-caudal locations on the tectum of axons from each naso-temporal origin in the retina are plotted as violin plots (with bars indicating the maximum and minimum datum in each group and shading indicating the data distribution).
+Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
+\textbf{A} We simulated the heterozygous EphA3 knock-in (EphA3 ki/+) of \citet{brown_topographic_2000} by selecting half of the retinal cells at random, and in these cells adding the value 0.8 to $r_0$ (which, as its expression increases towards the temporal retina, is our analogue for EphA).
+The affected cells experienced altered interaction with tectal ligand gradients and were shifted rostrally.
+The rostral shift was largest for nasal cells bound for the caudal tectum.
+These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
+Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (and hence the $r_0$-$L_0$ interaction is minimal here).
+However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis.
+\textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
+Here, we observed a separation into two maps, one for the wildtype cells and one for the manipulated cells, in agreement with \citet{brown_topographic_2000} (inset is from their Fig.\,5C).
+\textbf{C} Simulation of EphA3 ki/+ selective knock-in and EphA4 knock-down (all cells) from \citet{reber_relative_2004}.
+The knock-down was applied by reducing $r_0$ in all cells by 1.2.
+The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in the experimental result.
+In fact, the selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in of 0.8.}
 \label{f:GJeph}
@@
 
+% FIGURE 5
+\begin{figure}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_explanation.png}
+\caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
+To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrin ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA ($r_2$).
+\textbf{A} As described in the text, EphA ($r_0$) is expressed according to $1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position in this case.
+Knock-in is simulated simply by adding 0.8 to this function (giving the purple curve).
+Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
+\textbf{B} EphA4 expression is a new feature in this model extension.
+Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA can bind to EphA4, modelled here as a multiplicative interaction to give cis-bound EphA4 (green dashed curve) and un-bound EphA4 (green solid curve).
+We model genetic knock-down of EphA4 by subtracting a value of 2.1 from the constant EphA4 value, and re-apply the multiplication by retinal ephrinA to give the cyan curves.
+\textbf{NB: curves not quite right here - there should be a crossover of the filled cyan curve and the dotted red threshold near the T side; this is what happens in the sim.
+This it's due to an off-by-one discrepancy in x position.} \textbf{C} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}$.
+Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced with a combination of both knock-in and knock-down.
+\textbf{D} We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=1.6$ (red dotted line in A) and that $r_{A4}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
+Under these conditions, $r_2$ is reduced to one fifth of its normal value.
+\textbf{Note for Stuart: The graphs are as-generated by a Python script.
+I'll fix the legend boxes etc once they are final.}}
+\label{f:clustermech}
+\end{figure}
+
+%%% Caption for clustermechsize only model: \textbf{B} Model results.
+i) EphA3 knock-in result for ki=1.0.
+Inset shows experimental results from Fig.\,5B of \citet{brown_topographic_2000}.
+The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment.
+ii) The double knock-in result (Fig.\,5C of \citet{brown_topographic_2000}) agrees well with the experimental result, showing two, separated maps iii) When EphA is knocked in and EphA4 is simultaneously knocked-down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
+Although less dramatic than the experimental result of \citet{reber_relative_2004} the movement is in the correct direction, and the overlap of positions (the `collapse point') moves to about 90\% of the N--T axis.
+iv) Uniform knock-down of EphA4 in all cells does not perturb the mapping.
+
+% FIGURE 6
+\begin{figure}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
+\caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}.
+\textbf{A} EphA3 knock-in result for ki=0.8.
+The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment \textbf{Seb to check stats on this}.
+\textbf{B} The double knock-in result shows excellent agreement with the experimental result.
+\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked-down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
+Additionally, the combined knock-in and knock-down triggers the $r_2$ receptors to become one fifth as effective, leading to a degraded signal to move along gradients towards the caudal tectum.
+Consequently, the red curve is moved much more caudally, in agreement with \citet{reber_relative_2004}.
+\textbf{D} Uniform knock-down of EphA4 in all cells does not perturb the mapping to a significant extent.
+\textbf{TODO: Answer question as to why this curve is drawn rostrally and describe the discontinuity at 0.25 on N--T axis, at which the collapse mechanism is being engaged.
+Need to figure out the slight discrepancy between signals in Fig.\,\ref{f:clustermech} and those in the simulation.}}
+\label{f:r2collapse}
+\end{figure}
+
 % ANY SUPPLEMENTARY FIGURES GO AFTER THIS
-\renewcommand{\thefigure}{S\arabic{figure}}
-\setcounter{figure}{0}
+%\renewcommand{\thefigure}{S\arabic{figure}}
+%\setcounter{figure}{0}
 
+% FIGURE S1
 \begin{figure}
 \includegraphics[width=\linewidth]{./images/Tissuevisb.png}
-\caption{Ligand expression and gradients on the tectum for the
+\caption{(Supplementary) Ligand expression and gradients on the tectum for the
 graft-rotate 90 degrees simulated surgical manipulation}
@@
 
+% DITCH THIS ONE
+%
+% FIGURE S2
+%%\begin{figure}
+%%\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2_exit_true.png}
+%%% Note MANUAL reference to Fig. 2 here (otherwise it says Fig S2):
+%%\caption{Receptor-ligand model of chemotaxis alone. Panels A-H show results as in Fig\,\textbf{2}. Briefly:
+%%\textbf{A} Retinal cell position colour scheme.
+%%\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units.
+In this chemotaxis-only model, which differs from that shown in Fig\,\ref{f:GJ} only by having the parameter $m_X$ set to 0 in Eq.\,\ref{e:X}, variable interactions with the gradients cause each RGC to find a location that preserves the topographic ordering, but the absence of competition means that axons fail to extend across the full tectum.
+%%\textbf{E} Anticipated layout.
+%%\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time.
+%%\textbf{G} The arrangement of axon branches at t=1500.
+%%\textbf{H} Selected axon locations at t=1500, with path histories in grey.
+%%\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
+%%\label{f:G}
+%%\end{figure}
+
 \end{document}

### 2022-09-02 paper/main.tex
Friday WIP

@@
 %%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares (Fig.\,\ref{f:ex}B-D).
+We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
 The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
@@
 Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, are also expressed by RGCs, but these are not used for chemotaxis.)
+($N$ ligands, $l_i$, are also expressed by RGCs, but these are not used by this chemotaxis model.) 
 The expression of each receptor varies only with respect to one spatial dimension.
@@
 Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}B).
+We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
 
 Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Fig.\,\ref{f:expressions} summarises results from the literature).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}D).
+Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Fig.\,\ref{f:ex} summarises results from the literature).
+We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
 
@@
 Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (compare with the chemotaxis-only model in Fig.\,\ref{f:G}).
+The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
 Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
@@
 After finding an optimum set of parameters for the wildtype arrangement, we tested our model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
-Figs.~\ref{f:GJsurg}A and \ref{f:GJsurg}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
+Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
 %
@@
 Long connecting lines in the fish net show that a number of axons from near the edges of the tectum have been drawn into the graft region. 
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg}A \& B.
+The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
 We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red). 
@@
 %
-Fig.\,\ref{f:GJsurg}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped.
+Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. 
 There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
@@
 
-Fig.\,\ref{f:GJsurg}D shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
+Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
 According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
@@
 %
-Fig.\,\ref{f:GJsurg}E shows the result of a tectal ablation.
+Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
 Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
@@
 A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-Our model replicates this manipulation (Fig.\,\ref{f:GJsurg}F) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg}E; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
+Our model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
 %
@@
 Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg}G).
+These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
 
@@
 This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-We tested whether our model would reproduce these results.
+For our model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
+iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
 
@@
 The precise contribution of each subtype to the overall level is assumed not to be important. 
-In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we manipulated $r_0$ to recreate the Brown et al.~and Reber et al.~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 1.0 (heterozygote EphA3 ki/+) or 4.0 (homozygote EphA3 ki/ki).
-To include the additional knock-down of EphA4 from \citet{reber_relative_2004}, we subtracted 0.7 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, axon centroids and final rostral-caudal position on the tectum plotted versus the originating nasal-temporal retinal coordinate for time step 1500.
+In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
+In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
+To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
+Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
 The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
 In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue. 
-The map of treated cells on the tectum was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because a bootstrapped studentized t-test of the two sets of points indicates that they are drawn from statistically different distributions (p<0.0025).
+The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
+The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
+However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
+Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
+A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that they are drawn from statistically identical distributions.
 % While the convergence of the manipulated and wildtype lines was less abrupt than in the experimental result of \citet{brown_topographic_2000}, it was comparable with the result from \citet{simpson_simple_2011}.
+
 Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally.
-The maps were completely separated; as in the experiment, there was no collapse point.
-
+Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
+The maps were completely separated. As in the experiment, there was no collapse point.
 The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the tectal rostral-caudal ligand gradient. 
@@
 The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
+Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
 
 Figs.\,\ref{f:GJeph}C show the results for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines converge, but there is no map collapse (p<0.008).
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from our result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A because while $r_0$ is increased by 1 from the knock-in, the knock-down of 0.7 reduces these cells' interaction with tectal ephrin-A ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
+As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
+The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
+In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
+
 It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests some interaction effect between EphA4 and other EphA receptors that is not captured by our model.
-We address this in the next section.
+The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by this model.
+We address possible interaction effects in the next section.
 
-\subsection*{Two new mechanisms to address genetic results: Cluster size and $r_2$ collapse}
+\subsection*{Two mechanisms to address the genetic results: Cluster size and $r_2$ collapse}
 
@@
 We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different. 
-We redefined $r_{0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of free EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-Here, we assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
-The number of cis-bound EphA4 receptors is taken as $r_{A4}^{\mathrm{cis}} = 0.15\,l_0$ and $r_{\!\scriptscriptstyle A4, b} = 3.5 - r_{A4}^{\mathrm{cis}}$ where the value 0.15 is a binding weight and 3.5 represents the spatially homogeneous retinal   expression of EphA4 (Fig.\,\ref{f:clustermech}C).
+We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
+EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
+We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
+The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
+Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
 
@@
 We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$. 
-We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4,b}$, the density of EphA4 receptors on branch $b$:
+We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4}$, the density of EphA4 receptors:
 \begin{equation}
-    c_0 = 1/r_{\!\scriptscriptstyle A4,b}
+    c_0 = 1/r_{\!\scriptscriptstyle A4}
 \end{equation}
@@
 %(This is reminiscent of the divisive relationship suggested by the work of Brown, Reber and co-workers who showed that when the ratio of EphA in treated to EphA in untreated cells drops below a threshold, the mapping system directs the cells to the same tectal location.)
-Fig\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-Fig\,\ref{f:clustermech}D shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
+Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
+The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
+To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
+
+Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
 We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
-In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_1$, the effectiveness of $r_2$ is reduced to one fifth of its normal value.
+In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
+Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
+In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
+If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
+If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
 
-\textbf{Re-write this para for Fig.\,\ref{f:r2collapse}.}
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for transverse RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for transverse RGCs (although transverse RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the result in the un-modified model (Fig.\,\ref{GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/EphA4 kd cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
+Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
+The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
+The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
+Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
+The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
 This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
 This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift, the additional $r_2$ collapse mechanism, driven by increased EphA expression and knocked-down EphA4, is required.
-$r_2$ collapse occurs across most of the N--T axis in the knock-in/knock-down condition \textbf{Add a graph to Fig.\,\ref{f:clustermech} that shows the region affected}.
+To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
+Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
+Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin in temporal (above 0.56 N--T).
+This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
 The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
 Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with \textbf{just a couple of features for us to adequately explain}.
+This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
 
-%To more fully reproduce the experimental results of \citet{reber_relative_2004}, we refer to Fig.\,\ref{f:ephadiv}D.
-If the nasal-temporal signal curve for the knock-down condition was similar to the wildtype curve, and all curves tended towards a common signal strength for RGCs with a temporal origin, we would expect to retain the rostral shift in knock-in cells, an enhanced rostral shift for knock-in/knock-down cells and no shift for knock-down cells.
-
-%Now find the sums, starting from
-
-%\begin{equation}
-%    c_0 = 3*(0.2 + \frac{(r_{A4}^{cis}-0.8) %\exp(0.5(r_0-1.31))}{10\,(r_{\!\scriptscriptstyle A4})^3 + 2})
-%\end{equation}
-
-%There was a reason for the offset 0.2 which I need to remember/articulate. The 1.3 is an offset to ensure that the cluster size tends towards 0.2 for nasal RGCs.
+Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
 
@@
 The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development.
-The model also accounted for some aspects of the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}, but distinct aspects of the results were not accounted for, suggesting that additional mechanisms may be in play.
-\textbf{So we added to the model.}
+
+The model also accounted for some aspects of the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
+We identified four features of the genetic experimental results which should be reproduced and explained by an extended model. 
+These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
+We then showed that although the unmodified model successfully reproduced item ii, it failed with the other three features.
+
+To extend the model, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
 
@@
 The competition model, in which a close-range repulsion was enabled by receptor-ligand binding between RGC axons required the participation of only a single pair of opposed signalling molecules in the retina.
+
+\subsection*{Ephrins}
+A bit of discussion on my treatment of Ephrins, esp.~in the extended model with EphA4.
 
@@
 
+I've done sims that introduce noise, so find these and discuss. The headline is that the system is very robust to noise.
+
 \subsection*{The value of $s$}
@@
 Low resolution of gradient (limited by size of grid).
+
+No mechanism for $r_2$ collapse.
 
@@
 If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$. 
-We made 1000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL).
+We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL). 
 The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
@@
 \includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Simulated retinal and tectal signalling molecule expression.
-\textbf{A} A schematic of retino-tectal mapping.
+\caption{Schematic diagrams of retino-tectal mapping and gene expression.
+Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
 %
-\textbf{B} Retinal receptor expression, $r_i$.
-Four receptors are expressed and displayed here in two dual-colour maps.
-$r_0$ and $r_1$ are red and blue; $r_2$ and $r_3$ are cyan and magenta.
-$r_0$ increases as an exponential function of $-x$; $r_1$ increases with an exponential function of $-y$.
-$r_2$ and $r_3$ have the opposite sense.
-Retinal receptor expression is modelled with exponential functions because there is ample evidence to support this form.
-$N$: nasal, $T$: temporal, $D$: dorsal, $V$: ventral.
+\textbf{A} Tectal ligand expression $L_i$.
+Tectal ligand expression patterns are assumed to increase exponentially in one direction only across the tectal surface.
+$L_0$ (or ephrin A) increases in a caudal to rostral direction (i.e.
+with increasing $y$; $x$ and $y$ directions are indicated with arrows).
+$L_2$ increases in the opposing sense, with $-y$.
+$L_1$ and $L_3$ form counter-gradients in the $\pm x$ directions.
+\emph{L: lateral, M: medial, C: caudal, R: rostral}.
 %
-\textbf{C} Retinal ligand expression, $l_i$, is expressed in complimentary gradients \citep{hornberger_modulation_1999}.
+\textbf{B} Retinal receptor and ligand expressions.
+Four receptors, $r_i$, are expressed across the retina.
+$r_0$ increases as an exponential function of $-x$ ; $r_1$ increases with an exponential function of $-y$.
+$r_2$ and $r_3$ have the opposite sense and increase with positive $x$ and $y$, respectively.
+Retinal ligand expression, $l_i$, is expressed in two complimentary gradients \citep{hornberger_modulation_1999}.
+ \emph{N: nasal, T: temporal, D: dorsal, V: ventral}.
 %
-\textbf{D} Tectal ligand expression $L_i$.
-Tectal ligand expression is shown in an exponential form, although there is less evidence to determine whether tectal expression is exponential, linear or logarithmic.
-$L$: lateral, $M$: medial, $C$: caudal, $R$: rostral.
 }
@@
 \textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina. 
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results \textbf{expectation or prediction}.
+\textbf{E} Layout of axon centroids on the tectum observed in experimental results (that is, correctly ordered into a perfect map of the simulated, square retina). 
 \textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time. 
-\textbf{two subgraphs better}
 \textbf{G} The arrangement of axon branches at t=1500. 
@@
 
-% FIGURE 3 ./scripts/fig_GJ_surgical.sh
+% FIGURE 3 ./scripts/fig_GJ_surgical_grafts.svg (graphs generated with fig_GJ_surgical.sh)
 % Could split into retina and tectum or 
@@
 \begin{figure}
-{\centering
-\includegraphics[width=0.75\linewidth]{./images/fig_GJ_surgical.png}
-\par }
-\caption{Performance of the model with surgical manipulations.
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_grafts.png}
+\caption{Performance of the receptor-ligand model with surgical graft manipulations.
 Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
@@
 While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
-Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.
-\textbf{D.} The temporal half of the retina is removed in this manipulation.
-Axons from the nasal half grow onto an unmodified tectum.
-This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
-In this work, it is the competition that achieves this effect.
-In panels D-G, the axon centroid graphs are plotted as in A-C, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
-\textbf{E}.
-Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
-As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
-\textbf{F.} A `mismatch' experiment \citep{horder_retention_1971}.
-The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
-As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
-\textbf{G} Compound eye.
-If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
-The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
-\label{f:GJsurg}
+Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.}
+\label{f:GJsurg_grafts}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_ablations.png}
+\caption{Retinal and tectal ablation simulations.
+\textbf{A} The temporal half of the retina is removed in this manipulation.
+Axons from the nasal half grow onto an unmodified tectum.
+This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
+In this work, it is the competition that achieves this effect.
+The axon centroid graphs are plotted as in Fig.\,\ref{f:GJsurg_grafts}, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
+\textbf{B} Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
+As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
+\textbf{C} A `mismatch' experiment \citep{horder_retention_1971}.
+The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
+As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
+\textbf{D} Compound eye.
+If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
+The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
+\label{f:GJsurg_ablate}
+\end{figure}
+
+% FIGURE 5
+\begin{figure}
 \includegraphics[width=0.95\linewidth]{./images/EphA_manipulations.png}
-\caption{Three genetic manipulations simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-In each manipulation, the left graph shows axon branch agents at the end of a simulation ($t=1500$), the centre graph shows axon centroids (also at simulation end), with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
-In the right hand graph, the final rostro-caudal locations on the tectum of axons from each naso-temporal origin in the retina are plotted as violin plots (with bars indicating the maximum and minimum datum in each group and shading indicating the data distribution).
+\caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
+In each manipulation, the left graph shows axon branch agents (colour coded by retinal origin) at the end of a simulation ($t=1500$), the centre graph shows corresponding axon centroids, with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
+In the right hand graph, the final tectal rostro-caudal locations of axons from each naso-temporal origin in the retina are plotted as violin plots (bars indicate the max/min datum in each group and shading indicates the data distribution).
 Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
@@
 These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
-Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (and hence the $r_0$-$L_0$ interaction is minimal here).
-However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis.
+Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (thus the $r_0$-$L_0$ interaction is minimal).
+However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis, as seen in the inset graph.
 \textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
@@
 The knock-down was applied by reducing $r_0$ in all cells by 1.2.
-The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in the experimental result.
-In fact, the selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in of 0.8.}
+The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in experiments.
+Instead, selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in.}
 \label{f:GJeph}
@@
 
-% FIGURE 5
+% FIGURE 6
 \begin{figure}
@@
 \caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
-To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrin ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA ($r_2$).
-\textbf{A} As described in the text, EphA ($r_0$) is expressed according to $1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position in this case.
-Knock-in is simulated simply by adding 0.8 to this function (giving the purple curve).
+To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and for the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrinA ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA (i.e.~with $r_2$).
+\textbf{A} \emph{EphA/ephrinA expression.} As described in the text, EphA ($r_0 \equiv \sum_{i\neq 4}\mathrm{EphA}i$) is expressed according to $r_0 = 1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position.
+Knock-in is simulated by adding 0.8 to this function (giving the purple curve).
 Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
-\textbf{B} EphA4 expression is a new feature in this model extension.
-Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA can bind to EphA4, modelled here as a multiplicative interaction to give cis-bound EphA4 (green dashed curve) and un-bound EphA4 (green solid curve).
-We model genetic knock-down of EphA4 by subtracting a value of 2.1 from the constant EphA4 value, and re-apply the multiplication by retinal ephrinA to give the cyan curves.
-\textbf{NB: curves not quite right here - there should be a crossover of the filled cyan curve and the dotted red threshold near the T side; this is what happens in the sim.
-This it's due to an off-by-one discrepancy in x position.} \textbf{C} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}$.
-Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced with a combination of both knock-in and knock-down.
-\textbf{D} We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=1.6$ (red dotted line in A) and that $r_{A4}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
-Under these conditions, $r_2$ is reduced to one fifth of its normal value.
-\textbf{Note for Stuart: The graphs are as-generated by a Python script.
-I'll fix the legend boxes etc once they are final.}}
+A threshold, $h_0$, used by the $r_2$ collapse mechanism is shown as a red dotted line, with a green dash-dot line indicating the intersection of $\Sigma\mathrm{EphA}$ and $h_0$.
+\textbf{B} \emph{EphA4 expression.} Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA ($l_0$) can bind to EphA4 ($r_{A4}$), modelled here as a multiplicative interaction to give cis-bound EphA4, $r_{A4}^{\mathrm{cis}} = w_{A4} \times r_{A4} \times l_0$ (green dashed curve), where $w_{A4}$ is a binding weight described in the text.
+The remaining, un-bound, free EphA4 receptors, $r_{A4}^{\mathrm{free}} = r_{A4} - r_{A4}^{\mathrm{cis}}$ (green solid curve).
+We model genetic knock-down of EphA4 by subtracting 2.1 from $r_{A4}$, then recomputing $r_{A4}^{\mathrm{cis}}$ and $r_{A4}^{\mathrm{free}}$ to give the cyan curves.
+\textbf{C} \emph{Cluster size affects signal strength.} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}^{\mathrm{free}}$.
+Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced by combined knock-in and knock-down.
+This cluster size-modulated signal, $s_0$, is smaller in magnitude than the signal for $r_0$ in the original model (for which $s_0 \equiv r_0$).
+To compensate, $r_2$ is divided by 2, which gives the correct behaviour in a wildtype system.
+\textbf{D} \emph{$r_2$ collapse.} i) We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=2$ (red dotted line in A) and that $r_{A4}^{\mathrm{free}}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
+Under these conditions, $r_2$ is reduced to one fifth of the value it would otherwise take.
+ii) The regions of $r_2$ collapse are illustrated.
+No collapse occurs for wildtype EphA and EphA4 expressions; in cells with only EphA4 knock-down, $r_2$ collapse occurs for temporal cells (those more temporal than the dash-dot line in A).
+In cells with both EphA3 knock-in and EphA4 knock-down, the collapse mechanism occurs regardless of retinal origin.}
 \label{f:clustermech}
@@
 The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment.
-ii) The double knock-in result (Fig.\,5C of \citet{brown_topographic_2000}) agrees well with the experimental result, showing two, separated maps iii) When EphA is knocked in and EphA4 is simultaneously knocked-down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
+ii) The double knock-in result (Fig.\,5C of \citet{brown_topographic_2000}) agrees well with the experimental result, showing two, separated maps iii) When EphA is knocked in and EphA4 is simultaneously knocked down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
 Although less dramatic than the experimental result of \citet{reber_relative_2004} the movement is in the correct direction, and the overlap of positions (the `collapse point') moves to about 90\% of the N--T axis.
@@
 
-% FIGURE 6
+% FIGURE 7
 \begin{figure}
@@
 \textbf{A} EphA3 knock-in result for ki=0.8.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment \textbf{Seb to check stats on this}.
-\textbf{B} The double knock-in result shows excellent agreement with the experimental result.
-\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked-down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
-Additionally, the combined knock-in and knock-down triggers the $r_2$ receptors to become one fifth as effective, leading to a degraded signal to move along gradients towards the caudal tectum.
-Consequently, the red curve is moved much more caudally, in agreement with \citet{reber_relative_2004}.
-\textbf{D} Uniform knock-down of EphA4 in all cells does not perturb the mapping to a significant extent.
-\textbf{TODO: Answer question as to why this curve is drawn rostrally and describe the discontinuity at 0.25 on N--T axis, at which the collapse mechanism is being engaged.
-Need to figure out the slight discrepancy between signals in Fig.\,\ref{f:clustermech} and those in the simulation.}}
+The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment (inset).
+This is caused by the cluster size effect, in which the signal increase due to the knock-in is reduced for temporal RGCs compared with nasal cells (See Fig.\,\ref{f:clustermech}C, green/purple solid curves).
+%FOR TEXT: It is confirmed by an increased achieved significance level (ASL) from a bootstrapped, studentized t-test comparing the red and blue distributions at each N--T location. 
+\textbf{B} The cluster size effect does not change the double knock-in result, which shows excellent agreement with experiment (inset). 
+\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked down, cells on the red curve experience a combination of interactions: First, the increased interaction with ephrinA that gave the result in A is still in effect.
+Second, all cells on the red curve experience $r_2$ collapse, whereas only temporal RGCs on the blue curve experience this effect.
+$r_2$ receptors are more numerous on nasal RGCs and so the collapse in $r_2$ generates a substantial rostral movement of the red curve, in agreement with \citet{reber_relative_2004}.
+\textbf{D} Uniform knock-down of EphA4 in all cells (with no manipulation of EphA3) does not perturb the mapping to a significant extent, although the effect of the EphA4 knock-down on the cluster size mechanism causes a slight rostral shift in all cells.}
 \label{f:r2collapse}

### 2022-09-02 analysis/bs_manip_vs_wt.py

@@ -116,7 +116,7 @@ import h5py
 # simulations for the Genetic manipulation sims to create these.
 
 # 0: ORIGINAL model 1: Clustersize model 2: Clustersize + r2 collapse
-model_type = 2
+model_type = 0
 doitall = 0
 if doitall:
     filebases = ["../rcnt/j4_ee_GJ_best_1_eph_ki-wt_exit_true_steps_1500",

### 2022-09-02 misc/examineRetEph.py

@@ -122,6 +122,17 @@ def examineRetEph():
     ax1.plot (x, EphAx+ki, linestyle='-', color=clr_knockin, label='EphA3 knock-in ($r_0 + ki$)')
     ax1.plot ([0,1], [h_0, h_0], linestyle=':', color=clr_red, label='$\Sigma$EphA threshold ($h_0$)')
 
+    # Could look at cis/free EphAx in future, but avoid this complexity at present
+    study_EphAx_binding = 0
+    if study_EphAx_binding:
+        EphAx_cis = w_EphAx * EphAx * ephrinA
+        ax1.plot (x, EphAx_cis, linestyle='--', color=clr_wt, label='$\Sigma$EphA cis-bound')
+        ax1.plot (x, EphAx - EphAx_cis, linestyle='-.', color=clr_wt, label='$\Sigma$EphA free')
+        EphAx_ki_cis = w_EphAx * (EphAx+ki) * ephrinA
+        ax1.plot (x, EphAx_ki_cis, linestyle='--', color=clr_knockin, label='$\Sigma$EphA ki cis-bound')
+        ax1.plot (x, (EphAx+ki) - EphAx_ki_cis, linestyle='-.', color=clr_knockin, label='$\Sigma$EphA ki free')
+
+
     ax1.legend()
     ax1.set_xlabel('N {0} retina {0} T'.format(u"\u27f6"))
     ax1.set_ylabel('Expression')
@@ -170,6 +181,11 @@ def examineRetEph():
     ax21.plot (x, signal (EphAx,    clustersz_simple(EphAx, 0,  EphA4_free_kd, _p_epha4_kd)), linestyle='-', color=clr_knockdown)
     ax21.plot (x, signal (EphAx+ki, clustersz_simple(EphAx, ki, EphA4_free_kd, _p_epha4_kd)), linestyle='-', color=clr_knockdown)
     ax21.plot (x, signal (EphAx+ki, clustersz_simple(EphAx, ki, EphA4_free_kd, _p_epha4_kd)), linestyle='--', color=clr_knockin, dashes=(5, 5))
+
+    if study_EphAx_binding:
+        ax21.plot (x, signal (EphAx - EphAx_cis,    clustersz_simple(EphAx - EphAx_cis, 0,  EphA4_free,    _p_epha4)), linestyle='-.', color=clr_wt)
[…]

### 2022-09-02 paper/main.sh

@@ -1,0 +1,6 @@
+#!/bin/sh
+
+xelatex main.tex
+bibtex main.aux
+xelatex main.tex
+xelatex main.tex

### 2022-09-02 sim/agent/tissue.h

@@ -149,7 +149,7 @@ struct guidingtissue : public tissue<T>
     T EphA4_const_expression = T{3.5};
     T EphA4_current_expression = T{3.5};
     T EphA4_knockdown_expression = T{2.5};
-    T w_EphAx = T{0.25};
+    T w_EphAx = T{0.25}; // NB: Only used to calculate w_EphA4 = 0.611 * w_EphAx.
 
     //! Each receptor has a 2D gradient field, hence 2*N values here
     morph::vVector<morph::Vector<T,2*N>> rcpt_grad;

### 2022-09-26 paper/main.tex
Monday WIP on paper discussion

@@
 
-% If no page numbering wanted:
-%\pagenumbering{gobble}
-
 % A trick to get the bibliography to show up with 1. 2. etc in place of [1], [2] etc.:
@@
 
-% Define title, empty date and authors
 \title {
-Time to stop; agent-based modelling of chemoaffinity with competition
+Differential chemotaxis with competition as a mapping mechanism in the brain
 }
@@
 
-\emph{We show that a true chemoaffinity model based on gradient-following (via receptor-ligand interactions) combined with receptor-ligand-mediated axonal repulsion, in which information is only transferred by nearest-neighbour interactions, can account for a broad range of experimental manipulations of the retinotectal system.}
+\emph{The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
+There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
+Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
+However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
+Sperry's chemoaffinity theory provides an account of the wildtype pattern.
+Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
+Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
+We present a two-dimensional model of the retinotectal projection resembling that of Gierer, in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and show that this receptor-ligand model accounts for the full range of surgical manipulations.
+We then extend the detail with which the receptor-ligand interactions are treated to account for more recent genetic manipulation experiments.
+}
 
@@
 
-% Could use chemoaffinity as a coverall theory, comprising chemospecificity and chemotaxis.
-
 Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
@@
 This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.  
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of gene expression spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
+An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
 
@@
 In this system a two-dimensional map of the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-The retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
+This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
 This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
@@
 %
-The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002} studies.
+The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
 Two classes of ephrins have been identified, labelled `A' and `B'.
@@
 Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-These models have also helped establish how competition between axons for space in the target tissue can additionally account for many of the known effects of surgical manipulation on map development.
+The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
 
 Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
+%\emph{PDEs too complex (can't do maths analysis), so need agent based. Can represent more biological detail and ask how that affects the systems level.}
 These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue. 
 In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-% Not sure about this:
-However as the number of terms involved in PDEs necessarily grows large in systems representing differential chemotaxis, these systems quickly become impossible to analyse, making this potential advantage of such modelling redundant.
+However as the number of terms involved in PDEs necessarily grows large in systems representing differential chemotaxis, their analytical tractability quickly disappears, making the potential advantage of such modelling redundant.
 Additionally, the numerical simulation of a PDE system can become intractable, because each axonal projection must be represented by several state variables, each of which require solution across the entire spatial domain (which may consist of many thousands of elements).
-% in james_modelling we did 41 projections. I found that trying to build a retinotectal system with about 100 projections was already intractable.
 
-% But know we can motivate agent based modelling because *simulation* of PDEs for differential chemotaxis is intractable.
 An alternative simulation approach is to treat each of $n$ axons as an agent that interacts with its environment and with the other $n-1$ axons. 
@@
 We found that agent based simulations of $O(10^3)$ interacting axons can be achieved on a standard desktop computer whereas an equivalent PDE system would fail with about 100 axon projections.
+A further advantage of an agent based simulation is that the addition of new biological details is easy to implement and comes an a minimal computational cost, whereas in a PDE system, each new detail requires at least one state variable, which must be solved across the entire domain.
 
-\emph{PDEs too complex (can't do maths analysis), so need agent based. Can represent more biological detail and ask how that affects the systems level.}
-
-\emph{List the questions that emerge from the discussion above}
-
-\emph{Can Gierer model explain ALL the manipulations that S and G did?}
-
-\emph{Allows us to make a specific representation of Gierer's model in which the receptor-ligand mechanism is elaborated.}
-
-\textbf{xxxxxxxxxxxxxxx}
-
-Computational models have shown that in addition to chemotaxis, local competitive interactions between growing axons can account for a wide range of results from experiments in which the factors known to influence the growth of axons have been surgically or genetically manipulated \citep{prestige_role_1975,simpson_simple_2011,suetterlin_target-independent_2014}.
-
-However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemoaffinity model, that axons sample and evaluate positional information only \emph{locally}, has yet to be tested directly.
-%
-Consider, for example, an important model by Goodhill \& Simpson, which shows how [a certain set of assumptions about the form of the] competition between axons can account for patterns of innervation in the developing tectum that result from ablating/rotating sections of the retina \citep{simpson_simple_2011}.
-%   
-The formulation of this model includes a term that at every time-step moves the simulated axons in the direction of a pre-specified target location. 
-Thus the model implicitly assumes a global supervisor, capable of evaluating the current state of each axon, computing a direction vector to correct its trajectory, and communicating that information to each axon.
-%   
-This model was constructed primarily to establish the effects of competition, and to this end is elegantly formulated.
-%
-Nevertheless, the key assumptions of the chemoaffinity model, and specifically the proposal that axons find their way by evaluating only local information (chemotaxis), are yet to be investigated directly.
-
-The aim of the current study is therefore to construct a model that represents the assumptions of the chemotaxis model explicitly and to investigate how this mechanism behaves when combined with competitive interactions.
-
-One advantage of such a model is that it can be used to investigate the sensitivity and robustness of local interactions in establishing topographic maps to various sources of noise.
-
-Indeed, our results show that the local mechanisms of axonal information-processing that are assumed by the chemoaffinity model provide a remarkably robust solution to the problem of topographic map development.
+% List questions that emerge from the above
+For these reasons, we sought to implement an agent-based differential chemotaxis model of the retinotectal projection to ask: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
+ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
+and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
 
@@
 In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \mathbf{M}$ with the movement vector $\mathbf{M}$ given by
-\begin{equation} \label{e:mv2}
+\begin{equation} \label{e:M}
  \mathbf{M} = \mathbf{G} +  \mathbf{X} + \mathbf{B}.
@@
 Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, are also expressed by RGCs, but these are not used by this chemotaxis model.) 
+($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
 The expression of each receptor varies only with respect to one spatial dimension.
@@
 $r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
+Thus the model is of \emph{differential chemotaxis}.
 %
@@
 Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Fig.\,\ref{f:ex} summarises results from the literature).
+Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Supplementary Fig.\,\ref{f:experimental_expressions} summarises results from the literature).
 We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
 
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ includes EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ includes the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
+While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
 
@@
 %
-$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance of $b$.
+$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
 $m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
@@
 %
-$r_{\!_X}$ is the receptor-ligand interaction radius, $l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$.
+$l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$. 
 $F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
@@
 
-An assumption made in these papers is that all EphA receptor subtypes contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each subtype to the overall level is assumed not to be important. 
+An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
+The precise contribution of each sub-type to the overall level is assumed not to be important. 
 In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
@@
 
-The previous approach assumed that all EphA receptor subtypes (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
+The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
 This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
@@
 
-It is known that EphA receptors form clusters whose presence is necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
+It is known that EphA receptors form clusters, that these are necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
 It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
@@
 
-% Summarize the findings...
-
+% Summarize the findings
 Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
@@
 The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development. 
+Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
+Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
 
-The model also accounted for some aspects of the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by an extended model. 
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced item ii, it failed with the other three features.
+However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
+We identified four features of the genetic experimental results which should be reproduced and explained by the model. 
+These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
+We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
 
-To extend the model, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
+To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
+First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
+A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
+In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
+We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
+This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
+However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA. 
+A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
+We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
+We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
 
-%Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
-%Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations. Finish summarising results...
+% Place findings in context
 
-% Place findings in context... (starting with the next section)
-
-\subsection*{On chemotaxis}
+\subsection*{Chemotaxis}
 
@@
 This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-Gierer recognised that tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-He addressed this problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and noting that this was equivalent to a more direct fibre-fibre competition.
+%Gierer recognised that tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
+%He addressed this problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and noting that this was equivalent to a more direct fibre-fibre competition.
 
-Because of the discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
+Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
 Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
@@
 
-%\citet{nakamoto_topographically_1996} discussed gradient based models  stems in a large part from \citet{gierer_directional_1987}.
 Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
@@
 Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$.
+Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
 With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain, with examples of opposing pairs including XX and YY \citep{cang_developmental_2013}. 
-Adopting a pair of opposing gradients is mathematically reminiscent of/actually implements the models of \citet{gierer_model_1983} (which proposed the existence of potential functions) and \citet{fraser_differential_1980}.
+There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
+Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
+RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
+Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
+Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
+However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
+In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum. 
+The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}. 
+Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
+Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
+%Adopting a pair of opposing gradients is mathematically reminiscent both of the model of \citet{gierer_model_1983} (which proposed the existence of potential functions) and of \citet{fraser_differential_1980} which is another model incorporating gradient based movement and competition.
 
-\emph{(Place this:)} The model required opposing pairs of graded signalling molecules for the chemotaxis component of the model.
-The competition model, in which a close-range repulsion was enabled by receptor-ligand binding between RGC axons required the participation of only a single pair of opposed signalling molecules in the retina.
+We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
+\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level.
+This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
+It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:M} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
 
-\subsection*{Ephrins}
-A bit of discussion on my treatment of Ephrins, esp.~in the extended model with EphA4.
+% \citet{nakamoto_topographically_1996} discussed gradient based models briefly, citing \citet{gierer_directional_1987}.
 
-\subsection*{Not ephrins}
+%\subsection*{Chemotaxis in isolation}
+%
+%The model described in Eq.\,\ref{e:M} can be used to examine the behaviour of a system in which only chemotaxis operates, by setting the parameter $m_{\!_X}=0$. 
+%Retaining the same value for $m_{\!_G}=0.0012$ results in the wildtype map spanning only about half the tectal surface, although the neighbour relationships in the map are all correct. 
+%The chemotaxis-only model \emph{can} be tuned so that the map spans the entire tectum. 
+%This can be achieved by modifying the functional form for the tectal ligand expression. 
+%We verified that this could be done (by changing $L_i$ to $L_i=1.05 + 0.25 \exp(1.1 u) $). 
+%Applying the simulated surgical manipulations demonstrated significant shortcomings of the chemotaxis-only model, as had already been recognised by Gierer.
 
-Placeholder section.
-
-A number of other signalling molecules have been identified, which are be expressed in gradients that are similar to those of the ephrins. 
-The expression of \emph{Wnt} and its ligand \emph{Ryk} matches that of Eph B and ephrin B \citep{schmitt_wntryk_2006}. 
-Wnt is expressed at a high level towards the medial tectum; Ryk is expressed in a low to high gradient from the dorsal to ventral retina.
-Wnt-Ryk signalling induces repulsion away from areas of high Ryk expression, so RGCs from ventral retina should be repelled from the medial tectum. 
-This acts in direct opposition to the effect of ephrin B signalling, which induces attraction of ventral RGCs towards the medial tectum \citep{hindges_ephb_2002}.
 % In discussion will go into why that fails to provide a counter gradient for EphB/ephrin B. See expression_profiles.svg.
 
-Neogenin, RGM (repulsive guidance molecule) and netrin are a further set of signalling molecules that have been shown to act in the retinotectal system.
-Neogenin is a receptor which is found in a graded distribution in the retina.
-
-\subsection*{Chemotaxis in isolation}
-
-Describe results, shown in supplementary figure of what happens with only chemotaxis.
-
-\subsection*{On competition}
+\subsection*{Competition}
 
@@
 
-Note that $\mathbf{X}$ would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$, if there was a single $l_i$ expressed in the retina and its form was $1/r_i$.
-However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}C).
+Note that if there was a \emph{single} $l_i$ expressed in the retina and its form was $1/r_i$, then $\mathbf{X}$ would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
+However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
 If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
 
-Why does competition improve the surgical manipulation behaviour? Genetic manipulations?
+Competition appears to be necessary for all of the surgical manipulations. 
+It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
+It also assists with the graft rotation and graft swap manipulations \textbf{by...}.
 
-\subsection*{Single axon behaviour}
+The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
+This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
+If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map. 
+If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
+This range of competition is similar to that found by \citet{simpson_simple_2011}. 
+It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
 
-The chemoaffinity part of this model ensures that single axons will be drawn to approximately the correct location. 
+% Single axon behaviour
+\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
+They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targetting of these cells would occur.
+The chemotaxis component of our model ensures that single axons will be drawn to approximately the correct location.
 The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is in agreement with \citet{gosse_retinotopic_2008}... \emph{Do a similar linear fit to our models match Gosse Fig 2d/e}.
+This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
+
+%\subsection*{Genetic manipulations}
+% An assumption of the simulation of genetic manipulations was that there is no difference between the EphA receptor sub-types.
 
@@
 
-Discuss the form of tectal ligand expression that is required by this model if receptor-ligand interactions are attractive (under the condition that retinal expression is exponential).
-In this case, the expression should follow a logarithmic increase.
-This ties in with the finding that only one pair of receptor-ligand interactions is required to mediate competition; this leaves the other axis free to be either repulsive, or attractive (as EphB).
-
-\subsection*{Genetic manipulations}
-
- An assumption of the simulation of genetic manipulations was that there is no difference between the EphA receptor sub-types.
+In the results, we present a model in which all receptor-ligand interactions are repulsive.
+However, EphB/ephrinB interactions are largely attractive in nature.
+It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
+When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
+This is compatible with the finding that only one pair of repulsive receptor-ligand interactions was required to mediate competition; this left the other axis free to be either repulsive, or attractive (as EphB).
 
@@
 
-I've done sims that introduce noise, so find these and discuss. The headline is that the system is very robust to noise.
+To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
+Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
+We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
+Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (E4).
+Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determeined.
+The graphs show the expression values on a normalised scale. 
+We isolate a section that can be approximated by a linear function, obtain the best fit and determine the RMS error from the fit. The mean error is approximately 0.06.
+We therefore use 0.06 as the standard deviation for normal distributions used to generate random numbers to inject into the simulation.
+Need to figure out if 0.06 is correct for each of the three noises - is the full range of the retinal receptor and tectal ligand expressions 1? Don't think so.
 
-\subsection*{The value of $s$}
+We add noise in three places.
+i) At each simulation step, add noise to the measurement of the tectal ligand gradient, so that $\nabla L_i(\mathbf{x}) $ becomes $\nabla L_i(\mathbf{x}) + N(0,0.06)$.
+ii) Add noise to the specification of the retinal receptor expression so that  $r_i(x)$ becomes $r_i(x) + N(0, 0.06)$ and iii) Add noise to the specification of the tectal ligand expression so that $l_i(x)$ becomes $l_i(x) + N(0, 0.06)$.
+(ii) and (iii) are computed once only, at the start of the simulation.
+Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined.
 
-This is worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
+%\subsection*{The value of $s$}
+%This is worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
 Esesentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
@@
 % Any unexpected results?
-\subsection*{Other models}
+%\subsection*{Other models}
 
-Have covered Gierer in introduction.
+%Have covered Gierer in introduction.
 Also, be sure to discuss \citet{naoki_revisiting_2017}, which is a gradient model with just two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level.
@@
 No mechanism for $r_2$ collapse.
+
+This study ignores the mechanism in species X, by which axons grow the full length of the tectum, then make branches at an appropriate location.
 
@@
 
+% FIGURE S2
+\begin{figure}
+\includegraphics[width=\linewidth]{./images/expression_profiles_EphBs.png}
+\caption{(Supplementary) EphB and ephrin B gene expression.
+Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature \textbf{A} Retinal expression of EphB2.
+We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application.
+The resulting, normalized expression is shown in the graph for the E4 data, with an inset showing corresponding data for the E6 data.
+In the right-hand graph, we show an expansion of a portion of the E4 data to which we applied a linear fit to determine the RMS deviation from the mean to estimate the noise in the receptor signal.
+The mean RMS error was 0.06.
+\textbf{B} Stalefish screenshots and expression graphs for retinal ephrinB1 and ephrinB2 expression, using data from Fig.\,1 of \citet{braisted_graded_1997}.
+\textbf{C} Screenshots and expression graphs for tectal ephrinB1 expression, using data from Fig.\,4 of \citet{braisted_graded_1997}.}
+\label{f:expr_ephb}
+\end{figure}
+
+% FIGURE S3
+\begin{figure}
+\includegraphics[width=0.9\linewidth]{./images/noise_contribution.png}
+\caption{(Supplementary) Response of the model to noise. \textbf{A} \textbf{B} \textbf{C}\textbf{D}}
+\label{f:noise}
+\end{figure}
+
 % DITCH THIS ONE

### 2022-09-26 scripts/fig_GJ_noise.sh

@@ -1,10 +1,13 @@
 #!/bin/sh
 
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.0 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.2 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.4 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.6 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=0.8 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=1.0 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=1.5 -co:exit=true
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=1000 -co:noise_gain=2.0 -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=2000 -co:noise_gain=0.06 -co:ret_rcpt_noise_gain=0.0  -co:tec_lgnd_noise_gain=0.0 # -co:exit=true
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_wt_figcomp2.json -co:steps=2000 -co:noise_gain=0.0  -co:ret_rcpt_noise_gain=0.06 -co:tec_lgnd_noise_gain=0.0 # -co:exit=true
[…]

### 2022-09-26 sim/agent/branch.h

@@ -67,10 +67,8 @@ public:
         morph::Vector<T, 2*N> lg = true_lgnd_grad;
         // ...unless this if is entered into
         if (add_simple_gradient_noise == true && this->noise_gain > T{0}) {
-            this->rn.randomize(); // random uniform, 0->1
-            this->rn -= T{0.5};  // symmetric about 0
-            this->rn *= T{2};   // +- 1
-            lg += this->rn * this->noise_gain;
+            this->rn.randomizeN (T{0}, this->noise_gain); // Randomize normally
+            lg += this->rn;
         }
         return lg;
     }

### 2022-09-27 paper/main.tex
Changes to paper which is now almost in submittable form

@@
 Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection resembling that of Gierer, in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to account for more recent genetic manipulation experiments.
-}
-
-%\begin{itemize}
-%    \item Sperry's chemospecificity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development.
-%    \item A key component of the theory is that growing axons find their way into position by chemotaxis, i.e., by evaluating local information in gradients of gene expression that span the target tissue
-%    \item Computational models have shown that in addition to chemotaxis, local competitive interactions between growing axons can account for a wide range of results from experiments in which the factors known to influence the growth of axons have been surgically or genetically manipulated
-%    \item However, while such models have \emph{assumed} the effects of chemotaxis, the key assumption of the chemospecificity model, that axons sample and evaluate positional information only \emph{locally}, is yet to be tested directly.
-%    \item Consider, for example, an important model by Goodhill \& Simpson, which shows how [a certain set of assumptions about the form of the] competition between axons can account for patterns of innervation in the developing tectum that result from ablating/rotating sections of the retina.
-%    \item The formulation of this model includes a term that at every time-step moves the simulated axons in the direction of a pre-specified target location.
-Thus the model implicitly assumes a global supervisor, capable of evaluating the current state of each axon, computing a direction vector to correct its trajectory, and communicating that information to each axon.
-%    \item This model was constructed primarily to establish the effects of competition, and to this end is elegantly formulated.
-%    \item Nevertheless, the key assumptions of the chemospecificity model, and specifically the proposal that axons find their way via evaluating only local information (chemotaxis), are yet to be investigated directly.
-%    \item The aim of the current study is therefore to construct a model that represents the assumptions of the chemospecificity model explicitly. 
-%    \item One advantage of such a model is that it can be used to investigate the sensitivity and robustness of local interactions in establishing topographic maps to various sources of noise.
-%    \item Indeed, our simulation results show that the local mechanisms of axonal information-processing that are assumed by the chemospecificity model provide a remarkably robust solution to the problem of topographic map development.
-%\end{itemize}
+We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
+We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.}
 
@@
 
-Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity (\textbf{find citation}).
+Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
 In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
@@
 Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-%\emph{PDEs too complex (can't do maths analysis), so need agent based. Can represent more biological detail and ask how that affects the systems level.}
 These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue. 
 In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However as the number of terms involved in PDEs necessarily grows large in systems representing differential chemotaxis, their analytical tractability quickly disappears, making the potential advantage of such modelling redundant.
-Additionally, the numerical simulation of a PDE system can become intractable, because each axonal projection must be represented by several state variables, each of which require solution across the entire spatial domain (which may consist of many thousands of elements).
-
-An alternative simulation approach is to treat each of $n$ axons as an agent that interacts with its environment and with the other $n-1$ axons. 
-The computation of such a system (see for example, \citet{simpson_simple_2011}) is dominated by the $O(n^2)$ axon-axon interactions and so a limit will be approached as $n\rightarrow \infty$. 
-We found that agent based simulations of $O(10^3)$ interacting axons can be achieved on a standard desktop computer whereas an equivalent PDE system would fail with about 100 axon projections.
-A further advantage of an agent based simulation is that the addition of new biological details is easy to implement and comes an a minimal computational cost, whereas in a PDE system, each new detail requires at least one state variable, which must be solved across the entire domain.
-
-% List questions that emerge from the above
-For these reasons, we sought to implement an agent-based differential chemotaxis model of the retinotectal projection to ask: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
+However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
+Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
+With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
 ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
@@
 Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (Supplementary Fig.\,\ref{f:experimental_expressions} summarises results from the literature).
+Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
 We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
@@
 Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, lead to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fish net show that a number of axons from near the edges of the tectum have been drawn into the graft region. 
+The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
+Long connecting lines in the fish net show that a number of axons from near the edges of the tectum were drawn into the graft region. 
 The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
@@
 There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-% Note: the graph shows the pattern metric for a region that covers BOTH grafts and for the region that is not the two grafts.
 
@@
 The fish-net plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal tectal axis plotted against its position on the nasal-temporal retinal axis in red triangles, with the experimental positions plotted in blue squares.
+In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
 %
@@
 For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours. 
-By enhancing the quantity of EphA receptors expressed by half of all retino-tectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
+By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
 In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5). 
@@
 The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
+In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
 In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
@@
 Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that they are drawn from statistically identical distributions.
-% While the convergence of the manipulated and wildtype lines was less abrupt than in the experimental result of \citet{brown_topographic_2000}, it was comparable with the result from \citet{simpson_simple_2011}.
+A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
 
@@
 The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the tectal rostral-caudal ligand gradient. 
+The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient. 
 The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
@@
 
-Figs.\,\ref{f:GJeph}C show the results for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
+Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
 As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
 The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
+In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
 
 It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by this model.
+The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the differential chemotaxis with competition model.
 We address possible interaction effects in the next section.
@@
 Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).  
-%(This is reminiscent of the divisive relationship suggested by the work of Brown, Reber and co-workers who showed that when the ratio of EphA in treated to EphA in untreated cells drops below a threshold, the mapping system directs the cells to the same tectal location.)
 Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
@@
 Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin in temporal (above 0.56 N--T).
+Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
 This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
@@
 
-% Summarize the findings
+% Summarize the findings and place them in context
 Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
@@
 
-% Place findings in context
-
 \subsection*{Chemotaxis}
@@
 This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-%Gierer recognised that tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-%He addressed this problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and noting that this was equivalent to a more direct fibre-fibre competition.
 
@@
 Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-%Adopting a pair of opposing gradients is mathematically reminiscent both of the model of \citet{gierer_model_1983} (which proposed the existence of potential functions) and of \citet{fraser_differential_1980} which is another model incorporating gradient based movement and competition.
 
 We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level.
+\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
 This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
 It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:M} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-% \citet{nakamoto_topographically_1996} discussed gradient based models briefly, citing \citet{gierer_directional_1987}.
-
-%\subsection*{Chemotaxis in isolation}
-%
-%The model described in Eq.\,\ref{e:M} can be used to examine the behaviour of a system in which only chemotaxis operates, by setting the parameter $m_{\!_X}=0$. 
-%Retaining the same value for $m_{\!_G}=0.0012$ results in the wildtype map spanning only about half the tectal surface, although the neighbour relationships in the map are all correct. 
-%The chemotaxis-only model \emph{can} be tuned so that the map spans the entire tectum. 
-%This can be achieved by modifying the functional form for the tectal ligand expression. 
-%We verified that this could be done (by changing $L_i$ to $L_i=1.05 + 0.25 \exp(1.1 u) $). 
-%Applying the simulated surgical manipulations demonstrated significant shortcomings of the chemotaxis-only model, as had already been recognised by Gierer.
-
-% In discussion will go into why that fails to provide a counter gradient for EphB/ephrin B. See expression_profiles.svg.
 
@@
 
-Note that if there was a \emph{single} $l_i$ expressed in the retina and its form was $1/r_i$, then $\mathbf{X}$ would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
+Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{X}$ from Eq.\,\ref{e:M} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
 However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
@@
 It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations \textbf{by...}.
+It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
 
@@
 If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This range of competition is similar to that found by \citet{simpson_simple_2011}. 
+This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}. 
 It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
@@
 
-%\subsection*{Genetic manipulations}
-% An assumption of the simulation of genetic manipulations was that there is no difference between the EphA receptor sub-types.
-
 \subsection*{Attractive receptor-ligand interactions}
@@
 When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-This is compatible with the finding that only one pair of repulsive receptor-ligand interactions was required to mediate competition; this left the other axis free to be either repulsive, or attractive (as EphB).
+In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
+Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
 
@@
 We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (E4).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determeined.
+Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
+Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
 The graphs show the expression values on a normalised scale. 
-We isolate a section that can be approximated by a linear function, obtain the best fit and determine the RMS error from the fit. The mean error is approximately 0.06.
-We therefore use 0.06 as the standard deviation for normal distributions used to generate random numbers to inject into the simulation.
-Need to figure out if 0.06 is correct for each of the three noises - is the full range of the retinal receptor and tectal ligand expressions 1? Don't think so.
-
-We add noise in three places.
-i) At each simulation step, add noise to the measurement of the tectal ligand gradient, so that $\nabla L_i(\mathbf{x}) $ becomes $\nabla L_i(\mathbf{x}) + N(0,0.06)$.
-ii) Add noise to the specification of the retinal receptor expression so that  $r_i(x)$ becomes $r_i(x) + N(0, 0.06)$ and iii) Add noise to the specification of the tectal ligand expression so that $l_i(x)$ becomes $l_i(x) + N(0, 0.06)$.
+We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
+The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
+Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
+i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
+ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
+iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
 (ii) and (iii) are computed once only, at the start of the simulation.
 Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined. 
+It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A). 
+The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern. 
+Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions. 
+Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
 
 %\subsection*{The value of $s$}
-%This is worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Esesentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Any unexpected results?
-%\subsection*{Other models}
-
-%Have covered Gierer in introduction.
-Also, be sure to discuss \citet{naoki_revisiting_2017}, which is a gradient model with just two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level.
+%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
+Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
 
 % Limitations/weaknesses of research
-\subsection*{Limitations}
+\subsection*{Predictions and Limitations}
 
-The sensitivity of the model to the interaction radius - receptor-ligand binding really quite close range, but the interaction radius in this model is required to be of the order of the target axon-axon spacing.
+We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth. 
+We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
+By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
+We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
 
-Low resolution of gradient (limited by size of grid).
-
-No mechanism for $r_2$ collapse.
-
-This study ignores the mechanism in species X, by which axons grow the full length of the tectum, then make branches at an appropriate location.
+Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
+In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
+Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
+Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
+Adapting this framework to address an axon branching model is left for future work.
 
@@
 \subsection*{Border effect} % border effect description
-The border effect is also based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
+The border effect ($\mathbf{B}$ in Eq.\,\ref{e:M}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
 For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
@@
 \includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Schematic diagrams of retino-tectal mapping and gene expression.
+\caption{Schematic diagrams of retinotectal mapping and gene expression.
 Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
@@
 
-%%% Caption for clustermechsize only model: \textbf{B} Model results.
-i) EphA3 knock-in result for ki=1.0.
-Inset shows experimental results from Fig.\,5B of \citet{brown_topographic_2000}.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment.
-ii) The double knock-in result (Fig.\,5C of \citet{brown_topographic_2000}) agrees well with the experimental result, showing two, separated maps iii) When EphA is knocked in and EphA4 is simultaneously knocked down, nasal axons experience an enhanced signal to move along gradients towards rostral tectum and so the red curve is drawn further rostrally.
-Although less dramatic than the experimental result of \citet{reber_relative_2004} the movement is in the correct direction, and the overlap of positions (the `collapse point') moves to about 90\% of the N--T axis.
-iv) Uniform knock-down of EphA4 in all cells does not perturb the mapping.
-
 % FIGURE 7
@@
 \includegraphics[width=\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Ligand expression and gradients on the tectum for the
-graft-rotate 90 degrees simulated surgical manipulation}
+\caption{(Supplementary) Visualisation of ligand expression and gradients on the tectum for the
+graft-rotate 90 degrees simulated surgical manipulations.}
 \label{f:trot90}
@@
 \includegraphics[width=0.9\linewidth]{./images/noise_contribution.png}
-\caption{(Supplementary) Response of the model to noise. \textbf{A} \textbf{B} \textbf{C}\textbf{D}}
+\caption{(Supplementary) Response of the model to noise.
+\textbf{A} Adding 3\% gradient measurement noise (normally distributed, standard deviation, $\sigma = 0.19$) has negligible effect on the pattern.
+\textbf{B} 3\% retinal receptor noise ($\sigma = 0.08$).
+\textbf{C} 3\% tectal ligand noise ($\sigma = 0.08$).
+\textbf{D} Combined effect of retinal receptor, tectal ligand and gradient measurement noise is dominated by the tectal ligand noise.}
 \label{f:noise}
@@
 
-% DITCH THIS ONE
-%
-% FIGURE S2
-%%\begin{figure}
-%%\includegraphics[width=\linewidth]{./images/j4_ee_G_wt_fig2_exit_true.png}
-%%% Note MANUAL reference to Fig. 2 here (otherwise it says Fig S2):
-%%\caption{Receptor-ligand model of chemotaxis alone. Panels A-H show results as in Fig\,\textbf{2}. Briefly:
-%%\textbf{A} Retinal cell position colour scheme.
-%%\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units.
-In this chemotaxis-only model, which differs from that shown in Fig\,\ref{f:GJ} only by having the parameter $m_X$ set to 0 in Eq.\,\ref{e:X}, variable interactions with the gradients cause each RGC to find a location that preserves the topographic ordering, but the absence of competition means that axons fail to extend across the full tectum.
-%%\textbf{E} Anticipated layout. 
-%%\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time. 
-%%\textbf{G} The arrangement of axon branches at t=1500. 
-%%\textbf{H} Selected axon locations at t=1500, with path histories in grey. 
-%%\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-%%\label{f:G}
-%%\end{figure}
-
 \end{document}

### 2022-09-29 paper/main.tex
Latest state of paper, plus add paper pngs now they're almost there

@@
-% Use the article doc class, with an 11 pt basic font size
-%\documentclass[11pt, a4paper, draft]{article}
-\documentclass[11pt, a4paper]{article}
+\documentclass[9pt,lineno]{elife}
 % Nice citations
-\usepackage{natbib}
+%\usepackage{natbib}
 % Set margins
-\usepackage[margin=2.5cm]{geometry}
+%\usepackage[margin=2.5cm]{geometry}
 % Multilingual support
-\usepackage[english]{babel}
+%\usepackage[english]{babel}
 % Nice mathematics and Left-right harpoons for kinetic equations
@@
 % Control over maketitle
-\usepackage{titling,titlesec}
+%\usepackage{titling,titlesec}
 % Ability to use colour in text
-\usepackage[usenames]{color}
-\definecolor{grey}{rgb}{0.6, 0.6, 0.63}
-\definecolor{black}{rgb}{0, 0, 0}
+\usepackage{color}
 % For the \degree symbol
@@
 % Caption font size to be small.
-\usepackage[font=small,labelfont=bf]{caption}
+%\usepackage[font=small,labelfont=bf]{caption}
 
-% Make a dot for the dot product, call it vcdot for 'vector calculus dot'. Bigger than \cdot, smaller than \bullet.
-\makeatletter
-\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
-\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
-\makeatother
-
-\def\firstAuthorLast{James}
-\def\Address{\\
-\affiliationListFont Department of Psychology,\\
- \affiliationListFont  The University of Sheffield, Sheffield, UK \\
-}
-\def\corrAuthor{Seb James}
-\def\corrAddress{Department of Psychology, The University of Sheffield,
-  Western Bank, Sheffield, S10 2TP, UK}
-\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address}
-%  \corrAuthorFont $^{*}$ Correspondence: {seb.james@sheffield.ac.uk}}
-
-% A trick to get the bibliography to show up with 1. 2. etc in place of [1], [2] etc.:
-\makeatletter
-\renewcommand\@biblabel[1]{#1.}
-\makeatother
-
-% reduce separation between bibliography items if not using natbib:
-\let\OLDthebibliography\thebibliography
-\renewcommand\thebibliography[1]{
-  \OLDthebibliography{#1}
-  \setlength{\parskip}{0pt}
-  \setlength{\itemsep}{0pt plus 0.3ex}
+\title {
+Competitive differential chemotaxis can explain the self-organisation of topographic maps
 }
 
-\title {
-Differential chemotaxis with competition as a mapping mechanism in the brain
-}
-\date{} % No date please
-\author{\Authors}
+\author[1*]{Sebastian~S.~James}
+\author[1]{Stuart~P.~Wilson}
+\affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
+\corr{seb.james@sheffield.ac.uk}{SSJ}
+%% END OF PREAMBLE
 
-%% END OF PREAMBLE
+% Seb to fix: Expected not experiment (in figs), Fig 8 - monochromes. Fig 1 - monochromes
+% Stuart to do: Intro/Start of results modifications
 
@@
 
-\emph{The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
+\begin{abstract}
+The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
 There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
@@
 We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.}
+We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
+\end{abstract}
 
@@
 
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \mathbf{M}$ with the movement vector $\mathbf{M}$ given by
-\begin{equation} \label{e:M}
- \mathbf{M} = \mathbf{G} +  \mathbf{X} + \mathbf{B}.
+In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
+\begin{equation} \label{e:dX}
+ \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
 \end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{X}$ is the movement due to competitive interactions with other branches. 
+$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
 $\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
 
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{X}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found an $\mathbf{X}$ comprised of two components; a simple distance-based competition between \emph{all} branches ($\mathbf{C}$), along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon ($\mathbf{I}$).
+\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
+Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
 We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
@@
 
- We modelled competition as a distance-based competitive movement vector, $\mathbf{X}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
+ We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
 %
 \begin{equation} \label{e:X}
-\mathbf{X} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
+\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
 \end{equation}
@@
 
-We used a combination of pattern metrics to define the objective function. $\epsilon$ is the mean distance error of the axon centroids, given by:
+We used a combination of pattern metrics to define the objective function.
+$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
 %
-\begin{equation}\label{e:eps}
-\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-\end{equation}
+%\begin{equation}\label{e:eps}
+%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
+%\end{equation}
 %
@@
 However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other. 
-For this reason, we also used the number of crossings in the fish net plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
+For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
 The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
@@
 Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fish net' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
+Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
 Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
@@
 
-After finding an optimum set of parameters for the wildtype arrangement, we tested our model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
+After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
 Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
@@
 The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fish net show that a number of axons from near the edges of the tectum were drawn into the graft region. 
+Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region. 
 The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
@@
 The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}). 
-The fish-net plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
+The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
 In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
@@
 Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-Our model clearly replicates this result, indicated by both fish-net and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
+The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
 %
 A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-Our model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
+The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
 %
@@
 This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For our model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
+For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
 iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
@@
 The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In our model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
+In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
 In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
@@
 It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the differential chemotaxis with competition model.
+The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
 We address possible interaction effects in the next section.
@@
 This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:M} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
+It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
 
@@
 
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{X}$ from Eq.\,\ref{e:M} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
+Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
+However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
 If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
@@
 \citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targetting of these cells would occur.
-The chemotaxis component of our model ensures that single axons will be drawn to approximately the correct location.
+They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
+The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
 The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
@@
 
-% Bibliography here
-\selectlanguage{English}
-\bibliographystyle{apalike}
-\bibliography{RetinoTectal}
-
-\appendix
-
-\section*{Method details}
+\section{Method details}
 
 \subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:M}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
+The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
 For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
@@
 
-\subsection*{Bootstrap analyses of R-C tectal position vs.~N-T retinal origin}
+\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
 
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R-C positions obtained for the $n=20$ different dorso-ventral origins for each N-T location.
+To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
 The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
@@
 
+\section{Data Availability}
+
+All data and code required to reproduce the results in this work are available in the public repository 
+\url{https://github.com/sebjameswml/RetinoTectal}.
+The software was built using the library morphologica \citep{james_morphologica_2022}.
+
+\section{Acknowledgements}
+
+This work was supported by grant 220020516 from the James S. McDonnell Foundation.
+
+\bibliography{RetinoTectal}
+
 %
@@
 % Expand panel A to be able to lose B-D.
+% Full page width
 \begin{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
 \caption{Receptor-ligand model of chemotaxis and competition. 
@@
 \label{f:GJ}
+\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_grafts.png}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
 \caption{Performance of the receptor-ligand model with surgical graft manipulations.
@@
 \label{f:GJsurg_grafts}
+\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_ablations.png}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
 \caption{Retinal and tectal ablation simulations.
@@
 \label{f:GJsurg_ablate}
+\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations.png}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
 \caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_explanation.png}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
 \caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
@@
 \begin{figure}
-\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
+\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
 \caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}. 
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the tectum for the
-graft-rotate 90 degrees simulated surgical manipulations.}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
+\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
+graft-rotate 90 degrees surgical manipulations. 
+\textbf{A} Ligand expression. 
+The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps. 
+The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$. 
+Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the 
+The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively. 
+\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A. 
+For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis). 
+The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
+Full saturation indicates a positive gradient; white indicates a negative gradient.
+For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis. 
+Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
+\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
 \label{f:trot90}
+\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/expression_profiles_EphBs.png}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
 \caption{(Supplementary) EphB and ephrin B gene expression. 
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature \textbf{A} Retinal expression of EphB2.
+Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature. 
+\textbf{A} Retinal expression of EphB2. 
 We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application. 
@@
 \label{f:expr_ephb}
+\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.9\linewidth]{./images/noise_contribution.png}
+\begin{fullwidth}
+\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
 \caption{(Supplementary) Response of the model to noise.
@@
 \label{f:noise}
+\end{fullwidth}
 \end{figure}

### 2023-08-25 paper/main.tex
Updates paper from changes made on Overleaf in Jan 2023

@@
 \title {
-Competitive differential chemotaxis can explain the self-organisation of topographic maps
+Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
 }
@@
 
-Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-%
-The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'. 
+Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
+Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
+Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
+
+One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
+A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
+
+A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
+\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
+A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
+
+The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
+Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
+Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
+
+The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
+According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
+As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
+Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
+The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
+
+These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
+Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
+
+The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
+However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
+Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
+
+In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
+For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
+
+Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
+
+The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
+However, a major limitation is that chemotaxis is not implemented explicitly.
+Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
+This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
+
+An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
+If so, such a model would represent the most complete account of retinotectal map self-organization to date.
+
+%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
+
+%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
 The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-%
-This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.  
+
+%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
 An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
 
-Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
+%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
+
+%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
 This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
@@
 
-In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
+%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
 % And sounds?
-%
-The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
+
+%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
 Two classes of ephrins have been identified, labelled `A' and `B'.
@@
 
-Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins. 
+%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
 The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
@@
 ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
+
+%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
 He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
 
-Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
+%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
 The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
 
-Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
+%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
 These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
@@
 All data and code required to reproduce the results in this work are available in the public repository 
-\url{https://github.com/sebjameswml/RetinoTectal}.
+{https://github.com/sebjameswml/RetinoTectal}.
 The software was built using the library morphologica \citep{james_morphologica_2022}.

### 2023-08-25 paper/main.tex
Progress on paper following meeting with Stuart

@@
-\documentclass[9pt,lineno]{elife}
+\documentclass[9pt,lineno,draft]{elife}
 % Nice citations
@@
 \cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
+A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition
+% Stuart: I think the model is Koulakov's, not Triplett's.
+In Triplett2011 they give evidence for a competition model (like Koulakov's) by reducing the number of RGCs in the Math5 mutant mouse.
+*Hjorth* shows that the Koulakov model is the best one at reproducing the collapse in the EphA3 ki experiment.
+by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
 
@@
 If so, such a model would represent the most complete account of retinotectal map self-organization to date.
+
+** Maybe we also need to say that a second objective to to express the model
+exclusively in terms of receptor-ligand interactions.
 
@@
 \section{Results}
-\subsection*{A model of chemotaxis and competition}
 
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
+%% * receptors are expressed by RGCs as f(x,y)=... (Eq. 1)
+
+We adopt the same form for the expression of receptors on retinal ganglion cells as
+\citet{simpson_simple_2011}:
+\begin{equation} \label{e:f}
+  f(x,y) = 1.05 + 0.26 \exp(2.3 u),
+\end{equation}
+where $u$ gives the direction with which the expression increases and may be substituted by $(1-x)$, $(1-y)$, $x$ or $y$.
+We defined $x$ as an axis from temporal to the nasal retina and $y$ from ventral to dorsal retina.
+
+%% * Simpson and Goodhill modelled according to deltax = B + C + G (Eq 2)
+In \citet{simpson_simple_2011}'s model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
+% Maybe re-order the terms?
 \begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
+ \Delta \mathbf{x} = \mathbf{B} + \mathbf{C} + \mathbf{G}.
 \end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
+$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue (\citet{holt_target_1998}---see method details).
+%
+%% * They had C composed of two terms, distance based and a component based on EphA expression
+%
+$\mathbf{C}$ is the movement due to competitive interactions with other branches
+%
+and $\mathbf{G}$ is the movement vector given rise to by the chemotaxis effect.
 
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
+%% * To represent the mechanism for comp. explicitly, we introduced receptor AND ligand expressions
 
-\subsection*{A gradient based model of chemotaxis}
+%%We reformulated this so that the distance based repulsion is also mediated via receptor ligand interactions. % No need to justify yet - can go in disucssion
+%
 
-We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
+%%%%%%%% Remove/coment out %%%%%%%%%%%%%%%%%
+% No need for this rhetorical stuff
+%\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
 
-\begin{equation}\label{e:G}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
-$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
-%
-We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
-%
-$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
+%Their model, which successfully accounts for a wide range of experimental
+%manipulations, foundxxx *Their model had a* $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
+% Rewrite next little bit
+% Say the global supervisor comes into the expression G, but having G a vector
+% that always points towards...
+% This may come later
 
-%%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
+%We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
+%%%%%%%%%%%%%%%%%%%%%%%
 
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
-The expression of each receptor varies only with respect to one spatial dimension.
-%
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-%
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
+%  Maybe lose the sub sub sections
+\subsubsection*{A signalling model for competition}
 
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
+Simpson \& Goodhill's model had a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
 
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
+%Although the model incorporated EphA expression, the distance-based competitive interaction was given no mechanism of operation.
 
-\subsection*{A signalling model for competition}
+To provide an explicit mechanism of operation for the competition, we modelled it as a sum of interactions that are `switched on' via receptor-ligand signalling, meaning that some axons would have no competitive interaction, while others would repel strongly.
+This single-component competitive movement vector for a branch $b$ is
 
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
-%
 \begin{equation} \label{e:X}
@@
 %
-$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
+$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to $b$.
 %
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
+$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
+$r_{\!_X}$ is the receptor-ligand interaction radius.
+%
+The distance-based weighting can be interpreted as being related to the probability with which a ligand may bind to a receptor and mediate the interaction.
+%
+$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
 %
@@
 
-\subsection*{Simulation}
+\subsubsection*{A gradient based model of chemotaxis}
+
+% Say the global supervisor comes into the expression G, but having G a vector
+% that always points towards...
+% This may come later
+
+* The S\&G chemotaxis was a placeholder because they chose G to be a vector
+  pointing to a predetermined target location. In order to make this a local
+  model we do the following....
+
+We also imagine ligand expression on the tectum (L) - describe these here.
+
+We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
+
+\begin{equation}\label{e:G}
+\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
+\end{equation}
+%
+where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
+$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
+%
+We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
+%
+$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
+
+%%%%% Describing the simulation TAKE THIS TO METHOD DETAILS
+We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}.
+The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
+One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
+%%%%%%%%%%%%%
+
+% Keep, but say M grown cones. (M=4 specified in method details)
+% ALSO: Put M near competition equation where it's used!
+Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
+
+The expression of each receptor varies only with respect to one spatial dimension.
+
+% Thi sis in the righ tplace here
+We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
+$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
+Thus the model is of \emph{differential chemotaxis}.
+
+% NO NEED FOR This sentence/bit
+Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
+%
+We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
+
+Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
+Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
+We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
+
+While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
+%%%%%%%%%%%%%%%%%
+
+\subsection*{Simulation} % MOST OF THIS GOES TO METHOD DETAILS, just keep
+                         % first sentence, have list of values used in sims, then 'see method details',
 
@@
 
+% Results proper here Leave this as is for now
+\subsection*{Retinotectal map self-organization}
+
 Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps.
@@
 
+% Stuart to update the text here
 \subsection*{Surgical manipulations}
 
-After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
+After finding an optimum set of parameters for the wildtype arrangement, we
+tested the model with a number of simulations of surgical manipulations
+%(for reviews, see \citet{udin_formation_1988} and
+%\citet{goodhill_retinotectal_1999})
+.
 Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
@@
 
+% Stuart to work on...
 \subsection*{Genetic manipulations}
 
-In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
-explain. 
+% Context that might need to go in the Discussion
+%In the early 2000s, a series of genetic experiments added to the set of
+%manipulations that a model of the retinotectal system would need to explain.
+
+% Modify following so that the discursive stuff goes in the discussion.
 First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum.
@@
 \includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
+%% Not two mechanisms as the story. Story is that Fig 5 is ok, but not ace,
+%% and we have the flexibility to propose mechanisms that fix.
 \caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.

### 2023-09-15 paper/draft1/main.tex
Adds an archive of draft1, to refer back to.

--- 2022-09-29 paper/main.tex
@@
-\documentclass[9pt,lineno]{elife}
+% Use the article doc class, with an 11 pt basic font size
+\documentclass[11pt, a4paper, draft]{article}
+%\documentclass[11pt, a4paper]{article}
 % Nice citations
-%\usepackage{natbib}
+\usepackage{natbib}
 % Set margins
-%\usepackage[margin=2.5cm]{geometry}
+\usepackage[margin=2.5cm]{geometry}
 % Multilingual support
-%\usepackage[english]{babel}
+\usepackage[english]{babel}
 % Nice mathematics and Left-right harpoons for kinetic equations
@@
 % Control over maketitle
-%\usepackage{titling,titlesec}
+\usepackage{titling,titlesec}
 % Ability to use colour in text
-\usepackage{color}
+\usepackage[usenames]{color}
+\definecolor{grey}{rgb}{0.6, 0.6, 0.63}
+\definecolor{black}{rgb}{0, 0, 0}
 % For the \degree symbol
@@
 % Caption font size to be small.
-%\usepackage[font=small,labelfont=bf]{caption}
+\usepackage[font=small,labelfont=bf]{caption}
+
+% Make a dot for the dot product, call it vcdot for 'vector calculus dot'. Bigger than \cdot, smaller than \bullet.
+\makeatletter
+\newcommand*\vcdot{\mathpalette\vcdot@{.35}}
+\newcommand*\vcdot@[2]{\mathbin{\vcenter{\hbox{\scalebox{#2}{$\m@th#1\bullet$}}}}}
+\makeatother
+
+\def\firstAuthorLast{James}
+\def\Address{\\
+\affiliationListFont Department of Psychology,\\
+ \affiliationListFont  The University of Sheffield, Sheffield, UK \\
+}
+\def\corrAuthor{Seb James}
+\def\corrAddress{Department of Psychology, The University of Sheffield,
+  Western Bank, Sheffield, S10 2TP, UK}
+\def\Authors{\authorListFont Sebastian~S.~James, Stuart~P.~Wilson  \Address}
+%  \corrAuthorFont $^{*}$ Correspondence: {seb.james@sheffield.ac.uk}}
+
+% A trick to get the bibliography to show up with 1. 2. etc in place of [1], [2] etc.:
+\makeatletter
+\renewcommand\@biblabel[1]{#1.}
+\makeatother
+
+% reduce separation between bibliography items if not using natbib:
+\let\OLDthebibliography\thebibliography
+\renewcommand\thebibliography[1]{
+  \OLDthebibliography{#1}
+  \setlength{\parskip}{0pt}
+  \setlength{\itemsep}{0pt plus 0.3ex}
+}
 
 \title {
-Competitive differential chemotaxis can explain the self-organisation of topographic maps
+Differential chemotaxis with competition as a mapping mechanism in the brain
 }
+\date{} % No date please
+\author{\Authors}
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
-\affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
 %% END OF PREAMBLE
@@
 
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
+\emph{The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
 There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
@@
 We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
+We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.}
 
@@
 
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
-\begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
+In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \mathbf{M}$ with the movement vector $\mathbf{M}$ given by
+\begin{equation} \label{e:M}
+ \mathbf{M} = \mathbf{G} +  \mathbf{X} + \mathbf{B}.
 \end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
+$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{X}$ is the movement due to competitive interactions with other branches. 
 $\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
 
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
+\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{X}$ was studied.
+Their model, which successfully accounts for a wide range of experimental manipulations, found an $\mathbf{X}$ comprised of two components; a simple distance-based competition between \emph{all} branches ($\mathbf{C}$), along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon ($\mathbf{I}$).
 We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
@@
 
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
+ We modelled competition as a distance-based competitive movement vector, $\mathbf{X}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
 %
 \begin{equation} \label{e:X}
-\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
+\mathbf{X} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
 \end{equation}
@@
 
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
+We used a combination of pattern metrics to define the objective function. $\epsilon$ is the mean distance error of the axon centroids, given by:
 %
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
+\begin{equation}\label{e:eps}
+\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
+\end{equation}
 %
@@
 A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
+The differential chemotaxis with competition model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
 %
@@
 It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
+The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the differential chemotaxis with competition model.
 We address possible interaction effects in the next section.
@@
 This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
+It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:M} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
 
@@
 
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
+Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{X}$ from Eq.\,\ref{e:M} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
+However, because we intended for $\mathbf{X}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
 If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
@@
 
-\section{Method details}
+\section*{Method details}
 
 \subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
+The border effect ($\mathbf{B}$ in Eq.\,\ref{e:M}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
 For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
@@
 
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
+\subsection*{Bootstrap analyses of R-C tectal position vs.~N-T retinal origin}
 
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
+To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R-C positions obtained for the $n=20$ different dorso-ventral origins for each N-T location.
 The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
@@
 
-\section{Data Availability}
+\section*{Data Availability}
 
 All data and code required to reproduce the results in this work are available in the public repository 
-\url{https://github.com/sebjameswml/RetinoTectal}.
+{https://github.com/sebjameswml/RetinoTectal}.
 The software was built using the library morphologica \citep{james_morphologica_2022}.
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 
@@
 
+% Bibliography here
+\selectlanguage{English}
+\bibliographystyle{apalike}
 \bibliography{RetinoTectal}
+
+\appendix
 
@@
 % Expand panel A to be able to lose B-D.
-% Full page width
 \begin{figure}
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
+\includegraphics[width=\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
 \caption{Receptor-ligand model of chemotaxis and competition. 
@@
 \label{f:GJ}
-\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_grafts.png}
 \caption{Performance of the receptor-ligand model with surgical graft manipulations.
@@
 \label{f:GJsurg_grafts}
-\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
+\includegraphics[width=\linewidth]{./images/fig_GJ_surgical_ablations.png}
 \caption{Retinal and tectal ablation simulations.
@@
 \label{f:GJsurg_ablate}
-\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations.png}
 \caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_explanation.png}
 \caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
@@
 \begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
+\includegraphics[width=0.95\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
 \caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}. 
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
-graft-rotate 90 degrees surgical manipulations. 
-\textbf{A} Ligand expression. 
-The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps. 
-The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$. 
-Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the 
-The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively. 
-\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A. 
-For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis). 
-The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
-Full saturation indicates a positive gradient; white indicates a negative gradient.
-For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis. 
-Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
-\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
+\includegraphics[width=\linewidth]{./images/Tissuevisb.png}
+\caption{(Supplementary) Visualisation of ligand expression and gradients on the tectum for the
+graft-rotate 90 degrees simulated surgical manipulations.}
 \label{f:trot90}
-\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
+\includegraphics[width=\linewidth]{./images/expression_profiles_EphBs.png}
 \caption{(Supplementary) EphB and ephrin B gene expression.
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature. 
-\textbf{A} Retinal expression of EphB2. 
+Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature \textbf{A} Retinal expression of EphB2.
 We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application.
@@
 \label{f:expr_ephb}
-\end{fullwidth}
 \end{figure}
@@
 \begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
+\includegraphics[width=0.9\linewidth]{./images/noise_contribution.png}
 \caption{(Supplementary) Response of the model to noise.
@@
 \label{f:noise}
-\end{fullwidth}
 \end{figure}

### 2023-09-15 paper/draft1/zerodraft.tex
Adds an archive of draft1, to refer back to.

--- 2022-03-07 paper/paper.tex
@@
  centroids \textbf{C} The pattern metric $\eta$ is zero, as it was in
- Fig.\,\ref{f:G}F and $\epsilon$ approaches a much lower value as $t \to \infty.}
+ Fig.\,\ref{f:G}F and $\epsilon$ approaches a much lower value as $t \to \infty$.}
 \label{f:G0}

### 2023-09-15 paper/main.tex
Completes update of initial Results section in which model is described

@@
 \newcommand{\dvrg}{\nabla\vcdot\nabla}
+
+\definecolor{colcmnt}        {rgb} {0.7098, 0.0745, 0.0431}
+\newcommand{\redcmnt}[1]{\textcolor{colcmnt}{#1}}
+
 %% END aliases
@@
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
+\author[1]{Sebastian~S.~James}
+\author[1*]{Stuart~P.~Wilson}
 \affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
+\corr{s.wilson@sheffield.ac.uk}{SW}
 %% END OF PREAMBLE
@@
 \cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition
+A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that \redcmnt{one focused on competition
 % Stuart: I think the model is Koulakov's, not Triplett's.
@@
 *Hjorth* shows that the Koulakov model is the best one at reproducing the collapse in the EphA3 ki experiment.
-by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
+by \cite{Triplett2011}} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
 
-The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
+\redcmnt{The model of \cite{Triplett2011} describes maps as minima of an energy function}, whose structure can be obtained via a stochastic optimisation process.
 Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
@@
 
-We adopt the same form for the expression of receptors on retinal ganglion cells as
+Each of $n$ retinal ganglion cells (RGC) projects $M$ growth cones (referred to as \emph{branches}) which carry a set of $N$ different receptor types, expressed at levels determined by the cell soma's location on the retinal grid.
+RGC growth cones and cells on the tectum are also assumed to express $N$ different ligand types which interact with RGC receptors.
+%
+We adopt the same form for the expression of receptors by retinal ganglion cells as
 \citet{simpson_simple_2011}:
@@
 %% * Simpson and Goodhill modelled according to deltax = B + C + G (Eq 2)
-In \citet{simpson_simple_2011}'s model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
+In Simpson \& GoodHill's model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
 % Maybe re-order the terms?
@@
 Simpson \& Goodhill's model had a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-
 %Although the model incorporated EphA expression, the distance-based competitive interaction was given no mechanism of operation.
-
 To provide an explicit mechanism of operation for the competition, we modelled it as a sum of interactions that are `switched on' via receptor-ligand signalling, meaning that some axons would have no competitive interaction, while others would repel strongly.
@@
 %
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to $b$.
+$m_{\!_X}$ is a scalar parameter and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
+$r_{\!_X}$ is the receptor-ligand interaction radius.
+$nM$ is the total number of branches in the system and $\hat{\mathbf{x}}_{kb}$ is the unit vector from branch $k$ to $b$.
 %
-$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$r_{\!_X}$ is the receptor-ligand interaction radius.
+$W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$).
 %
@@
 %
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
+$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$).
+$Q$ is set to 1 only if at least one repulsive signal exceeds a positive signal threshold, $s$, and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$:
 %
@@
 $F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
-$s$ is a positive signal threshold for interaction.
 With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
-$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.
 
@@
 
-* The S\&G chemotaxis was a placeholder because they chose G to be a vector
-  pointing to a predetermined target location. In order to make this a local
-  model we do the following....
-
-We also imagine ligand expression on the tectum (L) - describe these here.
+\citet{simpson_simple_2011} chose a placeholder mechanism for the chemotaxis effect, using a global supervisor to set $\mathbf{G}$ to a vector which would always point to the predetermined target location for each branch.
+In order to make this a local model we replaced the global supervisor with an interaction between RGC receptors and ligands expressed on the tectum.
+We used forms for the ligand expression, $L(\mathbf{x})$, which matched those used for the RGC receptor expression.
 
@@
 
-%%%%% Describing the simulation TAKE THIS TO METHOD DETAILS
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}.
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-%%%%%%%%%%%%%
-
-% Keep, but say M grown cones. (M=4 specified in method details)
-% ALSO: Put M near competition equation where it's used!
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-
-The expression of each receptor varies only with respect to one spatial dimension.
-
-% Thi sis in the righ tplace here
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
+Each of the $M$ RGC growth cones carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
+%
+%The expression of each receptor varies only with respect to one spatial dimension.
+%
+We defined 4 receptor expression gradients (each varying only wrt one spatial dimension) arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
 $r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
@@
 
-% NO NEED FOR This sentence/bit
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
+% NO NEED FOR This sentence/bit (keep comment in case for discussion)
+%Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
 %
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
+% Already said this:
+%We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
 
@@
 
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
+%While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
 %%%%%%%%%%%%%%%%%
 
-\subsection*{Simulation} % MOST OF THIS GOES TO METHOD DETAILS, just keep
-                         % first sentence, have list of values used in sims, then 'see method details',
+\subsubsection*{Simulation}
 
@@
 We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other.
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
+The parameters found were $m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$ (see method details).
 
@@
 
+\subsubsection*{Outtakes}
+
+Previously in results:
+
+While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
+
 \subsection*{Chemotaxis}
@@
 
+\subsection*{Simulation domain}
+We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}.
+The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
+One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
+
+% Make sure this ain't a copy of results
+Each retinal ganglion cell (RGC) projected $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
+
+The expression of each receptor was varied only with respect to one spatial dimension.
+
+% Thi sis in the right place here
+We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
+$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
+Thus the model is of \emph{differential chemotaxis}.
+
 \subsection*{Initial conditions}
@@
 % end initial conditions
+
+\subsection*{Receptor and Ligand expression forms}
+
+A section to add more detail on the functions used, if necessary.
+
+\subsection*{Parameter finding}
+
+We performed a simulated annealing optimization to find three free parameters; $m_{\!_X}$, $r_{\!_X}$ and $s$ which would balance a chosen evel of chemotaxis strength given by $m_{\!_G}=0.00120$.
+We minimized an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
+
+We used a combination of pattern metrics to define the objective function.
+$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
+%
+%\begin{equation}\label{e:eps}
+%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
+%\end{equation}
+%
+where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
+$n$ is the number of RGC axons in the system.
+%
+$\epsilon$ tends to 0 as the pattern approaches the target layout.
+However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other.
+For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
+The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
+The parameters returned by the simulated annealing were
+$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
 

### 2024-07-04 scripts/paper_figures.sh
Updates to images (following switch to HSV colour map

@@ -26,7 +26,10 @@ rm -rf ./rcnt/*.h5
 # 3. EphA3 ki/+ and EphA4 +/-
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-wt.json -co:exit=true -co:steps=1500
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_kiki-wt.json -co:exit=true -co:steps=1500
+# e_eph_ki-kd_lite.json has less knockdown than e_eph_ki-kd.json (1.2 vs. 2.1 at time of writing)
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-kd_lite.json -co:exit=true -co:steps=1500
+
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-kd.json -co:exit=true -co:steps=1500
 
 # Extended GJ model (EphA cluster size) plus the r2 collapse condition (EphA3 and EphA4 manipulations)
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1_EphA4_r2collapse.json configs/simpler/e_eph_ki-wt.json -co:exit=true -co:steps=1500

### 2024-07-04 analysis/bs_manip_vs_wt.py
Results updates

@@ -115,49 +115,36 @@ import h5py
 # These are RC-vs-NT graphs generated by the simulation. Run the three
 # simulations for the Genetic manipulation sims to create these.
 
-# 0: ORIGINAL model 1: Clustersize model 2: Clustersize + r2 collapse
+# 0: GJ model 1: GJ + Clustersize model 2: GJ + Clustersize + r2 collapse
 model_type = 0
-doitall = 0
+doitall = 1     # overrrides model_type
+
+# Original GJ Model (Figure 5 in paper draft)
+filebases_0 = ["../rcnt/j4_ee_GJ_best_1_eph_ki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_eph_kiki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_eph_ki-kd_exit_true_steps_1500"]
+
+# Clustersize-only Model (not currently generated)
+filebases_1 = ["../rcnt/j4_ee_GJ_best_1_cluster_eph_ki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_cluster_eph_kiki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_cluster_eph_ki-kd_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_cluster_eph_wt-kd_exit_true_steps_1500"]
+
+# Extended GJ Model. Clustersize plus r2 collapse (Fig 7 in paper draft)
+filebases_2 = ["../rcnt/j4_ee_GJ_best_1_EphA4_r2collapse_eph_ki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_EphA4_r2collapse_eph_kiki-wt_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_EphA4_r2collapse_eph_ki-kd_exit_true_steps_1500",
+               "../rcnt/j4_ee_GJ_best_1_EphA4_r2collapse_eph_wt-kd_exit_true_steps_1500"]
 if doitall:
-    filebases = ["../rcnt/j4_ee_GJ_best_1_eph_ki-wt_exit_true_steps_1500",
-                         "../rcnt/j4_ee_GJ_best_1_eph_kiki-wt_exit_true_steps_1500",
[…]

### 2024-07-04 misc/examineRetEph.py

@@ -43,7 +43,7 @@ def examineRetEph():
 
     # Knockin and knockdown, which should be copied from e_eph_ki-wt.json and e_eph_ki-kd.json (
     ki = .8
-    kd = 2.1
+    kd = 1.9
     kiki = 3.2
 
     ## Binding affinity for EphAx. prop. to 1/K_D. see Monschau et al

### 2024-07-04 scripts/paper_figures.sh

@@ -26,9 +26,6 @@ rm -rf ./rcnt/*.h5
 # 3. EphA3 ki/+ and EphA4 +/-
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-wt.json -co:exit=true -co:steps=1500
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_kiki-wt.json -co:exit=true -co:steps=1500
-# e_eph_ki-kd_lite.json has less knockdown than e_eph_ki-kd.json (1.2 vs. 2.1 at time of writing)
-./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-kd_lite.json -co:exit=true -co:steps=1500
-
 ./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-kd.json -co:exit=true -co:steps=1500
 
 # Extended GJ model (EphA cluster size) plus the r2 collapse condition (EphA3 and EphA4 manipulations)

### 2024-07-04 scripts/rerun_kikd.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+#
+# Generate all the paper figure graph elements, one after the other.
+#
+
+# First clear out some of rcnt
+rm -f ./rcnt/j4_ee_GJ_best_1_EphA4_r2collapse_eph_ki-kd_exit_true_steps_1500_*.h5
+rm -f ./rcnt/j4_ee_GJ_best_1_eph_ki-kd_exit_true_steps_1500_*.h5
+
+# GJ model. EphA3 and EphA4 manipulations
+# 1. EphA3 ki/+ and EphA4 +/+ (i.e. EphA4 is wildtype/unmodified)
+# 2. EphA3 ki/ki and EphA4 +/+
+# 3. EphA3 ki/+ and EphA4 +/-
+
+#./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-wt.json -co:exit=true -co:steps=1500
+#./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_kiki-wt.json -co:exit=true -co:steps=1500
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1.json configs/simpler/e_eph_ki-kd.json -co:exit=true -co:steps=1500
+
+# Extended GJ model (EphA cluster size) plus the r2 collapse condition (EphA3 and EphA4 manipulations)
+#./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1_EphA4_r2collapse.json configs/simpler/e_eph_ki-wt.json -co:exit=true -co:steps=1500
+#./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1_EphA4_r2collapse.json configs/simpler/e_eph_kiki-wt.json -co:exit=true -co:steps=1500
+./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1_EphA4_r2collapse.json configs/simpler/e_eph_ki-kd.json -co:exit=true -co:steps=1500
+#./build/sim/agent/agent1 configs/simpler/m_ee_GJ_best_1_EphA4_r2collapse.json configs/simpler/e_eph_wt-kd.json -co:exit=true -co:steps=1500

### 2024-07-04 paper/neural_2024/poster_text/abs.tex
Text for poster

--- 2023-08-25 paper/main.tex
@@
-\documentclass[9pt,lineno]{elife}
+\documentclass[9pt]{elife} % no lineno or draft
 % Nice citations
@@
 
-\title {
-Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
-}
+\title {none}
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
+\author[1]{Sebastian~S.~James}
+\author[1*]{Stuart~P.~Wilson}
 \affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
+\corr{s.wilson@sheffield.ac.uk}{SW}
 %% END OF PREAMBLE
 
-% Seb to fix: Expected not experiment (in figs), Fig 8 - monochromes. Fig 1 - monochromes
-% Stuart to do: Intro/Start of results modifications
+\begin{document}
 
-\begin{document}
-%\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
+The axons of retinal ganglion cells form an ordered map of the retina
+in the optic tectum. Surgical and genetic experiments have revealed
+several key mechanisms that contribute to retinotectal map
+development, and computational modelling studies have demonstrated how
+maps can be shaped by their interactions. Models derived from Sperry's
+classic chemoaffinity theory can reproduce retinotopy, and Gierer
+later showed how an extension to include competition between growing
+axons could also account for a subset of the disordered maps that have
+resulted from surgical manipulation. Further extensions have each
+accounted for additional datapoints, but none has been able to explain
+the full range of maps observed experimentally. Here we present a
+unified model of retinotectal development, assuming only local
+interactions between growing axons, which can account for the full
+range of maps resulting from surgical manipulation. This model extends
+the `potential functions' of Gierer's original to describe
+self-organisation across a 2D cortical sheet, and it represents a
+coordination of axonal growth by orthogonal pairs of opposing ligand
+gradients in the tectum and corresponding receptor gradients in the
+retina. In addition, we present progress towards accounting for the
+full range of maps that have resulted from genetic manipulation, by
+incorporating a mechanistic description of receptor ligand
+interactions.
 
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
-There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
-Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
-However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
-Sperry's chemoaffinity theory provides an account of the wildtype pattern.
-Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
-Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
-
-\section{Introduction}
-
-Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
-Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
-Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
-
-One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
-A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
-
-A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
-\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
-
-The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
-Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
-Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
-
-The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
-According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
-As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
-Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
-The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
-
-These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
-Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
-
-The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
-However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
-Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
-
-In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
-For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
-
-Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
-
-The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
-However, a major limitation is that chemotaxis is not implemented explicitly.
-Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
-This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
-
-An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
-If so, such a model would represent the most complete account of retinotectal map self-organization to date.
-
-%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-
-%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
-The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-
-%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-
-%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-
-%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
-This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
-
-%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
-% And sounds?
-
-%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
-Two classes of ephrins have been identified, labelled `A' and `B'.
-Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
-Ephrin B acts on medial-lateral ordering with an attractive interaction.
-Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
-
-%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
-The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
-Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
-Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
-He presented gradient following as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-This model forms a mathematical expression of Sperry's theory.
-Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
-ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-
-%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
-He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
-
-%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
-
-%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
-In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
-Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
-With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
-ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
-and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
-
-\section{Results}
-\subsection*{A model of chemotaxis and competition}
-
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
-\begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
-\end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
-
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
-
-\subsection*{A gradient based model of chemotaxis}
-
-We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
-
-\begin{equation}\label{e:G}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
-$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
-%
-We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
-%
-$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
-
-%%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
-The expression of each receptor varies only with respect to one spatial dimension.
-%
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-%
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
-
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
-
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-
-\subsection*{A signalling model for competition}
-
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
-%
-\begin{equation} \label{e:X}
-\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
-\end{equation}
-%
-$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
-%
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
-%
-\begin{equation}
-Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
-                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
-                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_X} \\
-                 1 & \mathrm{otherwise.}
-     \end{cases}
-\end{equation}
-%
-$l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$. 
-$F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
-$s$ is a positive signal threshold for interaction.
-With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
-$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.
-
-\subsection*{Simulation}
-
-The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other. 
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
-
-Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps. 
-Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red.
-Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
-Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
-The axons retain the colour of their originating cell on the retina.
-Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
-Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
-In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation. 
-The number of line crossings in this plot gives an indication of the disorder in the arrangement.
-Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
-By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
-Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
-Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
-For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E. 
-Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.
-
-\subsection*{Surgical manipulations}
-
-After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
-Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
-%
-To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column). 
-After the manipulation, we re-computed the tectal ligand expression gradients.
-Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region. 
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
-We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red). 
-The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04). 
-The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
-The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
-%
-Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. 
-There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-
-Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
-According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}). 
-The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
-%
-Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
-Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
-Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
-%
-A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
-%
-The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}.
-Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
-
-\subsection*{Genetic manipulations}
-
-In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
-explain. 
-First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum. 
-For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours. 
-By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
-In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5). 
-This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
-iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
-
-An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
-To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
-The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
-In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue. 
-The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
-The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
-Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
-
-Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
-The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient. 
-The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
-Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
-
-Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
-
-It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
-We address possible interaction effects in the next section.
-
-\subsection*{Two mechanisms to address the genetic results: Cluster size and $r_2$ collapse}
-
-The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
-This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
-We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different. 
-We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
-We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
-The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
-Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
-
-It is known that EphA receptors form clusters, that these are necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
-It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
-Additionally, EphA4 receptors make `side attachments' to clusters of EphA receptors, possibly regulating the cluster size \citep{nikolov_ephephrin_2013}. 
-%
-We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$. 
-We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4}$, the density of EphA4 receptors:
-\begin{equation}
-    c_0 = 1/r_{\!\scriptscriptstyle A4}
-\end{equation}
-%
-Eq.\,\ref{e:G} was modified to include the cluster size:
-%
-\begin{equation}\label{e:Gcs}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-
-$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
-Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).  
-Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
-To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
-
-Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
-We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
-In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
-Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
-In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
-If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
-If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
-
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
-This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
-This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
-Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
-This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
-The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
-Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
-
-Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
-
-\section{Discussion}
-
-% Summarize the findings and place them in context
-Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
-If not, does the addition of competition resolve the remaining problems? 
-Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
-To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
-The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold. 
-These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals. 
-The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development. 
-Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
-Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
-
-However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by the model. 
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
-
-To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
-First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
-A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
-In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
-We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
-This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
-However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA. 
-A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
-We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
-We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
-
-\subsection*{Chemotaxis}
-
-The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-
-Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
-Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
-This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
-
-Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
-Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
-Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
-With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
-Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
-RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
-Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
-Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
-However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
-In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum. 
-The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}. 
-Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
-Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-
-We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
-This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-\subsection*{Competition}
-
-Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
-We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive.
-An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.
-
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
-If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
-
-Competition appears to be necessary for all of the surgical manipulations. 
-It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
-
-The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
-This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
-If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map. 
-If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}. 
-It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
-
-% Single axon behaviour
-\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
-The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
-The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
-
-\subsection*{Attractive receptor-ligand interactions}
-
-In the results, we present a model in which all receptor-ligand interactions are repulsive.
-However, EphB/ephrinB interactions are largely attractive in nature.
-It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
-When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
-Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
-
-\subsection*{Robustness to noise}
-
-To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
-Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
-We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
-The graphs show the expression values on a normalised scale. 
-We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
-The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
-Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
-i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
-ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
-iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
-(ii) and (iii) are computed once only, at the start of the simulation.
-Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined. 
-It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A). 
-The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern. 
-Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions. 
-Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
-
-%\subsection*{The value of $s$}
-%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Limitations/weaknesses of research
-\subsection*{Predictions and Limitations}
-
-We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth. 
-We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
-By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
-We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
-
-Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
-In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
-Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
-Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
-Adapting this framework to address an axon branching model is left for future work.
-
-% Restate significant findings and implications
-
-\section{Method details}
-
-\subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
-For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
-%
-\begin{equation}\label{e:B}
-B_x = \begin{cases}
-        \frac{1}{2} (r-x)     & x<r \\
-        \frac{1}{2} (1-r-x)    & x>1-r
-\end{cases}
-B_y = \begin{cases}
-        \frac{1}{2} (r-y)      & y<r \\
-        \frac{1}{2} (1-r-y)    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
-% end border effect description
-
-\subsection*{Initial conditions}
-Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation. 
-Each RGC axon $k$ was assigned a random initial position coordinate:
-\begin{equation}\label{e:ic}
-\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$.
-Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
-% end initial conditions
-
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
-
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
-The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
-Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
-The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
-%
-$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
-%
-where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
-
-A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
-The transformations are defined as:
-%
-$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
-%
-where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
-$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$. 
-The Studentized difference of the means of the resamples is computed:
-
-$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
-
-If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
-If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$. 
-We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL). 
-The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
-
-The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
-%% End bootstrap desc
-
-\section{Data Availability}
-
-All data and code required to reproduce the results in this work are available in the public repository 
-{https://github.com/sebjameswml/RetinoTectal}.
-The software was built using the library morphologica \citep{james_morphologica_2022}.
-
-\section{Acknowledgements}
-
-This work was supported by grant 220020516 from the James S. McDonnell Foundation.
-
-\bibliography{RetinoTectal}
-
-%
-% FIGURES
-%
-
-% FIGURE 1
-% Expand panel A to be able to lose B-D.
-% Full page width
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Schematic diagrams of retinotectal mapping and gene expression.
-Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
-%
-\textbf{A} Tectal ligand expression $L_i$.
-Tectal ligand expression patterns are assumed to increase exponentially in one direction only across the tectal surface.
-$L_0$ (or ephrin A) increases in a caudal to rostral direction (i.e.
-with increasing $y$; $x$ and $y$ directions are indicated with arrows).
-$L_2$ increases in the opposing sense, with $-y$.
-$L_1$ and $L_3$ form counter-gradients in the $\pm x$ directions.
-\emph{L: lateral, M: medial, C: caudal, R: rostral}.
-%
-\textbf{B} Retinal receptor and ligand expressions.
-Four receptors, $r_i$, are expressed across the retina.
-$r_0$ increases as an exponential function of $-x$ ; $r_1$ increases with an exponential function of $-y$.
-$r_2$ and $r_3$ have the opposite sense and increase with positive $x$ and $y$, respectively.
-Retinal ligand expression, $l_i$, is expressed in two complimentary gradients \citep{hornberger_modulation_1999}.
- \emph{N: nasal, T: temporal, D: dorsal, V: ventral}.
-%
-}
-\label{f:ex}
-\end{figure}
-
-% FIGURE 2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
-\caption{Receptor-ligand model of chemotaxis and competition. 
-\textbf{A} The origin position on the retina gives axons a colour according to a dual-colour map.
-Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
-\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina. 
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results (that is, correctly ordered into a perfect map of the simulated, square retina). 
-\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time. 
-\textbf{G} The arrangement of axon branches at t=1500. 
-\textbf{H} Selected axon locations at t=1500, with path histories in grey. 
-\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-\label{f:GJ}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 3 ./scripts/fig_GJ_surgical_grafts.svg (graphs generated with fig_GJ_surgical.sh)
-% Could split into retina and tectum or 
-% a: rotation/graft (as hand waved by Gierer)
-% b: ablation (as simulated in 1D by Gierer)
-% c: mismatch/compound (as extensions of Gierer)
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
-\caption{Performance of the receptor-ligand model with surgical graft manipulations.
-Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
-\textbf{A.} Graft rotation.
-An 8$\times$8 square patch of tectum has been rotated by 90$\degree$, altering the tectal ligand expression pattern.
-High gradients form around the new discontinuities at the edge of the patch (see Fig\,\ref{f:trot90}), driving axons to follow the rotation of the patch.
-The left graph shows the axon centroid positions after 1500 time steps, the smaller inset graph shows the (idealised) anticipated layout seen in experiments \citep{levine_deployment_1974}.
-The right-hand graph shows the time-course of the pattern quality metric, $\epsilon$ (i.e.~how closely the centroid pattern matches the inset) for the entire tectum (blue), the graft region (yellow) and the surround (red).
-The axons within the graft are slightly less well ordered than those outside.
-\textbf{B} Shows the same type of graft rotation as \textbf{A} for a 180$\degree$ rotation \citep{yoon_retention_1973}.
-\textbf{C.} A simulated graft swap experiment.
-Two 12$\times$4 rectangular regions of tectum are cut out and grafted back in swapped positions \citep{hope_arrow_1976,gaze_visuotectal_1983}.
-While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
-Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.}
-\label{f:GJsurg_grafts}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 4
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
-\caption{Retinal and tectal ablation simulations.
-\textbf{A} The temporal half of the retina is removed in this manipulation.
-Axons from the nasal half grow onto an unmodified tectum.
-This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
-In this work, it is the competition that achieves this effect.
-The axon centroid graphs are plotted as in Fig.\,\ref{f:GJsurg_grafts}, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
-\textbf{B} Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
-As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
-\textbf{C} A `mismatch' experiment \citep{horder_retention_1971}.
-The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
-As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
-\textbf{D} Compound eye.
-If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
-The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
-\label{f:GJsurg_ablate}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 5
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-In each manipulation, the left graph shows axon branch agents (colour coded by retinal origin) at the end of a simulation ($t=1500$), the centre graph shows corresponding axon centroids, with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
-In the right hand graph, the final tectal rostro-caudal locations of axons from each naso-temporal origin in the retina are plotted as violin plots (bars indicate the max/min datum in each group and shading indicates the data distribution).
-Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-\textbf{A} We simulated the heterozygous EphA3 knock-in (EphA3 ki/+) of \citet{brown_topographic_2000} by selecting half of the retinal cells at random, and in these cells adding the value 0.8 to $r_0$ (which, as its expression increases towards the temporal retina, is our analogue for EphA).
-The affected cells experienced altered interaction with tectal ligand gradients and were shifted rostrally.
-The rostral shift was largest for nasal cells bound for the caudal tectum.
-These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
-Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (thus the $r_0$-$L_0$ interaction is minimal).
-However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis, as seen in the inset graph.
-\textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
-Here, we observed a separation into two maps, one for the wildtype cells and one for the manipulated cells, in agreement with \citet{brown_topographic_2000} (inset is from their Fig.\,5C).
-\textbf{C} Simulation of EphA3 ki/+ selective knock-in and EphA4 knock-down (all cells) from \citet{reber_relative_2004}.
-The knock-down was applied by reducing $r_0$ in all cells by 1.2.
-The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in experiments.
-Instead, selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in.}
-\label{f:GJeph}
-\end{figure}
-
-% FIGURE 6
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
-\caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
-To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and for the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrinA ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA (i.e.~with $r_2$).
-\textbf{A} \emph{EphA/ephrinA expression.} As described in the text, EphA ($r_0 \equiv \sum_{i\neq 4}\mathrm{EphA}i$) is expressed according to $r_0 = 1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position.
-Knock-in is simulated by adding 0.8 to this function (giving the purple curve).
-Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
-A threshold, $h_0$, used by the $r_2$ collapse mechanism is shown as a red dotted line, with a green dash-dot line indicating the intersection of $\Sigma\mathrm{EphA}$ and $h_0$.
-\textbf{B} \emph{EphA4 expression.} Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA ($l_0$) can bind to EphA4 ($r_{A4}$), modelled here as a multiplicative interaction to give cis-bound EphA4, $r_{A4}^{\mathrm{cis}} = w_{A4} \times r_{A4} \times l_0$ (green dashed curve), where $w_{A4}$ is a binding weight described in the text.
-The remaining, un-bound, free EphA4 receptors, $r_{A4}^{\mathrm{free}} = r_{A4} - r_{A4}^{\mathrm{cis}}$ (green solid curve).
-We model genetic knock-down of EphA4 by subtracting 2.1 from $r_{A4}$, then recomputing $r_{A4}^{\mathrm{cis}}$ and $r_{A4}^{\mathrm{free}}$ to give the cyan curves.
-\textbf{C} \emph{Cluster size affects signal strength.} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}^{\mathrm{free}}$.
-Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced by combined knock-in and knock-down.
-This cluster size-modulated signal, $s_0$, is smaller in magnitude than the signal for $r_0$ in the original model (for which $s_0 \equiv r_0$).
-To compensate, $r_2$ is divided by 2, which gives the correct behaviour in a wildtype system.
-\textbf{D} \emph{$r_2$ collapse.} i) We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=2$ (red dotted line in A) and that $r_{A4}^{\mathrm{free}}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
-Under these conditions, $r_2$ is reduced to one fifth of the value it would otherwise take.
-ii) The regions of $r_2$ collapse are illustrated.
-No collapse occurs for wildtype EphA and EphA4 expressions; in cells with only EphA4 knock-down, $r_2$ collapse occurs for temporal cells (those more temporal than the dash-dot line in A).
-In cells with both EphA3 knock-in and EphA4 knock-down, the collapse mechanism occurs regardless of retinal origin.}
-\label{f:clustermech}
-\end{figure}
-
-% FIGURE 7
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
-\caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}. 
-\textbf{A} EphA3 knock-in result for ki=0.8.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment (inset).
-This is caused by the cluster size effect, in which the signal increase due to the knock-in is reduced for temporal RGCs compared with nasal cells (See Fig.\,\ref{f:clustermech}C, green/purple solid curves).
-%FOR TEXT: It is confirmed by an increased achieved significance level (ASL) from a bootstrapped, studentized t-test comparing the red and blue distributions at each N--T location. 
-\textbf{B} The cluster size effect does not change the double knock-in result, which shows excellent agreement with experiment (inset). 
-\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked down, cells on the red curve experience a combination of interactions: First, the increased interaction with ephrinA that gave the result in A is still in effect.
-Second, all cells on the red curve experience $r_2$ collapse, whereas only temporal RGCs on the blue curve experience this effect.
-$r_2$ receptors are more numerous on nasal RGCs and so the collapse in $r_2$ generates a substantial rostral movement of the red curve, in agreement with \citet{reber_relative_2004}.
-\textbf{D} Uniform knock-down of EphA4 in all cells (with no manipulation of EphA3) does not perturb the mapping to a significant extent, although the effect of the EphA4 knock-down on the cluster size mechanism causes a slight rostral shift in all cells.}
-\label{f:r2collapse}
-\end{figure}
-
-% ANY SUPPLEMENTARY FIGURES GO AFTER THIS
-%\renewcommand{\thefigure}{S\arabic{figure}}
-%\setcounter{figure}{0}
-
-% FIGURE S1
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
-graft-rotate 90 degrees surgical manipulations. 
-\textbf{A} Ligand expression. 
-The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps. 
-The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$. 
-Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the 
-The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively. 
-\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A. 
-For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis). 
-The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
-Full saturation indicates a positive gradient; white indicates a negative gradient.
-For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis. 
-Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
-\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
-\label{f:trot90}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
-\caption{(Supplementary) EphB and ephrin B gene expression. 
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature. 
-\textbf{A} Retinal expression of EphB2. 
-We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application. 
-The resulting, normalized expression is shown in the graph for the E4 data, with an inset showing corresponding data for the E6 data. 
-In the right-hand graph, we show an expansion of a portion of the E4 data to which we applied a linear fit to determine the RMS deviation from the mean to estimate the noise in the receptor signal. 
-The mean RMS error was 0.06. 
-\textbf{B} Stalefish screenshots and expression graphs for retinal ephrinB1 and ephrinB2 expression, using data from Fig.\,1 of \citet{braisted_graded_1997}. 
-\textbf{C} Screenshots and expression graphs for tectal ephrinB1 expression, using data from Fig.\,4 of \citet{braisted_graded_1997}.}
-\label{f:expr_ephb}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S3
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
-\caption{(Supplementary) Response of the model to noise.
-\textbf{A} Adding 3\% gradient measurement noise (normally distributed, standard deviation, $\sigma = 0.19$) has negligible effect on the pattern. 
-\textbf{B} 3\% retinal receptor noise ($\sigma = 0.08$). 
-\textbf{C} 3\% tectal ligand noise ($\sigma = 0.08$). 
-\textbf{D} Combined effect of retinal receptor, tectal ligand and gradient measurement noise is dominated by the tectal ligand noise.}
-\label{f:noise}
-\end{fullwidth}
-\end{figure}
+%\bibliography{RetinoTectal}
 

### 2024-07-04 paper/neural_2024/poster_text/fig1.tex
Text for poster

--- 2023-08-25 paper/main.tex
@@
 
-\title {
-Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
-}
+\title {none}
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
+\author[1]{Sebastian~S.~James}
+\author[1*]{Stuart~P.~Wilson}
 \affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
+\corr{s.wilson@sheffield.ac.uk}{SW}
 %% END OF PREAMBLE
 
-% Seb to fix: Expected not experiment (in figs), Fig 8 - monochromes. Fig 1 - monochromes
-% Stuart to do: Intro/Start of results modifications
+\begin{document}
 
-\begin{document}
-%\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
-
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
-There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
-Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
-However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
-Sperry's chemoaffinity theory provides an account of the wildtype pattern.
-Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
-Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
-
-\section{Introduction}
-
-Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
-Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
-Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
-
-One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
-A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
-
-A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
-\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
-
-The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
-Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
-Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
-
-The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
-According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
-As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
-Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
-The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
-
-These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
-Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
-
-The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
-However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
-Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
-
-In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
-For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
-
-Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
-
-The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
-However, a major limitation is that chemotaxis is not implemented explicitly.
-Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
-This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
-
-An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
-If so, such a model would represent the most complete account of retinotectal map self-organization to date.
-
-%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-
-%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
-The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-
-%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-
-%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-
-%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
-This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
-
-%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
-% And sounds?
-
-%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
-Two classes of ephrins have been identified, labelled `A' and `B'.
-Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
-Ephrin B acts on medial-lateral ordering with an attractive interaction.
-Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
-
-%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
-The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
-Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
-Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
-He presented gradient following as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-This model forms a mathematical expression of Sperry's theory.
-Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
-ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-
-%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
-He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
-
-%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
-
-%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
-In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
-Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
-With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
-ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
-and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
-
-\section{Results}
-\subsection*{A model of chemotaxis and competition}
-
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
-\begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
-\end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
-
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
-
-\subsection*{A gradient based model of chemotaxis}
-
-We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
-
-\begin{equation}\label{e:G}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
-$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
-%
-We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
-%
-$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
-
-%%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
-The expression of each receptor varies only with respect to one spatial dimension.
-%
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-%
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
-
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
-
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-
-\subsection*{A signalling model for competition}
-
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
-%
-\begin{equation} \label{e:X}
-\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
-\end{equation}
-%
-$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
-%
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
-%
-\begin{equation}
-Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
-                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
-                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_X} \\
-                 1 & \mathrm{otherwise.}
-     \end{cases}
-\end{equation}
-%
-$l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$. 
-$F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
-$s$ is a positive signal threshold for interaction.
-With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
-$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.
-
-\subsection*{Simulation}
-
-The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other. 
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
-
-Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps. 
-Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red.
-Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
-Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
-The axons retain the colour of their originating cell on the retina.
-Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
-Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
-In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation. 
-The number of line crossings in this plot gives an indication of the disorder in the arrangement.
-Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
-By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
-Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
-Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
-For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E. 
-Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.
-
-\subsection*{Surgical manipulations}
-
-After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
-Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
-%
-To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column). 
-After the manipulation, we re-computed the tectal ligand expression gradients.
-Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region. 
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
-We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red). 
-The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04). 
-The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
-The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
-%
-Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. 
-There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-
-Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
-According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}). 
-The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
-%
-Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
-Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
-Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
-%
-A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
-%
-The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}.
-Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
-
-\subsection*{Genetic manipulations}
-
-In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
-explain. 
-First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum. 
-For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours. 
-By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
-In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5). 
-This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
-iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
-
-An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
-To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
-The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
-In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue. 
-The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
-The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
-Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
-
-Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
-The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient. 
-The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
-Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
-
-Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
-
-It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
-We address possible interaction effects in the next section.
-
-\subsection*{Two mechanisms to address the genetic results: Cluster size and $r_2$ collapse}
-
-The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
-This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
-We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different. 
-We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
-We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
-The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
-Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
-
-It is known that EphA receptors form clusters, that these are necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
-It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
-Additionally, EphA4 receptors make `side attachments' to clusters of EphA receptors, possibly regulating the cluster size \citep{nikolov_ephephrin_2013}. 
-%
-We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$. 
-We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4}$, the density of EphA4 receptors:
-\begin{equation}
-    c_0 = 1/r_{\!\scriptscriptstyle A4}
-\end{equation}
-%
-Eq.\,\ref{e:G} was modified to include the cluster size:
-%
-\begin{equation}\label{e:Gcs}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-
-$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
-Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).  
-Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
-To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
-
-Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
-We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
-In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
-Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
-In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
-If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
-If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
-
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
-This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
-This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
-Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
-This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
-The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
-Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
-
-Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
-
-\section{Discussion}
-
-% Summarize the findings and place them in context
-Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
-If not, does the addition of competition resolve the remaining problems? 
-Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
-To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
-The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold. 
-These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals. 
-The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development. 
-Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
-Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
-
-However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by the model. 
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
-
-To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
-First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
-A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
-In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
-We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
-This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
-However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA. 
-A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
-We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
-We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
-
-\subsection*{Chemotaxis}
-
-The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-
-Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
-Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
-This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
-
-Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
-Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
-Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
-With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
-Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
-RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
-Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
-Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
-However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
-In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum. 
-The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}. 
-Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
-Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-
-We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
-This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-\subsection*{Competition}
-
-Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
-We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive.
-An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.
-
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
-If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
-
-Competition appears to be necessary for all of the surgical manipulations. 
-It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
-
-The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
-This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
-If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map. 
-If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}. 
-It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
-
-% Single axon behaviour
-\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
-The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
-The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
-
-\subsection*{Attractive receptor-ligand interactions}
-
-In the results, we present a model in which all receptor-ligand interactions are repulsive.
-However, EphB/ephrinB interactions are largely attractive in nature.
-It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
-When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
-Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
-
-\subsection*{Robustness to noise}
-
-To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
-Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
-We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
-The graphs show the expression values on a normalised scale. 
-We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
-The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
-Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
-i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
-ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
-iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
-(ii) and (iii) are computed once only, at the start of the simulation.
-Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined. 
-It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A). 
-The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern. 
-Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions. 
-Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
-
-%\subsection*{The value of $s$}
-%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Limitations/weaknesses of research
-\subsection*{Predictions and Limitations}
-
-We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth. 
-We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
-By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
-We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
-
-Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
-In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
-Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
-Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
-Adapting this framework to address an axon branching model is left for future work.
-
-% Restate significant findings and implications
-
-\section{Method details}
-
-\subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
-For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
-%
-\begin{equation}\label{e:B}
-B_x = \begin{cases}
-        \frac{1}{2} (r-x)     & x<r \\
-        \frac{1}{2} (1-r-x)    & x>1-r
-\end{cases}
-B_y = \begin{cases}
-        \frac{1}{2} (r-y)      & y<r \\
-        \frac{1}{2} (1-r-y)    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
-% end border effect description
-
-\subsection*{Initial conditions}
-Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation. 
-Each RGC axon $k$ was assigned a random initial position coordinate:
-\begin{equation}\label{e:ic}
-\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$.
-Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
-% end initial conditions
-
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
-
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
-The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
-Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
-The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
-%
-$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
-%
-where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
-
-A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
-The transformations are defined as:
-%
-$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
-%
-where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
-$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$. 
-The Studentized difference of the means of the resamples is computed:
-
-$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
-
-If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
-If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$. 
-We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL). 
-The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
-
-The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
-%% End bootstrap desc
-
-\section{Data Availability}
-
-All data and code required to reproduce the results in this work are available in the public repository 
-{https://github.com/sebjameswml/RetinoTectal}.
-The software was built using the library morphologica \citep{james_morphologica_2022}.
-
-\section{Acknowledgements}
-
-This work was supported by grant 220020516 from the James S. McDonnell Foundation.
-
-\bibliography{RetinoTectal}
-
-%
-% FIGURES
-%
-
-% FIGURE 1
-% Expand panel A to be able to lose B-D.
-% Full page width
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Schematic diagrams of retinotectal mapping and gene expression.
+\textbf{Figure 1.} Schematic diagrams of retinotectal mapping and gene expression.
 Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
@@
 %
-}
-\label{f:ex}
-\end{figure}
 
-% FIGURE 2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
-\caption{Receptor-ligand model of chemotaxis and competition. 
-\textbf{A} The origin position on the retina gives axons a colour according to a dual-colour map.
-Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
-\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina. 
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results (that is, correctly ordered into a perfect map of the simulated, square retina). 
-\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time. 
-\textbf{G} The arrangement of axon branches at t=1500. 
-\textbf{H} Selected axon locations at t=1500, with path histories in grey. 
-\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-\label{f:GJ}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 3 ./scripts/fig_GJ_surgical_grafts.svg (graphs generated with fig_GJ_surgical.sh)
-% Could split into retina and tectum or 
-% a: rotation/graft (as hand waved by Gierer)
-% b: ablation (as simulated in 1D by Gierer)
-% c: mismatch/compound (as extensions of Gierer)
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
-\caption{Performance of the receptor-ligand model with surgical graft manipulations.
-Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
-\textbf{A.} Graft rotation.
-An 8$\times$8 square patch of tectum has been rotated by 90$\degree$, altering the tectal ligand expression pattern.
-High gradients form around the new discontinuities at the edge of the patch (see Fig\,\ref{f:trot90}), driving axons to follow the rotation of the patch.
-The left graph shows the axon centroid positions after 1500 time steps, the smaller inset graph shows the (idealised) anticipated layout seen in experiments \citep{levine_deployment_1974}.
-The right-hand graph shows the time-course of the pattern quality metric, $\epsilon$ (i.e.~how closely the centroid pattern matches the inset) for the entire tectum (blue), the graft region (yellow) and the surround (red).
-The axons within the graft are slightly less well ordered than those outside.
-\textbf{B} Shows the same type of graft rotation as \textbf{A} for a 180$\degree$ rotation \citep{yoon_retention_1973}.
-\textbf{C.} A simulated graft swap experiment.
-Two 12$\times$4 rectangular regions of tectum are cut out and grafted back in swapped positions \citep{hope_arrow_1976,gaze_visuotectal_1983}.
-While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
-Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.}
-\label{f:GJsurg_grafts}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 4
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
-\caption{Retinal and tectal ablation simulations.
-\textbf{A} The temporal half of the retina is removed in this manipulation.
-Axons from the nasal half grow onto an unmodified tectum.
-This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
-In this work, it is the competition that achieves this effect.
-The axon centroid graphs are plotted as in Fig.\,\ref{f:GJsurg_grafts}, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
-\textbf{B} Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
-As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
-\textbf{C} A `mismatch' experiment \citep{horder_retention_1971}.
-The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
-As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
-\textbf{D} Compound eye.
-If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
-The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
-\label{f:GJsurg_ablate}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 5
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-In each manipulation, the left graph shows axon branch agents (colour coded by retinal origin) at the end of a simulation ($t=1500$), the centre graph shows corresponding axon centroids, with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
-In the right hand graph, the final tectal rostro-caudal locations of axons from each naso-temporal origin in the retina are plotted as violin plots (bars indicate the max/min datum in each group and shading indicates the data distribution).
-Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-\textbf{A} We simulated the heterozygous EphA3 knock-in (EphA3 ki/+) of \citet{brown_topographic_2000} by selecting half of the retinal cells at random, and in these cells adding the value 0.8 to $r_0$ (which, as its expression increases towards the temporal retina, is our analogue for EphA).
-The affected cells experienced altered interaction with tectal ligand gradients and were shifted rostrally.
-The rostral shift was largest for nasal cells bound for the caudal tectum.
-These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
-Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (thus the $r_0$-$L_0$ interaction is minimal).
-However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis, as seen in the inset graph.
-\textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
-Here, we observed a separation into two maps, one for the wildtype cells and one for the manipulated cells, in agreement with \citet{brown_topographic_2000} (inset is from their Fig.\,5C).
-\textbf{C} Simulation of EphA3 ki/+ selective knock-in and EphA4 knock-down (all cells) from \citet{reber_relative_2004}.
-The knock-down was applied by reducing $r_0$ in all cells by 1.2.
-The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in experiments.
-Instead, selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in.}
-\label{f:GJeph}
-\end{figure}
-
-% FIGURE 6
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
-\caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
-To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and for the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrinA ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA (i.e.~with $r_2$).
-\textbf{A} \emph{EphA/ephrinA expression.} As described in the text, EphA ($r_0 \equiv \sum_{i\neq 4}\mathrm{EphA}i$) is expressed according to $r_0 = 1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position.
-Knock-in is simulated by adding 0.8 to this function (giving the purple curve).
-Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
-A threshold, $h_0$, used by the $r_2$ collapse mechanism is shown as a red dotted line, with a green dash-dot line indicating the intersection of $\Sigma\mathrm{EphA}$ and $h_0$.
-\textbf{B} \emph{EphA4 expression.} Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA ($l_0$) can bind to EphA4 ($r_{A4}$), modelled here as a multiplicative interaction to give cis-bound EphA4, $r_{A4}^{\mathrm{cis}} = w_{A4} \times r_{A4} \times l_0$ (green dashed curve), where $w_{A4}$ is a binding weight described in the text.
-The remaining, un-bound, free EphA4 receptors, $r_{A4}^{\mathrm{free}} = r_{A4} - r_{A4}^{\mathrm{cis}}$ (green solid curve).
-We model genetic knock-down of EphA4 by subtracting 2.1 from $r_{A4}$, then recomputing $r_{A4}^{\mathrm{cis}}$ and $r_{A4}^{\mathrm{free}}$ to give the cyan curves.
-\textbf{C} \emph{Cluster size affects signal strength.} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}^{\mathrm{free}}$.
-Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced by combined knock-in and knock-down.
-This cluster size-modulated signal, $s_0$, is smaller in magnitude than the signal for $r_0$ in the original model (for which $s_0 \equiv r_0$).
-To compensate, $r_2$ is divided by 2, which gives the correct behaviour in a wildtype system.
-\textbf{D} \emph{$r_2$ collapse.} i) We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=2$ (red dotted line in A) and that $r_{A4}^{\mathrm{free}}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
-Under these conditions, $r_2$ is reduced to one fifth of the value it would otherwise take.
-ii) The regions of $r_2$ collapse are illustrated.
-No collapse occurs for wildtype EphA and EphA4 expressions; in cells with only EphA4 knock-down, $r_2$ collapse occurs for temporal cells (those more temporal than the dash-dot line in A).
-In cells with both EphA3 knock-in and EphA4 knock-down, the collapse mechanism occurs regardless of retinal origin.}
-\label{f:clustermech}
-\end{figure}
-
-% FIGURE 7
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
-\caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}. 
-\textbf{A} EphA3 knock-in result for ki=0.8.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment (inset).
-This is caused by the cluster size effect, in which the signal increase due to the knock-in is reduced for temporal RGCs compared with nasal cells (See Fig.\,\ref{f:clustermech}C, green/purple solid curves).
-%FOR TEXT: It is confirmed by an increased achieved significance level (ASL) from a bootstrapped, studentized t-test comparing the red and blue distributions at each N--T location. 
-\textbf{B} The cluster size effect does not change the double knock-in result, which shows excellent agreement with experiment (inset). 
-\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked down, cells on the red curve experience a combination of interactions: First, the increased interaction with ephrinA that gave the result in A is still in effect.
-Second, all cells on the red curve experience $r_2$ collapse, whereas only temporal RGCs on the blue curve experience this effect.
-$r_2$ receptors are more numerous on nasal RGCs and so the collapse in $r_2$ generates a substantial rostral movement of the red curve, in agreement with \citet{reber_relative_2004}.
-\textbf{D} Uniform knock-down of EphA4 in all cells (with no manipulation of EphA3) does not perturb the mapping to a significant extent, although the effect of the EphA4 knock-down on the cluster size mechanism causes a slight rostral shift in all cells.}
-\label{f:r2collapse}
-\end{figure}
-
-% ANY SUPPLEMENTARY FIGURES GO AFTER THIS
-%\renewcommand{\thefigure}{S\arabic{figure}}
-%\setcounter{figure}{0}
-
-% FIGURE S1
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
-graft-rotate 90 degrees surgical manipulations. 
-\textbf{A} Ligand expression. 
-The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps. 
-The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$. 
-Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the 
-The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively. 
-\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A. 
-For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis). 
-The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
-Full saturation indicates a positive gradient; white indicates a negative gradient.
-For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis. 
-Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
-\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
-\label{f:trot90}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
-\caption{(Supplementary) EphB and ephrin B gene expression. 
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature. 
-\textbf{A} Retinal expression of EphB2. 
-We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application. 
-The resulting, normalized expression is shown in the graph for the E4 data, with an inset showing corresponding data for the E6 data. 
-In the right-hand graph, we show an expansion of a portion of the E4 data to which we applied a linear fit to determine the RMS deviation from the mean to estimate the noise in the receptor signal. 
-The mean RMS error was 0.06. 
-\textbf{B} Stalefish screenshots and expression graphs for retinal ephrinB1 and ephrinB2 expression, using data from Fig.\,1 of \citet{braisted_graded_1997}. 
-\textbf{C} Screenshots and expression graphs for tectal ephrinB1 expression, using data from Fig.\,4 of \citet{braisted_graded_1997}.}
-\label{f:expr_ephb}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S3
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
-\caption{(Supplementary) Response of the model to noise.
-\textbf{A} Adding 3\% gradient measurement noise (normally distributed, standard deviation, $\sigma = 0.19$) has negligible effect on the pattern. 
-\textbf{B} 3\% retinal receptor noise ($\sigma = 0.08$). 
-\textbf{C} 3\% tectal ligand noise ($\sigma = 0.08$). 
-\textbf{D} Combined effect of retinal receptor, tectal ligand and gradient measurement noise is dominated by the tectal ligand noise.}
-\label{f:noise}
-\end{fullwidth}
-\end{figure}
+\bibliography{RetinoTectal}
 

### 2024-07-04 paper/neural_2024/poster_text/intro.tex
Text for poster

--- 2024-07-04 paper/neural_2024/poster_text/abs.tex
@@
 
-The axons of retinal ganglion cells form an ordered map of the retina
-in the optic tectum. Surgical and genetic experiments have revealed
-several key mechanisms that contribute to retinotectal map
-development, and computational modelling studies have demonstrated how
-maps can be shaped by their interactions. Models derived from Sperry's
-classic chemoaffinity theory can reproduce retinotopy, and Gierer
-later showed how an extension to include competition between growing
-axons could also account for a subset of the disordered maps that have
-resulted from surgical manipulation. Further extensions have each
-accounted for additional datapoints, but none has been able to explain
-the full range of maps observed experimentally. Here we present a
-unified model of retinotectal development, assuming only local
-interactions between growing axons, which can account for the full
-range of maps resulting from surgical manipulation. This model extends
-the `potential functions' of Gierer's original to describe
-self-organisation across a 2D cortical sheet, and it represents a
-coordination of axonal growth by orthogonal pairs of opposing ligand
-gradients in the tectum and corresponding receptor gradients in the
-retina. In addition, we present progress towards accounting for the
-full range of maps that have resulted from genetic manipulation, by
-incorporating a mechanistic description of receptor ligand
+\section{Introduction}
+The axons of retinal ganglion cells form an ordered map of the retina in the
+optic tectum. Surgical and genetic experiments have revealed several key
+mechanisms that contribute to retinotectal map development, and computational
+modelling studies have demonstrated how maps can be shaped by their
+interactions. Models derived from Sperry's classic chemoaffinity theory can
+reproduce retinotopy, and Gierer later showed how an extension to include
+competition between growing axons could also account for a subset of the
+disordered maps that have resulted from surgical manipulation. Further
+extensions have each accounted for additional datapoints, but none has been
+able to explain the full range of maps observed experimentally. Here we
+present a unified model of retinotectal development, assuming only local
+interactions between growing axons, which can account for the full range of
+maps resulting from surgical manipulation. This model extends the `potential
+functions' of Gierer's original to describe self-organisation across a 2D
+cortical sheet, and it represents a coordination of axonal growth by
+orthogonal pairs of opposing ligand gradients in the tectum and corresponding
+receptor gradients in the retina. In addition, we present progress towards
+accounting for the full range of maps that have resulted from genetic
+manipulation, by incorporating a mechanistic description of receptor ligand
 interactions.
 
-%\bibliography{RetinoTectal}
+One important class of theories assumes that map formation is driven by
+chemotaxis, influenced by pairs of anti-parallel expression gradients in the
+tectum, with the balance of influences in each axon determined by a receptor
+density specific to its origin in the retina
+(\citealp{Gierer1983,Gierer1987,Simpson2011}; see also
+\citealp{Karbowski2004,James2020}). A second class of theories assumes that a
+single gradient is primarily responsible for chemotaxis and that maps reflect
+additional processes such as competition \citep{Triplett2011}, Hebbian
+activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction
+\citep{Prestige1975,Willshaw2006}.
+
+A key test of the validity of recent models is their ability to account for
+data from a set of genetic manipulation studies by
+\cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons
+from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a
+separate, rostrally shifted topographic map, which becomes indistinct from
+that formed by untreated cells, at a `collapse point' on the rostral
+side. \cite{Sterratt2013} found that a mixture of counter-gradients and an
+additional competition term provided the best account of these data. A
+follow-up study in which a range of different computational models were
+systematically compared \citep{hjorth_quantitative_2015} found that one
+focused on competition by Koulakov \citep{Triplett2011} could best account for
+several such experimental findings, and that the other types of model could
+not explain the collapse point.
+
+The explanatory power of Koulakov's `switching' models stems from their
+simplicity and degree of abstraction, revealing the map pattern that optimises
+the constraints represented by a given energy function. However, while
+switching models are evaluated through a numerical procedure (i.e., iterative
+switching), that procedure should not necessarily be interpreted as an
+explicit representation of the process of development
+\citep{Wilson2015}.
+
+We turned back to mechanistic models, taking as our starting point the model of
+\cite{Simpson2011}, which incorporates chemotaxis, competitive interactions
+and receptor-ligand interactions. This model used a global supervisor to
+provide the chemotactic component of axon movement. We set about determining
+whether we could re-cast the model into one in which simulated axon growth is
+guided by only local information. If such a model could account for the full
+range of experimental results, it would represent the most complete account of
+retinotectal map self-organization to date.
+
+\bibliography{RetinoTectal}
 

### 2024-07-04 paper/neural_2024/poster_text/results.tex
Text for poster

--- 2023-09-15 paper/main.tex
@@
-\documentclass[9pt,lineno,draft]{elife}
+\documentclass[9pt]{elife} % no lineno or draft
 % Nice citations
@@
 
-\title {
-Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
-}
+\title {none}
 
@@
 %\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
 
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
-There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
-Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
-However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
-Sperry's chemoaffinity theory provides an account of the wildtype pattern.
-Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
-Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
-
-\section{Introduction}
-
-Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
-Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
-Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
-
-One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
-A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
-
-A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
-\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that \redcmnt{one focused on competition
-% Stuart: I think the model is Koulakov's, not Triplett's.
-In Triplett2011 they give evidence for a competition model (like Koulakov's) by reducing the number of RGCs in the Math5 mutant mouse.
-*Hjorth* shows that the Koulakov model is the best one at reproducing the collapse in the EphA3 ki experiment.
-by \cite{Triplett2011}} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
-
-\redcmnt{The model of \cite{Triplett2011} describes maps as minima of an energy function}, whose structure can be obtained via a stochastic optimisation process.
-Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
-Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
-
-The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
-According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
-As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
-Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
-The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
-
-These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
-Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
-
-The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
-However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
-Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
-
-In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
-For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
-
-Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
-
-The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
-However, a major limitation is that chemotaxis is not implemented explicitly.
-Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
-This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
-
-An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
-If so, such a model would represent the most complete account of retinotectal map self-organization to date.
-
-** Maybe we also need to say that a second objective to to express the model
-exclusively in terms of receptor-ligand interactions.
-
-%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-
-%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
-The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-
-%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-
-%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-
-%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
-This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
-
-%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
-% And sounds?
-
-%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
-Two classes of ephrins have been identified, labelled `A' and `B'.
-Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
-Ephrin B acts on medial-lateral ordering with an attractive interaction.
-Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
-
-%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
-The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
-Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
-Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
-He presented gradient following as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-This model forms a mathematical expression of Sperry's theory.
-Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
-ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-
-%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
-He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
-
-%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
-
-%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
-In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
-Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
-With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
-ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
-and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
-
-\section{Results}
+\section{A receptor-ligand model of chemotaxis}
 
@@
 $\mathbf{C}$ is the movement due to competitive interactions with other branches
-%
 and $\mathbf{G}$ is the movement vector given rise to by the chemotaxis effect.
 
-%% * To represent the mechanism for comp. explicitly, we introduced receptor AND ligand expressions
-
-%%We reformulated this so that the distance based repulsion is also mediated via receptor ligand interactions. % No need to justify yet - can go in disucssion
-%
-
-%%%%%%%% Remove/coment out %%%%%%%%%%%%%%%%%
-% No need for this rhetorical stuff
-%\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-
-%Their model, which successfully accounts for a wide range of experimental
-%manipulations, foundxxx *Their model had a* $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-% Rewrite next little bit
-% Say the global supervisor comes into the expression G, but having G a vector
-% that always points towards...
-% This may come later
-
-%We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
-%%%%%%%%%%%%%%%%%%%%%%%
-
 %  Maybe lose the sub sub sections
-\subsubsection*{A signalling model for competition}
+\subsection*{A signalling model for competition}
 
 Simpson \& Goodhill's model had a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-%Although the model incorporated EphA expression, the distance-based competitive interaction was given no mechanism of operation.
 To provide an explicit mechanism of operation for the competition, we modelled it as a sum of interactions that are `switched on' via receptor-ligand signalling, meaning that some axons would have no competitive interaction, while others would repel strongly.
@@
 
-\subsubsection*{A gradient based model of chemotaxis}
-
-% Say the global supervisor comes into the expression G, but having G a vector
-% that always points towards...
-% This may come later
-
-\citet{simpson_simple_2011} chose a placeholder mechanism for the chemotaxis effect, using a global supervisor to set $\mathbf{G}$ to a vector which would always point to the predetermined target location for each branch.
-In order to make this a local model we replaced the global supervisor with an interaction between RGC receptors and ligands expressed on the tectum.
-We used forms for the ligand expression, $L(\mathbf{x})$, which matched those used for the RGC receptor expression.
-
-We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
-
-\begin{equation}\label{e:G}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
-$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
-%
-We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
-%
-$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
-
-Each of the $M$ RGC growth cones carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-%
-%The expression of each receptor varies only with respect to one spatial dimension.
-%
-We defined 4 receptor expression gradients (each varying only wrt one spatial dimension) arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-
-% NO NEED FOR This sentence/bit (keep comment in case for discussion)
-%Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-%
-% Already said this:
-%We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
-
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
-
-%While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-%%%%%%%%%%%%%%%%%
-
-\subsubsection*{Simulation}
-
-The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-The parameters found were $m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$ (see method details).
-
-% Results proper here Leave this as is for now
-\subsection*{Retinotectal map self-organization}
-
-Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps.
-Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red.
-Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
-Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
-The axons retain the colour of their originating cell on the retina.
-Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints.
-Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
-In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation.
-The number of line crossings in this plot gives an indication of the disorder in the arrangement.
-Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
-By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
-Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
-Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
-For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E.
-Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.
-
-% Stuart to update the text here
-\subsection*{Surgical manipulations}
-
-After finding an optimum set of parameters for the wildtype arrangement, we
-tested the model with a number of simulations of surgical manipulations
-%(for reviews, see \citet{udin_formation_1988} and
-%\citet{goodhill_retinotectal_1999})
-.
-Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
-%
-To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column).
-After the manipulation, we re-computed the tectal ligand expression gradients.
-Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region.
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B.
-We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red).
-The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04).
-The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
-The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
-%
-Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped.
-There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-
-Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
-According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}).
-The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
-%
-Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
-Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
-Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
-%
-A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
-%
-The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}.
-Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
-
-% Stuart to work on...
-\subsection*{Genetic manipulations}
-
-% Context that might need to go in the Discussion
-%In the early 2000s, a series of genetic experiments added to the set of
-%manipulations that a model of the retinotectal system would need to explain.
-
-% Modify following so that the discursive stuff goes in the discussion.
-First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum.
-For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours.
-By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map.
-In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5).
-This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
-iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
-
-An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each sub-type to the overall level is assumed not to be important.
-In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki).
-To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
-The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
-In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue.
-The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
-The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
-Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
-
-Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
-The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient.
-The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
-Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
-
-Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$).
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
-
-It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged.
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
-We address possible interaction effects in the next section.
-
-\subsection*{Two mechanisms to address the genetic results: Cluster size and $r_2$ collapse}
-
-The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
-This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
-We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different.
-We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
-We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
-The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
-Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
-
-It is known that EphA receptors form clusters, that these are necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
-It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
-Additionally, EphA4 receptors make `side attachments' to clusters of EphA receptors, possibly regulating the cluster size \citep{nikolov_ephephrin_2013}.
-%
-We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$.
-We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4}$, the density of EphA4 receptors:
-\begin{equation}
-    c_0 = 1/r_{\!\scriptscriptstyle A4}
-\end{equation}
-%
-Eq.\,\ref{e:G} was modified to include the cluster size:
-%
-\begin{equation}\label{e:Gcs}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-
-$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
-Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).
-Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
-To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
-
-Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
-We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
-In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
-Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
-In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
-If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
-If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
-
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
-This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
-This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
-Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
-This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
-The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
-Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
-
-Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
-
-\section{Discussion}
-
-% Summarize the findings and place them in context
-Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
-If not, does the addition of competition resolve the remaining problems?
-Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
-To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
-The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold.
-These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals.
-The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development.
-Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate.
-Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
-
-However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by the model.
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
-
-To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
-First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
-A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
-In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
-We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
-This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
-However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA.
-A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
-We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
-We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
-
-\subsubsection*{Outtakes}
-
-Previously in results:
-
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-
-\subsection*{Chemotaxis}
-
-The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-
-Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
-Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
-This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
-
-Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
-Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
-Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
-With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
-Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
-RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
-Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
-Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
-However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
-In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum.
-The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}.
-Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
-Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-
-We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen.
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
-This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here.
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-\subsection*{Competition}
-
-Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
-We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive.
-An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.
-
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
-If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
-
-Competition appears to be necessary for all of the surgical manipulations.
-It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
-
-The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
-This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
-If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map.
-If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}.
-It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
-
-% Single axon behaviour
-\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
-The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
-The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
-
-\subsection*{Attractive receptor-ligand interactions}
-
-In the results, we present a model in which all receptor-ligand interactions are repulsive.
-However, EphB/ephrinB interactions are largely attractive in nature.
-It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
-When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
-Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
-
-\subsection*{Robustness to noise}
-
-To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
-Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
-We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
-The graphs show the expression values on a normalised scale.
-We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
-The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
-Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
-i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
-ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
-iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
-(ii) and (iii) are computed once only, at the start of the simulation.
-Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined.
-It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A).
-The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern.
-Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions.
-Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
-
-%\subsection*{The value of $s$}
-%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Limitations/weaknesses of research
-\subsection*{Predictions and Limitations}
-
-We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth.
-We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
-By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
-We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
-
-Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
-In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
-Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
-Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
-Adapting this framework to address an axon branching model is left for future work.
-
-% Restate significant findings and implications
-
-\section{Method details}
-
-\subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
-For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
-%
-\begin{equation}\label{e:B}
-B_x = \begin{cases}
-        \frac{1}{2} (r-x)     & x<r \\
-        \frac{1}{2} (1-r-x)    & x>1-r
-\end{cases}
-B_y = \begin{cases}
-        \frac{1}{2} (r-y)      & y<r \\
-        \frac{1}{2} (1-r-y)    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
-% end border effect description
-
-\subsection*{Simulation domain}
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}.
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-
-% Make sure this ain't a copy of results
-Each retinal ganglion cell (RGC) projected $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-
-The expression of each receptor was varied only with respect to one spatial dimension.
-
-% Thi sis in the right place here
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-
-\subsection*{Initial conditions}
-Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation.
-Each RGC axon $k$ was assigned a random initial position coordinate:
-\begin{equation}\label{e:ic}
-\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$.
-Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
-% end initial conditions
-
-\subsection*{Receptor and Ligand expression forms}
-
-A section to add more detail on the functions used, if necessary.
-
-\subsection*{Parameter finding}
-
-We performed a simulated annealing optimization to find three free parameters; $m_{\!_X}$, $r_{\!_X}$ and $s$ which would balance a chosen evel of chemotaxis strength given by $m_{\!_G}=0.00120$.
-We minimized an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other.
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
-
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
-
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
-The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
-Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
-The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
-%
-$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
-%
-where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
-
-A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
-The transformations are defined as:
-%
-$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
-%
-where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
-$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$.
-The Studentized difference of the means of the resamples is computed:
-
-$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
-
-If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
-If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$.
-We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL).
-The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
-
-The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
-%% End bootstrap desc
-
-\section{Data Availability}
-
-All data and code required to reproduce the results in this work are available in the public repository
-{https://github.com/sebjameswml/RetinoTectal}.
-The software was built using the library morphologica \citep{james_morphologica_2022}.
-
-\section{Acknowledgements}
-
-This work was supported by grant 220020516 from the James S. McDonnell Foundation.
-
 \bibliography{RetinoTectal}
 
-%
-% FIGURES
-%
-
-% FIGURE 1
-% Expand panel A to be able to lose B-D.
-% Full page width
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Schematic diagrams of retinotectal mapping and gene expression.
-Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
-%
-\textbf{A} Tectal ligand expression $L_i$.
-Tectal ligand expression patterns are assumed to increase exponentially in one direction only across the tectal surface.
-$L_0$ (or ephrin A) increases in a caudal to rostral direction (i.e.
-with increasing $y$; $x$ and $y$ directions are indicated with arrows).
-$L_2$ increases in the opposing sense, with $-y$.
-$L_1$ and $L_3$ form counter-gradients in the $\pm x$ directions.
-\emph{L: lateral, M: medial, C: caudal, R: rostral}.
-%
-\textbf{B} Retinal receptor and ligand expressions.
-Four receptors, $r_i$, are expressed across the retina.
-$r_0$ increases as an exponential function of $-x$ ; $r_1$ increases with an exponential function of $-y$.
-$r_2$ and $r_3$ have the opposite sense and increase with positive $x$ and $y$, respectively.
-Retinal ligand expression, $l_i$, is expressed in two complimentary gradients \citep{hornberger_modulation_1999}.
- \emph{N: nasal, T: temporal, D: dorsal, V: ventral}.
-%
-}
-\label{f:ex}
-\end{figure}
-
-% FIGURE 2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
-\caption{Receptor-ligand model of chemotaxis and competition.
-\textbf{A} The origin position on the retina gives axons a colour according to a dual-colour map.
-Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
-\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina.
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results (that is, correctly ordered into a perfect map of the simulated, square retina).
-\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time.
-\textbf{G} The arrangement of axon branches at t=1500.
-\textbf{H} Selected axon locations at t=1500, with path histories in grey.
-\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-\label{f:GJ}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 3 ./scripts/fig_GJ_surgical_grafts.svg (graphs generated with fig_GJ_surgical.sh)
-% Could split into retina and tectum or
-% a: rotation/graft (as hand waved by Gierer)
-% b: ablation (as simulated in 1D by Gierer)
-% c: mismatch/compound (as extensions of Gierer)
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
-\caption{Performance of the receptor-ligand model with surgical graft manipulations.
-Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
-\textbf{A.} Graft rotation.
-An 8$\times$8 square patch of tectum has been rotated by 90$\degree$, altering the tectal ligand expression pattern.
-High gradients form around the new discontinuities at the edge of the patch (see Fig\,\ref{f:trot90}), driving axons to follow the rotation of the patch.
-The left graph shows the axon centroid positions after 1500 time steps, the smaller inset graph shows the (idealised) anticipated layout seen in experiments \citep{levine_deployment_1974}.
-The right-hand graph shows the time-course of the pattern quality metric, $\epsilon$ (i.e.~how closely the centroid pattern matches the inset) for the entire tectum (blue), the graft region (yellow) and the surround (red).
-The axons within the graft are slightly less well ordered than those outside.
-\textbf{B} Shows the same type of graft rotation as \textbf{A} for a 180$\degree$ rotation \citep{yoon_retention_1973}.
-\textbf{C.} A simulated graft swap experiment.
-Two 12$\times$4 rectangular regions of tectum are cut out and grafted back in swapped positions \citep{hope_arrow_1976,gaze_visuotectal_1983}.
-While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
-Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.}
-\label{f:GJsurg_grafts}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 4
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
-\caption{Retinal and tectal ablation simulations.
-\textbf{A} The temporal half of the retina is removed in this manipulation.
-Axons from the nasal half grow onto an unmodified tectum.
-This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
-In this work, it is the competition that achieves this effect.
-The axon centroid graphs are plotted as in Fig.\,\ref{f:GJsurg_grafts}, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
-\textbf{B} Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
-As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
-\textbf{C} A `mismatch' experiment \citep{horder_retention_1971}.
-The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
-As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
-\textbf{D} Compound eye.
-If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
-The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
-\label{f:GJsurg_ablate}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 5
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-In each manipulation, the left graph shows axon branch agents (colour coded by retinal origin) at the end of a simulation ($t=1500$), the centre graph shows corresponding axon centroids, with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
-In the right hand graph, the final tectal rostro-caudal locations of axons from each naso-temporal origin in the retina are plotted as violin plots (bars indicate the max/min datum in each group and shading indicates the data distribution).
-Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-\textbf{A} We simulated the heterozygous EphA3 knock-in (EphA3 ki/+) of \citet{brown_topographic_2000} by selecting half of the retinal cells at random, and in these cells adding the value 0.8 to $r_0$ (which, as its expression increases towards the temporal retina, is our analogue for EphA).
-The affected cells experienced altered interaction with tectal ligand gradients and were shifted rostrally.
-The rostral shift was largest for nasal cells bound for the caudal tectum.
-These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
-Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (thus the $r_0$-$L_0$ interaction is minimal).
-However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis, as seen in the inset graph.
-\textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
-Here, we observed a separation into two maps, one for the wildtype cells and one for the manipulated cells, in agreement with \citet{brown_topographic_2000} (inset is from their Fig.\,5C).
-\textbf{C} Simulation of EphA3 ki/+ selective knock-in and EphA4 knock-down (all cells) from \citet{reber_relative_2004}.
-The knock-down was applied by reducing $r_0$ in all cells by 1.2.
-The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in experiments.
-Instead, selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in.}
-\label{f:GJeph}
-\end{figure}
-
-% FIGURE 6
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
-%% Not two mechanisms as the story. Story is that Fig 5 is ok, but not ace,
-%% and we have the flexibility to propose mechanisms that fix.
-\caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
-To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and for the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrinA ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA (i.e.~with $r_2$).
-\textbf{A} \emph{EphA/ephrinA expression.} As described in the text, EphA ($r_0 \equiv \sum_{i\neq 4}\mathrm{EphA}i$) is expressed according to $r_0 = 1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position.
-Knock-in is simulated by adding 0.8 to this function (giving the purple curve).
-Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
-A threshold, $h_0$, used by the $r_2$ collapse mechanism is shown as a red dotted line, with a green dash-dot line indicating the intersection of $\Sigma\mathrm{EphA}$ and $h_0$.
-\textbf{B} \emph{EphA4 expression.} Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA ($l_0$) can bind to EphA4 ($r_{A4}$), modelled here as a multiplicative interaction to give cis-bound EphA4, $r_{A4}^{\mathrm{cis}} = w_{A4} \times r_{A4} \times l_0$ (green dashed curve), where $w_{A4}$ is a binding weight described in the text.
-The remaining, un-bound, free EphA4 receptors, $r_{A4}^{\mathrm{free}} = r_{A4} - r_{A4}^{\mathrm{cis}}$ (green solid curve).
-We model genetic knock-down of EphA4 by subtracting 2.1 from $r_{A4}$, then recomputing $r_{A4}^{\mathrm{cis}}$ and $r_{A4}^{\mathrm{free}}$ to give the cyan curves.
-\textbf{C} \emph{Cluster size affects signal strength.} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}^{\mathrm{free}}$.
-Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced by combined knock-in and knock-down.
-This cluster size-modulated signal, $s_0$, is smaller in magnitude than the signal for $r_0$ in the original model (for which $s_0 \equiv r_0$).
-To compensate, $r_2$ is divided by 2, which gives the correct behaviour in a wildtype system.
-\textbf{D} \emph{$r_2$ collapse.} i) We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=2$ (red dotted line in A) and that $r_{A4}^{\mathrm{free}}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
-Under these conditions, $r_2$ is reduced to one fifth of the value it would otherwise take.
-ii) The regions of $r_2$ collapse are illustrated.
-No collapse occurs for wildtype EphA and EphA4 expressions; in cells with only EphA4 knock-down, $r_2$ collapse occurs for temporal cells (those more temporal than the dash-dot line in A).
-In cells with both EphA3 knock-in and EphA4 knock-down, the collapse mechanism occurs regardless of retinal origin.}
-\label{f:clustermech}
-\end{figure}
-
-% FIGURE 7
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
-\caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}.
-\textbf{A} EphA3 knock-in result for ki=0.8.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment (inset).
-This is caused by the cluster size effect, in which the signal increase due to the knock-in is reduced for temporal RGCs compared with nasal cells (See Fig.\,\ref{f:clustermech}C, green/purple solid curves).
-%FOR TEXT: It is confirmed by an increased achieved significance level (ASL) from a bootstrapped, studentized t-test comparing the red and blue distributions at each N--T location.
-\textbf{B} The cluster size effect does not change the double knock-in result, which shows excellent agreement with experiment (inset).
-\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked down, cells on the red curve experience a combination of interactions: First, the increased interaction with ephrinA that gave the result in A is still in effect.
-Second, all cells on the red curve experience $r_2$ collapse, whereas only temporal RGCs on the blue curve experience this effect.
-$r_2$ receptors are more numerous on nasal RGCs and so the collapse in $r_2$ generates a substantial rostral movement of the red curve, in agreement with \citet{reber_relative_2004}.
-\textbf{D} Uniform knock-down of EphA4 in all cells (with no manipulation of EphA3) does not perturb the mapping to a significant extent, although the effect of the EphA4 knock-down on the cluster size mechanism causes a slight rostral shift in all cells.}
-\label{f:r2collapse}
-\end{figure}
-
-% ANY SUPPLEMENTARY FIGURES GO AFTER THIS
-%\renewcommand{\thefigure}{S\arabic{figure}}
-%\setcounter{figure}{0}
-
-% FIGURE S1
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
-graft-rotate 90 degrees surgical manipulations.
-\textbf{A} Ligand expression.
-The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps.
-The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$.
-Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the
-The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively.
-\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A.
-For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis).
-The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
-Full saturation indicates a positive gradient; white indicates a negative gradient.
-For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis.
-Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
-\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
-\label{f:trot90}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
-\caption{(Supplementary) EphB and ephrin B gene expression.
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature.
-\textbf{A} Retinal expression of EphB2.
-We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application.
-The resulting, normalized expression is shown in the graph for the E4 data, with an inset showing corresponding data for the E6 data.
-In the right-hand graph, we show an expansion of a portion of the E4 data to which we applied a linear fit to determine the RMS deviation from the mean to estimate the noise in the receptor signal.
-The mean RMS error was 0.06.
-\textbf{B} Stalefish screenshots and expression graphs for retinal ephrinB1 and ephrinB2 expression, using data from Fig.\,1 of \citet{braisted_graded_1997}.
-\textbf{C} Screenshots and expression graphs for tectal ephrinB1 expression, using data from Fig.\,4 of \citet{braisted_graded_1997}.}
-\label{f:expr_ephb}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S3
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
-\caption{(Supplementary) Response of the model to noise.
-\textbf{A} Adding 3\% gradient measurement noise (normally distributed, standard deviation, $\sigma = 0.19$) has negligible effect on the pattern.
-\textbf{B} 3\% retinal receptor noise ($\sigma = 0.08$).
-\textbf{C} 3\% tectal ligand noise ($\sigma = 0.08$).
-\textbf{D} Combined effect of retinal receptor, tectal ligand and gradient measurement noise is dominated by the tectal ligand noise.}
-\label{f:noise}
-\end{fullwidth}
-\end{figure}
-
 \end{document}

### 2024-07-04 paper/neural_2024/poster_text/results2.tex
Text for poster

--- 2023-08-25 paper/main.tex
@@
 
-\title {
-Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
-}
+\title {none}
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
+\author[1]{Sebastian~S.~James}
+\author[1*]{Stuart~P.~Wilson}
 \affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
+\corr{s.wilson@sheffield.ac.uk}{SW}
 %% END OF PREAMBLE
 
-% Seb to fix: Expected not experiment (in figs), Fig 8 - monochromes. Fig 1 - monochromes
-% Stuart to do: Intro/Start of results modifications
-
 \begin{document}
-%\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
-
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
-There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
-Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
-However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
-Sperry's chemoaffinity theory provides an account of the wildtype pattern.
-Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
-Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
-
-\section{Introduction}
-
-Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
-Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
-Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
-
-One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
-A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
-
-A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
-\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
-
-The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
-Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
-Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
-
-The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
-According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
-As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
-Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
-The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
-
-These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
-Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
-
-The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
-However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
-Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
-
-In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
-For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
-
-Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
-
-The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
-However, a major limitation is that chemotaxis is not implemented explicitly.
-Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
-This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
-
-An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
-If so, such a model would represent the most complete account of retinotectal map self-organization to date.
-
-%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-
-%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
-The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-
-%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-
-%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-
-%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
-This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
-
-%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
-% And sounds?
-
-%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
-Two classes of ephrins have been identified, labelled `A' and `B'.
-Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
-Ephrin B acts on medial-lateral ordering with an attractive interaction.
-Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
-
-%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
-The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
-Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
-Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
-He presented gradient following as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-This model forms a mathematical expression of Sperry's theory.
-Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
-ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-
-%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
-He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
-
-%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
-
-%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
-In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
-Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
-With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
-ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
-and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
-
-\section{Results}
-\subsection*{A model of chemotaxis and competition}
-
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
-\begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
-\end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
-
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
 
 \subsection*{A gradient based model of chemotaxis}
+
+\citet{simpson_simple_2011} chose a placeholder mechanism for the chemotaxis effect, using a global supervisor to set $\mathbf{G}$ to a vector which would always point to the predetermined target location for each branch.
+In order to make this a local model we replaced the global supervisor with an interaction between RGC receptors and ligands expressed on the tectum.
+We used forms for the ligand expression, $L(\mathbf{x})$, which matched those used for the RGC receptor expression.
 
@@
 
-%%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
-The expression of each receptor varies only with respect to one spatial dimension.
+Each of the $M$ RGC growth cones carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
 %
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
+We defined 4 receptor expression gradients (each varying only wrt one spatial dimension) arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
+$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,1B).
 Thus the model is of \emph{differential chemotaxis}.
-%
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
 
 Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
-
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-
-\subsection*{A signalling model for competition}
-
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
-%
-\begin{equation} \label{e:X}
-\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
-\end{equation}
-%
-$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
-%
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
-%
-\begin{equation}
-Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
-                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
-                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_X} \\
-                 1 & \mathrm{otherwise.}
-     \end{cases}
-\end{equation}
-%
-$l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$. 
-$F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
-$s$ is a positive signal threshold for interaction.
-With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
-$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.
-
-\subsection*{Simulation}
-
-The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other. 
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
-
-Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps. 
-Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red.
-Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
-Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
-The axons retain the colour of their originating cell on the retina.
-Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
-Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
-In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation. 
-The number of line crossings in this plot gives an indication of the disorder in the arrangement.
-Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
-By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
-Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
-Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
-For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E. 
-Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.
-
-\subsection*{Surgical manipulations}
-
-After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
-Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
-%
-To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column). 
-After the manipulation, we re-computed the tectal ligand expression gradients.
-Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region. 
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
-We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red). 
-The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04). 
-The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
-The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
-%
-Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. 
-There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-
-Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
-According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}). 
-The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
-%
-Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
-Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
-Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
-%
-A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
-%
-The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}.
-Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
-
-\subsection*{Genetic manipulations}
-
-In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
-explain. 
-First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum. 
-For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours. 
-By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
-In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5). 
-This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
-iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
-
-An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
-To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
-The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
-In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue. 
-The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
-The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
-Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
-
-Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
-The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient. 
-The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
-Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
-
-Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
-
-It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
-We address possible interaction effects in the next section.
-
-\subsection*{Two mechanisms to address the genetic results: Cluster size and $r_2$ collapse}
-
-The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
-This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
-We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different. 
-We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
-We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
-The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
-Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
-
-It is known that EphA receptors form clusters, that these are necessary for ligand attachments to induce signal transmission and that cluster size affects signal size \citep{nikolov_ephephrin_2013}.
-It is also known that that some receptor sub-types may compete for ligand attachment \citep{fiore_regulation_2019}.
-Additionally, EphA4 receptors make `side attachments' to clusters of EphA receptors, possibly regulating the cluster size \citep{nikolov_ephephrin_2013}. 
-%
-We focused on the idea that signal strength is related to receptor cluster size and defined $c_i$, the effective cluster size for receptor $i$. 
-We set $c_i=1$ for $i=1$, $2$ or $3$ and investigated a simple relationship between EphA cluster size, $c_0$, and $r_{\!\scriptscriptstyle A4}$, the density of EphA4 receptors:
-\begin{equation}
-    c_0 = 1/r_{\!\scriptscriptstyle A4}
-\end{equation}
-%
-Eq.\,\ref{e:G} was modified to include the cluster size:
-%
-\begin{equation}\label{e:Gcs}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-
-$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
-Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).  
-Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
-To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
-
-Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
-We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
-In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
-Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
-In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
-If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
-If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
-
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
-This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
-This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
-Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
-This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
-The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
-Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
-
-Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
-
-\section{Discussion}
-
-% Summarize the findings and place them in context
-Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
-If not, does the addition of competition resolve the remaining problems? 
-Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
-To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
-The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold. 
-These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals. 
-The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development. 
-Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
-Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
-
-However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by the model. 
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
-
-To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
-First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
-A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
-In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
-We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
-This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
-However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA. 
-A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
-We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
-We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
-
-\subsection*{Chemotaxis}
-
-The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-
-Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
-Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
-This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
-
-Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
-Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
-Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
-With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
-Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
-RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
-Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
-Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
-However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
-In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum. 
-The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}. 
-Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
-Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-
-We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
-This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-\subsection*{Competition}
-
-Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
-We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive.
-An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.
-
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
-If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
-
-Competition appears to be necessary for all of the surgical manipulations. 
-It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
-
-The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
-This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
-If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map. 
-If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}. 
-It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
-
-% Single axon behaviour
-\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
-The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
-The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
-
-\subsection*{Attractive receptor-ligand interactions}
-
-In the results, we present a model in which all receptor-ligand interactions are repulsive.
-However, EphB/ephrinB interactions are largely attractive in nature.
-It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
-When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
-Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
-
-\subsection*{Robustness to noise}
-
-To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
-Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
-We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
-The graphs show the expression values on a normalised scale. 
-We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
-The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
-Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
-i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
-ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
-iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
-(ii) and (iii) are computed once only, at the start of the simulation.
-Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined. 
-It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A). 
-The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern. 
-Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions. 
-Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
-
-%\subsection*{The value of $s$}
-%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Limitations/weaknesses of research
-\subsection*{Predictions and Limitations}
-
-We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth. 
-We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
-By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
-We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
-
-Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
-In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
-Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
-Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
-Adapting this framework to address an axon branching model is left for future work.
-
-% Restate significant findings and implications
-
-\section{Method details}
-
-\subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
-For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
-%
-\begin{equation}\label{e:B}
-B_x = \begin{cases}
-        \frac{1}{2} (r-x)     & x<r \\
-        \frac{1}{2} (1-r-x)    & x>1-r
-\end{cases}
-B_y = \begin{cases}
-        \frac{1}{2} (r-y)      & y<r \\
-        \frac{1}{2} (1-r-y)    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
-% end border effect description
-
-\subsection*{Initial conditions}
-Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation. 
-Each RGC axon $k$ was assigned a random initial position coordinate:
-\begin{equation}\label{e:ic}
-\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$.
-Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
-% end initial conditions
-
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
-
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
-The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
-Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
-The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
-%
-$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
-%
-where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
-
-A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
-The transformations are defined as:
-%
-$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
-%
-where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
-$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$. 
-The Studentized difference of the means of the resamples is computed:
-
-$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
-
-If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
-If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$. 
-We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL). 
-The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
-
-The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
-%% End bootstrap desc
-
-\section{Data Availability}
-
-All data and code required to reproduce the results in this work are available in the public repository 
-{https://github.com/sebjameswml/RetinoTectal}.
-The software was built using the library morphologica \citep{james_morphologica_2022}.
-
-\section{Acknowledgements}
-
-This work was supported by grant 220020516 from the James S. McDonnell Foundation.
+Although several studies model tectal ligand expression with exponential
+functions \citep{koulakov_stochastic_2004}, the experimental evidence for the
+form of tectal ligand expression is more ambiguous than that for retinal
+receptor expression (For example data, see Fig.\,9 which shows expression patterns for EphB/ephrinB).
+We kept in mind the possibility that tectal ligand expression may be modelled
+by another function, but initially, we set it to the same exponential used for
+retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in
+place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$
+and $L_3$ (Fig.\,1).
 
@@
 
-%
-% FIGURES
-%
-
-% FIGURE 1
-% Expand panel A to be able to lose B-D.
-% Full page width
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/expressions_fig.png}
-\caption{Schematic diagrams of retinotectal mapping and gene expression.
-Coloured lines indicate that temporal retinal ganglion cells project to the rostral part of the contra-lateral tectum; nasal cells to caudal tectum; dorsal cells find their way to the lateral tectum and ventral retinal cells project to the medial tectum.
-%
-\textbf{A} Tectal ligand expression $L_i$.
-Tectal ligand expression patterns are assumed to increase exponentially in one direction only across the tectal surface.
-$L_0$ (or ephrin A) increases in a caudal to rostral direction (i.e.
-with increasing $y$; $x$ and $y$ directions are indicated with arrows).
-$L_2$ increases in the opposing sense, with $-y$.
-$L_1$ and $L_3$ form counter-gradients in the $\pm x$ directions.
-\emph{L: lateral, M: medial, C: caudal, R: rostral}.
-%
-\textbf{B} Retinal receptor and ligand expressions.
-Four receptors, $r_i$, are expressed across the retina.
-$r_0$ increases as an exponential function of $-x$ ; $r_1$ increases with an exponential function of $-y$.
-$r_2$ and $r_3$ have the opposite sense and increase with positive $x$ and $y$, respectively.
-Retinal ligand expression, $l_i$, is expressed in two complimentary gradients \citep{hornberger_modulation_1999}.
- \emph{N: nasal, T: temporal, D: dorsal, V: ventral}.
-%
-}
-\label{f:ex}
-\end{figure}
-
-% FIGURE 2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/j4_ee_GJ_best_1_wt_fig2_exit_true_steps_1500.png}
-\caption{Receptor-ligand model of chemotaxis and competition. 
-\textbf{A} The origin position on the retina gives axons a colour according to a dual-colour map.
-Axons in panels B-H that are red are those which originated in the most naso-ventral corner of the square simulated retina.
-\textbf{B}-\textbf{D} Axon centroid positions on the simulated tectum (dashed line) at time 0, 40 and 1500 units. Lines show adjacency relationships between the axons' origin positions on the retina. 
-\textbf{E} Layout of axon centroids on the tectum observed in experimental results (that is, correctly ordered into a perfect map of the simulated, square retina). 
-\textbf{F} The sum-of-squared difference error, $\epsilon$, and the number-of-crossings error metric, $\eta$ of the tectal axon arrangement plotted with respect to simulation time. 
-\textbf{G} The arrangement of axon branches at t=1500. 
-\textbf{H} Selected axon locations at t=1500, with path histories in grey. 
-\emph{Abbreviations: D dorsal, V ventral, T temporal, N nasal, C caudal, R rostral, M medial, L lateral.}}
-\label{f:GJ}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 3 ./scripts/fig_GJ_surgical_grafts.svg (graphs generated with fig_GJ_surgical.sh)
-% Could split into retina and tectum or 
-% a: rotation/graft (as hand waved by Gierer)
-% b: ablation (as simulated in 1D by Gierer)
-% c: mismatch/compound (as extensions of Gierer)
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_grafts.png}
-\caption{Performance of the receptor-ligand model with surgical graft manipulations.
-Here, we made a number of simulated surgical manipulations of the tectum and retina and applied the same model shown in Fig\,\ref{f:GJ}.
-\textbf{A.} Graft rotation.
-An 8$\times$8 square patch of tectum has been rotated by 90$\degree$, altering the tectal ligand expression pattern.
-High gradients form around the new discontinuities at the edge of the patch (see Fig\,\ref{f:trot90}), driving axons to follow the rotation of the patch.
-The left graph shows the axon centroid positions after 1500 time steps, the smaller inset graph shows the (idealised) anticipated layout seen in experiments \citep{levine_deployment_1974}.
-The right-hand graph shows the time-course of the pattern quality metric, $\epsilon$ (i.e.~how closely the centroid pattern matches the inset) for the entire tectum (blue), the graft region (yellow) and the surround (red).
-The axons within the graft are slightly less well ordered than those outside.
-\textbf{B} Shows the same type of graft rotation as \textbf{A} for a 180$\degree$ rotation \citep{yoon_retention_1973}.
-\textbf{C.} A simulated graft swap experiment.
-Two 12$\times$4 rectangular regions of tectum are cut out and grafted back in swapped positions \citep{hope_arrow_1976,gaze_visuotectal_1983}.
-While there is disorder in the simulated pattern of axon centroids, a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement.
-Note that here, the pattern metric shown for `Graft' (yellow curve) is computed from a square covering both grafts.}
-\label{f:GJsurg_grafts}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 4
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/fig_GJ_surgical_ablations.png}
-\caption{Retinal and tectal ablation simulations.
-\textbf{A} The temporal half of the retina is removed in this manipulation.
-Axons from the nasal half grow onto an unmodified tectum.
-This manipulation is problematic for the pure chemoaffinity theory, because axons from the middle of the retina now grow towards the rostral side of the tectum; the topographic map of the remaining retina expands across the entire tectum.
-In this work, it is the competition that achieves this effect.
-The axon centroid graphs are plotted as in Fig.\,\ref{f:GJsurg_grafts}, but right hand graph indicates the rostro-caudal tectal position of each axon centroid as a function only of its naso-temporal origin on the retina.
-\textbf{B} Axons from an unmanipulated retina grow onto a tectum whose caudal half has been removed.
-As in experiments \citep{yoon_reorganization_1971,finlay_orderly_1979}, the axons `crush up' to form a complete map of the retina over the half tectum.
-\textbf{C} A `mismatch' experiment \citep{horder_retention_1971}.
-The axons from the surviving nasal half of an ablated retina, which would usually map to the caudal half of the tectum arrange themselves into a map on the \emph{rostral} half of the tectum as the caudal portion has been removed.
-As with the other ablations, the model achieves this due to variable interactions with the tectal ligand gradient and competitive interactions.
-\textbf{D} Compound eye.
-If two nasal retina halves are grafted together, two overlaid maps form on the tectum \citep{gaze_retino-tectal_1963}.
-The simulation reproduces this result because axons from corresponding locations on the naso-temporal retinal axis interact equally with the tectal gradients and arrive in similar locations.}
-\label{f:GJsurg_ablate}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE 5
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations.png}
-\caption{Three genetic manipulations applied to the receptor-ligand model simulating those performed by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-In each manipulation, the left graph shows axon branch agents (colour coded by retinal origin) at the end of a simulation ($t=1500$), the centre graph shows corresponding axon centroids, with the colour blue indicating axons with no selective manipulation and red indicating those axons with a selective genetic knock-in of EphA3.
-In the right hand graph, the final tectal rostro-caudal locations of axons from each naso-temporal origin in the retina are plotted as violin plots (bars indicate the max/min datum in each group and shading indicates the data distribution).
-Inset graphs show experimental results from  \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-\textbf{A} We simulated the heterozygous EphA3 knock-in (EphA3 ki/+) of \citet{brown_topographic_2000} by selecting half of the retinal cells at random, and in these cells adding the value 0.8 to $r_0$ (which, as its expression increases towards the temporal retina, is our analogue for EphA).
-The affected cells experienced altered interaction with tectal ligand gradients and were shifted rostrally.
-The rostral shift was largest for nasal cells bound for the caudal tectum.
-These experience the strongest interaction with $L_0$ (which is highest in caudal tectum).
-Temporal cells with the manipulation experience a smaller change in their final location because $L_0$ expression is low at their target location (thus the $r_0$-$L_0$ interaction is minimal).
-However, there is no evidence of the observed map collapse at around 0.8 on the N--T axis, as seen in the inset graph.
-\textbf{B} To increase the magnitude of the manipulation of EphA3, \citet{brown_topographic_2000} added a homozygous EphA3 knock-in (modelled as a knock-in of 3.2) which pushed manipulated cells further towards the rostral tectum.
-Here, we observed a separation into two maps, one for the wildtype cells and one for the manipulated cells, in agreement with \citet{brown_topographic_2000} (inset is from their Fig.\,5C).
-\textbf{C} Simulation of EphA3 ki/+ selective knock-in and EphA4 knock-down (all cells) from \citet{reber_relative_2004}.
-The knock-down was applied by reducing $r_0$ in all cells by 1.2.
-The system fails to reproduce either the map collapse at around 0.9 on the N--T axis, nor the strong rostral pull for selective knock-in cells seen in experiments.
-Instead, selective knock-in cells are drawn in the \emph{caudal} direction, because the knock-down of 1.2 undoes the effect of the knock-in.}
-\label{f:GJeph}
-\end{figure}
-
-% FIGURE 6
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_explanation.png}
-\caption{Two new mechanisms to account for the genetic manipulation results: cluster-size and $r_2$ collapse.
-To account for map collapse (\citet{brown_topographic_2000} and Fig.\,\ref{f:GJeph}A), and for the significant rostral termination of knock-in/knock-down cells (\citet{reber_relative_2004} and Fig.\,\ref{f:GJeph}C), we assumed that i) retinal EphA interactions with tectal ephrinA ligands are modulated by EphA4 receptors through a `side-binding' mechanism and ii) EphA3 and EphA4 receptors interact with the receptors which mediate the chemotaxis effect that \emph{opposes} EphA (i.e.~with $r_2$).
-\textbf{A} \emph{EphA/ephrinA expression.} As described in the text, EphA ($r_0 \equiv \sum_{i\neq 4}\mathrm{EphA}i$) is expressed according to $r_0 = 1.05 + 0.26 \exp(2.3 u)$ (green curve) where $u$ is the N--T retinal position.
-Knock-in is simulated by adding 0.8 to this function (giving the purple curve).
-Retinal ephrinA expression is a reversed version of EphA expression (green dotted curve).
-A threshold, $h_0$, used by the $r_2$ collapse mechanism is shown as a red dotted line, with a green dash-dot line indicating the intersection of $\Sigma\mathrm{EphA}$ and $h_0$.
-\textbf{B} \emph{EphA4 expression.} Overall EphA4 expression across the retina is constant (green dotted line, value 3.5) but retinal ephrinA ($l_0$) can bind to EphA4 ($r_{A4}$), modelled here as a multiplicative interaction to give cis-bound EphA4, $r_{A4}^{\mathrm{cis}} = w_{A4} \times r_{A4} \times l_0$ (green dashed curve), where $w_{A4}$ is a binding weight described in the text.
-The remaining, un-bound, free EphA4 receptors, $r_{A4}^{\mathrm{free}} = r_{A4} - r_{A4}^{\mathrm{cis}}$ (green solid curve).
-We model genetic knock-down of EphA4 by subtracting 2.1 from $r_{A4}$, then recomputing $r_{A4}^{\mathrm{cis}}$ and $r_{A4}^{\mathrm{free}}$ to give the cyan curves.
-\textbf{C} \emph{Cluster size affects signal strength.} The signal strength, $s_0$, for an interaction between EphA receptors and a tectal ligand gradient is modelled as $s_0 = r_0 \times c_0$ where $c_0$, the cluster size, is $1/r_{A4}^{\mathrm{free}}$.
-Shown for wildtype, knock-in and knock-down experiments, the cluster-size modulated signal $s_0$ increases for knock-in and for knock-down manipulations and is further enhanced by combined knock-in and knock-down.
-This cluster size-modulated signal, $s_0$, is smaller in magnitude than the signal for $r_0$ in the original model (for which $s_0 \equiv r_0$).
-To compensate, $r_2$ is divided by 2, which gives the correct behaviour in a wildtype system.
-\textbf{D} \emph{$r_2$ collapse.} i) We assume that the strength of receptor interactions for $r_2$ collapses under the conditions that $r_0$ exceeds a threshold $h_0=2$ (red dotted line in A) and that $r_{A4}^{\mathrm{free}}$ is simultaneously below a threshold $h_{A4}=1.1626$ (red dotted line in B).
-Under these conditions, $r_2$ is reduced to one fifth of the value it would otherwise take.
-ii) The regions of $r_2$ collapse are illustrated.
-No collapse occurs for wildtype EphA and EphA4 expressions; in cells with only EphA4 knock-down, $r_2$ collapse occurs for temporal cells (those more temporal than the dash-dot line in A).
-In cells with both EphA3 knock-in and EphA4 knock-down, the collapse mechanism occurs regardless of retinal origin.}
-\label{f:clustermech}
-\end{figure}
-
-% FIGURE 7
-\begin{figure}
-\includegraphics[width=\linewidth]{./images/EphA_manipulations_clustertheory_r2collapse.png}
-\caption{Model results with the mechanisms described in Fig.\,\ref{f:clustermech}. 
-\textbf{A} EphA3 knock-in result for ki=0.8.
-The graph appears similar to that in Fig.\,\ref{f:GJeph}A but crucially, the distributions of tectal locations begin to overlap at about 80\% of the N--T axis, in agreement with experiment (inset).
-This is caused by the cluster size effect, in which the signal increase due to the knock-in is reduced for temporal RGCs compared with nasal cells (See Fig.\,\ref{f:clustermech}C, green/purple solid curves).
-%FOR TEXT: It is confirmed by an increased achieved significance level (ASL) from a bootstrapped, studentized t-test comparing the red and blue distributions at each N--T location. 
-\textbf{B} The cluster size effect does not change the double knock-in result, which shows excellent agreement with experiment (inset). 
-\textbf{C} When EphA is knocked in and EphA4 is simultaneously knocked down, cells on the red curve experience a combination of interactions: First, the increased interaction with ephrinA that gave the result in A is still in effect.
-Second, all cells on the red curve experience $r_2$ collapse, whereas only temporal RGCs on the blue curve experience this effect.
-$r_2$ receptors are more numerous on nasal RGCs and so the collapse in $r_2$ generates a substantial rostral movement of the red curve, in agreement with \citet{reber_relative_2004}.
-\textbf{D} Uniform knock-down of EphA4 in all cells (with no manipulation of EphA3) does not perturb the mapping to a significant extent, although the effect of the EphA4 knock-down on the cluster size mechanism causes a slight rostral shift in all cells.}
-\label{f:r2collapse}
-\end{figure}
-
-% ANY SUPPLEMENTARY FIGURES GO AFTER THIS
-%\renewcommand{\thefigure}{S\arabic{figure}}
-%\setcounter{figure}{0}
-
-% FIGURE S1
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/Tissuevisb.png}
-\caption{(Supplementary) Visualisation of ligand expression and gradients on the simulated, square tectum for the
-graft-rotate 90 degrees surgical manipulations. 
-\textbf{A} Ligand expression. 
-The expression of four ligand types, $L_i$,  $i = 0$--$3$, is visualised with two duochrome maps. 
-The lefthand map shows values for ligands $L_0$ and $L_1$ with red saturation indicating the $L_0$ expression value and green indicating the expression of $L_1$. 
-Elements which have high expression of both $L_0$ and $L_1$ are coloured yellow; those with expression of neither are black. The $8\times8$ square in the centre of the 
-The righthand map shows values for $L_2$ and $L_3$, using magenta and cyan, respectively. 
-\textbf{B} Visualisations of the $x$ components of the ligand gradients computed from the maps in A. 
-For each ligand $L_i$, the gradient $\nabla_x(L_i)$ is shown (where $x$ indicates the M--L axis). 
-The $x$ gradient is computed discretely as the mean difference between expression values in the neighbouring elements along the $x$ axis.
-Full saturation indicates a positive gradient; white indicates a negative gradient.
-For the surrounding, non-graft tissue, only the $L_1$ (green) and $L_3$ (cyan) $x$ gradients vary along the M--L axis. 
-Strong discontinuities in $\nabla_x$ are seen on the medial and lateral edges of the patch.
-\textbf{C} Corresponding visualisations of $\nabla_y(L_i)$, with strong discontinuities in this gradient along the caudal and rostral edges of the square graft.}
-\label{f:trot90}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S2
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/expression_profiles_EphBs.png}
-\caption{(Supplementary) EphB and ephrin B gene expression. 
-Here, we used Stalefish \citep{james_comparing_2022} to quantitatively analyse gene expression of EphB2 and ephrinB1/B2 from images published in the literature. 
-\textbf{A} Retinal expression of EphB2. 
-We analysed Fig.\,7 of \citet{holash_polarized_1995} using `dual curves' to delineate the expression region in the Stalefish application. 
-The resulting, normalized expression is shown in the graph for the E4 data, with an inset showing corresponding data for the E6 data. 
-In the right-hand graph, we show an expansion of a portion of the E4 data to which we applied a linear fit to determine the RMS deviation from the mean to estimate the noise in the receptor signal. 
-The mean RMS error was 0.06. 
-\textbf{B} Stalefish screenshots and expression graphs for retinal ephrinB1 and ephrinB2 expression, using data from Fig.\,1 of \citet{braisted_graded_1997}. 
-\textbf{C} Screenshots and expression graphs for tectal ephrinB1 expression, using data from Fig.\,4 of \citet{braisted_graded_1997}.}
-\label{f:expr_ephb}
-\end{fullwidth}
-\end{figure}
-
-% FIGURE S3
-\begin{figure}
-\begin{fullwidth}
-\includegraphics[width=0.95\linewidth]{./images/noise_contribution.png}
-\caption{(Supplementary) Response of the model to noise.
-\textbf{A} Adding 3\% gradient measurement noise (normally distributed, standard deviation, $\sigma = 0.19$) has negligible effect on the pattern. 
-\textbf{B} 3\% retinal receptor noise ($\sigma = 0.08$). 
-\textbf{C} 3\% tectal ligand noise ($\sigma = 0.08$). 
-\textbf{D} Combined effect of retinal receptor, tectal ligand and gradient measurement noise is dominated by the tectal ligand noise.}
-\label{f:noise}
-\end{fullwidth}
-\end{figure}
-
 \end{document}

### 2024-07-05 paper/neural_2024/poster_text/intro.tex
Adds poster for Neural Computation 2024

--- 2024-07-04 paper/neural_2024/poster_text/abs.tex
@@
 
-The axons of retinal ganglion cells form an ordered map of the retina
-in the optic tectum. Surgical and genetic experiments have revealed
-several key mechanisms that contribute to retinotectal map
-development, and computational modelling studies have demonstrated how
-maps can be shaped by their interactions. Models derived from Sperry's
-classic chemoaffinity theory can reproduce retinotopy, and Gierer
-later showed how an extension to include competition between growing
-axons could also account for a subset of the disordered maps that have
-resulted from surgical manipulation. Further extensions have each
-accounted for additional datapoints, but none has been able to explain
-the full range of maps observed experimentally. Here we present a
-unified model of retinotectal development, assuming only local
-interactions between growing axons, which can account for the full
-range of maps resulting from surgical manipulation. This model extends
-the `potential functions' of Gierer's original to describe
-self-organisation across a 2D cortical sheet, and it represents a
-coordination of axonal growth by orthogonal pairs of opposing ligand
-gradients in the tectum and corresponding receptor gradients in the
-retina. In addition, we present progress towards accounting for the
-full range of maps that have resulted from genetic manipulation, by
-incorporating a mechanistic description of receptor ligand
-interactions.
+%\section{Introduction}
 
-%\bibliography{RetinoTectal}
+\subsection{A summary of this work}
+
+\begin{itemize}
+\item Retinal cells grow axons that form an ordered map in the optic tectum
+\item Sperry, Gierer, Prestige, Willshaw and many others have all sought to model this
+  developmental process to better understand how brains wire themselves up
+\item Surgical and genetic manipulation experiments have provided a wide range
+  of developmental responses that these models should be able to describe
+\item We want to make a model based on local interactions that can account for
+  all of the experimental results
+\end{itemize}
+
+\subsection{A brief history of retinotectal studies}
+
+(For reviews see \citealp{Prestige1975,Goodhill2005})
+
+\begin{itemize}
+\item 1940s: Sperry's surgical experimental and theoretical work
+  \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}
+\item 1970s: Marker-induction models \citep{Prestige1975,Willshaw2006}
+\item 1980s: Chemotaxis models \citep{Gierer1983,Gierer1987}, branch arrow
+  models \citep{overton_extended_1982}
+\item 1990s: Discovery of the ephrin signalling molecules \citep{nakamoto_topographically_1996}
+\item 2000s: Important genetic manipulation studies
+  \citep{brown_topographic_2000,reber_relative_2004}; Hebbian activity models \citep{Tsigankov2006,Tsigankov2010}
+\item 2010s: Competition models \citep{Triplett2011,Simpson2011}; comparative
+  studies \citep{hjorth_quantitative_2015}
+\item Recently: Re-evaluation of chemoaffinity \citep{Sterratt2013,naoki_revisiting_2017}
+\end{itemize}
+
+One of the more abstract models \citep{Triplett2011} was shown by
+\cite{hjorth_quantitative_2015} to be the only one able to explain the `EphA3
+collapse point'. However, although the model is evaluated numerically (by
+iterative switching), this procedure is not necessarily an explicit
+representation of the process of development \citep{Wilson2015}. We wanted to
+find a mechanistic model which can be directly related to the behaviour of
+developing axon processes. Starting with a modified version of \cite{James2020},
+and influenced by the \cite{Simpson2011}
+we built a Gierer-like gradient climbing model in which we attempted to
+express all the interactions of the axon branches in terms of receptor-ligand
+signalling.
+
+\bibliography{RetinoTectal}
 

### 2024-07-05 paper/neural_2024/poster_text/results2.tex
Adds poster for Neural Computation 2024

@@
 %
-We defined 4 receptor expression gradients (each varying only wrt one spatial dimension) arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,1B).
+We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
+$r_2$ was orthogonal to $r_3$ (Fig.\,1B).
 Thus the model is of \emph{differential chemotaxis}.
 
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential
-functions \citep{koulakov_stochastic_2004}, the experimental evidence for the
-form of tectal ligand expression is more ambiguous than that for retinal
-receptor expression (For example data, see Fig.\,9 which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled
-by another function, but initially, we set it to the same exponential used for
-retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in
-place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$
-and $L_3$ (Fig.\,1).
+%Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in
+%orthogonal pairs of gradients.  Several studies model tectal ligand expression
+%with exponential functions \citep{koulakov_stochastic_2004}. Although the
+%experimental evidence to support this is lean, we also set $L_i$ to the same
+%exponential used for retinal expression.
 

### 2024-07-05 paper/neural_2024/poster_text/speculate.tex
Adds poster for Neural Computation 2024

--- 2023-08-25 paper/main.tex
@@
 
-\title {
-Competitive differential chemotaxis can explain the self-organisation of retinotopic maps
-}
+\title {none}
 
-\author[1*]{Sebastian~S.~James}
-\author[1]{Stuart~P.~Wilson}
+\author[1]{Sebastian~S.~James}
+\author[1*]{Stuart~P.~Wilson}
 \affil[1]{Department of Psychology, The University of Sheffield, Sheffield, United Kingdom.}
-\corr{seb.james@sheffield.ac.uk}{SSJ}
+\corr{s.wilson@sheffield.ac.uk}{SW}
 %% END OF PREAMBLE
 
-% Seb to fix: Expected not experiment (in figs), Fig 8 - monochromes. Fig 1 - monochromes
-% Stuart to do: Intro/Start of results modifications
-
 \begin{document}
-%\setlength{\droptitle}{-1.8cm} % move the title up a suitable amount
-\maketitle
-
-\begin{abstract}
-The retinotectal projection, in which the axons of retinal ganglion cells form an ordered map in the optic tectum, is an important developmental system that has been studied for many decades.
-There is a large body of evidence, consisting of surgical and genetic manipulations, that give insight into how it behaves.
-Experimental work has uncovered a number of mechanisms that are involved in aspects of the mapping.
-However, a single, consistent theoretical framework has yet to describe all the patterns that are formed.
-Sperry's chemoaffinity theory provides an account of the wildtype pattern.
-Gierer formalised and extended this theory to incorporate competition to account for a subset of surgical manipulation experiments, but it was never tested against the full range of surgical manipulations.
-Other theoretical and computational studies have been partially successful in accounting for surgical manipulations, but to date, no single theory, operating only by local interactions, has accounted for every manipulation.
-We present a two-dimensional model of the retinotectal projection in which orthogonal pairs of opposing tectal ligand gradients combined with retinal gradients of receptors form Gierer's potential functions and then show that this receptor-ligand model accounts for the full range of surgical manipulations.
-We then extend the detail with which the receptor-ligand interactions are treated to successfully account for more recent genetic manipulation experiments.
-\end{abstract}
-
-\section{Introduction}
-
-Since the pioneering work of Sperry \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}, progress on understanding topographic patterning of brain connectivity has been made through a synergy of experimental and computational neuroscience, focussed on the projection of retinal ganglion cell axons into the optic tectum.
-Since computers first became available as scientific tools, a wide range of theories have been developed and tested through formal algorithms and computer simulation work.
-Models can be distinguished on a number of grounds, such as the level of abstraction at which they describe the process of map development, and the assumptions they represent about the contributions of chemotaxis, molecular interactions between axons, and competition for space in the destination tissue (for reviews see \citealp{Prestige1975,Goodhill2005}).
-
-One important class of theories assumes that map formation is driven by chemotaxis, influenced by pairs of anti-parallel expression gradients in the tectum, with the balance of influences in each axon determined by a receptor density specific to its origin in the retina (\citealp{Gierer1983,Gierer1987,Simpson2011}; see also \citealp{Karbowski2004,James2020}).
-A second class of theories assumes that a single gradient is primarily responsible for chemotaxis and that maps reflect additional processes such as competition \citep{Triplett2011}, Hebbian activity \citep{Tsigankov2006,Tsigankov2010}, or marker induction \citep{Prestige1975,Willshaw2006}.
-
-A key test of the validity of recent models is their ability to account for data from a set of genetic manipulation studies by \cite{brown_topographic_2000} and \cite{reber_relative_2004}, in which axons from a subset of retinal ganglia cells treated with a Eph-A3 knock-in form a separate, rostrally shifted topographic map, which becomes indistinct from that formed by untreated cells, at a ‘collapse point’ on the rostral side.
-\cite{Sterratt2013} found that a mixture of counter-gradients and an additional competition term provided the best account of these data.
-A follow-up study in which a range of different computational models were systematically compared \citep{hjorth_quantitative_2015} found that one focused on competition by \cite{Triplett2011} could best account for several such experimental findings, and that the other types of model could not explain the collapse point.
-
-The model of \cite{Triplett2011} describes maps as minima of an energy function, whose structure can be obtained via a stochastic optimisation process.
-Accordingly, the energy is the sum of three terms, relating to the influences of chemical gradients, axon density (competition), and activity-dependent growth processes.
-Maps that optimise this function can be obtained by starting with a random map and then substituting connections at random, accepting those substitutions with a probability that increases if they reduce the energy, and repeating until the algorithm converges on a pattern of connections corresponding to a minimal energy state.
-
-The ability of this model to explain the collapse point in the Eph-A3 knockin studies can be understood in terms of its relationship to simpler models developed previously, with which it shares (via the activity-dependent term in the energy function) similar assumptions \citep{Koulakov2004}.
-According to this earlier model, pairs of adjacent axons on the tectum are evaluated iteratively and the probability of their positions being switched increases with an energy defined by the product of the difference in Eph receptor density and the difference in ephrin ligand density.
-As the retinal receptor and tectum ligand densities express anti-parallel gradients, a positive energy is translated into a pairwise switch of axon positions with high probability.
-Eph knock-ins are assigned higher receptor density values, and the map collapse point emerges when differences in receptor density (due to retina position and treatment condition) are offset by differences in ephrin ligand density in the tectum.
-The two terms are balanced at a point that shifts further in the direction of cells with a more temporal retina origin for a stronger (i.e., homozygous) knock-in, as observed experimentally.
-
-These models inherit from thermodynamic models used to investigate phase transitions in statistical physics, the concept of a `temperature', via a parameter that weights the switching probabilities and effectively determines the sharpness of the collapse between treated and untreated maps.
-Interestingly, at high temperatures (sharp switching) the model of \cite{Koulakov2004} becomes formally equivalent to the `arrow’ model, one of the first models of retinotectal map self-organisation proposed  \citep{Hope1976}.
-
-The explanatory power of these `switching' models stems from their simplicity and degree of abstraction, revealing the map pattern that optimises the constraints represented by a given energy function.
-However, while switching models are evaluated through a numerical procedure (i.e., iterative switching), that procedure should not necessarily be interpreted as an explicit representation of the process of development \citep{Wilson2015}.
-Consider, for example, that from more realistic initial conditions than the uniform random map typically assumed, local energy minima quite unlike organised biological maps are more likely to be encountered.
-
-In a parallel strand of work the principles represented by switching models have been explored in models that represent the developmental process more explicitly.
-For example, \cite{Overton1982} showed how the net effects of iterative switching in discrete systems could be achieved via a continuous dynamics based on iteratively ‘nudging’ axons by a velocity vector that is weighted by the factors considered by the arrow model, including for example, functions of the proximity of growing axons to their neighbours.
-
-Continuing this strand of work, \cite{Simpson2011} have more recently extended the model of \cite{Overton1982} from its original description of axon growth in one spatial dimension to simulate growth in two dimensions, and they have shown that it can account for a wide range of data from genetic studies, as well as a wide range of surgical manipulation experiments; including the emergence of a map collapse point in simulated Eph-A3 knock-in experiments.
-
-The models of \cite{Overton1982} and \cite{Simpson2011} incorporate three basic ingredients; competition for space, receptor-based interactions between axons, and gradient-based guidance.
-However, a major limitation is that chemotaxis is not implemented explicitly.
-Instead, simulations involve providing growing axons with explicit directional information about their target destination at every timestep.
-This was an appropriate choice given the scientific goals of \cite{Simpson2011}, but for this reason the theory of retintopic map formation that is represented by their model cannot be tested comprehensively, and indeed this model was excluded from the systematic comparison of models by \cite{Hjorth2015} on that basis.
-
-An open question then, which we address in this paper, is whether the \cite{Simpson2011} model can still account for a wide range of experimental data from surgical and genetic experiments, when its assumptions about chemotaxis are represented explicitly, i.e., when simulated axon growth is guided by only local information?
-If so, such a model would represent the most complete account of retinotectal map self-organization to date.
-
-%Sperry's chemoaffinity model is an influential theory of how axons from neurons that originate in one tissue can find their way to specific locations in a target tissue during early brain development \citep{sperry_reestablishment_1942,sperry_visuomotor_1943,sperry_chemoaffinity_1963}.
-
-%The central proposal of the theory is that growing axons find their location in a destination tissue by means of `labels of a cytochemical nature'.
-The chemical affinity between labels in source and destination tissues allows axons to distinguish correct target locations from incorrect ones.
-
-%This key idea (called chemospecificity) does not, on its own, explain how an axon can find the correct \emph{direction} in which it should move to reach its target.
-An additional component of Sperry's theory \citep{stevens_handbook_1950,sperry_problems_1955,sperry_chemoaffinity_1963} suggested that the axon could determine this direction by evaluating local information in gradients of chemicals spanning the target tissue, giving rise to chemotaxis of the axonal growth cones.
-
-%Much of Sperry's work was carried out on the retinotectal system, which is ideal for the study of ordered axonal development in fish, amphibia and birds, because in these animals, patterning of axons across the tectum is largely complete prior to the onset of neural activity \citep{oleary_molecular_1999}.
-
-%In this system a two-dimensional map of  the retina is developed on the surface of the optic tectum such that the axons from adjacent retinal ganglion cells (RGCs) find their way to adjacent locations on the tectum.
-This retinotectal map is amenable to a number of surgical manipulations especially in amphibia, in which re-growth of the projection can be observed after surgical changes to the sensory apparatus (such as ablation or rotation of the retina) or the map substrate (the tectum).
-This allowed Sperry and other researchers to carry out a wide range of experimental manipulations, the results of which informed the chemoaffinity theory.
-
-%In the 1990s, the theory of chemoaffinity and chemotaxis was given robust support by the discovery of the ephrin ligands and their receptors \citep{cheng_complementary_1995,drescher_vitro_1995} which have been shown to form into graded expression fields in the retina \citep{braisted_graded_1997}, tectum \citep{braisted_graded_1997,feldheim_genetic_2000} as well as in other sensory systems, such as the somatosensory system \citep{vanderhaeghen_mapping_2000}.
-% And sounds?
-
-%The Ephrin ligands have a clear effect on axonal outgrowth, as shown in-vitro \citep{cheng_complementary_1995,drescher_vitro_1995,hansen_retinal_2004} and by in-vivo studies \citep{frisen_ephrin-a5_1998,rodger_transient_2000,mann_topographic_2002,hindges_ephb_2002}.
-Two classes of ephrins have been identified, labelled `A' and `B'.
-Ephrin A and its receptors mediate a repulsive interaction which contributes to an ordering of RGC axons on the tectum along the rostral-caudal axis.
-Ephrin B acts on medial-lateral ordering with an attractive interaction.
-Together, the ephrins are able to provide a coordinate system on the tectum to guide growing axons, but precisely how this positional information is used by axonal growth cones is not yet fully understood.
-
-%Quantitative models exploring axonal ordering were already in existence by the time of the discovery of the ephrins.
-The importance of chemical gradients to biological processes was already well known from the study of slime moulds \citep{bonner_evidence_1947}.
-Sperry himself discussed chemical gradients in relation to the problem of neuronal pathfinding \citep{sperry_problems_1955}, albeit qualitatively.
-Some time after Sperry's active period, but before the ephrins were identified, \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987} made one of the few quantitative investigations of a gradient-based model of the retinotectal projection.
-He presented gradient following as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-This model forms a mathematical expression of Sperry's theory.
-Because it is quantitative, it allowed Gierer to identify two features of the theory: i) A \emph{pair} of opposing gradients is needed to form a potential function in one dimension (and \emph{two} pairs to create a 2D potential function).
-ii) Tectal or retinal ablation would require $p$ to be re-specified to reproduce the experimental results that fibres from an ablated retina will generate a map across the entire tectum and fibres from an intact retina will form a compressed map on an ablated tectum \citep{attardi_preferential_1963,schmidt_retinal_1978,schmidt_expansion_1978}.
-
-%Gierer addressed the second problem by introducing a form of competition between fibres, postulating that $p$ would accumulate locally at a rate proportional to the fibre density in that location and he noted that this was equivalent to a more direct fibre-fibre competition.
-He demonstrated, in a one-dimensional computer simulation, that this form of competition would lead to the expected response to ablations \citep{gierer_model_1983}.
-
-%Gierer's simulations confirmed the validity of this basic idea, which we will refer to as \emph{differential chemotaxis}, for the problem of map formation with respect to a single spatial dimension \citep{gierer_model_1983,karbowski_model_2004}.
-The model helped establish how \emph{competition between axons} for space in the target tissue can additionally account for  map development after tissue ablation.
-
-%Simulations of differential chemotaxis and axonal competition have recently been extended to investigate map development in two spatial dimensions, revealing how these two ingredients can give rise to realistic patterns of synaptic innervation \citep{james_modelling_2020}.
-These models are formulated as systems of partial differential equations (PDE), and solved through numerical integration on an underlying lattice structure representing the target tissue.
-In general, this approach makes mathematical analysis of system dynamics tractable, and as such PDE modelling has played a central r\^ole in the development of theories of pattern formation (see \citet{krause_modern_2021} for recent examples).
-However, we found it numerically intractable to extend the model described in \citet{james_modelling_2020}, which simulated 41 whisker-specific projections, to the retinotectal system for which hundreds of projections are required (to demonstrate surgical manipulations).
-Instead, we sought to implement an \emph{agent-based} model of differential chemotaxis, in which each axon growth cone is an agent that interacts with its environment and with the other growth cone agents in the system.
-With this model, we asked: i) Can a two-dimensional implementation of Gierer's model explain the full range of surgical manipulations that have been described in the literature?
-ii) Can the model be developed such that only \emph{local} interactions are required to transfer information between agents?
-and iii) Can a differential chemotaxis model explain more recent results in which levels of EphA receptor sub-types are changed by genetic manipulation \citep{brown_topographic_2000,reber_relative_2004}?
-
-\section{Results}
-\subsection*{A model of chemotaxis and competition}
-
-In an agent-based model of chemotaxis and competition, the position, $\mathbf{x}_t$, of each branch $b$ on a simulated tectum can be updated according to $\mathbf{x}_{t+1} = \mathbf{x}_{t} + \Delta \mathbf{x}$ with the movement vector given by
-\begin{equation} \label{e:dX}
- \Delta \mathbf{x} = \mathbf{G} +  \mathbf{C} + \mathbf{B}.
-\end{equation}
-$\mathbf{G}$ is the movement due to the chemotaxis effect and $\mathbf{C}$ is the movement due to competitive interactions with other branches. 
-$\mathbf{B}$ is a movement applied close to the tissue boundary that ensures all branches remain within the tissue \citep{holt_target_1998}.
-
-\citet{simpson_simple_2011} investigated a model of this form in which a `placeholder' provided a chemotaxis-like movement for $\mathbf{G}$ and a form for $\mathbf{C}$ was studied.
-Their model, which successfully accounts for a wide range of experimental manipulations, found a $\mathbf{C}$ comprised of two components; a simple distance-based competition between \emph{all} branches, along with an axon-axon interaction based on the relative levels of the receptor EphA expressed by each interacting axon.
-We asked i) if the placeholder mechanism for $\mathbf{G}$, in which a `global supervisor' informs each branch of the vector along which it should travel to approach its target, could be replaced by a true chemotaxis mechanism based on graded signalling molecule expression and ii) if the competition mechanism found by Simpson and Goodhill was the simplest possible mechanism that could explain all of the surgical and genetic manipulations accounted for by their model.
-
-\subsection*{A gradient based model of chemotaxis}
-
-We assumed a purely linear receptor binding model, and set the chemotactic movement vector of the branch $b$ at location $\mathbf{x}$ on the tectum to be
-
-\begin{equation}\label{e:G}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,r_{i,b} \nabla L_i(\mathbf{x})
-\end{equation}
-%
-where $r_{i,b}$ is the receptor expression on branch $b$ for ligand-receptor pair $i$, $L_i$ is the expression of ligand $i$ on the tectum and $F_i$ denotes the direction of the interaction induced when a molecule of ligand $i$ binds to a receptor $i$ molecule.
-$F_i$ takes the value $-1$ for a repulsive interaction or $1$ for an attractive interaction.
-%
-We assumed that all receptor-ligand signalled interactions are repulsive ($F_i=-1, \forall i$), as for EphA-ephrinA coupling \citep{drescher_vitro_1995,nakamoto_topographically_1996}.
-%
-$m_{\!_G}$ is a scalar parameter which controls how much movement is generated for a given level of receptor-ligand gradient signalling.
-
-%%%%% Describing the simulation
-We simulated the retina and tectum as unit square regions and defined retinal receptor expression and tectal ligand expression across these squares in the directions defined in Fig.\,\ref{f:ex}. 
-The simulated retina possessed $n$=400 (20$\times$20) retinal ganglion cells; the tectum was similarly arranged as 20$\times$20 discrete, square elements, with defined levels of expression for each ligand.
-One spatial unit corresponds to 2~mm on the retina or tectum for mouse \citep{reber_relative_2004}.
-
-Each retinal ganglion cell (RGC) projects $M=4$ growth cones (referred to as \emph{branches}) which carry a set of $N$ receptors, $r_i$, indexed by $i$ and expressed at levels determined by the cell soma's location on the retinal grid.
-($N$ ligands, $l_i$, may also be expressed by RGCs, but these are not used by the chemotaxis model.) 
-The expression of each receptor varies only with respect to one spatial dimension.
-%
-We defined 4 receptor expression gradients arranged in orthogonal pairs with the gradient of $r_0$ being orthogonal to that of $r_1$.
-$r_2$, whose gradient is opposite to $r_0$, was orthogonal to $r_3$ (Fig.\,\ref{f:ex}B).
-Thus the model is of \emph{differential chemotaxis}.
-%
-Because there is convincing evidence that EphA and EphB receptors are expressed in exponentially increasing patterns \citep{reber_relative_2004,feldheim_genetic_2000,brown_topographic_2000,koulakov_stochastic_2004}, we use an exponential form for retinal receptor expressions, in common with other modelling studies~\citep{reber_relative_2004,koulakov_stochastic_2004,simpson_simple_2011}.
-We adopt the same precise form for the retinal receptor expression as \citet{simpson_simple_2011}, $f(x,y) = 1.05 + 0.26 \exp(2.3 u)$, where $u$ gives the direction with which the expression increases and is substituted by $(1-x)$ for $r_0$, $(1-y)$ for $r_1$, $x$ for $r_2$ and by $y$ for $r_3$ (Fig.\,\ref{f:ex}).
-
-Cells on the tectum express ligands, $L_i$, for the retinal receptors, also in orthogonal pairs of gradients.
-Although several studies model tectal ligand expression with exponential functions \citep{koulakov_stochastic_2004}, the experimental evidence for the form of tectal ligand expression is more ambiguous than that for retinal receptor expression (For example data, see Fig.\,\ref{f:expr_ephb} which shows expression patterns for EphB/ephrinB).
-We kept in mind the possibility that tectal ligand expression may be modelled by another function, but initially, we set it to the same exponential used for retinal expression and used the substitutions $y$, $x$, $(1-y)$ or $(1-x)$ in place of $u$ to obtain the ligand expression functions for $L_0$, $L_1$, $L_2$ and $L_3$ (Fig.\,\ref{f:ex}).
-
-While we do not explicitly name $r_0$, $r_1$, etc.~as \emph{EphA}, \emph{EphB}, the suggestion is that $r$ could include EphA, EphB, Ryk \citep{schmitt_wntryk_2006} and Neogenin \citep{rajagopalan_neogenin_2004} receptors and that $L$ could include the ephrin-A, ephrin-B, Wnt3 \citep{schmitt_wntryk_2006} and RGM \citep{monnier_rgm_2002} ligands, each of which has been shown to play a r\^ole in retinotectal map formation.
-
-\subsection*{A signalling model for competition}
-
- We modelled competition as a distance-based competitive movement vector, $\mathbf{C}$, which is a sum of interactions that are `switched on' via receptor-ligand signalling:
-%
-\begin{equation} \label{e:X}
-\mathbf{C} = \frac{m_{\!_X}}{|B_{b}|} \sum_k^{nM} \hat{\mathbf{x}}_{kb}\,W\,Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}).
-\end{equation}
-%
-$r_{\!_X}$ is the receptor-ligand interaction radius, $W$ is the distance based weighting ($W = 1-\frac{d_{kb}}{2r_{\!_X}}~\mathrm{if}~  d_{kb}\leq 2r_{\!_X}$, otherwise $0$) and $B_{b}$ is the set of branches that are within interaction distance ($2 r_{\!_X}$) of $b$.
-$m_{\!_X}$ is a scalar parameter, $nM$ is the total number of branches and $\hat{\mathbf{x}}_{kb}$ is the unit vector from $k$ to $b$.
-%
-$Q$ is a signalling threshold function which depends on the distance $d_{kb}$ between two branches $b$ and $k$, the receptor expression on $b$ ($\mathbf{r}_b$) and the ligand expression on $k$ ($\mathbf{l}_k$):
-%
-\begin{equation}
-Q(d_{kb}, \mathbf{r}_{b}, \mathbf{l}_{k}) = \begin{cases}
-                 0 & \mathrm{if}~-F_i\,r_{i,b}\,l_{i,k} <
-                 s,\,\forall{i}~\mathrm{or}~d_{kb} > 2r_{\!_X} \\
-                 1 & \mathrm{otherwise.}
-     \end{cases}
-\end{equation}
-%
-$l_{i,k}$ (the expression of ligand $i$ on branch $k$) is the $i^{\mathrm{th}}$ element of $\mathbf{l}_k$ and $r_{i,b}$ is the $i^{\mathrm{th}}$ element of $\mathbf{r}_b$. 
-$F_i$, defined in Eq.\,\ref{e:G}, is -1 if the $r_{i}\,l_{i}$ interaction signals repulsion and +1 if it signals attraction.
-$s$ is a positive signal threshold for interaction.
-With the constraint $s>0$ comes the assumption that attractive interactions will have no effect on the competitive movement vector---a receptor-ligand pair that signals attraction cannot cause $Q$ to take the value 1.
-$Q$ is set to 1 only if at least one repulsive signal exceeds $s$ and the distance from branch $b$ to branch $k$ is smaller than $2 r_{\!_X}$.
-
-\subsection*{Simulation}
-
-The model has four free parameters; $m_{\!_G}$, $m_{\!_X}$, $r_{\!_X}$ and $s$.
-We fixed $m_{\!_G}$ to the value $0.0012$ and used a numerical optimization process to find the competition parameters that would balance this level of chemotaxis.
-We performed a simulated annealing optimization to minimize an objective function for the wildtype pattern (Fig.\,\ref{f:GJ}E) at simulation time 1000, reasoning that evolution would optimize the arrangement of this phenotype.
-
-We used a combination of pattern metrics to define the objective function.
-$\epsilon$ is the mean distance error of the axon centroids, given by $\epsilon = \sqrt{\left(\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2\right)/n}$,
-%
-%\begin{equation}\label{e:eps}
-%\epsilon = \sqrt{\frac{\sum_k^n (\mathbf{x}_{k} - \mathbf{x}'_{k})^2}{n}},
-%\end{equation}
-%
-where $\mathbf{x}_{k}$ is the location of the centroid of the branches of axon $k$ and $\mathbf{x}'_{k}$ the target tectal location for axon $k$.
-$n$ is the number of RGC axons in the system.
-%
-$\epsilon$ tends to 0 as the pattern approaches the target layout.
-However, two patterns may have the same non-zero value for $\epsilon$, despite one being qualitatively more disordered than the other. 
-For this reason, we also used the number of crossings in the fishnet plot of the axon centroids (see Fig.\,\ref{f:GJ}D) as a second metric, called $\eta$, which tends towards an expected target number of crossings, $\eta_t$ as the pattern approaches the target layout.
-The annealing objective function, $\xi$, combined these metrics: $\xi = \epsilon^2 (1+\frac{\eta}{100})$.
-The parameters returned by the simulated annealing were
-$m_{\!_X} = 0.2078$, $r_{\!_X} = 0.0759$ and $s = 4.815$.
-
-Fig.\,\ref{f:GJ} shows the simulation results with these parameters after 1500 time steps. 
-Fig.\,\ref{f:GJ}A indicates the colour code used for retinal cell position; dorsal retinal cells are indicated by green, nasal cells by red.
-Lightness gives position along the relevant axis so that cells at the dorsal-nasal corner are yellow and those at the ventral-temporal corner are black.
-Fig.\,\ref{f:GJ}E shows the expected layout of retinal ganglion cell axons on the tectum.
-The axons retain the colour of their originating cell on the retina.
-Nasal cell axons should be arranged caudally; dorsal axons laterally; axons should be evenly spaced and the pattern should span the entire tectum.
-Fig.\,\ref{f:GJ}, panels B--D show `fishnet' plots of the average location of the 4 branches of each axon (the axon's centroid) at three timepoints. 
-Fig.\,\ref{f:GJ}B indicates the randomised placement of axon branches at the rostral end of the tectum as specified by the initial conditions (Eq.\,\ref{e:ic}).
-In these plots, lines are drawn between axons that are expected to arrange themselves adjacently to one another by the end of the simulation. 
-The number of line crossings in this plot gives an indication of the disorder in the arrangement.
-Fig.\,\ref{f:GJ}G shows the positions of the individual branches (i.e.~growth cones) at $t=1500$ that contribute to the plot in Fig.\,\ref{f:GJ}D.
-By this time, the topological order matches that shown in Fig.\,\ref{f:GJ}E.
-Opposing gradients have sorted the axons along the two axes to form a grid of cells.
-The arrangement spans the full tectum due to competition between the growth cones (in the chemotaxis-only model with $m_{\!_X}=0$, axons cover only a fraction of the tectal tissue).
-Fig.\,\ref{f:GJ}F shows the two pattern metrics; $\epsilon$ (distance error) and $\eta$ (number of crossings).
-For the wildtype pattern, $\eta_t=0$, as in Fig.\,\ref{f:GJ}E. 
-Fig.\,\ref{f:GJ}H shows selected axons (one from the centre and one from each corner of the retina) and their position history.
-
-\subsection*{Surgical manipulations}
-
-After finding an optimum set of parameters for the wildtype arrangement, we tested the model with a number of simulations of surgical manipulations (for reviews, see \citet{udin_formation_1988} and \citet{goodhill_retinotectal_1999}).
-Figs.~\ref{f:GJsurg_grafts}A and \ref{f:GJsurg_grafts}B show the results of tectal graft rotation manipulations in which a square patch of tectal tissue is cut out and rotated $90\degree$ or $180\degree$ \citep{chung_observations_1978}.
-%
-To simulate the surgery, we rotated an $8\times8$ square of the tectum shown in Fig\,\ref{f:ex}D, carrying the ligand expressions with the rotation (resulting in Fig.\,\ref{f:trot90}, left column). 
-After the manipulation, we re-computed the tectal ligand expression gradients.
-Fig.\,\ref{f:trot90} demonstrates that the most significant change in the gradient is around the border of the graft, due to discontinuities in the ligand expression.
-The manipulated gradients, interacting with unchanged axon branches, led to a qualitative rotation of axons that were destined for the graft area, indicated by a 90$\degree$ or 180$\degree$ `twist' of the axon centroid colours.
-Long connecting lines in the fishnet show that a number of axons from near the edges of the tectum were drawn into the graft region. 
-The pattern metric, $\epsilon$, is shown in the right hand graphs of Fig.\,\ref{f:GJsurg_grafts}A \& B. 
-We plot $\epsilon$ (the distance error) for the entire map (blue) along with the error for the graft region (gold) and the surround region (red). 
-The overall distance error for the rotation manipulations (0.182 or 0.2752)  is greater than for the wildtype system (0.04). 
-The pattern within the graft exhibits greater average distance error than for the surrounding tissue.
-The system does not achieve a perfect reproduction of the target layout suggested by the experimental result (which is shown in the inset) but it is in qualitative agreement.
-%
-Fig.\,\ref{f:GJsurg_grafts}C shows the result of a similar manipulation in which two grafts are cut out of the tectum and swapped. 
-There is considerable disorder in the simulated pattern of axon centroids but a line of red/yellow agents that would been located caudally has advanced rostrally and a number of green agents are more caudal than in the wildtype arrangement, consistent with experimental observations.
-
-Fig.\,\ref{f:GJsurg_ablate}A shows the result of a retinal ablation, in which the temporal half of the retina is surgically removed and the remaining RGC axons are allowed to develop on the unmanipulated tectum.
-According to a pure chemoaffinity theory, the surviving retinal axons should retain their chemical labels and they would be expected to find their usual locations on the tectum, resulting in a tectum which is only half populated with axons.
-The experimental observation is that the axons arrange themselves across the whole tectum, retaining their topographic order (inset, \citet{attardi_preferential_1963,schmidt_expansion_1978}). 
-The fishnet plot of the simulation indicates that the competition in the system successfully `stretches out' the surviving axons to achieve a close representation of the experimentally observed result.
-In this figure, the right-hand graph shows the simulated position of each axon centroid on the rostro-caudal (R--C) tectal axis plotted against its position on the nasal-temporal (N--T) retinal axis in red triangles, with the experimental positions plotted in blue squares.
-%
-Fig.\,\ref{f:GJsurg_ablate}B shows the result of a tectal ablation.
-Here the retina is left un-modified, but the caudal half of the tectum is surgically removed.
-Experimental observation \citep{yoon_reorganization_1971,sharma_reformation_1972} shows that RGC axons `crush up' on the remaining rostral section, retaining topographic order and creating a complete map on the available tectal tissue.
-The model clearly replicates this result, indicated by both fishnet and tectal/retinal axes graphs, because the chemotaxis component of the model is based on balanced, relative interactions between opposing gradients, rather than absolute labels or expression measurements.
-%
-A related ablation experiment is the so-called `mismatch', in which the temporal half of the retina is ablated along with the caudal part of the tectum to which it normally maps, again challenging the pure chemoaffinity model \citep{horder_retention_1971}.
-The competitive differential chemotaxis model replicates this manipulation (Fig.\,\ref{f:GJsurg_ablate}C) for the same reason as in the simple tectal ablation of Fig.\,\ref{f:GJsurg_ablate}B; opposing gradients remain on the surviving tectum allowing RGC axons to find their relative ordering.
-%
-The final surgical manipulation which we investigated was the `compound eye' experiment, in which two nasal retina halves are fused together \citep{gaze_retino-tectal_1963,fawcett_retinotectal_1982}.
-Within the compound retina there exist matched pairs of retinal ganglion cells, one from each nasal half, whose receptor and ligand expressions cannot be distinguished.
-These form overlaid maps, each of which spans the entire tectum (Fig.\,\ref{f:GJsurg_ablate}D).
-
-\subsection*{Genetic manipulations}
-
-In the early 2000s, a series of genetic experiments added to the set of manipulations that a model of the retinotectal system would need to
-explain. 
-First, \citet{brown_topographic_2000} showed that a selective RGC knock-in of EphA3 receptors led to an interesting modulation of the maps found in the tectum. 
-For the strongest knock-in (a homozygote manipulation) two maps were formed, with knock-in cells finding termination zones more rostral than their wildtype retinal neighbours. 
-By enhancing the quantity of EphA receptors expressed by half of all retinotectal axons, they demonstrated the importance of retinal receptor expression for the arrangement of the tectal map. 
-In the weaker, heterozygote manipulation, the maps were still separated, but the separation `collapsed' towards the rostral end of the tectum (see \citet{brown_topographic_2000}, Fig\,5). 
-This led to a second study showing that with an additional knock-down of EphA4 receptors in \emph{all} RGCs, the retinal origin of the collapse point was shifted temporally and all heterozygote EphA3 knock-in cells were pushed even further rostrally (\citet{reber_relative_2004}, their Fig.\,3e).
-For the model to successfully reproduce these results, it should therefore: i)~Demonstrate a complete separation of maps for a `large' knock-in of EphA3; ii)~Show a separation of maps for a smaller knock-in of EphA3 that is greater for axons of nasal origin.
-iii)~For the smaller knock-in, a collapse point at 76\% of the N--T axis; iv)~Show a considerable rostral shift of cells for which EphA3 is knocked in and EphA4 is knocked down; and v)~Show a collapse point for the knock-in/knock-down cells and knock-down only cells at around 88\% of the N--T axis.
-
-An assumption made in these papers is that all EphA receptor sub-types contribute to a summed EphA expression level, $\sum\mathrm{EphA}$, which increases from a low level at the nasal side of the retina, to a high level at the temporal side.
-The precise contribution of each sub-type to the overall level is assumed not to be important. 
-In this model, receptor $r_0$ increases in the retina in a nasal to temporal direction (Fig.\,\ref{f:ex}B) and so we treated $r_0$ as being equivalent to $\sum\mathrm{EphA}$ and manipulated it to attempt a recreation of the \citet{brown_topographic_2000}~and \citet{reber_relative_2004}~experiments.
-In half of the RGCs, we knocked EphA3 in by adding to the value of $r_0$ in these cells either 0.8 (heterozygote EphA3 ki/+) or 3.2 (homozygote EphA3 ki/ki). 
-To include the additional heterozygote knock-down of EphA4 (EphA4 +/-) from \citet{reber_relative_2004}, we subtracted 1.2 from $r_0$ in \emph{all} RGCs.
-Fig.\,\ref{f:GJeph}A shows, for the single EphA3 knock-in (EphA3 ki/+), branch positions, a plot of axon centroids and final rostral-caudal position on the tectum versus the originating nasal-temporal retinal coordinate for time step 1500.
-The left panel shows branches coloured in the red/green duochrome scheme according to their origin location on the retina.
-In the middle panel, axon centroids of RGCs which were treated with the knock-in manipulation are coloured red, wildtype cells are blue. 
-The right hand panel mapping final R--C position versus N--T origin  (and its inset which shows the experimental result) follow the same red/blue colour scheme.
-The map of treated cells (red) was shifted slightly rostrally, and the gap between the red and blue curves became smaller for more temporal RGCs.
-However, it can't be described as a map collapse (the term used in \citet{brown_topographic_2000}) because the curves remain separated for all N--T values.
-Regardless of the value of the ratio between treated and untreated cells, treated temporal axons are drawn to a significantly more rostral position than those without the knock-in.
-A bootstrapped studentized t-test of the two sets of points indicates a low probability, $p<10^{-4}$, that any of them were drawn from statistically identical distributions.
-
-Fig.\,\ref{f:GJeph}B shows the same trio of graphs for the homozygote double knock-in, EphA3 ki/ki.
-Here, the treated axons were pushed further towards the rostral end of the tectum and the untreated axons were pushed caudally due to competition with the treated cells.
-The maps were completely separated. As in the experiment, there was no collapse point.
-The rostral movement of treated axons occurred due to their increased expression of EphA ($r_0$), which increased the simulated interaction with the rostral-caudal tectal ligand gradient. 
-The opposing receptor expression ($r_2$) was unchanged by the manipulation and so those axons with increased $r_0$ expression were pulled to a more rostral position than they would otherwise have taken.
-The caudal movement of wildtype axons, whose receptor expression in Figs.\,\ref{f:GJeph}A-B was unchanged, occurred due to competition with the most caudal of the rostrally-drawn treated axons.
-Overall, the correspondence between the experimental result and the simulated homozygous knock-in is close.
-
-Fig.\,\ref{f:GJeph}C shows the result for the \citet{reber_relative_2004} manipulation, in which the EphA expression was selectively enhanced (by EphA3 knock-in) in 50\% of cells and knocked down (by EphA4 knock-down) in all cells.
-As in Fig.\,\ref{f:GJeph}A, the red and blue lines are separated, but there is no map collapse ($p<10^{-4}$). 
-The significant rostral movement of the treated axons seen in \citet{reber_relative_2004}, where not a single treated axon was found more than 30\% along the rostral-caudal axis, was absent from this result.
-In fact, the knock-in/knock-down cells find locations slightly more \emph{caudal} than those in Fig.\,\ref{f:GJeph}A, because while $r_0$ is increased by 0.8 from the knock-in, the knock-down of 1.2 reduces these cells' interaction with tectal ephrinA ($L_0$) compared with those in the EphA3 knock-in/EphA4 wildtype simulation.
-
-It is important to note that these results emerge purely from modifications to the receptor expression levels; the model is otherwise unchanged. 
-The lack of collapse points or the more rostral positions of knock-in/knock-down cells suggests that there is an interaction effect between EphA4 and other EphA receptors that has not been captured by the competitive differential chemotaxis model.
-We address possible interaction effects in the next section.
 
@@
 
-The previous approach assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
+In the previous model, we assumed that all EphA receptor sub-types (EphA3, EphA4, EphA5, etc) transmit the same signal to induce indistinguishable function.
 This assumption is difficult to support, given that the effect caused by knock-\emph{in} of EphA3 (a shift of nasal RGC axons towards the rostral tectum) is enhanced by the knock-\emph{down} of EphA4.
-We therefore changed our assumption to the following: Most EphA receptors share the same function (EphA3, EphA5 etc) but EphA4 is different. 
-We redefined $r_{\!\scriptscriptstyle 0,b}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,\ref{f:clustermech}A), and introduced $r_{\!\scriptscriptstyle A4,b}$ as the expression of EphA4 receptors (Fig.\,\ref{f:clustermech}B).
-EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$, dropping the branch subscript.
+We therefore changed our assumption to the following: Most EphA receptors
+share the same function (EphA3, EphA5 etc) but EphA4 is different.
+
+We redefined $r_{\!\scriptscriptstyle 0}$ as the expression of EphA receptors (excluding EphA4) on branch $b$ (Fig.\,6A), and introduced $r_{\!\scriptscriptstyle A4}$ as the expression of EphA4 receptors (Fig.\,6B).
+EphA4 receptors are expressed in a spatially inhomogeneous pattern~\citep{reber_relative_2004} for which we choose the value $r_{\!\scriptscriptstyle A4} = 3.5$.
 We assume that cis-interactions between retinal ephrinA ligands and retinal EphA4 receptors lead to bound and un-bound EphA4 receptors \citep{hornberger_modulation_1999}.
 The number of cis-bound EphA4 receptors is assumed to be $r_{\!\scriptscriptstyle A4}^{\mathrm{cis}} = w_{\!\scriptscriptstyle A4} \times r_{\!\scriptscriptstyle A4} \times l_0$ where $w_{\!\scriptscriptstyle A4} = 0.153$ is a binding weight modelling the probability that an EphA4 receptor will bind to a nearby retinal ephrinA ligand.
-Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,\ref{f:clustermech}C).
+Remaining, free EphA4 receptors are $r_{\!\scriptscriptstyle A4}^{\mathrm{free}} = r_{\!\scriptscriptstyle A4} - r_{\!\scriptscriptstyle A4}^{\mathrm{cis}}$ (Fig.\,6C).
 
@@
 %
-Eq.\,\ref{e:G} was modified to include the cluster size:
+Eq.\,5 was modified to include the cluster size:
 %
 \begin{equation}\label{e:Gcs}
-\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i,b} \nabla L_i(\mathbf{x})
+\mathbf{G} = m_{\!_G}\,\sum_i^N F_i\,c_i\,r_{i} \nabla L_i(\mathbf{x})
 \end{equation}
 
-$c_i\,r_{i,b}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i,b}$ only for $i=0$ for which (dropping the branch subscript) $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
+$c_i\,r_{i}$ is the effective signal strength, $s_i$, for receptor $i$ and differs from $r_{i}$ only for $i=0$ for which $s_0 = r_0/r_{\!\scriptscriptstyle A4}$.
 Under this relationship, the signal strength changes as $r_0$ increases (simulating the knock-in of EphA3) and also as $r_{\!\scriptscriptstyle A4}$ decreases (simulating EphA4 knock-down).
-Fig.\,\ref{f:clustermech}C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
-The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,\ref{e:G}, $r_{i,b} \equiv s_i$).
-To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,\ref{f:GJ}D.
+Fig.\,6C shows $s_0$ under knock-in, knock-down and combined knock-in/knock-down conditions.
+The average magnitude of the signal $s_0$, as defined above, is smaller than $r_0$, which is the effective signal in the original model (in Eq.\,5, $r_{i} \equiv s_i$).
+To counterbalance this reduction in the magnitude of the receptor-ligand interaction between $r_0$ and $l_0$, we reduced the value of $r_2$: $r_2 \rightarrow r_2/2$ which ensures that the cluster size model reproduces the wildtype result as in Fig.\,2D.
 
-Fig\,\ref{f:clustermech}Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
+Fig\,6Di shows the logic that is assumed to be enacted by the EphA/EphA4 and $r_2$ receptors to implement the $r_2$ collapse mechanism.
 We assume that there is a coupling between $r_0$, $r_{A4}$ expression and $r_2$ expression.
 In a region where $r_0$ exceeds some threshold, $h_0$, and $r_{A4}$ is \emph{below} another threshold, $h_{A4}$, the effectiveness of $r_2$ is further reduced by a factor of five: $r_2 \rightarrow r_2/5$.
-Fig\,\ref{f:clustermech}Dii shows the regions in which the collapse occurs.
-In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,\ref{f:clustermech}A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
+Fig\,6Dii shows the regions in which the collapse occurs.
+In the wildtype case, even though $\sum$EphA expression exceeds $h_0$ in the region above 0.56 on the N--T axis (Fig\,6A, solid green), free EphA4 expression is always above the threshold $h_{A4}$, so $r_2$ collapse does not occur.
 If EphA4 is knocked down, then in the temporal region above 0.56, $\sum$EphA exceeds $h_0$ and EphA4 is below $h_{A4}$ and so $r_2$ collapse occurs here.
-If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,\ref{f:clustermech}A, purple), and so $r_2$ collapse occurs for any N--T origin.
-
-Fig.\,\ref{f:r2collapse}A shows the behaviour of the new model in the EphA3 knock-in condition.
-The additional repulsion away from the caudal tectum caused by the increase in $s_0$ leads to a rostral shift of knock-in cells (half of the RGCs, randomly selected, receive the EphA3 knock-in).
-The map shows a `collapse point' at about 0.8 of the nasal-temporal axis at which cells' rostral caudal position distributions begin to overlap, which results from the convergence of wildtype and knock-in signals for temporal RGCs (Fig.\,\ref{f:clustermech}C) and the reducing importance of the EphA signal for temporal RGCs (although temporal RGCs have high EphA/$r_0$ expression, the rostral tectum, to which they are guided, has low levels of tectal ephrinA).
-Fig.\,\ref{f:r2collapse}B demonstrates that this effect is increased for a stronger knock-in, resulting in two completely separated maps (and no collapse point), reproducing the successful result of the un-modified model (Fig.\,\ref{f:GJeph}B).
-The goal of reproducing the significant caudal shift for EphA3 ki/+ / EphA4 +/- cells is achieved because nasal RGCs have low free EphA4 due to the knock-down and increased $r_0$ expression (Fig.\,\ref{f:r2collapse}C).
-This results in an enhanced signal, which increases the rostral to caudal movement of nasal cells (See dual-colour dashed curve in Fig.\,\ref{f:clustermech}C).
-This mechanism alone accounts for a slight shift of cells to around 0.6 on the R--C axis, but does not explain the shift to around 0.2 seen in the inset to Fig.\,\ref{f:r2collapse}C.
-To reproduce this shift requires the additional $r_2$ collapse mechanism, driven by the combination of increased EphA expression and knocked down EphA4.
-Cells in Fig.\,\ref{f:r2collapse}C that have both the EphA3 knock-in and the EphA4 knock-down (those on the red curve) experience $r_2$ collapse regardless of their N--T origin.
-Cells that only have EphA4 knock-down experience $r_2$ collapse only if their origin is temporal (above 0.56 N--T).
-This leads to the wide gap between the red and blue curves for nasal origins and a much smaller difference for temporal origins, in good agreement with the \citet{reber_relative_2004} result (inset).
-The alteration of the chemotaxis gradient effect that opposes EphA-driven movement is necessary to achieve the substantial movement of the curve for selective knock-in cells.
-Fig.\,\ref{f:r2collapse}D shows the prediction of this model for a genetic manipulation in which only the EphA4 knock-down alteration was made.
-This gives a largely unperturbed map with a slight kink around 0.56 resulting from temporal $r_2$ collapse and an overall slight rostral shift of the map due to the effect of the EphA4 knock-down on the cluster size and hence the signal $s_0$.
-
-Taken together, the results presented in Fig.\,\ref{f:r2collapse} demonstrate remarkable reproduction of the four features of this set of genetic manipulation results.
-
-\section{Discussion}
-
-% Summarize the findings and place them in context
-Can the assumptions implicit in Sperry's chemoaffinity theory account for the behaviour of the axons that grow from the retina into the optic tectum?
-If not, does the addition of competition resolve the remaining problems? 
-Can chemotaxis and competition be implemented by invoking a signalling system that behaves like the known signalling molecules (Ephs/ephrins, Ryk/Wnt3 \& Neogenin/RGM)?
-To answer these questions, we developed a model combining a gradient-following form of chemotaxis (Eq.\,\ref{e:G}) together with axon-axon competition (Eq.\,\ref{e:X}) in which information between axonal growth cone agents was only transferred by the local interactions between their receptors and ligands.
-The model contained parameters to i) control the balance of competition and chemotaxis, ii) set the interaction range of the competition and iii) set a competitive interaction threshold. 
-These parameters were optimized so that the model would reproduce the experimentally observed topographic patterning of growth cones seen in healthy animals. 
-The model then correctly recreated a broad range of patterns that occur when healthy animals are subjected to surgical manipulations to disrupt neuronal development. 
-Qualitative rotation was observed in graft rotation simulations; the patch swap simulation was similarly qualitatively accurate. 
-Retinal and tectal ablation simulations produced highly convincing results, as did mismatch and compound eye simulations.
-
-However, the model did not appear to account for the genetic manipulations applied by \citet{brown_topographic_2000} and \citet{reber_relative_2004}.
-We identified four features of the genetic experimental results which should be reproduced and explained by the model. 
-These were, briefly, i) map duplication for selective EphA3 homozygous knock-in, ii) map collapse (at 0.76 N--T) into a partial duplication for a weaker, heterozygous EphA3 knock-in, iii) a big rostral shift of EphA3 knock-in/EphA4 knock-down cells and iv) a collapse point for this case at about 0.9 N--T.
-We then showed that although the unmodified model successfully reproduced the first feature, it failed with the other three.
-
-To extend the model to reproduce the genetic manipulations, we examined the receptor-ligand interactions that drive simulated axons in the R--C directions.
-First, we looked at mechanisms that might reproduce the `map collapse' features of the experiments, in which nasal RGCs with a knock-in of EphA3 find more rostral tectal positions than their non-knock-in neighbours (map separation), but for temporal RGCs with the same amount of knock-in, no separation occurs.
-A `collapse point' is seen marking a transition between these behaviours at around 80\% of the N--T axis.
-In our receptor-ligand gradient-following model, the collapse point could occur if, beyond 80\% of the N--T axis, the signal magnitude between wildtype and knock-in cells was smaller than in the region below 80\% of the N--T axis.
-We proposed a mechanism based on the necessity of EphA clusters for effective signal transmission and the reported ability of EphA4 to `side-bind' to EphA clusters \citep{nikolov_ephephrin_2013}, assuming that side-binding regulates cluster size.
-This cluster size mechanism reduces the signal difference between wildtype and EphA4 knock-in RGCs as the N--T axis is traversed and leads to a convergence of wildtype and knock-in cells for temporal RGCs.
-However, it didn't cause the large rostral shift of EphA4 knock-in/EphA4 knock-down cells. To reproduce this feature, we examined the receptor-ligand pair \emph{opposing} the effect of EphA/ephrinA. 
-A strong rostral shift of cells could be caused by increased vigour in the mechanism causing cells to move rostrally (the EphA/ephrinA pairing), or equally by a strong \emph{decrease} in the vigour of the mechanism that moves cells caudally ($r_2$/$L_2$ in this model).
-We suggested that a genetic logic operation based on comparing EphA and EphA4 receptor levels against thresholds could cause the collapse of the $r_2$/$L_2$ mediated gradient climbing and reproduce the strong rostral movement of knock-in/knock-down cells observed in \citet{reber_relative_2004}.
-We demonstrated that these two mechanisms could closely reproduce the genetic manipulation experiments and found that they do not alter the other results (wildtype map reproduction and surgical manipulations).
-
-\subsection*{Chemotaxis}
-
-The properties of gradient-following models of retinotectal development were explored by \citet{gierer_development_1981,gierer_model_1983,gierer_directional_1987}, where they were presented as a special case of a general class of \emph{potential models}.
-In a potential model, the interactions between the properties of a growing axon originating from a location $\mathbf{u}$ on the retina and its environment at location $\mathbf{x}$ on the tectum set up some potential function, $p(\mathbf{x},\mathbf{u})$.
-The axon then grows or moves to minimize $p$, which causes it to find its target location.
-This model successfully creates the wildtype map because each axon originates with a different $\mathbf{u}$ and has its minimum $p(\mathbf{x},\mathbf{u})$ in the correct target location.
-
-Because of the subsequent discovery of the ephrins \citep{cheng_complementary_1995,drescher_vitro_1995}, we chose to model a gradient-following effect directly, rather than working with potential functions.
-Earlier works (e.g. \citet{hope_arrow_1976}, from which the model of \citet{simpson_simple_2011} was developed) had discussed, but not simulated or comprehensively analysed gradient-following.
-This study thus offers a novel treatment of gradient-following, acting together with competition in a two-dimensional agent-based model.
-
-Although the minimum number of gradients required to provide a coordinate system on the surface of a tissue such as the tectum is $N=2$, Gierer recognised that a gradient-following system would require opposing effects to prevent axons from gathering at the edges of the tissue \citep{gierer_model_1983}.
-Some form of `stopping mechanism' is required to allow axons to find target locations across the whole tissue.
-Gierer showed that opposing gradients could fulfil this requirement, which in his model allow for a minimum in a potential function to occur at an arbitrary location on a two-dimensional tissue.
-Following Gierer, we chose to model \emph{opposing pairs of orthogonal gradients} with $N=4$, calling this \emph{differential chemotaxis}.
-With opposing gradients, an axon which interacts most strongly with a gradient in the caudal direction would be drawn to the caudal side of the tectum, whereas an axon whose strongest interaction was to climb a gradient in the rostral direction would be drawn to the rostral side.
-There is evidence that multiple gene expression gradients are found in the brain \citep{cang_developmental_2013}, although examples of opposing pairs acting on the N--T retinal axis are difficult to identify.
-Molecules involved in patterning axons along the N--T axis include the EphA receptors and their ligands, as well as repulsive guidance molecule (RGM, \citet{monnier_rgm_2002}) and engrailed-2 (En-2 \citet{brunet_transcription_2005}).
-RGM and its ligand, neogenin are expressed in patterns complimenting EphA/ephrinA---a low rostral to high caudal expression of RGM and ephrin A on the tectum and a low nasal to high temporal expression of EphA/neogenin in the retina.
-Both receptor-ligand pairs induce repulsion and so appear to fulfil the same function.
-Tantalisingly, En-2 is expressed in the opposite sense on the tectum; low caudal to high rostral expression.
-However, \citet{brunet_transcription_2005} report that temporal RGC axons are attracted to high expression of En-2, and so this is a \emph{third} mechanism which sends temporal axons rostrally and nasal axons towards caudal tectum.
-In the retinal D--V axis, there \emph{does} exist a candidate to act as opposition to the EphB/ephrinB receptor-ligand pairing, which drives axons medially on the tectum. 
-The morphogen Wnt3 is expressed on the tectum in a pattern similar to that of ephrinB \citep{schmitt_wntryk_2006}. 
-Two receptor classes have been identified; Frizzled, a receptor mediating attraction/stimulation, is expressed in a uniform distribution across the retina whereas Ryk is a repulsive receptor for Wnt3 and is expressed in a dorsal-to-ventral increasing gradient \citep{schmitt_wntryk_2006}.
-Ryk is a higher affinity receptor for Wnt3 than Frizzled, and thus dorsal RGCs that are low in Ryk tend to bind Frizzled to Wnt3 and are attracted medially on the tectum, whereas ventral RGCs tend to bind Ryk to Wnt3 and are repelled laterally.
-
-We note that it is possible to restrict $N$ to 2, as long as some other stopping mechanism is chosen. 
-\citet{naoki_revisiting_2017} did just this in a gradient model with two orthogonal gradients and a proposition for a stopping mechanism based on a switch within growth cones from repulsion to attraction at a critical level (not unlike that of the Wnt3/Ryk/Frizzled system).
-This work concentrated on the mechanism by which the repulsion/attraction switch could occur and did not model system level behaviour as we do here. 
-It is possible that Naoki's mechanism would be compatible with this model, replacing $\mathbf{G}$ in Eq.\,\ref{e:dX} and it would be interesting to examine this mechanism within the current agent-based modelling framework in a future study.
-
-\subsection*{Competition}
-
-Because many of the ligands expressed on the tectum are also expressed by retinal ganglion cells, there exists the possibility for inter-RGC signalling via receptor-ligand binding (forward interactions).
-We chose to model this kind of interaction between RGCs, noting that if a forward interaction on the tectum leads to growth cones moving away from a region of high ligand expression (ephrin-A/EphA interactions act in this way), then the equivalent axon-axon interaction would also be expected to be repulsive.
-An interaction between one branch dense with receptors and another with a large number of ligands would have a high chance of repelling; an interaction between branches each with low levels of receptor/ligand expression would have a low chance of repelling.
-
-Note that if there was a \emph{single} ligand type, $l_i$, expressed in the retina and its form was $1/r_i$, then $\mathbf{C}$ from Eq.\,\ref{e:dX} would be equivalent to Simpson \& Goodhill's $\mathbf{I}_b$, which is a threshold-activated signal based on the ratio of $r_{i,b}/r_{i,k}$.
-However, because we intended for $\mathbf{C}$ to be the only mechanism for competition, we found that the minimum number of ligands that were required to be expressed on the retina was 2, expressed in opposing directions (Fig.\,\ref{f:ex}B).
-If only a single receptor-ligand pair participated in competition, then many axons experienced no competitive interaction and such a system was unable to reproduce experimental observations.
-
-Competition appears to be necessary for all of the surgical manipulations. 
-It is most obviously useful for the ablation experiments, allowing axons to spread out over the full tectum even if half of the RGCs have been removed.
-It also assists with the graft rotation and graft swap manipulations by pushing axons that have been shifted by gradient discontinuities at the graft edges towards empty regions of tectum to which they would not otherwise be drawn by chemotaxis.
-
-The range of competition, $r_{\!_X}=0.07$, found by optimization, was similar to the final desired axon-axon spacing ($0.05$) and much larger than the typical radius of a growth cone.
-This is notable because at least Ephs and ephrins are membrane-bound molecules \citep{cang_developmental_2013} and are not expected to act over long ranges.
-If $r_{\!_X}$ was reduced to a value approximating the growth cone radius, the pattern would not form and the competitive interactions would act to disorder the map. 
-If set to a very large value (such as a quarter of the tectal width) the longer range interactions cause the centre of the pattern to become thinned out, with more growth cones collecting at the tissue boundary.
-This `interaction distance for competition' is similar to that found by \citet{simpson_simple_2011}. 
-It suggests not that the range of competitive interaction between growth cones should be many multiples of their diameter, but rather that there should be enough interacting growth cones so that the expected spacing between growth cones in the map is equivalent to the range over which they can send signals to each other.
-
-% Single axon behaviour
-\citet{gosse_retinotopic_2008} argued that competition is not required to form the retinotectal map in zebrafish.
-They showed that in a system with no competition (achieved by transplanting single RGCs into mutants that would otherwise have none), only minor changes in the tectal targeting of these cells would occur.
-The chemotaxis component of the model ensures that single axons will be drawn to approximately the correct location.
-The lack of a competitive effect results in axons that would be targeted towards the edge of the tectal tissue finding their way to slightly more central locations, leading to a compressed, but topographically accurate map.
-This is not inconsistent with \citet{gosse_retinotopic_2008}, who found some differences in targeting behaviour between the different experimental conditions.
-
-\subsection*{Attractive receptor-ligand interactions}
-
-In the results, we present a model in which all receptor-ligand interactions are repulsive.
-However, EphB/ephrinB interactions are largely attractive in nature.
-It is possible to explore the behaviour of a model incorporating a single attractive receptor-ligand interaction.
-When we did so, we found that, if we retained an exponential expression of retinal receptors, we had to change our assumption about the expression of tectal ligands, replacing the exponential expression with a logarithmic one.
-In this model, the form of the tectal ligand must be compatible with the form of the retinal receptor expression.
-Because we find that only one pair of repulsive receptor-ligand interactions on the retina was required to mediate competition, the other axis was free to be either repulsive, or attractive like the EphBs.
-
-\subsection*{Robustness to noise}
-
-To determine if the system is robust to the quantity of noise present in the biological system, we first estimate noise from gene expression data.
-Fig.\,\ref{f:expr_ephb} shows expression data for EphB and ephrinB in the chick retina and tectum.
-We used the tool \emph{Stalefish} \citep{james_comparing_2022} to extract quantitative data from previously published histology images.
-Fig.\,\ref{f:expr_ephb}A shows expression of EphB2 in retina at embyonic day 4 (using data presented in Fig.\,7 of \citet{holash_polarized_1995}).
-Stalefish is used to place two curves around the retinal wall such that the mean expression from the stain can be determined.
-The graphs show the expression values on a normalised scale. 
-We isolated a section of the expression that can be approximated by a linear function, obtained the best fit and determined the RMS error from the fit.
-The mean error was approximately 0.06 or 6\% of the full range of values and we made the assumption that half of this error was measurement noise and applied 3\% random variation to retinal receptor expression in the simulation.
-Figs.\,\ref{f:expr_ephb}C shows a similar noise magnitude for tectal ligand expression (with data from \citet{braisted_graded_1997}) and so we chose to apply the same amount of noise to tectal ligand expression and tectal ligand gradient measurements in the model.
-i) At each simulation step, we added noise to the measurement of the tectal ligand gradient, which had a typical range of 0--6.34 so that $\nabla L_i(\mathbf{x}) \rightarrow \nabla L_i(\mathbf{x}) + N(0,0.19)$ where $N(0,\sigma)$ indicates a number drawn from a normal distribution of mean 0 and standard deviation $\sigma$.
-ii) We added noise to the specification of the retinal receptor expression (whose range is 2.67; see Fig.\,\ref{f:clustermech}A) so that $r_i(x) \rightarrow r_i(x) + N(0, 0.08)$.
-iii) Finally, we add noise to the specification of the tectal ligand expression (range 2.67) so that $l_i(x) \rightarrow l_i(x) + N(0, 0.08)$.
-(ii) and (iii) are computed once only, at the start of the simulation.
-Fig.\,\ref{f:noise} shows the results of running the simulation with each of these noise contributions individually, and with all three combined. 
-It indicates that the gradient following system was highly robust to gradient measurement noise. The pattern was largely unaffected by 3\% random variation (Fig.\,\ref{f:noise}A). 
-The pattern remained coherent even with random noise 5 times higher than this (results not shown). However, systematic errors in retinal receptor expression were more disordering to the pattern. 
-Fig.\,\ref{f:noise}B indicates that, at simulation end (t=2000), the pattern was broadly present, but there was some local disorder in axon positions. 
-Tectal ligand expression noise was still more disordering (Fig.\,\ref{f:noise}C) and appeared to dominate the other forms of noise, as when all three noise types are introduced, the final values for the pattern metric $\epsilon$ were similar in Figs.\,\ref{f:noise}C and \ref{f:noise}D.
-
-%\subsection*{The value of $s$}
-%This may be worth discussing, because the system finds that a value of $s$ which guarantees interactions always occur.
-Essentially, any growth cones that meet will interact unless they originate from almost exactly the same location on the tectum.
-
-% Limitations/weaknesses of research
-\subsection*{Predictions and Limitations}
-
-We have presented an agent-based model which can be used to investigate mechanisms at play in two-dimensional gradient-climbing developmental systems of axonal growth. 
-We used it to show that a gradient-following model with competition is able to reproduce most of the non-genetic experimental results that have been carried out on the retinotectal projection, giving support for a system in which \emph{orthogonal pairs} of counter-acting signalling gradients guide growth cones towards their targets.
-By examining genetic manipulations of EphA3 and EphA4, we were able to predict that a combination of a cluster-size mechanism and a coupling between EphA3/4 and the corresponding opposing signalling molecules would be required to explain these results.
-We also predict that, if EphB is expressed in an exponential pattern across the retina, then its ligand ephrinB should be expressed in a logarithmically increasing expression pattern on the tectum.
-
-Limitations of the model include the low resolution ($20\times20$) of the specification of the ligand and receptor expressions, which resulted in coarse gradients.
-In the current implementation, the resolution of the grid related to the number of axons simulated; each grid element is a `source location' on the retina.
-Another limitation of this model is that it does not address the pattern of axon growth seen in chicks and rodents, by which axons grow the full length of the tectum, then make branches at an appropriate location along the axon.
-Instead, it models development of simple growth cone agents which find their way to their targets as observed in frogs and fish \citep{oleary_molecular_1999}.
-Adapting this framework to address an axon branching model is left for future work.
-
-% Restate significant findings and implications
-
-\section{Method details}
-
-\subsection*{Border effect} % border effect description
-The border effect ($\mathbf{B}$ in Eq.\,\ref{e:dX}) was based on gradient-following by assuming that there is some other molecular signal which acts on all branches near the boundary of the tectal tissue.
-For a branch with position $(x,y)$, $\mathbf{B}$ is given by:
-%
-\begin{equation}\label{e:B}
-B_x = \begin{cases}
-        \frac{1}{2} (r-x)     & x<r \\
-        \frac{1}{2} (1-r-x)    & x>1-r
-\end{cases}
-B_y = \begin{cases}
-        \frac{1}{2} (r-y)      & y<r \\
-        \frac{1}{2} (1-r-y)    & y>1-r
-\end{cases}
-\end{equation}
-%
-Thus $\mathbf{B}$ is equivalent to the action of a repulsive signalling molecule expressed around the border of the tectum, whose expression increases quadratically outside the tectum and affects any branch touching (or outside) the boundary.
-% end border effect description
-
-\subsection*{Initial conditions}
-Branches were initially randomly distributed in a stripe at the rostral side of the tectum at the start of each simulation. 
-Each RGC axon $k$ was assigned a random initial position coordinate:
-\begin{equation}\label{e:ic}
-\mathbf{x}_{k,t=0} = (U(0,1), U(-0.2,0))
-\end{equation}
-where U(p,q) is a number selected from a random uniform distribution in the range $[p,q)$.
-Each of the 4 branches per RGC axon was given its parent axon's initial position, plus a randomly generated offset with coordinates derived from a normal distribution of mean 0 and standard deviation 0.1.
-% end initial conditions
-
-\subsection*{Bootstrap analyses of R--C tectal position vs.~N--T retinal origin}
-
-To test whether a manipulation affected the mean rostro-caudal position as a function of the origin location of the nasal-temporal retinal axis, we applied a \emph{Studentized, bootstrapped test of the equality of means} of two distributions of R--C positions obtained for the $n=20$ different dorso-ventral origins for each N--T location.
-The test follows Algorithm 16.2 of~\cite{efron_introduction_1993} and has the null hypothesis, H$_0$: \emph{the means are the same}.
-Consider two samples, $\textbf{z}$ and $\textbf{y}$ (both of size $n$); which are the positions obtained for the wildtype and manipulated cells of Fig.\,\ref{f:GJeph}.
-The \emph{observed value} of this test, $t(\textbf{x})$, is a Studentized (meaning the variances are accounted for) difference of the means of $\textbf{z}$ and $\textbf{y}$, given by:
-%
-$$t(\textbf{x}) =  \frac{\bar{z} - \bar{y}}{\sqrt{\sigma_{z}^2/n + \sigma_{y}^2/n} } $$
-%
-where $\textbf{x}$ is the combined sample formed by joining $\textbf{z}$ and $\textbf{y}$, $\sigma_z$ \& $\sigma_y$ are the standard deviations of $\textbf{z}$ \& $\textbf{y}$ and $\bar{z}$ \& $\bar{y}$ are the arithmetic means of $\textbf{z}$ \& $\textbf{y}$.
-
-A set of re-samples is now made from $\textbf{z}$ and $\textbf{y}$ \emph{after applying a transformation that assumes the null hypothesis is true}.
-The transformations are defined as:
-%
-$$ \tilde{\textbf{z}} = \textbf{z} - \bar{z} + \bar{x}; \quad \tilde{\textbf{y}} = \textbf{y} - \bar{y} + \bar{x} $$
-%
-where $\bar{x}$ is the mean of $\textbf{x}$. This shifts $\textbf{z}$ and $\textbf{y}$ to force their means to be equal.
-$\textbf{z}^*$ and $\textbf{y}^*$ are individual resamples from $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$. 
-The Studentized difference of the means of the resamples is computed:
-
-$$t(\textbf{x}^*) =  \frac{\bar{z}^* - \bar{y}^*}{\sqrt{\sigma_{z}^{*2}/n + \sigma_{y}^{*2}/n} } $$
-
-If the original means of $\textbf{z}$ and $\textbf{y}$ were genuinely very close, then $\tilde{\textbf{z}}$ and $\tilde{\textbf{y}}$ won't have been shifted very much and it is likely that $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$ with probability around 0.5.
-If they were not close, and the mean($\textbf{z}$) $\gg$ mean($\textbf{y}$), then very few $t(\textbf{x}^*)$ will exceed $t(\textbf{x})$. 
-We made 10000 $\textbf{x}^*$ resamples; the proportion of those for which $t(\textbf{x}^*) \geq t(\textbf{x})$ is the \emph{Achieved Significance Level} (ASL). 
-The smaller the ASL is, the less probable is H$_0$, and the more significant is the difference of the means.
-
-The test makes no assumption about the shape of the distributions which generated the samples, but it does assume that $\bar{z} > \bar{y}$.
-%% End bootstrap desc
-
-\section{Data Availability}
-
-All data and code required to reproduce the results in this work are available in the public repository 
-{https://github.com/sebjameswml/RetinoTectal}.
-The software was built using the library morphologica \citep{james_morphologica_2022}.
-
-\section{Acknowledgements}
-
-This work was supported by grant 220020516 from the James S. McDonnell Foundation.
+If, in addition to EphA4 knock-down, EphA3 is knocked in, $\sum$EphA expression rises above $h_0$ across the entire N--T axis (Fig\,6A, purple), and so $r_2$ collapse occurs for any N--T origin.
 

### 2024-07-05 misc/examineRetEph.py

@@ -211,6 +211,11 @@ def examineRetEph():
     ax22.plot (x, collapse2_ki,    linestyle='-.', color=clr_knockin)
     ax22.plot (x, collapse2_kikd,    linestyle='-.', color=clr_knockdown)
 
+    ax22.plot (x, EphAx,           linestyle='-', color=clr_wt)
+    ax22.plot (x, EphA4_free,           linestyle='-', color=clr_wt)
+    ax22.plot (x, (EphAx+ki),      linestyle='-', color=clr_knockin)
+    ax22.plot (x, (EphA4_free_kd),      linestyle='-', color=clr_knockdown)
+
     ax22.legend([filled_line_wt, filled_line_ki, (dotted_line1, dotted_line2)],
                 ['Coll. WT','$Coll. EphA ki', 'Coll. EphA ki/kd'])
 
